---
title: "I2 - integrated land crop model - ARID conditions"
author: "Andreas Angourakis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---


## General preparation

Define paths:

```{r}
data_dir <- "simdata"
plots_dir <- "plots"
```

Get file names in simulation data directory:
```{r}
file_names <- paste0(data_dir, "/", list.files(data_dir))
```

Define lookup functions to access data between tables using identifiers or 'keys':

```{r}
getRelatedVariableWithKey <- function(keyInCallingTable, keyInCalledTable, variableInCalledTable)
{
  return(
    sapply(keyInCallingTable, function(x) variableInCalledTable[match(x, keyInCalledTable)])
  )
}

getRelativeVariableWithTwoKeys <- function(firstKeyInCallingTable, secondKeyInCallingTable, 
                                       firstKeyInCalledTable, secondKeyInCalledTable, variableInCalledTable)
{
  temp <- firstKeyInCallingTable
  
  for (i in 1:length(firstKeyInCallingTable))
  {
    temp[i] <- variableInCalledTable[firstKeyInCalledTable == firstKeyInCallingTable & secondKeyInCalledTable == secondKeyInCallingTable]
  }
  
  return(temp)
}
```

Load crop table (assumed to always have the same name):

```{r}
cropTable <- read.csv(paste0(data_dir, "/I2_yield-exp_cropTable.csv"))
```

Reorder crops for a more logical display:

WARNING: doing this manually makes the script sensitive to the value of crop-selection in NetLogo. Thus, this chunk must be updated if crop-selection is changed before running experiments.

```{r}
cropTable$crop <- factor(cropTable$crop, levels = c("proso millet", "pearl millet", "rice", "barley", "wheat 1", "wheat 2"))
```

## Terrain example 1: no river

### Experiment definition

First set any parameters to be used as identifiers of the experiment batches to retrieved:

```{r}
terrainRandomSeed = 0
```

### Load data

Load the terrain data in the "yield" folder:

```{r}
terrainData <- read.csv(paste0(data_dir, "/I2_yield-exp_terrainRandomSeed=", terrainRandomSeed, ".csv"))
```

Load the parameters data:

```{r}
parsData <- read.csv(paste0(
  data_dir, 
  "/I2_yield-exp_pars_terrainRandomSeed=", 
  terrainRandomSeed, 
  "_type-of-experiment=user-defined_experiment-name=exp3.csv"))
```

Load all experiment output files in the "yield" folder, generated with this version (I2):

```{r}
nameFilter <- grepl(paste0(
  "I2_yield-exp_terrainRandomSeed=", 
  terrainRandomSeed, 
  "_"
  ), file_names)

yieldData <- do.call(rbind, lapply(file_names[nameFilter], read.csv))
```

Force the same order of crops used by cropTable:

```{r}
yieldData$crop <- factor(yieldData$crop, levels = levels(cropTable$crop))
```

### Preparations

Get vector of colours to represent crops:

```{r}
cropColours <- rainbow(nlevels(cropTable$crop), s = 0.8, v = 0.8, end = 0.85)
```

Precalculate ranges of ARID and yield:

```{r}
minARID = round(min(c(yieldData$p_soil_meanARID, yieldData$p_soil_meanARID_grow), na.rm = TRUE), digits = 2)
maxARID = round(max(c(yieldData$p_soil_meanARID, yieldData$p_soil_meanARID_grow), na.rm = TRUE), digits = 2)

minYield = round(min(yieldData$yield), digits = -1)
maxYield = round(max(yieldData$yield), digits = -1)

minPrecipitation_yearlyMean = round(min(parsData$precipitation_yearlyMean, na.rm = TRUE), digits = -2)
maxPrecipitation_yearlyMean = round(max(parsData$precipitation_yearlyMean, na.rm = TRUE), digits = -2)

minPrecipitation_dailyCum_plateauValue_yearlyMean = round(min(parsData$precipitation_dailyCum_plateauValue_yearlyMean, na.rm = TRUE), digits = 2)
maxPrecipitation_dailyCum_plateauValue_yearlyMean = round(max(parsData$precipitation_dailyCum_plateauValue_yearlyMean, na.rm = TRUE), digits = 2)
```

### Summary statistics per crop

```{r}
yieldData_summary <- tapply(as.numeric(as.character(yieldData$yield)), yieldData$crop, summary)

yieldData_summary <- data.frame(Reduce(rbind, yieldData_summary), row.names = names(yieldData_summary))
```

```{r}
knitr::kable(yieldData_summary)
```

```{r}
plotName = paste0(plots_dir, "I2/yieldPerCrop_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

par(mar = c(6,5,1,1))

boxplot(yield ~ factor(crop), data = yieldData,#[yieldData$yield > 0,], # show only non-zero yield 
        ylab = expression(paste("yield (", g/m^2, ")")), xlab = "",
        las = 2, 
        col = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

### Effect of precipitation on ARID

Plotting ARID vs precipitation yearly mean:

```{r}
plotName = paste0("plots/I2/ARIDvsPrecipitation_yearlyMean_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")


for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_yearlyMean),
         yieldData$p_soil_meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(p_soil_meanARID ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              p_soil_meanARID = yieldData[yieldData$crop == aCrop, "p_soil_meanARID"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_yearlyMean),
         yieldData$p_soil_meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(p_soil_meanARID_grow ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              p_soil_meanARID_grow = yieldData[yieldData$crop == aCrop, "p_soil_meanARID_grow"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean, by = 50),
     tck = 0, lwd = 0)

mtext("precipitation yearly mean (mm)", side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water, " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Plotting ARID vs precipitation seasonal balance (plateau value):

```{r}
plotName = paste0(plots_dir, "I2/ARIDvsPrecipitation_dailyCum_plateauValue_yearlyMean_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_dailyCum_plateauValue_yearlyMean),
         yieldData$p_soil_meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(p_soil_meanARID ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              p_soil_meanARID = yieldData[yieldData$crop == aCrop, "p_soil_meanARID"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_dailyCum_plateauValue_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_dailyCum_plateauValue_yearlyMean),
         yieldData$p_soil_meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(p_soil_meanARID_grow ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              p_soil_meanARID_grow = yieldData[yieldData$crop == aCrop, "p_soil_meanARID_grow"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_dailyCum_plateauValue_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean), c(0, 1), 
     ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean, by = 0.1),
     tck = 0, lwd = 0)

mtext("precipitation plateau value (i.e. winter / summer) yearly mean (mm)", side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[order(match(cropTable$crop, levels(cropTable$crop)))], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

### Effect of ARID on yield

Plotting absolute values of yield vs ARID:

```{r}
plotName = paste0(plots_dir, "I2/yieldVsARID_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 3), nrow = 2, ncol = 2, byrow = FALSE),
       widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minARID, maxARID), 
     c(minYield, maxYield),
     xlab = "mean ARID (current year)", 
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$p_soil_meanARID[yieldData$crop == aCrop],
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ p_soil_meanARID, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minARID, maxARID), 
     c(minYield, maxYield),
     xlab = "mean ARID (grow season)",
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$p_soil_meanARID_grow[yieldData$crop == aCrop],
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ p_soil_meanARID_grow, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[order(match(cropTable$crop, levels(cropTable$crop)))], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$p_soil_meanARID_grow[yieldData$crop == aCrop], y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ p_soil_meanARID_grow, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

### Yield by crop and precipitation conditions

Plotting absolute values of yield vs precipitation yearly mean:

```{r}
plotName = paste0(plots_dir, "I2/yieldVsprecipitation_yearlyMean_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2), nrow = 1, ncol = 2, byrow = FALSE),
       widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean), 
     c(minYield, maxYield),
     xlab = "precipitation yearly mean (mm)",
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable = parsData$precipitation_yearlyMean),
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              yield = yieldData[yieldData$crop == aCrop, "yield"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[order(match(cropTable$crop, levels(cropTable$crop)))], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable = parsData$precipitation_yearlyMean), 
        y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              yield = yieldData[yieldData$crop == aCrop, "yield"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_yearlyMean))))

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

Plotting absolute values of yield vs precipitation seasonal balance (plateau value):

```{r}
plotName = paste0(plots_dir, "I2/yieldVsprecipitation_dailyCum_plateauValue_yearlyMean_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2), nrow = 1, ncol = 2, byrow = FALSE),
       widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean), 
     c(minYield, maxYield),
     xlab = "precipitation plateau value (i.e. winter / summer) yearly mean (mm)",
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_dailyCum_plateauValue_yearlyMean),
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              yield = yieldData[yieldData$crop == aCrop, "yield"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_dailyCum_plateauValue_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = levels(yieldData$crop), 
       title = "Crop-cultivar", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                      keyInCalledTable = parsData$randomSeed, 
                                      variableInCalledTable =parsData$precipitation_dailyCum_plateauValue_yearlyMean), 
        y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ precipitation_yearlyMean,
                    data = data.frame(cbind(
                      yield = yieldData[yieldData$crop == aCrop, "yield"],
                      precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                           parsData$randomSeed, 
                                                                           parsData$precipitation_dailyCum_plateauValue_yearlyMean))))

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Terrain example 2: river

### Experiment definition

First set any parameters to be used as identifiers of the experiment batches to retrieved:

```{r}
terrainRandomSeed = 35
```

### Load data

Load the terrain data in the "yield" folder:

```{r}
terrainData <- read.csv(paste0(data_dir, "/I2_yield-exp_terrainRandomSeed=", terrainRandomSeed, ".csv"))
```

Load the parameters data:

```{r}
nameFilter <- grepl(paste0("I2_yield-exp_pars_terrainRandomSeed=", terrainRandomSeed, "_"), file_names)

parsData <- read.csv(file_names[nameFilter])
```

Load all experiment output files in the "yield" folder, generated with this version (I2):

```{r}
nameFilter <- grepl(paste0("I2_yield-exp_terrainRandomSeed=", terrainRandomSeed, "_"), file_names)

yieldData <- do.call(rbind, lapply(file_names[nameFilter], read.csv))
```

Force the same order of crops used by cropTable:

```{r}
yieldData$crop <- factor(yieldData$crop, levels = levels(cropTable$crop))
```

### Preparations

Get vector of colours to represent crops:

```{r}
cropColours <- rainbow(nlevels(cropTable$crop), s = 0.8, v = 0.8, end = 0.85)
```

Precalculate ranges of ARID and yield:

```{r}
minARID = round(min(c(yieldData$p_soil_meanARID, yieldData$p_soil_meanARID_grow), na.rm = TRUE), digits = 2)
maxARID = round(max(c(yieldData$p_soil_meanARID, yieldData$p_soil_meanARID_grow), na.rm = TRUE), digits = 2)

minYield = round(min(yieldData$yield), digits = -1)
maxYield = round(max(yieldData$yield), digits = -1)

minPrecipitation_yearlyMean = round(min(parsData$precipitation_yearlyMean, na.rm = TRUE), digits = -2)
maxPrecipitation_yearlyMean = round(max(parsData$precipitation_yearlyMean, na.rm = TRUE), digits = -2)

minPrecipitation_dailyCum_plateauValue_yearlyMean = round(min(parsData$precipitation_dailyCum_plateauValue_yearlyMean, na.rm = TRUE), digits = 2)
maxPrecipitation_dailyCum_plateauValue_yearlyMean = round(max(parsData$precipitation_dailyCum_plateauValue_yearlyMean, na.rm = TRUE), digits = 2)
```

### Summary statistics per crop

```{r}
yieldData_summary <- tapply(as.numeric(as.character(yieldData$yield)), yieldData$crop, summary)

yieldData_summary <- data.frame(Reduce(rbind, yieldData_summary), row.names = names(yieldData_summary))
```

```{r}
knitr::kable(yieldData_summary)
```

```{r}
plotName = paste0(plots_dir, "I2/yieldPerCrop_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

par(mar = c(6,5,1,1))

boxplot(yield ~ factor(crop), data = yieldData,#[yieldData$yield > 0,], # show only non-zero yield 
        ylab = expression(paste("yield (", g/m^2, ")")), xlab = "",
        las = 2, 
        col = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

### Effect of precipitation on ARID

Plotting ARID vs precipitation yearly mean:

```{r}
plotName = paste0(plots_dir, "I2/ARIDvsPrecipitation_yearlyMean_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")


for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_yearlyMean),
         yieldData$p_soil_meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(p_soil_meanARID ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              p_soil_meanARID = yieldData[yieldData$crop == aCrop, "p_soil_meanARID"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_yearlyMean),
         yieldData$p_soil_meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(p_soil_meanARID_grow ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              p_soil_meanARID_grow = yieldData[yieldData$crop == aCrop, "p_soil_meanARID_grow"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean, by = 50),
     tck = 0, lwd = 0)

mtext("precipitation yearly mean (mm)", side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[order(match(cropTable$crop, levels(cropTable$crop)))], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Plotting ARID vs precipitation seasonal balance (plateau value):

```{r}
plotName = paste0(plots_dir, "I2/ARIDvsPrecipitation_dailyCum_plateauValue_yearlyMean_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_dailyCum_plateauValue_yearlyMean),
         yieldData$p_soil_meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(p_soil_meanARID ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              p_soil_meanARID = yieldData[yieldData$crop == aCrop, "p_soil_meanARID"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_dailyCum_plateauValue_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_dailyCum_plateauValue_yearlyMean),
         yieldData$p_soil_meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(p_soil_meanARID_grow ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              p_soil_meanARID_grow = yieldData[yieldData$crop == aCrop, "p_soil_meanARID_grow"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_dailyCum_plateauValue_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean), c(0, 1), 
     ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean, by = 0.1),
     tck = 0, lwd = 0)

mtext("precipitation plateau value (i.e. winter / summer) yearly mean (mm)", side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[order(match(cropTable$crop, levels(cropTable$crop)))], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

### Effect of ARID on yield

Plotting absolute values of yield vs ARID:

```{r}
plotName = paste0(plots_dir, "I2/yieldVsARID_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 3), nrow = 2, ncol = 2, byrow = FALSE),
       widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minARID, maxARID), 
     c(minYield, maxYield),
     xlab = "mean ARID (current year)", 
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$p_soil_meanARID[yieldData$crop == aCrop],
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ p_soil_meanARID, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minARID, maxARID), 
     c(minYield, maxYield),
     xlab = "mean ARID (grow season)",
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$p_soil_meanARID_grow[yieldData$crop == aCrop],
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ p_soil_meanARID_grow, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[order(match(cropTable$crop, levels(cropTable$crop)))], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$p_soil_meanARID_grow[yieldData$crop == aCrop], y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ p_soil_meanARID_grow, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

### Yield by crop and precipitation conditions

Plotting absolute values of yield vs precipitation yearly mean:

```{r}
plotName = paste0(plots_dir, "I2/yieldVsprecipitation_yearlyMean_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2), nrow = 1, ncol = 2, byrow = FALSE),
       widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean), 
     c(minYield, maxYield),
     xlab = "precipitation yearly mean (mm)",
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable = parsData$precipitation_yearlyMean),
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              yield = yieldData[yieldData$crop == aCrop, "yield"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[order(match(cropTable$crop, levels(cropTable$crop)))], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable = parsData$precipitation_yearlyMean), 
        y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              yield = yieldData[yieldData$crop == aCrop, "yield"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_yearlyMean))))

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

Plotting absolute values of yield vs precipitation seasonal balance (plateau value):

```{r}
plotName = paste0(plots_dir, "I2/yieldVsprecipitation_dailyCum_plateauValue_yearlyMean_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2), nrow = 1, ncol = 2, byrow = FALSE),
       widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_dailyCum_plateauValue_yearlyMean, maxPrecipitation_dailyCum_plateauValue_yearlyMean), 
     c(minYield, maxYield),
     xlab = "precipitation plateau value (i.e. winter / summer) yearly mean (mm)",
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable =parsData$precipitation_dailyCum_plateauValue_yearlyMean),
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ precipitation_yearlyMean,
            data = data.frame(cbind(
              yield = yieldData[yieldData$crop == aCrop, "yield"],
              precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                   parsData$randomSeed, 
                                                                   parsData$precipitation_dailyCum_plateauValue_yearlyMean)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = levels(yieldData$crop), 
       title = "Crop-cultivar", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                      keyInCalledTable = parsData$randomSeed, 
                                      variableInCalledTable =parsData$precipitation_dailyCum_plateauValue_yearlyMean), 
        y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ precipitation_yearlyMean,
                    data = data.frame(cbind(
                      yield = yieldData[yieldData$crop == aCrop, "yield"],
                      precipitation_yearlyMean = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                           parsData$randomSeed, 
                                                                           parsData$precipitation_dailyCum_plateauValue_yearlyMean))))

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

### Yield by crop and river flow and length conditions

Plotting absolute values of yield vs precipitation yearly mean:

```{r}
plotName = paste0(plots_dir, "I2/yieldVsriverWaterPerFlowAccumulation_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2), nrow = 1, ncol = 2, byrow = FALSE),
       widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean), 
     c(minYield, maxYield),
     xlab = "precipitation yearly mean (mm)",
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable = parsData$riverWaterPerFlowAccumulation),
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ riverWaterPerFlowAccumulation,
            data = data.frame(cbind(
              yield = yieldData[yieldData$crop == aCrop, "yield"],
              riverWaterPerFlowAccumulation = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                        parsData$randomSeed, 
                                                                        parsData$riverWaterPerFlowAccumulation)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[order(match(cropTable$crop, levels(cropTable$crop)))], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                      keyInCalledTable = parsData$randomSeed, 
                                      variableInCalledTable = parsData$riverWaterPerFlowAccumulation), 
        y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ riverWaterPerFlowAccumulation,
                    data = data.frame(cbind(
                      yield = yieldData[yieldData$crop == aCrop, "yield"],
                      riverWaterPerFlowAccumulation = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                                parsData$randomSeed, 
                                                                                parsData$riverWaterPerFlowAccumulation))))

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

Plotting absolute values of yield vs river length (r:

```{r}
plotName = paste0(plots_dir, "I2/yieldVsriverWaterPerFlowAccumulation_terrainRandomSeed=", terrainRandomSeed, ".png")

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2), nrow = 1, ncol = 2, byrow = FALSE),
       widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minPrecipitation_yearlyMean, maxPrecipitation_yearlyMean), 
     c(minYield, maxYield),
     xlab = "precipitation yearly mean (mm)",
     ylab = expression(paste("yield (", g/m^2, ")")))

for (aCrop in levels(yieldData$crop))
{
  points(getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                   keyInCalledTable = parsData$randomSeed, 
                                   variableInCalledTable = terrainData$),
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ riverWaterPerFlowAccumulation,
            data = data.frame(cbind(
              yield = yieldData[yieldData$crop == aCrop, "yield"],
              riverWaterPerFlowAccumulation = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                        parsData$randomSeed, 
                                                                        parsData$riverWaterPerFlowAccumulation)))),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[order(match(cropTable$crop, levels(cropTable$crop)))], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = getRelatedVariableWithKey(keyInCallingTable =  yieldData$randomSeed[yieldData$crop == aCrop], 
                                      keyInCalledTable = parsData$randomSeed, 
                                      variableInCalledTable = parsData$riverWaterPerFlowAccumulation), 
        y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ riverWaterPerFlowAccumulation,
                    data = data.frame(cbind(
                      yield = yieldData[yieldData$crop == aCrop, "yield"],
                      riverWaterPerFlowAccumulation = getRelatedVariableWithKey(yieldData$randomSeed[yieldData$crop == aCrop], 
                                                                                parsData$randomSeed, 
                                                                                parsData$riverWaterPerFlowAccumulation))))

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```
