---
title: "Simple crop model with multipcropping - plotting land unit yield"
author: "Andreas Angourakis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---


```{r}
require(stringr)
```

## Load data

Load all experiment output files in the "yield" folder, generated with this version ("Simple crop model").

```{r}
file_names <- paste0("yield/", list.files("yield"))

isThisVersion <- grepl("SIMPLE-crop-model_withMulticropping_yield", file_names)

yieldData <- do.call(rbind, lapply(file_names[isThisVersion], read.csv))
```

## Preparations

Reconstruct cropTable content:

```{r}
cropTableNames <- c("crop", "T_sum", "HI", "I_50A", "I_50B", "T_base", "T_opt", "RUE", "I_50maxH", "I_50maxW", "T_heat", "T_extreme", "S_CO2", "S_water", "sowingDay", "harvestDay")

cropTable <- data.frame(matrix(ncol=length(cropTableNames),nrow=0, dimnames=list(NULL, cropTableNames)))

for (aCrop in levels(yieldData$crop))
{
  cropTable <- rbind(cropTable, yieldData[match(aCrop, yieldData$crop), cropTableNames])
}
```

Get vector of colours to represent crops:

```{r}
cropColours <- rainbow(nlevels(cropTable$crop), s = 0.8, end = 0.9)
```

extract map dimensions:

```{r}
minX = min(yieldData$x)
maxX = max(yieldData$x)
mapWidth = maxX - minX + 1

minY = min(yieldData$y)
maxY = max(yieldData$y)
mapHeight = maxY - minY + 1
```

## Mean yield per crop

Build spatial yield matrices:

```{r}
meanYield_byPosAndCrop <- list()

for (aCrop in levels(cropTable$crop))
{
  meanYield_byPosAndCrop[[aCrop]] <- matrix(rep(NA, mapWidth * mapHeight), nrow = mapHeight, ncol = mapWidth)
  
  for (i in minX:maxX)
  {
    for (j in minY:maxY)
    {
      meanYield_byPosAndCrop[[aCrop]][j+1, i+1] <- mean(yieldData$yield[yieldData$crop == aCrop & yieldData$x == i & yieldData$y == j])
    }
  }
}

```

Plot mean yield per crop in grid map:

```{r}
cleanCropName <- function(x) trimws(gsub("[[:punct:]]", "", x), 'l')

plotName = paste0("plots/SIMPLE-crop-model_withMulticropping/meanYieldRaster_byCrop.png")

grScale = 2
  
png(plotName, width = grScale * 200 * nlevels(cropTable$crop), height = grScale * 200)

layout(matrix(1:(nlevels(cropTable$crop)), nrow = 1, byrow = FALSE))

par(cex.main = 1.5 * grScale, cex.axis = 0.6 * grScale)

for (aCrop in levels(cropTable$crop))
{
  colgrp <- findInterval(c(meanYield_byPosAndCrop[[aCrop]]),
                         seq(min(c(meanYield_byPosAndCrop[[aCrop]])),
                             max(c(meanYield_byPosAndCrop[[aCrop]])),
                             length.out = 10))
  colfunc <- colorRampPalette(c("white", cropColours[match(aCrop, cropTable$crop)]))
  collist <- colfunc(length(unique(colgrp))) 
  
  image(meanYield_byPosAndCrop[[aCrop]],
        col = collist,
        axes = FALSE, xaxt='n', yaxt='n', ann=FALSE)
  
  title(main = cleanCropName(aCrop))
  
  axis(2, at = seq(0, 1, length.out = mapHeight), labels = minY:maxY)
  axis(1, at = seq(0, 1, length.out = mapWidth), labels = minX:maxX)
}

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Alternative using data frame structure (instead of separate matrices) and plotting using more elemental base graphic functions. This alternative is equivalent to the previous one but takes longer to process.

```{r}
meanYield_byPosAndCrop_v2 <- list(
  crop = c(),
  x = c(),
  y = c(),
  meanYield = c()
)

for (aCrop in levels(cropTable$crop))
{
  for (i in minX:maxX)
  {
    for (j in minY:maxY)
    { 
      meanYield_byPosAndCrop_v2$crop <- c(meanYield_byPosAndCrop_v2$crop, aCrop)
      meanYield_byPosAndCrop_v2$x <- c(meanYield_byPosAndCrop_v2$x, i)
      meanYield_byPosAndCrop_v2$y <- c(meanYield_byPosAndCrop_v2$y, j)
      
      meanYield_byPosAndCrop_v2$meanYield <- c(meanYield_byPosAndCrop_v2$meanYield,
                                            mean(yieldData$yield[yieldData$crop == aCrop & yieldData$x == i & yieldData$y == j]))
    }
  }
}

meanYield_byPosAndCrop_v2 <- as.data.frame(meanYield_byPosAndCrop_v2)
```


```{r}
cleanCropName <- function(x) trimws(gsub("[[:punct:]]", "", x), 'l')

plotName = paste0("plots/SIMPLE-crop-model_withMulticropping/meanYieldRaster_byCrop_v2.png")

grScale = 2
  
png(plotName, width = grScale * 200 * nlevels(cropTable$crop), height = grScale * 200)

layout(matrix(1:(nlevels(cropTable$crop)), nrow = 1))

for (aCrop in levels(cropTable$crop))
{
  colgrp <- findInterval(meanYield_byPosAndCrop_v2$meanYield[meanYield_byPosAndCrop_v2$crop == aCrop],
                         seq(min(meanYield_byPosAndCrop_v2$meanYield[meanYield_byPosAndCrop_v2$crop == aCrop]),
                             max(meanYield_byPosAndCrop_v2$meanYield[meanYield_byPosAndCrop_v2$crop == aCrop]),
                             length.out = 10))
  colfunc <- colorRampPalette(c("white", cropColours[match(aCrop, cropTable$crop)]))
  collist <- colfunc(length(unique(colgrp))) 
  
  par(cex.main = 1.5 * grScale, cex.axis = 0.6 * grScale)
  
  plot(NA, ylim = c(0.5, mapHeight + 0.5), xlim = c(0.5, mapWidth + 0.5), ann = FALSE, axes = FALSE)
  
  title(main = cleanCropName(aCrop))
  
  axis(2, at = minY:maxY, labels = minY:maxY)
  axis(1, at = minX:maxX, labels = minX:maxX)

  symbols(meanYield_byPosAndCrop_v2$x[meanYield_byPosAndCrop_v2$crop == aCrop], 
          meanYield_byPosAndCrop_v2$y[meanYield_byPosAndCrop_v2$crop == aCrop],
          squares = rep(1, mapHeight * mapWidth),
          add = TRUE,
          inches = FALSE,
          fg = collist[colgrp],
          bg = collist[colgrp])
}

dev.off()


```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Alternative using `heatmap`: 
NOTE: this function is incompatible with `layout` so plots must be generated separately; it is also hard to customise because of many inaccessible defaults which are set for biostatistics (i.e. genetic data, with lateral dendrograms)).

```{r}

cleanCropName <- function(x) gsub("[[:punct:]]", "", x)

for (aCrop in levels(cropTable$crop))
{
  plotName = paste0("plots/SIMPLE-crop-model_withMulticropping/meanYieldRaster_", cleanCropName(aCrop), ".png")

  grScale = 2
  
  png(plotName, width = grScale * 500, height = grScale * 300)
  
  par(mar = c(3,1,10,1), cex.main = grScale)
  heatmap(meanYield_byPosAndCrop[[aCrop]], 
        Rowv = NA, Colv = NA, 
        main = aCrop,
        labRow = minY:maxY, labCol = minX:maxX,
        cexRow = 0.8 * grScale, cexCol = 0.8 * grScale)
  
  dev.off()
}
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

