---
title: "Simple crop model - analysis of yield experiments"
author: "Andreas Angourakis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}

```

## Load data

Load all experiment output files in the "yield" folder, generated with this version ("Simple crop model").

```{r}
file_names <- paste0("yield/", list.files("yield"))

isThisVersion <- grepl("SIMPLE-crop-model_yield", file_names)

yieldData <- do.call(rbind, lapply(file_names[isThisVersion], read.csv))
```

Preview the data structure

```{r}
knitr::kable(head(yieldData))
```

## All crops

Visualise simulation results for all crops included in Zhao et al. 2019 (except those taking more than a year to reach maturity; i.e. banana, cotton, peanut and cassava). NOTE: banana is already not used for simulations.

## Preparations

Reconstruct cropTable content:

```{r}
cropTableNames <- c("crop", "T_sum", "HI", "I_50A", "I_50B", "T_base", "T_opt", "RUE", "I_50maxH", "I_50maxW", "T_heat", "T_extreme", "S_CO2", "S_water", "sowingDay", "harvestDay")

cropTable <- data.frame(matrix(ncol=length(cropTableNames),nrow=0, dimnames=list(NULL, cropTableNames)))

for (aCrop in levels(yieldData$crop))
{
  cropTable <- rbind(cropTable, yieldData[match(aCrop, yieldData$crop), cropTableNames])
}
```

Filter out cassava, cotton and peanut:

```{r}
notAnnualCrops <- c(" 'cassava'", " 'cotton'", " 'peanut'")
notAnnual <- yieldData$crop == " 'cassava'" | yieldData$crop == " 'cotton'" | yieldData$crop == " 'peanut'"

yieldData <- yieldData[!notAnnual,]

# crop factor variable needs resetting levels
yieldData$crop <- factor(yieldData$crop)
```

Get vector of colours to represent crops:

```{r}
cropColours <- rainbow(nlevels(yieldData$crop), s = 0.8, end = 0.9)
```

## Impact of weather variables

### Temperature

```{r}
temperatureParNames <- c("temperature_annualMaxAt2m","temperature_annualMinAt2m",
                         "temperature_meanDailyFluctuation",
                         "temperature_dailyLowerDeviation","temperature_dailyUpperDeviation")
temperatureParUnits <- c("(C)", "(C)", "(C)", "(C)", "(C)")

plotName = "plots/yieldVsTemperaturePars.png"

grScale = 2
fontRescale = 0
fontRescaleDay = 0

nPlotRows = nlevels(yieldData$crop)
nPlotColumns = length(temperatureParNames)

png(plotName, width = grScale * 500, height = grScale * 500)

layout(matrix(1:((nPlotRows + 1) * (nPlotColumns + 1)), ncol = nPlotColumns + 1, byrow = T),
       widths = c(3, rep(10, nPlotColumns)),
       heights = c(rep(10, nPlotRows), 3))

# plot rows

for (aCrop in levels(yieldData$crop))
{
  # row title
  par(mar = c(0, 0, 0, 0))
  plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
  text(x = 0.5, y = 0.5, font = 4, 
       cex = grScale * (0.6 + fontRescale), srt = 90,
       labels = aCrop)
  # plots
  par(mar = c(1, 1, 0.1, 0.1), 
      cex.axis = grScale * (0.6 + fontRescale))
  for (aParName in temperatureParNames)
  {
    xaxt = ifelse(aParName == temperatureParNames[1],'s', 'n')
    yaxt = ifelse(aCrop == nlevels(yieldData$crop),'s', 'n')
    plot(yieldData[yieldData$crop == aCrop, aParName], yieldData$yield[yieldData$crop == aCrop],
         xlab = "", xaxt = xaxt, ylab = "", yaxt = yaxt)
  }
}

# last row
par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

for (aParName in temperatureParNames)
{
  plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
  text(x = 0.5, y = 0.5, font = 4, 
     cex = grScale * (0.6 + fontRescale), 
     labels = paste(aParName, temperatureParUnits[match(aParName, temperatureParNames)]))
}

dev.off()

```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

### Solar radiation

```{r}
solarParNames <- c("solar_annualMax", "solar_annualMin", "solar_meanDailyFluctuation")
solarParUnits <- rep(expression(paste("    Solar\nRadiation (", MJ/m^-2, ")")), 3)

plotName = "plots/yieldVsSolarPars.png"

grScale = 2
fontRescale = 0
fontRescaleDay = 0

nPlotRows = nlevels(yieldData$crop)
nPlotColumns = length(solarParNames)

png(plotName, width = grScale * 500, height = grScale * 500)

layout(matrix(1:((nPlotRows + 1) * (nPlotColumns + 1)), ncol = nPlotColumns + 1, byrow = T),
       widths = c(3, rep(10, nPlotColumns)),
       heights = c(rep(10, nPlotRows), 3))

# plot rows

for (aCrop in levels(yieldData$crop))
{
  # row title
  par(mar = c(0, 0, 0, 0))
  plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
  text(x = 0.5, y = 0.5, font = 4, 
       cex = grScale * (0.6 + fontRescale), srt = 90,
       labels = aCrop)
  # plots
  par(mar = c(1, 1, 0.1, 0.1), 
      cex.axis = grScale * (0.6 + fontRescale))
  for (aParName in solarParNames)
  {
    xaxt = ifelse(aParName == solarParNames[1],'s', 'n')
    yaxt = ifelse(aCrop == nlevels(yieldData$crop),'s', 'n')
    plot(yieldData[yieldData$crop == aCrop, aParName], yieldData$yield[yieldData$crop == aCrop],
         xlab = "", xaxt = xaxt, ylab = "", yaxt = yaxt)
  }
}

# last row
par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

for (aParName in solarParNames)
{
  plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
  text(x = 0.5, y = 0.5, font = 4, 
     cex = grScale * (0.6 + fontRescale), 
     labels = paste(aParName, solarParUnits[match(aParName, solarParNames)]))
}

dev.off()

```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

### Precipitation

```{r}
precipitationParNames <- c(
  "precipitation_yearlyMean", "precipitation_yearlySd",
  "precipitation_dailyCum_plateauValue_yearlyMean", "precipitation_dailyCum_plateauValue_yearlySd",
  "precipitation_dailyCum_inflection1_yearlyMean", "precipitation_dailyCum_inflection1_yearlySd", 
  "precipitation_dailyCum_rate1_yearlyMean", "precipitation_dailyCum_rate1_yearlySd",
  "precipitation_dailyCum_inflection2_yearlyMean", "precipitation_dailyCum_inflection2_yearlySd", 
  "precipitation_dailyCum_rate2_yearlyMean", "precipitation_dailyCum_rate2_yearlySd"
  )
precipitationParUnits <- c("(mm)", "(mm)", rep("", length(precipitationParNames) - 2))

plotName = "plots/yieldVsPrecipitation.png"

grScale = 2
fontRescale = 0
fontRescaleDay = 0

nPlotRows = nlevels(yieldData$crop)
nPlotColumns = length(precipitationParNames)

png(plotName, width = grScale * 500, height = grScale * 500)

layout(matrix(1:((nPlotRows + 1) * (nPlotColumns + 1)), ncol = nPlotColumns + 1, byrow = T),
       widths = c(3, rep(10, nPlotColumns)),
       heights = c(rep(10, nPlotRows), 3))

# plot rows

for (aCrop in levels(yieldData$crop))
{
  # row title
  par(mar = c(0, 0, 0, 0))
  plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
  text(x = 0.5, y = 0.5, font = 4, 
       cex = grScale * (0.6 + fontRescale), srt = 90,
       labels = aCrop)
  # plots
  par(mar = c(1, 1, 0.1, 0.1), 
      cex.axis = grScale * (0.6 + fontRescale))
  for (aParName in precipitationParNames)
  {
    xaxt = ifelse(aParName == precipitationParNames[1],'s', 'n')
    yaxt = ifelse(aCrop == nlevels(yieldData$crop),'s', 'n')
    plot(yieldData[yieldData$crop == aCrop, aParName], yieldData$yield[yieldData$crop == aCrop],
         xlab = "", xaxt = xaxt, ylab = "", yaxt = yaxt)
  }
}

# last row
par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

for (aParName in precipitationParNames)
{
  plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
  text(x = 0.5, y = 0.5, font = 4, 
     cex = grScale * (0.6 + fontRescale), 
     labels = paste(aParName, precipitationParUnits[match(aParName, precipitationParNames)]))
}

dev.off()

```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

