---
title: "Simple crop model with spatial diversity - analysis of yield experiments"
author: "Andreas Angourakis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

## Load data

Load all experiment output files in the "yield" folder, generated with this version ("Simple crop model with spatial diversity").

```{r}
file_names <- paste0("yield/", list.files("yield"))

isThisVersion <- grepl("SIMPLE-crop-model_withSpatialDiversity_yield", file_names)

yieldData <- do.call(rbind, lapply(file_names[isThisVersion], read.csv))
```

Preview the data structure

```{r}
knitr::kable(head(yieldData))
```

## All crops

Visualise simulation results for all crops included in Zhao et al. 2019 (except those taking more than a year to reach maturity; i.e. banana, cotton, peanut and cassava). NOTE: banana is already not used for simulations.

## Preparations

Reconstruct cropTable content:

```{r}
cropTableNames <- c("crop", "T_sum", "HI", "I_50A", "I_50B", "T_base", "T_opt", "RUE", "I_50maxH", "I_50maxW", "T_heat", "T_extreme", "S_CO2", "S_water", "sowingDay", "harvestDay")

cropTable <- data.frame(matrix(ncol=length(cropTableNames),nrow=0, dimnames=list(NULL, cropTableNames)))

for (aCrop in levels(yieldData$crop))
{
  cropTable <- rbind(cropTable, yieldData[match(aCrop, yieldData$crop), cropTableNames])
}
```

Filter out cassava, cotton and peanut:

```{r}
notAnnualCrops <- c(" 'cassava'", " 'cotton'", " 'peanut'")
notAnnual <- yieldData$crop == " 'cassava'" | yieldData$crop == " 'cotton'" | yieldData$crop == " 'peanut'"

yieldData <- yieldData[!notAnnual,]

# crop factor variable needs resetting levels
yieldData$crop <- factor(yieldData$crop)
```

Get vector of colours to represent crops:

```{r}
cropColours <- rainbow(nlevels(yieldData$crop), s = 0.8, end = 0.9)
```

Precalculate ranges of ARID and yield:

```{r}
minARID = round(min(c(yieldData$meanARID, yieldData$meanARID_grow), na.rm = TRUE), digits = 2)
maxARID = round(max(c(yieldData$meanARID, yieldData$meanARID_grow), na.rm = TRUE), digits = 2)

minYield = round(min(yieldData$yield), digits = -1)
maxYield = round(max(yieldData$yield), digits = -1)

minElevation = round(min(yieldData$elevation, na.rm = TRUE), digits = -1)
maxElevation = round(max(yieldData$elevation, na.rm = TRUE), digits = -1)

minDC = round(min(yieldData$DC, na.rm = TRUE), digits = 2)
maxDC = round(max(yieldData$DC, na.rm = TRUE), digits = 2)

minWHC = round(min(yieldData$WHC, na.rm = TRUE), digits = 2)
maxWHC = round(max(yieldData$WHC, na.rm = TRUE), digits = 2)

minAlbedo = round(min(yieldData$albedo, na.rm = TRUE), digits = 2)
maxAlbedo = round(max(yieldData$albedo, na.rm = TRUE), digits = 2)
```

## Summary statistics per crop

```{r}
yieldData_summary <- tapply(as.numeric(as.character(yieldData$yield)), yieldData$crop, summary)

yieldData_summary <- data.frame(Reduce(rbind, yieldData_summary), row.names = names(yieldData_summary))
```

```{r}
knitr::kable(yieldData_summary)
```

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/yieldPerCrop.png"

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

par(mar = c(6,5,1,1))

boxplot(yield ~ factor(crop), data = yieldData,#[yieldData$yield > 0,], # show only non-zero yield 
        ylab = expression(paste("yield (", g/m^2, ")")), xlab = "",
        las = 2, 
        col = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

## Effect of elevation on ARID

Plotting ARID vs elevation:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/ARIDvsElevation.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minElevation, maxElevation),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$elevation[yieldData$crop == aCrop],
         yieldData$meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID ~ elevation, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minElevation, maxElevation),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$elevation[yieldData$crop == aCrop],
         yieldData$meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID_grow ~ elevation, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minElevation, maxElevation), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minElevation, maxElevation, by = 5),
     tck = 0, lwd = 0)

mtext("elevation (m)", side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[!cropTable$crop %in% notAnnualCrops], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$elevation[yieldData$crop == aCrop], y = yieldData$meanARID_grow[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(meanARID_grow ~ elevation, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Effect of drainage coefficient (DC) on ARID

Plotting ARID vs DC:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/ARIDvsDC.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minDC, maxDC),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$DC[yieldData$crop == aCrop],
         yieldData$meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID ~ DC, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minDC, maxDC),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$DC[yieldData$crop == aCrop],
         yieldData$meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID_grow ~ DC, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minDC, maxDC), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minDC, maxDC, by = 0.05),
     tck = 0, lwd = 0)

mtext(expression("drainage coefficient " (m^3*m^-3) ), side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[!cropTable$crop %in% notAnnualCrops], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$DC[yieldData$crop == aCrop], y = yieldData$meanARID_grow[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(meanARID_grow ~ DC, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Effect of water holding capacity (WHC) on ARID

Plotting ARID vs DC:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/ARIDvsWHC.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minWHC, maxWHC),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$WHC[yieldData$crop == aCrop],
         yieldData$meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID ~ WHC, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minWHC, maxWHC),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$WHC[yieldData$crop == aCrop],
         yieldData$meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID_grow ~ WHC, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minWHC, maxWHC), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minWHC, maxWHC, by = 0.05),
     tck = 0, lwd = 0)

mtext(expression("water holding capacity " (m^3*m^-3) ), side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[!cropTable$crop %in% notAnnualCrops], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$WHC[yieldData$crop == aCrop], y = yieldData$meanARID_grow[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(meanARID_grow ~ WHC, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Effect of albedo on ARID

Plotting ARID vs DC:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/ARIDvsAlbedo.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minAlbedo, maxAlbedo),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$albedo[yieldData$crop == aCrop],
         yieldData$meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID ~ albedo, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minAlbedo, maxAlbedo),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$albedo[yieldData$crop == aCrop],
         yieldData$meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID_grow ~ albedo, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minAlbedo, maxAlbedo), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minAlbedo, maxAlbedo, by = 0.05),
     tck = 0, lwd = 0)

mtext("albedo", side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[!cropTable$crop %in% notAnnualCrops], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$albedo[yieldData$crop == aCrop], y = yieldData$meanARID_grow[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(meanARID_grow ~ albedo, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Effect of albedo on yield

Plotting ARID vs DC:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/yieldvsAlbedo.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2), nrow = 1, ncol = 2), widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minAlbedo, maxAlbedo),
     c(minYield, maxYield),
     xlab = "albedo",
     ylab = "yield")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$albedo[yieldData$crop == aCrop],
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ albedo, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[!cropTable$crop %in% notAnnualCrops], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$albedo[yieldData$crop == aCrop], y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ albedo, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Wheat vs rice

### Preparations

Create filter for wheat and rice:

```{r}
wheatAndRice <- c(" 'wheat 1'", " 'wheat 2'", " 'rice'")

isWheatAndRice <- yieldData$crop == " 'wheat 1'" | yieldData$crop == " 'wheat 2'" | yieldData$crop == " 'rice'"
```

Apply filter:

```{r}
yieldData <- yieldData[isWheatAndRice,]

# crop factor variable needs resetting levels
yieldData$crop <- factor(yieldData$crop)
```

Get vector of colours to represent crops:

```{r}
cropColours <- rainbow(nlevels(yieldData$crop), s = 0.8, end = 0.9)
```

Precalculate ranges of ARID and yield:

```{r}
minARID = round(min(c(yieldData$meanARID, yieldData$meanARID_grow), na.rm = TRUE), digits = 2)
maxARID = round(max(c(yieldData$meanARID, yieldData$meanARID_grow), na.rm = TRUE), digits = 2)

minYield = round(min(yieldData$yield), digits = -1)
maxYield = round(max(yieldData$yield), digits = -1)

minElevation = round(min(yieldData$elevation, na.rm = TRUE), digits = -1)
maxElevation = round(max(yieldData$elevation, na.rm = TRUE), digits = -1)

minDC = round(min(yieldData$DC, na.rm = TRUE), digits = 2)
maxDC = round(max(yieldData$DC, na.rm = TRUE), digits = 2)

minWHC = round(min(yieldData$WHC, na.rm = TRUE), digits = 2)
maxWHC = round(max(yieldData$WHC, na.rm = TRUE), digits = 2)

minAlbedo = round(min(yieldData$albedo, na.rm = TRUE), digits = 2)
maxAlbedo = round(max(yieldData$albedo, na.rm = TRUE), digits = 2)
```

## Summary statistics per crop

```{r}
yieldData_summary <- tapply(as.numeric(as.character(yieldData$yield)), yieldData$crop, summary)

yieldData_summary <- data.frame(Reduce(rbind, yieldData_summary), row.names = names(yieldData_summary))
```

```{r}
knitr::kable(yieldData_summary)
```

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/yieldPerCrop_wheatVsRice.png"

grScale = 2
fontRescale = 0
fontRescaleDay = 0

png(plotName, width = grScale * 500, height = grScale * 300)

par(mar = c(6,5,1,1))

boxplot(yield ~ factor(crop), data = yieldData,#[yieldData$yield > 0,], # show only non-zero yield 
        ylab = expression(paste("yield (", g/m^2, ")")), xlab = "",
        las = 2, 
        col = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

## Effect of elevation on ARID

Plotting ARID vs elevation:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/ARIDvsElevation_wheatVsRice.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minElevation, maxElevation),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$elevation[yieldData$crop == aCrop],
         yieldData$meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID ~ elevation, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minElevation, maxElevation),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$elevation[yieldData$crop == aCrop],
         yieldData$meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID_grow ~ elevation, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minElevation, maxElevation), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minElevation, maxElevation, by = 5),
     tck = 0, lwd = 0)

mtext("elevation (m)", side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[cropTable$crop %in% wheatAndRice], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$elevation[yieldData$crop == aCrop], y = yieldData$meanARID_grow[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(meanARID_grow ~ elevation, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Effect of drainage coefficient (DC) on ARID

Plotting ARID vs drainage coefficient:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/ARIDvsDC_wheatVsRice.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minDC, maxDC),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$DC[yieldData$crop == aCrop],
         yieldData$meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID ~ DC, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minDC, maxDC),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$DC[yieldData$crop == aCrop],
         yieldData$meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID_grow ~ DC, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minDC, maxDC), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minDC, maxDC, by = 0.05),
     tck = 0, lwd = 0)

mtext(expression("drainage coefficient " (m^3*m^-3) ), side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[cropTable$crop %in% wheatAndRice], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$DC[yieldData$crop == aCrop], y = yieldData$meanARID_grow[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(meanARID_grow ~ DC, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Effect of water holding capacity (WHC) on ARID

Plotting ARID vs water holding capacity:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/ARIDvsWHC_wheatVsRice.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minWHC, maxWHC),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$WHC[yieldData$crop == aCrop],
         yieldData$meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID ~ WHC, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minWHC, maxWHC),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$WHC[yieldData$crop == aCrop],
         yieldData$meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID_grow ~ WHC, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minWHC, maxWHC), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minWHC, maxWHC, by = 0.05),
     tck = 0, lwd = 0)

mtext(expression("water holding capacity " (m^3*m^-3) ), side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[cropTable$crop %in% wheatAndRice], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$WHC[yieldData$crop == aCrop], y = yieldData$meanARID_grow[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(meanARID_grow ~ WHC, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Effect of albedo on ARID

Plotting ARID vs albedo:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/ARIDvsAlbedo_wheatVsRice.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2, 3, 4, 4, 4), nrow = 3, ncol = 2),
       heights = c(10, 10, 2), widths = c(10, 3))

par(mar = c(1,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minAlbedo, maxAlbedo),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (current year)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$albedo[yieldData$crop == aCrop],
         yieldData$meanARID[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID ~ albedo, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

plot(c(minAlbedo, maxAlbedo),
     c(minARID, maxARID),
     xaxt ='n',
     xlab = "",
     ylab = "mean ARID (grow seasons)")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$albedo[yieldData$crop == aCrop],
         yieldData$meanARID_grow[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(meanARID_grow ~ albedo, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(1, 1, 1, 0.1))

plot(c(minAlbedo, maxAlbedo), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

axis(3, 
     at = seq(minAlbedo, maxAlbedo, by = 0.05),
     tck = 0, lwd = 0)

mtext("albedo", side = 1, line = -2)

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[cropTable$crop %in% wheatAndRice], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$albedo[yieldData$crop == aCrop], y = yieldData$meanARID_grow[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(meanARID_grow ~ albedo, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

## Effect of albedo on yield

Plotting yield vs albedo:

```{r}
plotName = "plots/SIMPLE-crop-model_withSpatialDiversity/yieldvsAlbedo_wheatVsRice.png"

grScale = 2
fontRescale = 0.5

png(plotName, width = grScale * 500, height = grScale * 300)

layout(matrix(c(1, 2), nrow = 1, ncol = 2), widths = c(10, 3))

par(mar = c(5,5,1,1), cex.lab = 0.8 * grScale)

plot(c(minAlbedo, maxAlbedo),
     c(minYield, maxYield),
     xlab = "albedo",
     ylab = "yield")

for (aCrop in levels(yieldData$crop))
{
  points(yieldData$albedo[yieldData$crop == aCrop],
         yieldData$yield[yieldData$crop == aCrop],
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         pch = 20
  )
  abline(lm(yield ~ albedo, data = yieldData[yieldData$crop == aCrop,]),
         col = cropColours[match(aCrop, levels(yieldData$crop))],
         lwd = 2)
}

par(mar = c(0, 0, 0, 0), cex = 0.8 * grScale)

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

legend(x = 0, y = 1, 
       legend = stringi::stri_c(cropTable$S_water[cropTable$crop %in% wheatAndRice], " (", levels(yieldData$crop), ")"), 
       title = "S_water (Crop-cultivar)", 
       fill = cropColours)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Calculate correlation and linear models fitness:

```{r}
linearModels.table <- data.frame(
  levels(yieldData$crop),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop)),
  numeric(nlevels(yieldData$crop))
)

names(linearModels.table) <- c("crop", "correlation_Pearson", "intercept", "intercept_p", "speed", "speed_p", "adjrsquared")

for (aCrop in levels(yieldData$crop))
{
  linearModels.table$correlation_Pearson[linearModels.table$crop == aCrop] <- 
    cor(x = yieldData$albedo[yieldData$crop == aCrop], y = yieldData$yield[yieldData$crop == aCrop], use = "pairwise.complete.obs")
  
  linearModel <- lm(yield ~ albedo, data = yieldData[yieldData$crop == aCrop,])

  linearModels.table$intercept[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,1]
  linearModels.table$intercept_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[1,4]
  linearModels.table$speed[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,1]
  linearModels.table$speed_p[linearModels.table$crop == aCrop] <- summary(linearModel)$coefficients[2,4]
  linearModels.table$adjrsquared[linearModels.table$crop == aCrop] <- summary(linearModel)$adj.r.squared
}
```

```{r}
knitr::kable(linearModels.table)
```

