---
title: "Household Demography model"
output: html_notebook
---

## loading simulation data

```{r}
endStates <- read.csv("output/3-householdDemography v.1.1 - kinship tabu for new couples exp-endstates-table.csv", skip = 6)
```

## Survival of populations

```{r}
hist(endStates$totalIndividuals)
```

### Initial household population

This parameter has slight positive effect on population survival:

```{r}
summary(endStates$initialNumHouseholds[endStates$totalIndividuals == 0])
summary(endStates$initialNumHouseholds[endStates$totalIndividuals > 0])
```

```{r}
boxplot(endStates$initialNumHouseholds ~ (endStates$totalIndividuals > 0))
```

### Mortality factors

Coale-Demeny Life Tables Model level parameter has great effect:

```{r}
summary(endStates$cdmlt.level[endStates$totalIndividuals == 0])
summary(endStates$cdmlt.level[endStates$totalIndividuals > 0])
```

```{r}
boxplot(endStates$cdmlt.level ~ (endStates$totalIndividuals > 0))
```

but region parameter has a smaller effect. East tends to survive more often.  

```{r}
table(endStates$coale.demeny.region, endStates$totalIndividuals > 0)
```

```{r}
boxplot(data = endStates, totalIndividuals ~ coale.demeny.region)
```

### Fertility factors

All have a visible effect, but particularly c1.fert and sigma1.fert. 

```{r}
layout(matrix(1:4, nrow = 2, byrow = T))
boxplot(endStates$c1.fert ~ (endStates$totalIndividuals > 0), main = "c1.fert")
boxplot(endStates$mu.fert ~ (endStates$totalIndividuals > 0), main = "mu.fert")
boxplot(endStates$sigma1.fert ~ (endStates$totalIndividuals > 0), main = "sigma1.fert")
boxplot(endStates$sigma2.fert ~ (endStates$totalIndividuals > 0), main = "sigma2.fert")
```

### Nuptiality factors

Residential rule showing no clear effect:

```{r}
table(endStates$residence.rule, endStates$totalIndividuals > 0)
```

The effect of the acceptable kinship degree is also unclear: 

```{r}
boxplot(endStates$acceptable.kinship.degree.for.couples ~ (endStates$totalIndividuals > 0))
```

All nuptiality parameters for both men and women present a small, yet apparently consistent effect. c1 and sigma parameters have a positive effect while mu has a smaller negative effect.

```{r}
layout(matrix(1:8, nrow = 2, byrow = T))
boxplot(endStates$c1.women ~ (endStates$totalIndividuals > 0), main = "c1.women")
boxplot(endStates$mu.women ~ (endStates$totalIndividuals > 0), main = "mu.women")
boxplot(endStates$sigma1.women ~ (endStates$totalIndividuals > 0), main = "sigma1.women")
boxplot(endStates$sigma2.women ~ (endStates$totalIndividuals > 0), main = "sigma2.women")
boxplot(endStates$c1.men ~ (endStates$totalIndividuals > 0), main = "c1.men")
boxplot(endStates$mu.men ~ (endStates$totalIndividuals > 0), main = "mu.men")
boxplot(endStates$sigma1.men ~ (endStates$totalIndividuals > 0), main = "sigma1.men")
boxplot(endStates$sigma2.men ~ (endStates$totalIndividuals > 0), main = "sigma2.men")
```

## Timing of survival

```{r}
hist(endStates$X.step.)
```

```{r}
table(endStates$X.step. > 0)
```

```{r}
plot(endStates$initialNumHouseholds ~ endStates$X.step.)
```


