---
title: "Demostration of annual sinusoid curve with fluctuations of the Weather model"
author: "Andreas Angourakis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---

## Load source file {-}

```{r}
source("source/weatherModel.R")
```

## Functions {-}

Generic functions for generating sinusoid curves with fluctuations depending on minValue, maxValue, length of year in days, and hemisphere (north/south) and fluctuation.

```{r}
getAnnualSinusoidWithFluctuation
```
```{r}
getDayValueInAnnualSinusoidWithFluctuation
```
```{r}
getDayValueInAnnualSinusoid
```

## Demonstration {-}

Set up six variations of parameter settings (i.e. minValue, maxValue, southHemisphere and fluctuation), assuming length of year of 365 days:

```{r}
yearLengthInDays_sim = 365

southHemisphereValues <- c(FALSE, TRUE)

parValuesAnnualSinusoid <- rbind(
  c(0.1, 1.5, 0.31),
  c(-0.5, 3.3, 0.73),
  c(1.5, 2.7, 0.06),
  c(2.1, 4.2, 0.25),
  c(-1.6, 5, 1),
  c(4, 4.5, 0.02)
)

minMinValue = min(parValuesAnnualSinusoid[,1] - parValuesAnnualSinusoid[,3])
maxMaxValue = max(parValuesAnnualSinusoid[,2] + parValuesAnnualSinusoid[,3])
```

Create a colour palette for plotting:

```{r}
numColdColours = floor(nrow(parValuesAnnualSinusoid) / 2)
numWarmColours = ceiling(nrow(parValuesAnnualSinusoid) / 2)
colorPaletteValues <- cbind(
  # hue
  h = c(
    seq(198.6, 299.4, length.out = numColdColours),
    seq(5.15, 67.5, length.out = numWarmColours)
  ) / 360,
  # saturation
  s = c(
    seq(61.6, 75.3, length.out = numColdColours),
    seq(67, 77.8, length.out = numWarmColours)
  ) / 100,
  # value
  v = c(
    seq(95.2, 76.4, length.out = numColdColours),
    seq(73.7, 86.4, length.out = numWarmColours)
  ) / 100
)

# format the values a HSV readable for plotting
colorPalette <- c()

for (i in 1:nrow(parValuesAnnualSinusoid))
{
  colorPalette <- c(colorPalette,
                    hsv(colorPaletteValues[i, "h"],
                        colorPaletteValues[i, "s"],
                        colorPaletteValues[i, "v"])
                    )
}
```

Plot only sinusoidal curves with mathematical annotation:

```{r}
grScale = 2

plotName = "annualSinusoidCurve.png"
  
png(plotName, width = grScale * 800, height = grScale * 480)

par(cex = grScale * 1.2)

plot(c(0, yearLengthInDays_sim * 1.5), # leave some space on the right side to display legend
     c(minMinValue, maxMaxValue * 1.45), # leave some space on top to display equation 
     type = "n", 
     main = "Annual sinusoid curve",
     xlab = "day of year",
     ylab = "output",
     cex.main = grScale
)

for (i in 1:nrow(parValuesAnnualSinusoid))
{
  curve <- getAnnualSinusoid(
    minValue = parValuesAnnualSinusoid[i, 1], 
    maxValue = parValuesAnnualSinusoid[i, 2],
    southHemisphere = FALSE,
    yearLengthInDays = yearLengthInDays_sim)
  
  lines((1:length(curve)) - 1, curve, 
        col = colorPalette[i], lwd = grScale * 3)
  
  legend(x = yearLengthInDays_sim * 1.05, 
         y = maxMaxValue * (0.9 - 0.2 * (i - 1)), 
         legend = substitute(
           paste("minValue = ", minValue, 
                 ", maxValue = ", maxValue), 
           list(
             minValue = parValuesAnnualSinusoid[i, 1], 
             maxValue = parValuesAnnualSinusoid[i, 2])), 
         col = colorPalette[i],
         lwd = grScale * 3, cex = 0.8,
         title = NULL, bty = "n")
  
  # add example with southHemisphere = TRUE
  # curve <- getAnnualSinusoid(
  #   minValue = parValuesAnnualSinusoid[i, 1], 
  #   maxValue = parValuesAnnualSinusoid[i, 2], 
  #   southHemisphere = TRUE,
  #   yearLengthInDays = yearLengthInDays_sim)
  # 
  # lines((1:length(curve)), curve, 
  #       col = colorPalette[i+nrow(parValuesAnnualSinusoid)], 
  #       lwd = grScale * 3)
  # 
  # legend(x = yearLengthInDays_sim, 
  #        y = maxMaxValue * (0.5 - 0.1 * (i - 1)), 
  #        legend = substitute(
  #          paste("minValue = ", minValue, 
  #                ", maxValue = ", maxValue), 
  #          list(
  #            minValue = parValuesAnnualSinusoid[i, 1], 
  #            maxValue = parValuesAnnualSinusoid[i, 2])), 
  #        col = colorPalette[i+nrow(parValuesAnnualSinusoid)],
  #        lwd = grScale * 3, cex = 0.8,
  #        title = NULL, bty = "n")
}
# text(x = yearLengthInDays_sim * 1.2, 
#      y = maxMaxValue * c(1.12, 0.52),
#      labels = c("southHemisphere = FALSE", "southHemisphere = TRUE"),
#      cex = 0.8)
text(x = yearLengthInDays_sim * 0.75, y = maxMaxValue * 1.2,
     expression(paste(
       "output = minValue + ", frac(maxValue - minValue, 2) * 
         bgroup("(", 1 + "sin" * 
                  bgroup("(", 360 * frac(dayOfYear - (dayOfYearWithLowestValue - yearLengthInDays), yearLengthInDays) - 90, ")"), 
                ")") 
     ))
     , cex = grScale * 0.4)
dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

Plot curves:

```{r}
#---------
grScale = 2
fontRescale = 0

plotName = "annualSinusoidCurve_full.png"

png(plotName, width = grScale * 1000, height = grScale * 600)

layout(matrix(c(1,  2,  3, 12, # titles
                4,  5,  6, 12, # annual sinusoids
                7,  8,  9, 12, # annual sinusoids with fluctuations
                10,11, 11, 12), # x-axis title 
              nrow = 4, ncol = 4, byrow = TRUE),
       widths = c(1, 10, 10, 6),
       heights = c(2, 10, 10, 2)
       )

par(cex = grScale * 1.2)

# 1-3: titles (southHemisphere)

par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.55, y = 0.5, font = 4, 
     cex = grScale * (0.6 + fontRescale),
     labels = "southHemisphere = FALSE")

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.55, y = 0.5, font = 4, 
     cex = grScale * (0.6 + fontRescale),
     labels = "southHemisphere = FALSE")

# 4-6: annual sinusoid y-axis title and plots

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, font = 4, 
     cex = grScale * (0.6 + fontRescale), 
     srt = 90,
     labels = "annual sinusoidal curve")

par(mar = c(2,2,0.1,0.1))

for (southHemisphere in southHemisphereValues)
{
  plot(c(1, yearLengthInDays_sim),
       c(minMinValue, maxMaxValue),
       type = "n", 
       xlab = "", ylab = ""
  )
  
  for (i in 1:nrow(parValuesAnnualSinusoid))
  {
    curve <- getAnnualSinusoid(
      minValue = parValuesAnnualSinusoid[i, 1], 
      maxValue = parValuesAnnualSinusoid[i, 2],
      yearLengthInDays = yearLengthInDays_sim,
      southHemisphere = southHemisphere
    )
    
    lines((1:length(curve)), curve, 
          col = colorPalette[i], lwd = grScale * 3)
  }
}

# 7-9: annual sinusoid with fluctuations y-axis title and plots

par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, font = 4,
     cex = grScale * (0.5 + fontRescale),
     srt = 90,
     labels = "annual sinusoidal curve\nwith fluctuations")

par(mar = c(2,2,0.1,0.1))

for (southHemisphere in southHemisphereValues)
{
  plot(c(1, yearLengthInDays_sim),
       c(minMinValue, maxMaxValue),
       type = "n",
       xlab = "", ylab = ""
  )

  for (i in 1:nrow(parValuesAnnualSinusoid))
  {
    curve <- getAnnualSinusoidWithFluctuation(
      minValue = parValuesAnnualSinusoid[i, 1],
      maxValue = parValuesAnnualSinusoid[i, 2],
      yearLengthInDays = yearLengthInDays_sim,
      southHemisphere = southHemisphere,
      fluctuation = parValuesAnnualSinusoid[i, 3]
    )

    lines((1:length(curve)), curve,
          col = colorPalette[i], lwd = grScale * 1)
  }
}

# 10-11: empty & x-axis title

par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.4, font = 4, cex = grScale * (0.6 + fontRescale),
     labels = "day of year")

# 12: legend

par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, nrow(parValuesAnnualSinusoid) + 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

xPos = 0.25
yPos = c(0.5, 0.1, -0.1)
jump = 1

for (i in 1:nrow(parValuesAnnualSinusoid))
{
  legend(x = 0,
         y = (yPos[1] + jump * i),
         legend = substitute(
           paste("minValue = ", minValue, ","),
           list(minValue = parValuesAnnualSinusoid[i, 1])),
         col = colorPalette[i],
         lwd = grScale * 6, cex = grScale * (0.5 + fontRescale),
         title = NULL,
         bty = "n")
  text(x = xPos, 
       y = (yPos[2] + jump * i),
       labels = substitute(
         paste("maxValue = ", maxValue, ","), 
         list(maxValue = parValuesAnnualSinusoid[i, 2])),
       cex = grScale * (0.5 + fontRescale), adj = 0)
  text(x = xPos, 
       y = (yPos[3] + jump * i),
       labels = substitute(
         paste("fluctuation = ", fluctuation), 
         list(fluctuation = parValuesAnnualSinusoid[i, 3])),
       cex = grScale * (0.5 + fontRescale), adj = 0)
}


dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```
