---
title: "Demostration of double logistic curve of the Weather model"
author: "Andreas Angourakis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---

## Load source file {-}

```{r}
source("source/weatherModel.R")
```

## Functions

Generic functions for generating double logistic curves to emulate cumulative annual precipitation curves, depending on start and end days of year, a plateau value (intermediate value), and two sets of point of inflection and growth rate.

```{r}
getPrecipitationOfYear
```

```{r}
getAnnualDoubleLogisticCurve
```

```{r}
getDayValueInAnnualDoubleLogistic
```

```{r}
escalonateCurve
```

```{r}
rescaleCurve
```

```{r}
getIncrementsFromCumulativeCurve
```

## Demonstration {-}

Set up six variations of parameter settings of the annual double logistic curve (i.e. plateauValue, inflection1, rate1, inflection2, rate2), the escalonation producing the annual cumulative precipitation curve (i.e. nSamples, maxSampleSize) and annualPrecipitation, assuming length of year of 365 days. Random generator seed used in escalonation is fixed:

```{r}
seed = 0

yearLengthInDays_sim = 365

parValuesDoubleLogistic <- rbind(
  # plateauValue, inflection1, rate1, inflection2, rate2
  c(0.01,         125,         0.3,   245,         0.22),
  c(0.15,         63,          0.55,  195,         0.6),
  c(0.5,          64,          0.05,  261,         0.12),
  c(0.45,         215,         0.01,  276,         0.39),
  c(0.6,          20,          0.38,  254,         0.04),
  c(0.85,         97,          0.24,  219,         0.17)
)

parValuesEscalonation <- rbind(
  # nSamples, maxSampleSize
  c(152, 22),
  c(220, 10),
  c(240, 6),
  c(168, 13),
  c(191, 9),
  c(205, 17)
)

annualSumValues <- c(410, 1050, 636, 320, 1280, 745)

```

Initialise data frames for holding curves:

```{r}
doubleLogisticCurves <- data.frame(
  matrix(1:(yearLengthInDays_sim * nrow(parValuesDoubleLogistic)),
         nrow = yearLengthInDays_sim, 
         ncol = nrow(parValuesDoubleLogistic))
)

escalonatedDoubleLogisticCurves <- data.frame(
  matrix(1:(yearLengthInDays_sim * nrow(parValuesDoubleLogistic)),
         nrow = yearLengthInDays_sim, 
         ncol = nrow(parValuesDoubleLogistic))
)

dailyPrecipitation <- data.frame(
  matrix(1:(yearLengthInDays_sim * nrow(parValuesDoubleLogistic)),
         nrow = yearLengthInDays_sim, 
         ncol = nrow(parValuesDoubleLogistic))
)
```

Create a colour palette for plotting:

```{r}
numColdColours = floor(nrow(parValuesDoubleLogistic) / 2)
numWarmColours = ceiling(nrow(parValuesDoubleLogistic) / 2)
colorPaletteValues <- cbind(
  # hue
  h = c(
    seq(198.6, 299.4, length.out = numColdColours),
    seq(5.15, 67.5, length.out = numWarmColours)
  ) / 360,
  # saturation
  s = c(
    seq(61.6, 75.3, length.out = numColdColours),
    seq(67, 77.8, length.out = numWarmColours)
  ) / 100,
  # value
  v = c(
    seq(95.2, 76.4, length.out = numColdColours),
    seq(73.7, 86.4, length.out = numWarmColours)
  ) / 100
)

# format the values a HSV readable for plotting
colorPalette <- c()

for (i in 1:nrow(parValuesDoubleLogistic))
{
  colorPalette <- c(colorPalette,
                    hsv(colorPaletteValues[i, "h"],
                        colorPaletteValues[i, "s"],
                        colorPaletteValues[i, "v"])
                    )
}
```

Plot only double logistic curves with mathematical annotation:

```{r}
grScale = 2
plotName = "annualDoubleLogisticCurve.png"
  
png(plotName, width = grScale * 800, height = grScale * 520)
par(cex = grScale * 1.2)
plot(c(0, yearLengthInDays_sim * 1.8), # leave some space on the right side to display legend
     c(0, 1.35), # leave some space on top to display equation 
     type = "n", 
     main = "Double logistic curve",
     xlab = "day of year",
     ylab = "output",
     cex.main = grScale
)
for (i in 1:nrow(parValuesDoubleLogistic))
{
  curve <- getAnnualDoubleLogisticCurve(
    plateauValue = parValuesDoubleLogistic[i, 1], 
    inflection1 = parValuesDoubleLogistic[i, 2], 
    rate1 = parValuesDoubleLogistic[i, 3], 
    inflection2 = parValuesDoubleLogistic[i, 4],
    rate2 = parValuesDoubleLogistic[i, 5],
    yearLengthInDays = yearLengthInDays_sim)
  
  lines((1:length(curve)) - 1, curve, 
        col = colorPalette[i], lwd = grScale * 3)
  
  legend(x = yearLengthInDays_sim * 1, 
         y = 1.05 * (1 - 0.18 * (i - 1)), 
         legend = substitute(
           paste("plateauValue = ", plateauValue, 
                 ", inflection1 = ", inflection1,
                 ", rate1 = ", rate1, ","), 
           list(plateauValue = parValuesDoubleLogistic[i, 1], 
                inflection1 = parValuesDoubleLogistic[i, 2],
                rate1 = parValuesDoubleLogistic[i, 3])),
         col = colorPalette[i],
         lwd = grScale * 3, cex = 0.8,
         title = NULL, bty = "n")
  
  text(x = yearLengthInDays_sim * 1.1, 
       y = 1.05 * (0.88 - 0.18 * (i - 1)),
       labels = substitute(
         paste("inflection2 = ", inflection2,
               ", rate2 = ", rate2), 
         list(inflection2 = parValuesDoubleLogistic[i, 4],
              rate2 = parValuesDoubleLogistic[i, 5])),
       adj = 0, cex = 0.8
       )
}
text(x = yearLengthInDays_sim * 0.9, y = 1.2,
     expression(paste(
       "output =  ", frac(plateauValue, (1 + "e"^(rate1 * (inflection1 - dayOfYear)))) + 
         frac(1 - plateauValue, (1 + "e"^(rate2 * (inflection2 - dayOfYear))))
     ))
     , cex = grScale * 0.7)
dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```


Plot all curves:

```{r}
grScale = 2
fontRescale = 0

plotName = "annualDoubleLogisticCurve_full.png"

png(plotName, width = grScale * 1000, height = grScale * 1000)

# this layout is assuming 6 cases
layout(matrix(c(20, 20, 20, 20, 20, 23, 23,
                1,   5,  5,  5,  5, 23, 23,
                21, 21, 21, 21, 21, 23, 23,
                2,   6,  6,  6,  6, 23, 23,
                22, 22, 22, 22, 22, 23, 23,
                3,   7,  8,  9, 10, 11, 12,
                3,  13, 14, 15, 16, 17, 18,
                4,  19, 19, 19, 19, 19, 19), 
              nrow = 8, ncol = 7, byrow = TRUE), 
       widths = c(2, rep(10, 6)),
       heights = c(3, 12, 3, 12, 3, 12, 0.01, 1))

par(cex = grScale * 1.2)

# 1st column: y-axes titles

# 1: y-axis title - double logistic curves

par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, font = 4, 
     cex = grScale * (0.7 + fontRescale), 
     srt = 90,
     labels = "daily cumulative value")

# 2: y-axis title - escalonated curves

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, font = 4, 
     cex = grScale * (0.7 + fontRescale), 
     srt = 90,
     labels = "daily cumulative value")

# 3: y-axis title - year daily precipitation

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, font = 4, 
     cex = grScale * (0.7 + fontRescale), 
     srt = 90,
     labels = "daily increment")

# 4: empty

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

# top plots

# 5: double logistic curves plot

par(mar = c(1, 2, 0.1, 1), 
    cex.axis = grScale * (0.7 + fontRescale))

plot(c(1, yearLengthInDays_sim),
     c(0, 1),
     type = "n", 
     xlab = "",
     ylab = ""
)

for (i in 1:nrow(parValuesDoubleLogistic))
{
  curve <- getAnnualDoubleLogisticCurve(
    plateauValue = parValuesDoubleLogistic[i, 1], 
    inflection1 = parValuesDoubleLogistic[i, 2], 
    rate1 = parValuesDoubleLogistic[i, 3], 
    inflection2 = parValuesDoubleLogistic[i, 4],
    rate2 = parValuesDoubleLogistic[i, 5],
    yearLengthInDays = yearLengthInDays_sim)
  
  lines((1:length(curve)), curve, 
        col = colorPalette[i], lwd = grScale * 3)
  
  points(c(parValuesDoubleLogistic[i, 2], 
           parValuesDoubleLogistic[i, 4]), 
         c(curve[parValuesDoubleLogistic[i, 2]], 
           curve[parValuesDoubleLogistic[i, 4]]),
         col = colorPalette[i], 
         pch = 19)
  
  doubleLogisticCurves[,i] <- curve
}

# 6: escalonated double logistic or annual cumulative precipitation

plot(c(1, yearLengthInDays_sim),
     c(0, 1),
     type = "n", 
     xlab = "",
     ylab = ""
)

for (i in 1:nrow(parValuesDoubleLogistic))
{
  curve <- escalonateCurve(
    curve = doubleLogisticCurves[,i],
    nSamples = parValuesEscalonation[i, 1],
    maxSampleSize = parValuesEscalonation[i, 2],
    seed = seed)
  
  lines((1:length(curve)), curve, 
        col = adjustcolor(colorPalette[i], alpha.f = 0.5), 
        lwd = grScale * 3)
  
  points(c(parValuesDoubleLogistic[i, 2], 
           parValuesDoubleLogistic[i, 4]), 
         c(curve[parValuesDoubleLogistic[i, 2]], 
           curve[parValuesDoubleLogistic[i, 4]]),
         col = adjustcolor(colorPalette[i], alpha.f = 0.5), 
         pch = 19)
  
  curve <- rescaleCurve(curve)
  
  lines((1:length(curve)), curve, 
        col = colorPalette[i], lwd = grScale * 3)
  
  points(c(parValuesDoubleLogistic[i, 2], 
           parValuesDoubleLogistic[i, 4]), 
         c(curve[parValuesDoubleLogistic[i, 2]], 
           curve[parValuesDoubleLogistic[i, 4]]),
         col = colorPalette[i], 
         pch = 19)
  
  escalonatedDoubleLogisticCurves[,i] <- curve
}

# 8-13: daily precipitation plots

par(mar = c(2, 2, 1, 1), 
    cex.axis = grScale * (0.6 + fontRescale))

for (i in 1:nrow(parValuesDoubleLogistic))
{
  dailyPrecipitation[,i] <- getIncrementsFromCumulativeCurve(
    cumulativeCurve = escalonatedDoubleLogisticCurves[,i]
  ) * annualSumValues[i]
}

# get overall maximum
maxDailyPrecipitation = max(dailyPrecipitation)

# plot in inverted order (plays nicer with legend)
for (i in nrow(parValuesDoubleLogistic):1)
{
  barplot(dailyPrecipitation[,i], 
        ylab = "",
        ylim = c(0, maxDailyPrecipitation),
        xlab = "", xaxt = 'n',
        col = colorPalette[i],
        border = colorPalette[i])
  
  points(c(parValuesDoubleLogistic[i, 2], 
           parValuesDoubleLogistic[i, 4]), 
         rep(maxDailyPrecipitation * 0.9, 2),
         col = colorPalette[i], 
         pch = 19)
}

# 14-19: x-axes of barplots 
# (built-in axes in barplot position is somehow affected by cex)

par(mar = c(0, 2, 0, 0.1))

for (i in 1:nrow(parValuesDoubleLogistic))
{
  plot(c(1, yearLengthInDays_sim), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
  
  axis(3, 
       at = 1:yearLengthInDays_sim,
       tck = 0, lwd = 0)
}

# 20: x-axis title

par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.4, font = 4, cex = grScale * (0.7 + fontRescale),
     labels = "day of year")

# 21-23: infographic bits between plots

arrowPointsX = c(1/3, 2/3, 2/3, 1, 0.5, 0, 1/3, 1/3)
arrowPointsY = c(1, 1, 0.5, 0.5, 0, 0.5, 0.5, 1)
arrowPosX = c(0.9, 1)
textPos = c(0.88, 0.4)

par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
polygon(x = arrowPosX[1] + (arrowPosX[2] - arrowPosX[1]) * arrowPointsX,
        y = arrowPointsY,
        col = rgb(0,0,0, alpha = 0.3),
        border = NA)
text(x = textPos[1], y = textPos[2], 
     font = 4, cex = grScale * (0.65 + fontRescale), adj = c(1, 0.5),
     labels = "getAnnualDoubleLogisticCurve(plateauValue, inflection1,\nrate1, inflection2, rate2)")

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
polygon(x = arrowPosX[1] + (arrowPosX[2] - arrowPosX[1]) * arrowPointsX,
        y = arrowPointsY,
        col = rgb(0,0,0, alpha = 0.3),
        border = NA)
text(x = textPos[1], y = textPos[2], 
     font = 4, cex = grScale * (0.65 + fontRescale), adj = c(1, 0.5),
     labels = "escalonateCurve(curve, nSamples, maxSampleSize)\nrescaleCurve(curve)")

plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
polygon(x = arrowPosX[1] + (arrowPosX[2] - arrowPosX[1]) * arrowPointsX,
        y = arrowPointsY,
        col = rgb(0,0,0, alpha = 0.3),
        border = NA)
text(x = textPos[1], y = textPos[2], 
     font = 4, cex = grScale * (0.65 + fontRescale), adj = c(1, 0.5),
     labels = "getIncrementsFromCumulativeCurve(curve) x annualSum")

# 24: legend

par(mar = c(0, 0, 0, 0))

plot(c(0, 1), c(0, nrow(parValuesDoubleLogistic) + 1), 
     ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')

# set y positions for each line in the first case (bottom)
yPos <- c(0.5, seq(0.1, -0.3, length.out = 3))
xPos = 0.55
jump = 1

for (i in 1:nrow(parValuesDoubleLogistic))
{
  legend(x = 0, 
         y = (yPos[1] + jump * i), 
         legend = substitute(
           paste("plateauValue = ", plateauValue, ", ",
                 "inflection1 = ", inflection1, ", "), 
           list(plateauValue = parValuesDoubleLogistic[i, 1], 
                inflection1 = parValuesDoubleLogistic[i, 2])), 
         col = colorPalette[i],
         lwd = grScale * 6, cex = grScale * (0.5 + fontRescale),
         title = NULL, 
         bty = "n")
  text(x = xPos, 
       y = (yPos[2] + jump * i),
       labels = substitute(
         paste("rate1 = ", rate1, ", ",
               "inflection2 = ", inflection2, ", ",
               "rate2 = ", rate2, ","), 
         list(rate1 = parValuesDoubleLogistic[i, 3],
              inflection2 = parValuesDoubleLogistic[i, 4],
              rate2 = parValuesDoubleLogistic[i, 5])),
       cex = grScale * (0.5 + fontRescale))
  text(x = xPos, 
       y = (yPos[3] + jump * i),
       labels = substitute(
         paste("nSamples = ", nSamples, ", ",
               "maxSampleSize = ", maxSampleSize), 
         list(nSamples = parValuesEscalonation[i, 1],
              maxSampleSize = parValuesEscalonation[i, 2])),
       cex = grScale * (0.5 + fontRescale))
  text(x = xPos, 
       y = (yPos[4] + jump * i),
       labels = substitute(
         paste("annualSum = ", annualSum), 
         list(annualSum = annualSumValues[i])),
       cex = grScale * (0.5 + fontRescale))
}

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```
