---
title: "Demostration of double logistic curve of the Weather model"
author: "Andreas Angourakis"
date: "27 January 2020"
output:
  pdf_document: default
  html_document:
    number_sections: yes
---

## Preparation

Declare generic function for generating double logistic curves depending on start and end days of year, a plateau value (intermediate value), and two sets of point of inflection and growth rate.

```{r}
doubleLogistic <- function(dayOfYear, 
                           plateauValue, inflection1, rate1, inflection2, rate2
                           )
{
  return(
    (plateauValue / (1 + exp((inflection1 - dayOfYear) * rate1))) + 
      ((1 - plateauValue) / (1 + exp((inflection2 - dayOfYear) * rate2)))
  )
}

doubleLogisticCurve <- function(start.x, end.x, 
                                plateauValue, inflection1, rate1, inflection2, rate2)
{
  curve <- c()
  for (i in start.x:end.x)
  {
    curve <- c(curve, doubleLogistic(i, plateauValue, inflection1, rate1, inflection2, rate2))
  }
  return(curve)
}
```

## Plot parameter exploration

Set up six variations of parameter settings (e.g. c(plateauValue, inflection1, rate1, inflection2, rate2)), assuming start.x = 1 and end.x = 365:

```{r}
start.x = 1
end.x = 365

parValues <- rbind(
  c(0.01, 125, 0.3,  245, 0.22),
  c(0.15,  63, 0.55, 195, 0.6),
  c(0.45, 215, 0.1,  276, 0.39),
  c(0.6,   20, 0.38, 254, 0.04),
  c(0.85,  97, 0.24, 219, 0.17)
)
```

Plot curves:

```{r}
grScale = 2

plotName = "DoubleLogisticCurve.png"
  
png(plotName, width = grScale * 800, height = grScale * 480)

par(cex = grScale * 1.2)

plot(c(start.x, end.x * 1.45), # leave some space on the right side to display legend
     c(0, 1.5), # leave some space on top to display equation 
     type = "n", 
     main = "Double logistic curve",
     xlab = "day of year",
     ylab = "output",
     cex.main = grScale
)

for (i in 1:nrow(parValues))
{
  curve <- doubleLogisticCurve(start.x = start.x, end.x = end.x, 
                               plateauValue = parValues[i, 1], 
                               inflection1 = parValues[i, 2], 
                               rate1 = parValues[i, 3], 
                               inflection2 = parValues[i, 4],
                               rate2 = parValues[i, 5])
  
  lines((1:length(curve)) - 1, curve, 
        col = i, lwd = grScale * 3)
  
  legend(x = end.x * 0.8, 
         y = 0.8 * (1 - 0.2 * (i - 1)), 
         legend = substitute(paste("plateauValue = ", plateauValue, 
                                   ", inflection1 = ", inflection1,
                                   ", rate1 = ", rate1, ","), 
                           list(plateauValue = parValues[i, 1], 
                                inflection1 = parValues[i, 2],
                                rate1 = parValues[i, 3])),
         col = i,
         lwd = grScale * 3, cex = 0.8,
         title = NULL, bty = "n")
  
  text(x = end.x * 0.9, 
       y = 0.8 * (0.83 - 0.2 * (i - 1)),
       labels = substitute(paste("inflection2 = ", inflection2,
                                 ", rate2 = ", rate2), 
                           list(inflection2 = parValues[i, 4],
                                rate2 = parValues[i, 5])),
       adj = 0, cex = 0.8
       )
}

text(x = end.x * 0.7, y = 1.3,
     expression(paste(
       "output =  ", frac(plateauValue, (1 + "e"^(rate1 * (inflection1 - dayOfYear)))) + 
         frac(1 - plateauValue, (1 + "e"^(rate2 * (inflection2 - dayOfYear))))
     ))
     , cex = grScale * 0.7)

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```
