---
title: "Demostration of the Weather model"
author: "Andreas Angourakis"
date: "12 August 2020"
output:
  html_document:
    number_sections: yes
  pdf_document: default
---

## Defining base generic functions

Declare generic function for geting points in sinusoid curves depending on minValue, maxValue, length of year in days, and a normal stochastic fluctuation. See extended explanation at [HTML with R walkthrough](https://htmlpreview.github.io/?https://github.com/Andros-Spica/indus-village-model/blob/master/01-weather/documentation/annualSinusoidCurve.html)

```{r}
getDayValueInAnnualSinusoidWithFluctuation <- function(minValue, maxValue, fluctuation, dayOfYear, yearLengthInDays)
{
  return(
    max(0, 
        rnorm(1, getDayValueInAnnualSinusoid(minValue, maxValue, dayOfYear, yearLengthInDays), fluctuation))
  )
}

getDayValueInAnnualSinusoid <- function(minValue, maxValue, dayOfYear, yearLengthInDays)
{
  amplitude = (maxValue - minValue) / 2
  
  return(
    minValue + amplitude * (1 + sin(3*pi/2 + 2*pi * dayOfYear / yearLengthInDays))
  )
}

```

Declare generic function for geting an year of daily precipitation. First, it generates a double logistic curve of range 0-1 depending on start and end days of year, a plateau value (intermediate value), and two sets of point of inflection and growth rate. See extended explanation at [HTML with R walkthrough](https://htmlpreview.github.io/?https://github.com/Andros-Spica/indus-village-model/blob/master/01-weather/documentation/doubleLogisticCurve.html). Then, the curve is broken down stochastically into steps, controlled by `nSample` and `maxSampleSize`, which better represents the discretionality of precipitation events. Finally, the cumulative daily value is transformed into daily increments, which is multiplied by the annualSum, returning daily precipitation values. 

```{r}
getPrecipitationOfYear <- function(
  plateauValue, inflection1, rate1, inflection2, rate2, 
  yearLengthInDays, nSamples, maxSampleSize, annualSum)
{
  precipitationOfYear <- c()
  
  precipitationOfYear <- doubleLogisticCurve(
    start.x = 1, end.x = yearLengthInDays,
    plateauValue = plateauValue,
    inflection1 = inflection1,
    rate1 = rate1,
    inflection2 = inflection2,
    rate2 = rate2
  )
  
  precipitationOfYear <- escalonateCurve(
    x = 1:length(precipitationOfYear), 
    y = precipitationOfYear,
    nSamples = nSamples, maxSampleSize = maxSampleSize)
  
  precipitationOfYear <- getIncrementsFromCumulativeCurve(precipitationOfYear) * annualSum
  
  return(precipitationOfYear)
}

doubleLogisticCurve <- function(start.x, end.x, 
                                plateauValue, inflection1, rate1, inflection2, rate2)
{
  curve <- c()
  for (i in start.x:end.x)
  {
    curve <- c(curve, doubleLogistic(i, plateauValue, inflection1, rate1, inflection2, rate2))
  }
  return(curve)
}

doubleLogistic <- function(dayOfYear, 
                           plateauValue, inflection1, rate1, inflection2, rate2
                           )
{
  return(
    (plateauValue / (1 + exp((inflection1 - dayOfYear) * rate1))) + 
      ((1 - plateauValue) / (1 + exp((inflection2 - dayOfYear) * rate2)))
  )
}

escalonateCurve <- function(x, y, nSamples, maxSampleSize)
{
  for (i in 1:nSamples)
  {
    # get a decreasing sample size proportionally to sample i
    thisSampleSize = ceiling(maxSampleSize * i / nSamples)
    # get random x coord point
    plateauPoint = round(runif(1, min = 1, max = length(x)))
    # set sample limits
    earliestNeighbour = max(1, plateauPoint - thisSampleSize)
    latestNeighbour = min(366, plateauPoint + thisSampleSize)
    # get mean of neighbourhood
    meanNeighbourhood = mean(y[x >= earliestNeighbour & x <= latestNeighbour])
    # assign mean to all days in neighbourhood
    for (j in earliestNeighbour:latestNeighbour)
    {
      y[x == j] <- meanNeighbourhood
    }
  }
  return(y)
}

getIncrementsFromCumulativeCurve <- function(x)
{
  incrementsX <- c()
  
  # incrementsX = x, if it is the first point
  incrementsX[1] = x[1]
  
  for (i in 2:length(x))
  {
    incrementsX[i] <- x[i] - x[i - 1]
  }
  
  return(incrementsX)
}

```

Load function to estimate reference evapotranspiration:

```{r}
source("source/estimateETr.R")
```

An auxiliar function to simplify the code:

```{r}

getLastItemInVector <- function(x)
{
  return(x[length(x)])
}

```

Load function for marking the end of each year:

```{r}
markEndYears <- function(lengthOfData, offset = 1){
  for (i in 1:lengthOfData)
  {
    if (i %% (365 * offset) == 0)
    {
      abline(v = i, lty = 3)
    }
  }
}
```

## Defining the model main procedures

```{r}

weatherModel.init <- function(
  yearLengthInDays = 365,
  albedo = 0.4,
  seed = 0,
  temperature_annualMaxAt2m = 40,
  temperature_annualMinAt2m = 15,
  temperature_meanDailyFluctuation = 5,
  temperature_dailyLowerDeviation = 5,
  temperature_dailyUpperDeviation = 5,
  solar_annualMax = 7,
  solar_annualMin = 3,
  solar_meanDailyFluctuation = 1,
  precip_yearlyMean = 400,
  precip_yearlySd = 130,
  precip_nSamples = 200,
  precip_maxSampleSize = 10,
  precip_plateauValue_yearlyMean = 0.1,
  precip_plateauValue_yearlySd = 0.05,
  precip_inflection1_yearlyMean = 40,
  precip_inflection1_yearlySd = 20,
  precip_rate1_yearlyMean = 0.15,
  precip_rate1_yearlySd = 0.02,
  precip_inflection2_yearlyMean = 200, 
  precip_inflection2_yearlySd = 20,
  precip_rate2_yearlyMean = 0.05,
  precip_rate2_yearlySd = 0.01
)
{
  set.seed(seed)
  
  weatherModel <- list()
  
  weatherModel$PARS <- list(
    yearLengthInDays = yearLengthInDays,
    albedo = albedo,
    
    temperature = list(
      annualMaxAt2m = temperature_annualMaxAt2m,
      annualMinAt2m = temperature_annualMinAt2m,
      meanDailyFluctuation = temperature_meanDailyFluctuation,
      dailyLowerDeviation = temperature_dailyLowerDeviation,
      dailyUpperDeviation = temperature_dailyUpperDeviation
    ),
    
    solar = list(
      annualMax = solar_annualMax,
      annualMin = solar_annualMin,
      meanDailyFluctuation = solar_meanDailyFluctuation
    ),
    
    precipitation = list(
      yearlyMean = precip_yearlyMean,
      yearlySd = precip_yearlySd,
      nSamples = precip_nSamples,
      maxSampleSize = precip_maxSampleSize,
      plateauValue_yearlyMean = precip_plateauValue_yearlyMean,
      plateauValue_yearlySd = precip_plateauValue_yearlySd,
      inflection1_yearlyMean = precip_inflection1_yearlyMean,
      inflection1_yearlySd = precip_inflection1_yearlySd,
      rate1_yearlyMean = precip_rate1_yearlyMean,
      rate1_yearlySd = precip_rate1_yearlySd,
      inflection2_yearlyMean = precip_inflection2_yearlyMean, 
      inflection2_yearlySd = precip_inflection2_yearlySd,
      rate2_yearlyMean = precip_rate2_yearlyMean,
      rate2_yearlySd = precip_rate2_yearlySd
    )
  )
  
  weatherModel$annualPrecipitationPars <- list(
    annualSum = c(),
    plateauValue = c(),
    inflection1 = c(),
    rate1 = c(),
    inflection2 = c(),
    rate2 = c()
  )
  
  weatherModel$daily <- list(
    currentYear = c(),
    currentDayOfYear = c(),
    temperature = c(),
    temperature_max = c(),
    temperature_min = c(),
    solarRadiation = c(),
    ETr = c(),
    precipitation = c()
  )
  
  return(weatherModel)
}

weatherModel.run <- function(weatherModel, numberOfYears)
{
  for (i in 1:numberOfYears)
  {
    weatherModel$annualPrecipitationPars$annualSum <- c(
      weatherModel$annualPrecipitationPars$annualSum,
      max(0, rnorm(1, weatherModel$PARS$precipitation$yearlyMean, weatherModel$PARS$precipitation$yearlySd))
    )
    weatherModel$annualPrecipitationPars$plateauValue <- c(
      weatherModel$annualPrecipitationPars$plateauValue,
      min(1, max(0, rnorm(1, weatherModel$PARS$precipitation$plateauValue_yearlyMean, weatherModel$PARS$precipitation$plateauValue_yearlySd)))
    )
    weatherModel$annualPrecipitationPars$inflection1 <- c(
      weatherModel$annualPrecipitationPars$inflection1,
      min(weatherModel$PARS$yearLengthInDays, max(1, rnorm(1, weatherModel$PARS$precipitation$inflection1_yearlyMean, weatherModel$PARS$precipitation$inflection1_yearlySd)))
    )
    weatherModel$annualPrecipitationPars$rate1 <- c(
      weatherModel$annualPrecipitationPars$rate1,
      max(0, rnorm(1, weatherModel$PARS$precipitation$rate1_yearlyMean, weatherModel$PARS$precipitation$rate1_yearlySd))
    )
    weatherModel$annualPrecipitationPars$inflection2 <- c(
      weatherModel$annualPrecipitationPars$inflection2,
      min(weatherModel$PARS$yearLengthInDays, max(1, rnorm(1, weatherModel$PARS$precipitation$inflection2_yearlyMean, weatherModel$PARS$precipitation$inflection2_yearlySd)))
    )
    weatherModel$annualPrecipitationPars$rate2 <- c(
      weatherModel$annualPrecipitationPars$rate2,
      max(0, rnorm(1, weatherModel$PARS$precipitation$rate2_yearlyMean, weatherModel$PARS$precipitation$rate2_yearlySd))
    )
    
    weatherModel$daily$precipitation <- c(
      weatherModel$daily$precipitation,
      getPrecipitationOfYear(
        plateauValue = getLastItemInVector(weatherModel$annualPrecipitationPars$plateauValue),
        inflection1 = getLastItemInVector(weatherModel$annualPrecipitationPars$inflection1),
        rate1 = getLastItemInVector(weatherModel$annualPrecipitationPars$rate1),
        inflection2 = getLastItemInVector(weatherModel$annualPrecipitationPars$inflection2),
        rate2 = getLastItemInVector(weatherModel$annualPrecipitationPars$rate2), 
        yearLengthInDays = weatherModel$PARS$yearLengthInDays,
        nSamples = weatherModel$PARS$precipitation$nSamples,
        maxSampleSize = weatherModel$PARS$precipitation$maxSampleSize,
        annualSum = getLastItemInVector(weatherModel$annualPrecipitationPars$annualSum)
      )
    )
    
    for (day in 1:weatherModel$PARS$yearLengthInDays)
    {
      weatherModel$daily$currentYear <- c(weatherModel$daily$currentYear, i)
      
      weatherModel$daily$currentDayOfYear <- c(weatherModel$daily$currentDayOfYear, day)
      
      weatherModel$daily$temperature <- c(
        weatherModel$daily$temperature,
        getDayValueInAnnualSinusoidWithFluctuation(
          minValue = weatherModel$PARS$temperature$annualMinAt2m, 
          maxValue = weatherModel$PARS$temperature$annualMaxAt2m, 
          fluctuation = weatherModel$PARS$temperature$meanDailyFluctuation, 
          dayOfYear = day, yearLengthInDays = weatherModel$PARS$yearLengthInDays
        )
      )
      
      weatherModel$daily$temperature_min <- c(
        weatherModel$daily$temperature_min,
        getLastItemInVector(weatherModel$daily$temperature) - weatherModel$PARS$temperature$dailyLowerDeviation
      )
      
      weatherModel$daily$temperature_max <- c(
        weatherModel$daily$temperature_max,
        getLastItemInVector(weatherModel$daily$temperature) + weatherModel$PARS$temperature$dailyUpperDeviation
      )
      
      weatherModel$daily$solarRadiation <- c(
        weatherModel$daily$solarRadiation,
        getDayValueInAnnualSinusoidWithFluctuation(
          minValue = weatherModel$PARS$solar$annualMin, 
          maxValue = weatherModel$PARS$solar$annualMax, 
          fluctuation = weatherModel$PARS$solar$meanDailyFluctuation, 
          dayOfYear = day, yearLengthInDays = weatherModel$PARS$yearLengthInDays
        )
      )
      
      weatherModel$daily$ETr <- c(
        weatherModel$daily$ETr,
        estimateETr(
          R_s = getLastItemInVector(weatherModel$daily$solarRadiation), 
          Temp = getLastItemInVector(weatherModel$daily$temperature), 
          maxTemp = getLastItemInVector(weatherModel$daily$temperature_max), 
          minTemp = getLastItemInVector(weatherModel$daily$temperature_min)
        )
      )
    }
  }
  
  return(weatherModel)
}

```

## Running the model

```{r}
weatherModel <- weatherModel.init()

weatherModel <- weatherModel.run(weatherModel, 5)
```

Show table with parameter values:

```{r}
parvalues <- c(
  weatherModel$PARS[[1]],
  weatherModel$PARS[[2]]
)
parNames <- c("yearLengthInDays", "albedo")

for (j in 3:length(weatherModel$PARS))
{
  parGroupName <- names(weatherModel$PARS)[j]
  for (i in 1:length(weatherModel$PARS[[j]]))
  {
    parvalues <- c(
      parvalues,
      weatherModel$PARS[[j]][[i]]
    )
    parName <- paste(parGroupName, names(weatherModel$PARS[[j]])[i], sep = " - ")
    parNames <- c(parNames, parName)
  }
}

parvalues <- cbind(parNames, parvalues)
knitr::kable(parvalues, 
             format = "html",
             col.names = c("parameter", "values"),
             align = c("l", "r"))
```

Plot time-series:

```{r}
lengthOfSimulation = length(weatherModel$daily$currentYear)

#--------
# To create a png file:
plotName = "weatherModelExample.png"

fontScale = c(2, 2, 1.2, # c(cex.lab, legend, cex.axis)
              3, 2) # side annotation name and "model"

png(plotName, width = 1000, height = 800)

layout(matrix(c(1:4), nrow = 4))

yLabs <- c(expression(paste("    Solar\nRadiation (", MJ/m^-2, ")")), 
           "Temperature (ºC)", 
           "Precipitation (mm)", 
           "ETr (mm)")
leftMargin <- 7


# 1. Solar radiation
par(mar = c(0.1, leftMargin, 1.1, 0.2), 
    cex.lab = fontScale[1],
    cex.axis = fontScale[3])

plot(1:lengthOfSimulation, weatherModel$daily$solarRadiation, type = "l",
     xlab = "", xaxt = 'n',
     ylab = yLabs[1])
markEndYears(lengthOfSimulation)

# 2. Temperature
plot(1:lengthOfSimulation, weatherModel$daily$temperature, type = "l",
     xlab = "", xaxt = 'n',
     ylab = yLabs[2])
lines(1:lengthOfSimulation, weatherModel$daily$temperature_max, col = "red")
lines(1:lengthOfSimulation, weatherModel$daily$temperature_min, col = "blue")

markEndYears(lengthOfSimulation)

# 3. Precipitation
par(mar = c(1.1, leftMargin, 2.1, 0.2))

barplot(weatherModel$daily$precipitation, 
        ylab = yLabs[3])
markEndYears(lengthOfSimulation, offset = 1.2) ; abline(v = 2190 * 1.2, lty = 3)
# not sure why, but barplot() x coordinates do not behave as in plot()

# 4. Reference evapotranspiration
par(mar = c(0.1, leftMargin, 1.1, 0.2))

plot(1:lengthOfSimulation, weatherModel$daily$ETr, type = "l",
        ylab = yLabs[4])
markEndYears(lengthOfSimulation, offset = 1.2) ; abline(v = 2190 * 1.2, lty = 3)
# not sure why, but barplot() x coordinates do not behave as in plot()

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

