---
title: "Household Demography model: sensitivity analysis"
output: html_notebook
---

```{r, message=FALSE}
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(gridExtra)
library(scales)
library(randomForest)
```

```{r}
plotDirectory <- "plots/"
```

## Figure. Example of response of the three life table models to parameter variation

Load functions that create/import Life Table models for fertility and nuptiality (Peristera-Kostaki model), and mortality (Coale-Demeny model):

```{r}
### fertility
# The following correspond to the first parametric model
# mentioned in page 147 of:
# # Peristeva and Kostaki, 2009
# "Modeling fertility in modern populations"
# Available from: https://dx.doi.org/10.4054/DemRes.2007.16.6

### nuptiality
# The following correspond to the first parametric model
# for fitting the age-specific distributions of marriages, mentioned in page 133 of:
# # Peristeva and Kostaki, 2015
# "A parametric model for estimating nuptiality patterns in modern populations"
# Available from: https://www.researchgate.net/publication/285457704_A_parametric_model_for_estimating_nuptiality_patterns_in_modern_populations [accessed Nov 27 2018].

# the equation is the same for both fertility and nuptiality.

generatePeristeraKostaki <- function(par.c1 = 0.8, par.mu = 25, par.sigma = c(10, 10) )
{
  curve <- c()
  for (i in 1:100)
  {
    sigma = par.sigma[1]
    if (i > par.mu)
    { sigma = par.sigma[2] }
    
    curve <- c(curve, par.c1 * exp (-1 * (((i - par.mu) / sigma) ^ 2)) )
  }
  
  return(curve)
}

### mortality
# from Coale-Demeny Model Life Tables generated with 'demoR' package
# demoR package version 0.6.0 (2018-09-13)
# by James Holland Jones and collaborators
# Their source:
# Coale, A., P. Demeny, and B. Vaughn. 1983. 
# Regional model life tables and stable populations. 
# 2nd ed. New York: Academic Press.

interpolatePerYear <- function(raw, ages = c(0.5, 1.5, 4, seq(8.5, 93.5, 5))) {
  
  perYear <- data.frame(matrix(numeric(0), nrow = 151, ncol = ncol(raw)))
  names(perYear) <- 1:ncol(perYear)
  
  for (i in 1:ncol(raw)) {
    perYear[, i] <- approx(ages, raw[, i], 
                           xout = 1:151, yleft = 0, yright = 1)$y
  }
  row.names(perYear) <- 0:150
  
  return(perYear)
}

generateCoaleDemenyLifeTable <- function(region = "north", sex = "F", level = 8){
  
  curve <- 0
  
  if (region == "north")
  {
    curve <- interpolatePerYear(t(demogR::cdmltn(sex = sex)$nqx))[,level]
  }
  if (region == "west")
  {
    curve <- interpolatePerYear(t(demogR::cdmltw(sex = sex)$nqx))[,level]
  }
  if (region == "east")
  {
    curve <- interpolatePerYear(t(demogR::cdmlte(sex = sex)$nqx))[,level]
  }
  if (region == "south")
  {
    curve <- interpolatePerYear(t(demogR::cdmlts(sex = sex)$nqx))[,level]
  }
  
  return(curve)
}
```

Plot the curves generated by small explorations of the parameter space of each model: 

```{r}
#--------
# To create a png file:
plotName = paste0(plotDirectory, "parameter_demonstration.png")

grScale = 2

png(plotName, width = grScale * 600, height = grScale * 600)
#---------
# alternatively, to create eps file:
# plotName = "Fig9.eps"
# 
# grScale = 1.8
# 
# extrafont::loadfonts(device = "postscript")
# grDevices::postscript(file = plotName,
#                       pointsize = 10,
#                       width = grScale * 6,
#                       height = grScale * 6,
#                       horizontal = FALSE,
#                       paper = "special",
#                       onefile = FALSE,
#                       family = "sans",
#                       colormodel = "cmyk")
#---------

layout(matrix(1:2, nrow = 2, ncol = 1))

par(cex = grScale * 1.2, family = "sans",
    mar = c(3.5, 4.1, 1.1, 0.2))

# 1. Peristera-Kostaki equation (Fertility and Nuptiality models)

plot(c(1, 100), c(0, 1), type = "n", 
     main = "Peristera-Kostaki equation",
     xlab = "AGE\n",
     ylab = ""
)
mtext(text = "Probability of event\n(giving birth, marrying)", 
      side = 2, line = 2, cex = grScale * 1.2)

# five variations of parameter settings
parValues <- rbind(
  c(0.8, 25, 10, 10),
  c(0.95, 20, 5, 7),
  c(0.7, 25, 12, 15),
  c(0.85, 18, 2, 15),
  c(0.75, 22, 8, 10)
)

for (i in 1:nrow(parValues))
{
  lines(1:100, 
        generatePeristeraKostaki(par.c1 = parValues[i, 1], par.mu = parValues[i, 2], 
                                 par.sigma = c(parValues[i, 3], parValues[i, 4])), 
        col = i, lwd = grScale * 3)
  
  legend(60, 1 - 0.1 * (i - 1), 
       legend = substitute(paste(c[1], " = ", pc1, ", ", mu, " = ", pmu, ", ",
                                 sigma[1], " = ", ps1, ", ", sigma[2], " = ", ps2), 
                           list(pc1 = parValues[i, 1], pmu = parValues[i, 2],
                                ps1 = parValues[i, 3], ps2 = parValues[i, 4])), 
       col = i,
       lwd = grScale * 3, cex = 0.8,
       title = NULL, bty = "n")
}

# 2. Mortality model (CoaleDemeny Life tables)

plot(c(1, 151), c(0, 1), type = "n", 
     main = "Coale-Demeny Model Life Tables",
     xlab = "AGE\n",
     ylab = "Probability of dying",
     xlim = c(0, 100)
)

# five variations of parameter settings
parValues <- rbind(
  c("north", "M", 1),
  c("west", "F", 8),
  c("east", "M", 12),
  c("south", "F", 17),
  c("north", "F", 25)
)

for (i in 1:nrow(parValues))
{
  lines(1:151, 
        generateCoaleDemenyLifeTable(region = parValues[i, 1], 
                                     sex = parValues[i, 2], 
                                     level = parValues[i, 3]), 
        col = i, lwd = grScale * 3)
  
  legend(1, 1 - 0.1 * (i - 1), 
       legend = paste("region = ", parValues[i, 1], ", ", 
                      "sex = ", parValues[i, 2], ", ",
                      "level = ", parValues[i, 3]),
       col = i, text.font = 3,
       lwd = grScale * 3, cex = 0.8,
       title = NULL, bty = "n")
}

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plotName)
```

## Multiparameter exploration of "householdDemography v.1.1 - kinship tabu for new couples"

### Generate parameter values with Sobol Sampling

```{r}
sobol_parameter_sample_dataFileName <- "data/sobol_parameter_values.RData"

# sobol_parameter_names <- c(
#   'initialNumHouseholds', 
#   'householdInitialAgeDistribution_min', 'householdInitialAgeDistribution_max',      
#   'maxCoupleCountDistribution_min', 'maxCoupleCountDistribution_max',
#   'acceptableKinshipDegreeForCouples', 
#   'cdmlt-level',
#   'c1-fert', 'mu-fert', 
#   'sigma1-fert', 'sigma2-fert',
#   'c1-women', 'mu-women', 
#   'sigma1-women', 'sigma2-women',
#   'c1-men', 'mu-men', 
#   'sigma1-men', 'sigma2-men'
# )
sobol_parameter_ranges <- list(
  initialNumHouseholds = c(2, 50),         
  householdInitialAgeDistribution_min = c(0, 5),        
  householdInitialAgeDistribution_max = c(6, 60),
  maxCoupleCountDistribution_min = c(1, 2),
  maxCoupleCountDistribution_max = c(3, 8),
  acceptableKinshipDegreeForCouples = c(0, 9),
  `cdmlt-level` = c(1, 25),
  `c1-fert` = c(0, 0.5),
  `mu-fert` = c(15, 30),
  `sigma1-fert` = c(1, 10),
  `sigma2-fert` = c(5, 20), 
  `c1-women` = c(0, 1),
  `mu-women` = c(15, 40),
  `sigma1-women` = c(1, 20),
  `sigma2-women` = c(1, 20), 
  `c1-men` = c(0, 1),
  `mu-men` = c(15, 40), 
  `sigma1-men` = c(1, 20), 
  `sigma2-men` = c(1, 20)
)
```

```{r, eval=FALSE}
library(randtoolbox)
set.seed(123)
n <- 2000  # Number of samples
k <- length(sobol_parameter_ranges)  # Number of parameters

# Generate Sobol sequence
sobol_sample_matrix <- sobol(n, dim = k, scrambling = 3)

get_values_from_sobol <- function(param_index){
  return(qunif(sobol_sample_matrix[,param_index], 
                     min = sobol_parameter_ranges[[param_index]][1], 
                     max = sobol_parameter_ranges[[param_index]][2]))
}

```

```{r, eval=FALSE}
# Transform to specific distributions
params_sobol <- data.frame(
  initialNumHouseholds = round(get_values_from_sobol(1)),
  householdInitialAgeDistribution_min = round(get_values_from_sobol(2)),
  householdInitialAgeDistribution_max = round(get_values_from_sobol(3)),
  maxCoupleCountDistribution_min = round(get_values_from_sobol(4)),
  maxCoupleCountDistribution_max = round(get_values_from_sobol(5)),
  acceptableKinshipDegreeForCouples = round(get_values_from_sobol(6)),
  `cdmlt-level` = round(get_values_from_sobol(7)),
      
  `c1-fert` = get_values_from_sobol(8),
  `mu-fert` = get_values_from_sobol(9),
  `sigma1-fert` = get_values_from_sobol(10),
  `sigma2-fert` = get_values_from_sobol(11),
      
  `c1-women` = get_values_from_sobol(12),
  `mu-women` = get_values_from_sobol(13),
  `sigma1-women` = get_values_from_sobol(14),
  `sigma2-women` = get_values_from_sobol(15),
  `c1-men` = get_values_from_sobol(16),
  `mu-men` = get_values_from_sobol(17),
  `sigma1-men` = get_values_from_sobol(18),
  `sigma2-men` = get_values_from_sobol(19)
)

head(params_sobol)
```

```{r, eval=FALSE}
save(params_sobol, file = sobol_parameter_sample_dataFileName)
```

```{r}
load(sobol_parameter_sample_dataFileName)
```

Visualising parameter space with `pairs()`:

```{r}
plotName = paste0(plotDirectory, "sobol_param_pairsPlot.png")

plotScale = 10

png(plotName,
    width = 100 * plotScale, height = 100 * plotScale)
pairs(params_sobol, pch='·')
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

Saving parameter configurations (experiments) in separate files:

```{r}
save_experiment <- function(experiment_dataframe, experiment_row, version_name) {
  write(paste(unlist(experiment_dataframe[experiment_row,], use.names = FALSE)), 
        file = paste0(dirname(getwd()), "experiments/", version_name, "/", experiment_row, ".txt"))
}

for (i in 1:nrow(params_sobol)) {
  save_experiment(params_sobol, i, "v1.1")
}
```

>Run experiments in NetLogo Behaviour Space using the parameters `type-of-experiment` and `exp-number`.

Loading simulation experiments' end states:

```{r}
endStates <- read.csv(paste0(dirname(getwd()), "simdata/v1.1/householdDemography v.1.1 - kinship tabu for new couples exp-endstates-table.csv"), skip = 6)
```

Process parameters in list format in NetLogo:

```{r}
extract_values_from_NetLogo_list <- function(char_vector) {
  # Remove brackets and split each string into two parts
  split_values <- gsub("\\[|\\]", "", char_vector) |>
    strsplit(" ")

  # Convert to numeric vectors
  first_values <- sapply(split_values, function(x) as.numeric(x[1]))
  second_values <- sapply(split_values, function(x) as.numeric(x[2]))

  # Return as a named list
  list(
    first = first_values,
    second = second_values
  )
}

household_initial_age_distribution_range <- extract_values_from_NetLogo_list(endStates$householdInitialAgeDistribution)
max_couple_count_distribution_range <- extract_values_from_NetLogo_list(endStates$maxCoupleCountDistribution)

endStates <- cbind(
  endStates,
  household_initial_age_distribution_min = household_initial_age_distribution_range[[1]],
  household_initial_age_distribution_max = household_initial_age_distribution_range[[2]],
  max_couple_count_distribution_min = max_couple_count_distribution_range[[1]],
  max_couple_count_distribution_max = max_couple_count_distribution_range[[2]]
)
```


Categorise end states according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

Visualise `totalIndividuals`:

```{r}
hist(endStates$totalIndividuals)
```

create new categorical variable:

```{r}
max_population = 10000 # population size used in NetLogo as a stop condition when there is considerable growth  
endStates$survival <- ifelse(endStates$totalIndividuals > 0, "survival", "extinction")
endStates$survival <- paste(endStates$survival, ifelse(endStates$totalIndividuals > max_population, "with growth", ""))
```

```{r}
table(endStates$survival)
```

```{r}
barplot(table(endStates$survival))
```

```{r}
params_indexes <- c(9, 12, 14:17, 20:27, 47:50)
param_names <- names(endStates)[params_indexes]

endStates_long <- endStates %>%
  pivot_longer(cols = all_of(param_names)) %>%
  mutate(name = factor(name, levels = param_names))  # Preserve order
```


```{r}
source("library/bifurcationPlot.R")

plotName = paste0(plotDirectory, "sobol_bifurcationPlot.png")

plotScale = 10

png(plotName,
    width = 150 * plotScale, height = 80 * plotScale)

ggplot(data = endStates_long, aes(x = value, y = totalIndividuals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm") +
  facet_wrap(vars(name), ncol = 6, scales = "free_x") +
  theme(
    strip.text = element_text(size = 1.2 * plotScale)
  )

dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

Train a Random Forest model to assess parameter importance:

```{r}
endStates_for_rf <- endStates[,c(params_indexes, 29)]

# Train Random Forest to predict simulation output
set.seed(123)
rf_model <- randomForest(totalIndividuals ~ ., data = endStates_for_rf, importance = TRUE, ntree = 500)

# Print model summary
print(rf_model)
```

Extract importance values and print as table:

```{r}
# Extract importance values
importance_values <- importance(rf_model)

# Convert to a data frame
importance_df <- data.frame(Parameter = rownames(importance_values),
                            MDA = importance_values[, 1], 
                            MDG = importance_values[, 2])

# Print importance scores
print(importance_df)
```

Plot feature importance colouring parameters by domain (general, mortality, fertility, nuptiality):

```{r}
source("library/variableImportancePlot.R")

plotName = paste0(plotDirectory, "sobol_variableImportancePlot.png")

color_variable_map <- rev(c("#0000a2", 
                        "#bc272d", 
                        rep("#e9c716", 4), 
                        rep("#50ad9f", 8),
                        rep("#0000a2", 4)
))

# png
png(filename = plotName, width = 129, height = 168, units="mm", res=1200)
variableImportancePlot(rf_model, plot_scale = 0.5, color_variable_map = color_variable_map, range_standardised = TRUE)
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

### Initial household population

This parameter has slight positive effect on population survival:

```{r}
summary(endStates$initialNumHouseholds[endStates$totalIndividuals == 0])
summary(endStates$initialNumHouseholds[endStates$totalIndividuals > 0])
```

```{r}
boxplot(endStates$initialNumHouseholds ~ (endStates$totalIndividuals > 0))
```

### Mortality factors

Coale-Demeny Life Tables Model level parameter has great effect:

```{r}
summary(endStates$cdmlt.level[endStates$totalIndividuals == 0])
summary(endStates$cdmlt.level[endStates$totalIndividuals > 0])
```

```{r}
boxplot(endStates$cdmlt.level ~ (endStates$totalIndividuals > 0))
```

but region parameter has a smaller effect. East tends to survive more often.  

```{r}
table(endStates$coale.demeny.region, endStates$totalIndividuals > 0)
```

```{r}
boxplot(data = endStates, totalIndividuals ~ coale.demeny.region)
```

### Fertility factors

All have a visible effect, but particularly c1.fert and sigma1.fert. 

```{r}
layout(matrix(1:4, nrow = 2, byrow = T))
boxplot(endStates$c1.fert ~ (endStates$totalIndividuals > 0), main = "c1.fert")
boxplot(endStates$mu.fert ~ (endStates$totalIndividuals > 0), main = "mu.fert")
boxplot(endStates$sigma1.fert ~ (endStates$totalIndividuals > 0), main = "sigma1.fert")
boxplot(endStates$sigma2.fert ~ (endStates$totalIndividuals > 0), main = "sigma2.fert")
```

### Nuptiality factors

Residential rule showing no clear effect:

```{r}
table(endStates$residence.rule, endStates$totalIndividuals > 0)
```

The effect of the acceptable kinship degree is also unclear: 

```{r}
boxplot(endStates$acceptable.kinship.degree.for.couples ~ (endStates$totalIndividuals > 0))
```

All nuptiality parameters for both men and women present a small, yet apparently consistent effect. c1 and sigma parameters have a positive effect while mu has a smaller negative effect.

```{r}
layout(matrix(1:8, nrow = 2, byrow = T))
boxplot(endStates$c1.women ~ (endStates$totalIndividuals > 0), main = "c1.women")
boxplot(endStates$mu.women ~ (endStates$totalIndividuals > 0), main = "mu.women")
boxplot(endStates$sigma1.women ~ (endStates$totalIndividuals > 0), main = "sigma1.women")
boxplot(endStates$sigma2.women ~ (endStates$totalIndividuals > 0), main = "sigma2.women")
boxplot(endStates$c1.men ~ (endStates$totalIndividuals > 0), main = "c1.men")
boxplot(endStates$mu.men ~ (endStates$totalIndividuals > 0), main = "mu.men")
boxplot(endStates$sigma1.men ~ (endStates$totalIndividuals > 0), main = "sigma1.men")
boxplot(endStates$sigma2.men ~ (endStates$totalIndividuals > 0), main = "sigma2.men")
```

## Timing of survival

```{r}
hist(endStates$X.step.)
```

```{r}
table(endStates$X.step. > 0)
```

```{r}
plot(endStates$initialNumHouseholds ~ endStates$X.step.)
```


