---
title: "Household Demography model: sensitivity analysis"
output: html_notebook
---

```{r, message=FALSE}
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(gridExtra)
library(scales)
library(randomForest)
```

```{r}
plotDirectory <- "plots/"
plot_file_format <- c("png", "eps")[1] # modify index number to change format
```

## Multiparameter exploration of "householdDemography v.1.1 - kinship tabu for new couples"

### Generate parameter values with Sobol Sampling

```{r}
sobol_parameter_sample_dataFileName <- "data/sobol_parameter_values.RData"

# sobol_parameter_names <- c(
#   'initialNumHouseholds', 
#   'householdInitialAgeDistribution_min', 'householdInitialAgeDistribution_max',      
#   'maxCoupleCountDistribution_min', 'maxCoupleCountDistribution_max',
#   'acceptableKinshipDegreeForCouples', 
#   'cdmlt-level',
#   'c1-fert', 'mu-fert', 
#   'sigma1-fert', 'sigma2-fert',
#   'c1-women', 'mu-women', 
#   'sigma1-women', 'sigma2-women',
#   'c1-men', 'mu-men', 
#   'sigma1-men', 'sigma2-men'
# )
sobol_parameter_ranges <- list(
  initialNumHouseholds = c(2, 50),         
  householdInitialAgeDistribution_min = c(0, 5),        
  householdInitialAgeDistribution_max = c(6, 60),
  maxCoupleCountDistribution_min = c(1, 2),
  maxCoupleCountDistribution_max = c(3, 8),
  acceptableKinshipDegreeForCouples = c(0, 9),
  `cdmlt-level` = c(1, 25),
  `c1-fert` = c(0, 0.5),
  `mu-fert` = c(15, 30),
  `sigma1-fert` = c(1, 10),
  `sigma2-fert` = c(5, 20), 
  `c1-women` = c(0, 1),
  `mu-women` = c(15, 40),
  `sigma1-women` = c(1, 20),
  `sigma2-women` = c(1, 20), 
  `c1-men` = c(0, 1),
  `mu-men` = c(15, 40), 
  `sigma1-men` = c(1, 20), 
  `sigma2-men` = c(1, 20)
)
```

```{r, eval=FALSE}
library(randtoolbox)
set.seed(123)
n <- 2000  # Number of samples
k <- length(sobol_parameter_ranges)  # Number of parameters

# Generate Sobol sequence
sobol_sample_matrix <- sobol(n, dim = k, scrambling = 3)

get_values_from_sobol <- function(param_index){
  return(qunif(sobol_sample_matrix[,param_index], 
                     min = sobol_parameter_ranges[[param_index]][1], 
                     max = sobol_parameter_ranges[[param_index]][2]))
}

```

```{r, eval=FALSE}
# Transform to specific distributions
params_sobol <- data.frame(
  initialNumHouseholds = round(get_values_from_sobol(1)),
  householdInitialAgeDistribution_min = round(get_values_from_sobol(2)),
  householdInitialAgeDistribution_max = round(get_values_from_sobol(3)),
  maxCoupleCountDistribution_min = round(get_values_from_sobol(4)),
  maxCoupleCountDistribution_max = round(get_values_from_sobol(5)),
  acceptableKinshipDegreeForCouples = round(get_values_from_sobol(6)),
  `cdmlt-level` = round(get_values_from_sobol(7)),
      
  `c1-fert` = get_values_from_sobol(8),
  `mu-fert` = get_values_from_sobol(9),
  `sigma1-fert` = get_values_from_sobol(10),
  `sigma2-fert` = get_values_from_sobol(11),
      
  `c1-women` = get_values_from_sobol(12),
  `mu-women` = get_values_from_sobol(13),
  `sigma1-women` = get_values_from_sobol(14),
  `sigma2-women` = get_values_from_sobol(15),
  `c1-men` = get_values_from_sobol(16),
  `mu-men` = get_values_from_sobol(17),
  `sigma1-men` = get_values_from_sobol(18),
  `sigma2-men` = get_values_from_sobol(19)
)

head(params_sobol)
```

```{r, eval=FALSE}
save(params_sobol, file = sobol_parameter_sample_dataFileName)
```

```{r}
load(sobol_parameter_sample_dataFileName)
```

Visualising parameter space with `pairs()`:

```{r}
plotName = paste0(plotDirectory, "sobol_param_pairsPlot.png")

plotScale = 10

png(plotName,
    width = 100 * plotScale, height = 100 * plotScale)
pairs(params_sobol, pch='·')
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

Saving parameter configurations (experiments) in separate files:

```{r}
save_experiment <- function(experiment_dataframe, experiment_row, version_name) {
  write(paste(unlist(experiment_dataframe[experiment_row,], use.names = FALSE)), 
        file = paste0(dirname(getwd()), "experiments/", version_name, "/", experiment_row, ".txt"))
}

for (i in 1:nrow(params_sobol)) {
  save_experiment(params_sobol, i, "v1.1")
}
```

>Run experiments in NetLogo Behaviour Space using the parameters `type-of-experiment` and `exp-number`.

Loading simulation experiments' end states:

```{r}
endStates <- read.csv("simdata/v1.1/householdDemography v.1.1 - kinship tabu for new couples exp-endstates-table.csv", skip = 6)
```

Process parameters in list format in NetLogo:

```{r}
extract_values_from_NetLogo_list <- function(char_vector) {
  # Remove brackets and split each string into two parts
  split_values <- gsub("\\[|\\]", "", char_vector) |>
    strsplit(" ")

  # Convert to numeric vectors
  first_values <- sapply(split_values, function(x) as.numeric(x[1]))
  second_values <- sapply(split_values, function(x) as.numeric(x[2]))

  # Return as a named list
  list(
    first = first_values,
    second = second_values
  )
}

household_initial_age_distribution_range <- extract_values_from_NetLogo_list(endStates$householdInitialAgeDistribution)
max_couple_count_distribution_range <- extract_values_from_NetLogo_list(endStates$maxCoupleCountDistribution)

endStates <- cbind(
  endStates,
  household_initial_age_distribution_min = household_initial_age_distribution_range[[1]],
  household_initial_age_distribution_max = household_initial_age_distribution_range[[2]],
  max_couple_count_distribution_min = max_couple_count_distribution_range[[1]],
  max_couple_count_distribution_max = max_couple_count_distribution_range[[2]]
)
```

Categorise end states according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

Visualise `totalIndividuals`:

```{r}
hist(endStates$totalIndividuals, breaks = 50)
```

create new categorical variable:

```{r}
max_population = 10000 # population size used in NetLogo as a stop condition when there is considerable growth  
endStates$survival <- ifelse(endStates$totalIndividuals > 0, "survival", "extinction")
endStates$survival <- paste(endStates$survival, ifelse(endStates$totalIndividuals > max_population, "with growth", ""))
```

```{r}
table(endStates$survival)
```

```{r}
barplot(table(endStates$survival))
```

```{r}
extract_age_structure <- function(char_value) {
  split_values <- gsub("\\[|\\]", "", char_value) |>
    strsplit(" ")
  return(as.numeric(split_values[[1]]))
}

extract_age_structure_batch <- function(char_vector) {
  age_struct_df <- char_vector
  for (i in 1:nrow(char_vector)) {
    
  }
}

survival_cases <- endStates %>% filter(survival != "extinction ")

selected_case <- survival_cases[500, ]

selected_case_men <- extract_age_structure(selected_case["menAgeStructure"])
selected_case_women <- extract_age_structure(selected_case["womenAgeStructure"])

selected_case[1, "totalIndividuals"]

layout(matrix(1:2, nrow = 1))
hist(selected_case_women, breaks = 101)
hist(selected_case_men, breaks = 101)
```

```{r}
# Put data into long format
age_struct_df <- tibble(
  age = c(selected_case_men, selected_case_women),
  sex = c(rep("Male", length(selected_case_men)), 
          rep("Female", length(selected_case_women)))
)

# Bin ages into groups (5-year bins, adjust as needed)
age_struct_df <- age_struct_df |>
  mutate(age_group = cut(age, breaks = seq(0, 95, 5), right = FALSE))

# Summarize counts per age group and sex
age_struct_df_summary <- age_struct_df |>
  count(sex, age_group)

# For pyramid: make men negative
age_struct_df_summary <- age_struct_df_summary |>
  mutate(count = ifelse(sex == "Male", -n, n))
```

```{r}
ggplot(age_struct_df_summary, aes(x = age_group, y = count, fill = sex)) +
  geom_col(width = 1, color = "white") +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  labs(
    title = "Population Age Pyramid",
    x = "Age group",
    y = "Population count"
  ) +
  theme_minimal()
```


Calculate the Peristera-Kostaki parameter ranges associated to survival:
```{r}
compute_param_envelope <- function(param_df, low=0.025, high=0.975){
  apply(param_df, 2, quantile, probs = c(low, high), na.rm=TRUE) %>%
    t() %>% as.data.frame() #%>% rename(lower = `2.5%`, upper = `97.5%`)
}
params_names <- gsub("-", ".", names(sobol_parameter_ranges)[7:19])
compute_param_envelope(endStates %>% filter(survival != "extinction") %>% select(params_names))
```

```{r}
params_indexes <- c(9, 12, 14:17, 20:27, 47:50)
param_names <- names(endStates)[params_indexes]

endStates_long <- endStates %>%
  pivot_longer(cols = all_of(param_names)) %>%
  mutate(name = factor(name, levels = param_names))  # Preserve order
```


```{r}
source("library/bifurcationPlot.R")

plotName = paste0(plotDirectory, "sobol_bifurcationPlot.png")

plotScale = 10

png(plotName,
    width = 150 * plotScale, height = 80 * plotScale)

ggplot(data = endStates_long, aes(x = value, y = totalIndividuals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm") +
  facet_wrap(vars(name), ncol = 6, scales = "free_x") +
  theme(
    strip.text = element_text(size = 1.2 * plotScale)
  )

dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

Train a Random Forest model to assess parameter importance:

```{r}
endStates_for_rf <- endStates[,c(params_indexes, 29)]

# Train Random Forest to predict simulation output
set.seed(123)
rf_model <- randomForest(totalIndividuals ~ ., data = endStates_for_rf, importance = TRUE, ntree = 500)

# Print model summary
print(rf_model)
```

Extract importance values and print as table:

```{r}
# Extract importance values
importance_values <- importance(rf_model)

# Convert to a data frame
importance_df <- data.frame(Parameter = rownames(importance_values),
                            MDA = importance_values[, 1], 
                            MDG = importance_values[, 2])

# Print importance scores
print(importance_df)
```

Plot feature importance colouring parameters by domain (general, mortality, fertility, nuptiality):

```{r}
source("library/variableImportancePlot.R")

plotName = paste0(plotDirectory, "sobol_variableImportancePlot.png")

color_variable_map <- rev(c("#0000a2", 
                        "#bc272d", 
                        rep("#e9c716", 4), 
                        rep("#50ad9f", 8),
                        rep("#0000a2", 4)
))

# png
png(filename = plotName, width = 129, height = 168, units="mm", res=1200)
variableImportancePlot(rf_model, plot_scale = 0.5, color_variable_map = color_variable_map, range_standardised = TRUE)
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

### Initial household population

This parameter has slight positive effect on population survival:

```{r}
summary(endStates$initialNumHouseholds[endStates$totalIndividuals == 0])
summary(endStates$initialNumHouseholds[endStates$totalIndividuals > 0])
```

```{r}
boxplot(endStates$initialNumHouseholds ~ (endStates$totalIndividuals > 0))
```

### Mortality factors

Coale-Demeny Life Tables Model level parameter has great effect:

```{r}
summary(endStates$cdmlt.level[endStates$totalIndividuals == 0])
summary(endStates$cdmlt.level[endStates$totalIndividuals > 0])
```

```{r}
boxplot(endStates$cdmlt.level ~ (endStates$totalIndividuals > 0))
```

but region parameter has a smaller effect. East tends to survive more often.  

```{r}
table(endStates$coale.demeny.region, endStates$totalIndividuals > 0)
```

```{r}
boxplot(data = endStates, totalIndividuals ~ coale.demeny.region)
```

### Fertility factors

All have a visible effect, but particularly c1.fert and sigma1.fert. 

```{r}
layout(matrix(1:4, nrow = 2, byrow = T))
boxplot(endStates$c1.fert ~ (endStates$totalIndividuals > 0), main = "c1.fert")
boxplot(endStates$mu.fert ~ (endStates$totalIndividuals > 0), main = "mu.fert")
boxplot(endStates$sigma1.fert ~ (endStates$totalIndividuals > 0), main = "sigma1.fert")
boxplot(endStates$sigma2.fert ~ (endStates$totalIndividuals > 0), main = "sigma2.fert")
```

### Nuptiality factors

Residential rule showing no clear effect:

```{r}
table(endStates$residence.rule, endStates$totalIndividuals > 0)
```

The effect of the acceptable kinship degree is also unclear: 

```{r}
boxplot(endStates$acceptable.kinship.degree.for.couples ~ (endStates$totalIndividuals > 0))
```

All nuptiality parameters for both men and women present a small, yet apparently consistent effect. c1 and sigma parameters have a positive effect while mu has a smaller negative effect.

```{r}
layout(matrix(1:8, nrow = 2, byrow = T))
boxplot(endStates$c1.women ~ (endStates$totalIndividuals > 0), main = "c1.women")
boxplot(endStates$mu.women ~ (endStates$totalIndividuals > 0), main = "mu.women")
boxplot(endStates$sigma1.women ~ (endStates$totalIndividuals > 0), main = "sigma1.women")
boxplot(endStates$sigma2.women ~ (endStates$totalIndividuals > 0), main = "sigma2.women")
boxplot(endStates$c1.men ~ (endStates$totalIndividuals > 0), main = "c1.men")
boxplot(endStates$mu.men ~ (endStates$totalIndividuals > 0), main = "mu.men")
boxplot(endStates$sigma1.men ~ (endStates$totalIndividuals > 0), main = "sigma1.men")
boxplot(endStates$sigma2.men ~ (endStates$totalIndividuals > 0), main = "sigma2.men")
```

## Timing of survival

```{r}
hist(endStates$X.step.)
```

```{r}
table(endStates$X.step. > 0)
```

```{r}
plot(endStates$initialNumHouseholds ~ endStates$X.step.)
```

### With patrilineal rule

Loading simulation experiments' end states:

```{r}
endStates_patri <- read.csv("simdata/v1.1/householdDemography v.1.1 - kinship tabu for new couples exp-endstates-table_patri.csv", skip = 6)
```

Process parameters in list format in NetLogo:

```{r}
extract_values_from_NetLogo_list <- function(char_vector) {
  # Remove brackets and split each string into two parts
  split_values <- gsub("\\[|\\]", "", char_vector) |>
    strsplit(" ")

  # Convert to numeric vectors
  first_values <- sapply(split_values, function(x) as.numeric(x[1]))
  second_values <- sapply(split_values, function(x) as.numeric(x[2]))

  # Return as a named list
  list(
    first = first_values,
    second = second_values
  )
}

household_initial_age_distribution_range <- extract_values_from_NetLogo_list(endStates_patri$household.initial.age.distribution)
max_couple_count_distribution_range <- extract_values_from_NetLogo_list(endStates_patri$max.couple.count.distribution)

endStates_patri <- cbind(
  endStates_patri,
  household_initial_age_distribution_min = household_initial_age_distribution_range[[1]],
  household_initial_age_distribution_max = household_initial_age_distribution_range[[2]],
  max_couple_count_distribution_min = max_couple_count_distribution_range[[1]],
  max_couple_count_distribution_max = max_couple_count_distribution_range[[2]]
)
```

Categorise end states according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

Visualise `totalIndividuals`:

```{r}
hist(endStates_patri$totalIndividuals, breaks = 50)
```

create new categorical variable:

```{r}
max_population = 10000 # population size used in NetLogo as a stop condition when there is considerable growth  
endStates_patri$survival <- ifelse(endStates_patri$totalIndividuals > 0, "survival", "extinction")
endStates_patri$survival <- paste(endStates_patri$survival, ifelse(endStates_patri$totalIndividuals > max_population, "with growth", ""))
```

```{r}
table(endStates_patri$survival)
```

```{r}
barplot(table(endStates_patri$survival))
```

```{r}
extract_age_structure <- function(char_value) {
  split_values <- gsub("\\[|\\]", "", char_value) |>
    strsplit(" ")
  return(as.numeric(split_values[[1]]))
}

extract_age_structure_batch <- function(char_vector) {
  age_struct_df <- char_vector
  for (i in 1:nrow(char_vector)) {
    
  }
}

survival_cases <- endStates_patri %>% filter(survival != "extinction ")

selected_case <- survival_cases[500, ]

selected_case_men <- extract_age_structure(selected_case["menAgeStructure"])
selected_case_women <- extract_age_structure(selected_case["womenAgeStructure"])

selected_case[1, "totalIndividuals"]

layout(matrix(1:2, nrow = 1))
hist(selected_case_women, breaks = 101)
hist(selected_case_men, breaks = 101)
```

```{r}
# Put data into long format
age_struct_df <- tibble(
  age = c(selected_case_men, selected_case_women),
  sex = c(rep("Male", length(selected_case_men)), 
          rep("Female", length(selected_case_women)))
)

# Bin ages into groups (5-year bins, adjust as needed)
age_struct_df <- age_struct_df |>
  mutate(age_group = cut(age, breaks = seq(0, 95, 5), right = FALSE))

# Summarize counts per age group and sex
age_struct_df_summary <- age_struct_df |>
  count(sex, age_group)

# For pyramid: make men negative
age_struct_df_summary <- age_struct_df_summary |>
  mutate(count = ifelse(sex == "Male", -n, n))
```

```{r}
ggplot(age_struct_df_summary, aes(x = age_group, y = count, fill = sex)) +
  geom_col(width = 1, color = "white") +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  labs(
    title = "Population Age Pyramid",
    x = "Age group",
    y = "Population count"
  ) +
  theme_minimal()
```


Calculate the Peristera-Kostaki parameter ranges associated to survival:
```{r}
compute_param_envelope <- function(param_df, low=0.025, high=0.975){
  apply(param_df, 2, quantile, probs = c(low, high), na.rm=TRUE) %>%
    t() %>% as.data.frame() #%>% rename(lower = `2.5%`, upper = `97.5%`)
}
params_names <- gsub("-", ".", names(sobol_parameter_ranges)[7:19])
compute_param_envelope(endStates_patri %>% filter(survival != "extinction") %>% select(params_names))
```

```{r}
params_indexes <- c(9, 12, 14:17, 20:27, 47:50)
param_names <- names(endStates_patri)[params_indexes]

endStates_patri_long <- endStates_patri %>%
  pivot_longer(cols = all_of(param_names)) %>%
  mutate(name = factor(name, levels = param_names))  # Preserve order
```


```{r}
source("library/bifurcationPlot.R")

plotName = paste0(plotDirectory, "sobol_bifurcationPlot_patri.png")

plotScale = 10

png(plotName,
    width = 150 * plotScale, height = 80 * plotScale)

ggplot(data = endStates_patri_long, aes(x = value, y = totalIndividuals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm") +
  facet_wrap(vars(name), ncol = 6, scales = "free_x") +
  theme(
    strip.text = element_text(size = 1.2 * plotScale)
  )

dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

Train a Random Forest model to assess parameter importance:

```{r}
endStates_patri_for_rf <- endStates_patri[,c(params_indexes, 29)]

# Train Random Forest to predict simulation output
set.seed(123)
rf_model <- randomForest(totalIndividuals ~ ., data = endStates_patri_for_rf, importance = TRUE, ntree = 500)

# Print model summary
print(rf_model)
```

Extract importance values and print as table:

```{r}
# Extract importance values
importance_values <- importance(rf_model)

# Convert to a data frame
importance_df <- data.frame(Parameter = rownames(importance_values),
                            MDA = importance_values[, 1], 
                            MDG = importance_values[, 2])

# Print importance scores
print(importance_df)
```

Plot feature importance colouring parameters by domain (general, mortality, fertility, nuptiality):

```{r}
source("library/variableImportancePlot.R")

plotName = paste0(plotDirectory, "sobol_variableImportancePlot_patri.png")

color_variable_map <- rev(c("#0000a2", 
                        "#bc272d", 
                        rep("#e9c716", 4), 
                        rep("#50ad9f", 8),
                        rep("#0000a2", 4)
))

# png
png(filename = plotName, width = 129, height = 168, units="mm", res=1200)
variableImportancePlot(rf_model, plot_scale = 0.5, color_variable_map = color_variable_map, range_standardised = TRUE)
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```


## Trajectories

>Run experiments in NetLogo Behaviour Space using average  parameter values for parametric demographic models within the configurations with surviving population in the former experiment.

Loading simulation experiments' end states:

```{r}
trajectories_v11 <- read.csv("simdata/v1.1/householdDemography v.1.1 - kinship tabu for new couples exp-trajectories-table.csv", skip = 6)
trajectories_v122 <- read.csv("simdata/v1.2.2/householdDemography v.1.2.2 - mortality regulated by density with weighted k and labour exp-trajectories-table.csv", skip = 6)
```

Process parameters in list format in NetLogo:

```{r}
extract_values_from_NetLogo_list <- function(char_vector) {
  # Remove brackets and split each string into two parts
  split_values <- gsub("\\[|\\]", "", char_vector) |>
    strsplit(" ")

  # Convert to numeric vectors
  first_values <- sapply(split_values, function(x) as.numeric(x[1]))
  second_values <- sapply(split_values, function(x) as.numeric(x[2]))

  # Return as a named list
  list(
    first = first_values,
    second = second_values
  )
}

household_initial_age_distribution_range_v11 <- extract_values_from_NetLogo_list(trajectories_v11$household.initial.age.distribution)
max_couple_count_distribution_range_v11 <- extract_values_from_NetLogo_list(trajectories_v11$max.couple.count.distribution)

household_initial_age_distribution_range_v122 <- extract_values_from_NetLogo_list(trajectories_v122$household.initial.age.distribution)
max_couple_count_distribution_range_v122 <- extract_values_from_NetLogo_list(trajectories_v122$max.couple.count.distribution)

trajectories_v11 <- cbind(
  trajectories_v11,
  household_initial_age_distribution_min = household_initial_age_distribution_range_v11[[1]],
  household_initial_age_distribution_max = household_initial_age_distribution_range_v11[[2]],
  max_couple_count_distribution_min = max_couple_count_distribution_range_v11[[1]],
  max_couple_count_distribution_max = max_couple_count_distribution_range_v11[[2]]
)

trajectories_v122 <- cbind(
  trajectories_v122,
  household_initial_age_distribution_min = household_initial_age_distribution_range_v122[[1]],
  household_initial_age_distribution_max = household_initial_age_distribution_range_v122[[2]],
  max_couple_count_distribution_min = max_couple_count_distribution_range_v122[[1]],
  max_couple_count_distribution_max = max_couple_count_distribution_range_v122[[2]]
)
```

```{r}
params_indexes <- c(3:21)
param_names <- names(trajectories_v11)[params_indexes]
params_v11 <- trajectories_v11 %>% 
  select(all_of(param_names)) %>% 
  distinct() %>% 
  t %>%
  apply(1, unique)
params_v11
```

```{r}
params_indexes <- c(3:21, 22:24)
param_names <- names(trajectories_v122)[params_indexes]
params_v122 <- trajectories_v122 %>% 
  select(all_of(param_names)) %>% 
  distinct() %>% 
  t %>%
  apply(1, unique)
params_v122
```

```{r}
plot_name <- file.path(plotDirectory, paste0("HouseholdDemography-trajectories_v11-v122.", plot_file_format))

if (plot_file_format == "png") {
  graphic_scale <- 1
  legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
  axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5

  png(plot_name, width = graphic_scale * 1600, height = graphic_scale * 800, res = 300, pointsize = 4)
} else if (plot_file_format == "eps") {
  graphic_scale = 1.2
  # font_rescale = -0.2
  # axis_text_rescale = -0.1
  # margin_text_rescale = -0.5

  extrafont::loadfonts(device = "postscript")
  grDevices::cairo_ps(filename = plot_name ,
                      pointsize = 12,
                      width = graphic_scale * 8,
                      height = graphic_scale * 4,
                      onefile = FALSE,
                      family = "sans"
                      )
}

p1 <- ggplot(trajectories_v11[trajectories_v11$X.step. < 600,], aes(x = X.step., y = totalIndividuals, group = X.run.number., color = cdmlt.level)) +
  geom_line(linewidth = .1) +
  xlab("time (years)") +
  ylab("population (individuals)") +
  ylim(c(0, 7000)) +
  guides(color = "none") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10 * (graphic_scale + axis_title_rescale)),
        axis.text = element_text(size = 10 * (graphic_scale + axis_text_rescale)))

p2 <- ggplot(trajectories_v122[trajectories_v122$X.step. < 600,], aes(x = X.step., y = totalIndividuals, group = X.run.number., color = cdmlt.level)) +
  geom_line(linewidth = .1) +
  geom_hline(yintercept = as.numeric(params_v122$par_carrying.capacity)) +
  xlab("time (years)") +
  ylab("") +
  ylim(c(0, 7000)) +
  guides(color = guide_legend(title="Coale-Demeny\nLife Table Level", override.aes = list(linewidth = 2), reverse = TRUE)) +
  theme_minimal() +
  theme(axis.title = element_text(size = 10 * (graphic_scale + axis_title_rescale)),
        axis.text = element_text(size = 10 * (graphic_scale + axis_text_rescale)),
        legend.title = element_text(size = 10 * (graphic_scale + legend_text_rescale)),
        legend.text = element_text(size = 10 * (graphic_scale + legend_text_rescale)),
        legend.key.width = unit(2 * (graphic_scale + legend_key_rescale), 'lines'),
        legend.key.height = unit(2 * (graphic_scale + legend_key_rescale), 'lines'))

p3 <- grid.arrange(p1, p2, ncol = 2, widths = c(4, 5.1))

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plot_name)
```

