---
execute:
  echo: true
  warning: false
  message: false
params:
  show_diagnostics: false
---

# Base model with daily iteration (v.1.3)

## Setup

Set constant variables for this chapter.

```{r}
#| label: setup_constants
MODEL_VERSION <- "v1.3"
EXPORT_FIGURES <- TRUE
REMAKE_RF <- FALSE
REMAKE_DP <- FALSE
```

Load libraries, sources and set ggplot theme for the notebook.

```{r} 
#| label: setup_libraries
library(data.table)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(ggtext)
library(zoo)
library(stringr)
library(effectsize)
library(randomForest)
library(randtoolbox)
library(rngWELL)
library(pdp)
library(forcats)
library(patchwork)
library(scales)
```

```{r} 
#| label: setup_load_sources
source("library/source_folder.R")
source_folder("library")
source_folder("library/structure")
source_folder("library/sensitivity")
```

```{r} 
#| label: setup_ggplot_theme
theme_set(
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold"),
      plot.title = element_text(face = "bold")
    )
)
```

## Experiment design

Sobol sampling is used to generate a diverse set of parameter combinations across the defined parameter space. This allows us to explore how different parameter configurations affect the model's outcomes and to identify which parameters are most influential in driving population survival or extinction. The actual sampling was done with `generate_sobol.R` script, which generates a sample with 2000 parameter combinations across the parameter ranges defined in `parameter_ranges.R`. These were generated previously with `library/generate_sobol.R`.

```{r} 
#| label: sobol_load
# Load object as Rdata file
load("data/v13_sobol_parameter_values.RData")
```

Sobol sampling is applied to the following parameter ranges:

```{r} 
#| label: sobol_ranges_display
parameter_ranges_v13 <- as.data.frame(parameter_ranges$v1.3)

knitr::kable(
  parameter_ranges_v13 |>
    as.matrix() |>
    t() |>
    as.data.frame() |>
    setNames(c("min", "max"))
)
```

```{r} 
#| label: sobol_pairs 
#| include: !expr params$show_diagnostics 
#| fig.cap: "Parameter ranges explored in the Sobol sampling for v1.3."
# Visualising parameter space with `pairs()`:
plotName = paste0("diagnostics/", MODEL_VERSION, "_sobol_param_pairsPlot.png")

knitr::include_graphics(plotName)
```

The generated parameter combinations were then used to run simulations in NetLogo's Behaviour Space, for each combination of `residence-rule` ("matrilineal-matrilocal" and "patrilineal-patrilocal") and `coale-demeny-region` ("north", "south", "east", and "west"), varying the `SEED` parameter (n = 3) to introduce stochasticity, for a total of 48000 simulations. Moreover, all runs assume a fixed `maturity-age-in-years` of 15 years, a year length of 365 days (no leap years), and a gestation period of 280 days (40 weeks). Remember that `maturity-age-in-years` determines a conservative age solely for social maturity, used as a criterion for household viability, while marital and reproductive maturity is regulated probabilistically through the Peristera-Kostaki parameters in the fertility and nuptiality submodels.

## Data loading and preprocessing

Load simulation end states generated with the Sobol-sampled parameter combinations and sampled trajectories datasets. Preprocess datasets to adapt from NetLogo's specific format and categorise end states and trajectories. 

```{r}
#| label: preprocessing_vars
growth_threshold = 5000 # endstates with population equal or higher than this are considered "Growth" cases
```

```{r} 
#| label: preprocessing
endstates <- load_endstates(MODEL_VERSION) |>
  preprocess_simdata(
    model_version = MODEL_VERSION,
    growth_threshold = growth_threshold
  )
```

```{r} 
#| label: preprocessing_trajectories
file_name <- "data/v13_trajectories.RData"
if (!file.exists(file_name))
{
  sampled_trajectories <- load_and_preprocess_trajectories_v13_v14x_single(
    MODEL_VERSION,
    growth_threshold = growth_threshold,
    sample_interval_days = 365,
    step_start = 1,
    step_end = 182500, # max-iterations is 500 years or 182500 days
    ignored_variables = c(
        "womenAgeStructure",
        "menAgeStructure"
    ),
    keep_variables = c(
        "model_version",
        "run_unique_id",
        "run_number",
        "residence_rule",
        "max_iterations",
        "max_iterations_in_days",
        "max_population", 
        names(parameter_ranges_v13), 
        "coale_demeny_region", 
        "maturity_age_in_years", 
        "gestation_period_in_days", 
        "year_length_in_days",
        "step",
        "currentYear",
        "currentDayOfYear",
        "log_totalIndividuals",
        "totalIndividuals",
        "totalHouseholds",
        "survival"
    )
  )
  
  save_trajectories(
    sampled_trajectories,
    model_version = MODEL_VERSION
  )
}
```

```{r}
#| label: preprocessing_trajectories_symmetric
rm(sampled_trajectories)
file_name <- "data/v13_trajectories_symmetric.RData"
if (file.exists(file_name))
{
  sampled_trajectories_symmetric <- readRDS(file_name)
} else {
  sampled_trajectories_symmetric <- load_and_preprocess_trajectories_v13_v14x_single(
    MODEL_VERSION,
    suffix = "symmetric",
    growth_threshold = growth_threshold,
    sample_interval_days = 365,
    step_start = 1,
    step_end = 182500, # max-iterations is 500 years or 182500 days
    ignored_variables = c(
        "womenAgeStructure",
        "menAgeStructure"
    ),
    keep_variables = c(
        "model_version",
        "run_unique_id",
        "run_number",
        "residence_rule",
        "max_iterations",
        "max_iterations_in_days",
        "max_population", 
        names(parameter_ranges_v13), 
        "coale_demeny_region", 
        "maturity_age_in_years", 
        "gestation_period_in_days", 
        "year_length_in_days",
        "step",
        "currentYear",
        "currentDayOfYear",
        "log_totalIndividuals",
        "totalIndividuals",
        "totalHouseholds",
        "survival"
    )
  )
  
  save_trajectories(
    sampled_trajectories_symmetric,
    model_version = MODEL_VERSION,
    suffix = "symmetric"
  )
}
```

```{r}
#| label: preprocessing_trajectories_cdmlt_level
rm(sampled_trajectories_symmetric)
file_name <- "data/v13_trajectories_cdmlt_level.RData"
if (file.exists(file_name))
{
  sampled_trajectories_cdmlt_level <- readRDS(file_name)
} else {
  sampled_trajectories_cdmlt_level <- load_and_preprocess_trajectories_v13_v14x_single(
    MODEL_VERSION,
    suffix = "cdmlt-level",
    growth_threshold = growth_threshold,
    sample_interval_days = 365,
    step_start = 1,
    step_end = 182500, # max-iterations is 500 years or 182500 days
    ignored_variables = c(
        "womenAgeStructure",
        "menAgeStructure"
    ),
    keep_variables = c(
        "model_version",
        "run_unique_id",
        "run_number",
        "residence_rule",
        "max_iterations",
        "max_iterations_in_days",
        "max_population", 
        names(parameter_ranges_v13), 
        "coale_demeny_region", 
        "maturity_age_in_years", 
        "gestation_period_in_days", 
        "year_length_in_days",
        "step",
        "currentYear",
        "currentDayOfYear",
        "log_totalIndividuals",
        "totalIndividuals",
        "totalHouseholds",
        "survival"
    )
  )
  
  save_trajectories(
    sampled_trajectories_cdmlt_level,
    model_version = MODEL_VERSION,
    suffix = "cdmlt_level"
  )
}
```

```{r}
#| label: preprocessing_trajectories_c1_fert
rm(sampled_trajectories_cdmlt_level)
file_name <- "data/v13_trajectories_c1_fert.RData"
if (file.exists(file_name))
{
  sampled_trajectories_c1_fert <- readRDS(file_name)
} else {
  sampled_trajectories_c1_fert <- load_and_preprocess_trajectories_v13_v14x_single(
    MODEL_VERSION,
    suffix = "c1-fert",
    growth_threshold = growth_threshold,
    sample_interval_days = 365,
    step_start = 1,
    step_end = 182500, # max-iterations is 500 years or 182500 days
    ignored_variables = c(
        "womenAgeStructure",
        "menAgeStructure"
    ),
    keep_variables = c(
        "model_version",
        "run_unique_id",
        "run_number",
        "residence_rule",
        "max_iterations",
        "max_iterations_in_days",
        "max_population", 
        names(parameter_ranges_v13), 
        "coale_demeny_region", 
        "maturity_age_in_years", 
        "gestation_period_in_days", 
        "year_length_in_days",
        "step",
        "currentYear",
        "currentDayOfYear",
        "log_totalIndividuals",
        "totalIndividuals",
        "totalHouseholds",
        "survival"
    )
  )
  
  save_trajectories(
    sampled_trajectories_c1_fert,
    model_version = MODEL_VERSION,
    suffix = "c1_fert"
  )
}
```

As with v1.1, the variable `survival` is created during preprocessing to classify whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`)..

Because of the sheer number of daily iterations required to approach comparable time scales of previous versions, trajectories datasets are simply too large to be loaded conventionally in R. For this reason, we need to use the function argument `ignored_variables` (alternatively also `selected_variables` is available) to skip loading unnecessary variables, as well as potentially restricting the span of the simulation data to a subset of iteration steps (`start_step`, `end_step`). The different trajectory-type data frames must be loaded, preprocessed, saved as RData, and clear, so that only one is in memory simultaneously.

### Surviving population structure

```{r} 
#| label: age_structure_prepare
group_vars = c(
  "survival",
  "residence_rule",
  "coale_demeny_region",
  "mortality_regime",
  "fertility_intensity",
  "fertility_timing",
  "fertility_spread_pre_peak",
  "fertility_spread_post_peak",
  "marriage_timing_women",
  "marriage_timing_men",
  "marriage_intensity_women",
  "marriage_intensity_men",
  "marriage_spread_pre_peak_women",
  "marriage_spread_pre_peak_men",
  "marriage_spread_post_peak_women",
  "marriage_spread_post_peak_men",
  "amenorrhea_length"
)

age_structures <- endstates |>

  filter(
    survival != "Extinction"
  ) |>
  
  characterise_mort_fert_nupt_regimes() |>

  prepare_age_structure(
    age_breaks = seq(0, 100, 5),
    normalize = TRUE,
    group_vars = group_vars
  )

age_structure_indicators <- age_structures |>
  compute_age_structure_indicators(group_vars = group_vars)

endstates <- endstates |>
  left_join(
    age_structure_indicators |>
      select(
        run_unique_id,
        median_age,
        youth_share,
        elderly_share,
        dependency_ratio
      ),
    by = "run_unique_id"
  )
```

Age structures of surviving populations are prepared by categorising them according to the same grouping variables used in the sensitivity analysis, including categorical parameters (i.e., `residence_rule`, `coale_demeny_region`) and parameters converted into binomial categories based on their median value, exclusively those regulating demographic probabilites (e.g., `mortality_regime`, `fertility_intensity`, etc.).

Furthermore, this selection includes the categorical variable `survival` that distinguishes between populations that became extinct, persisted without reaching the maximum population size, and reached the maximum population size. 
This allows us to compare the age structures of surviving populations across different demographic regimes and residence systems, and to identify patterns or differences associated with population persistence or growth.

Moreover, a set of indicators is computed for each age structure, including median age, dependency ratios, and measures of age-structure irregularity, which can provide further insights into the demographic characteristics of surviving populations and how they relate to different parameter regimes and residence systems. 
These indicators can help identify specific features of the age structure associated with population persistence or growth, and understand how different demographic regimes and residence systems may influence the age structure of surviving populations. Namely, the indicators computed include:

- `median_age`: the age that divides the population into two halves, with 50% of the population being younger and 50% being older.  
- `youth_share`: the ratio of the population aged 0-14 to the population aged 15-64, which indicates the proportion of dependents in the population.  
- `reproductive_share`: the ratio of the population aged 15-64 to the population aged 65 and over, which indicates the proportion of reproductive-age individuals in the population.  
- `elderly_share`: the ratio of the population aged 65 and over to the population aged 15-64, which indicates the proportion of elderly dependents in the population.  
- `total_dependency_ratio`: the ratio of the population aged 0-14 and 65 and over to the population aged 15-64, which indicates the overall proportion of dependent individuals in the population.  

## Overview of end states and dynamics

### Population persistence

We start with an overview of end states across residence rules and the categorical "regions" of the Coale-Demeny model (`coale-demeny-region`), which are the two categorical parameters explored outside the Sobol sampling. The histogram shows the distribution of final population sizes for each residence rule, with different colours indicating whether the population became extinct or survived. The vertical line represents the maximum population size defined in the model, which helps visually assess how many simulations reached or exceeded this threshold. The x-axis can be set to a logarithmic scale to better visualise differences in population sizes among those that survived.

```{r} 
#| label: endstates_distrib 
#| fig.cap: "Distribution of final population sizes across residence rules and Coale-Demeny regions."
log_max_population <- max(endstates$log_totalIndividuals)

endstates_distrib_size_plot <- endstates |>
  plot_endstate_variable_distribution(
    main_var = "log_totalIndividuals",
    fill_var = "survival",
    facet_var = c("coale_demeny_region", "residence_rule"),
    bins = 40,
    x_axis_label = "Final population size",
    y_axis_label = "Simulation count",
    threshold_values = log_max_population,
    thresholds_linetype = 2,
    x_axis_log10 = TRUE,
    x_max = log_max_population,
    y_axis_log10 = TRUE
  )
endstates_distrib_size_plot
```

```{r}
#| label: endstates_distrib_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    endstates_distrib_size_plot,
    "v13_endstates_distrib"
  )
}
```

```{r} 
#| label: outcome_prop 
#| fig.cap: "Proportion of simulations across survival category, residence rules and Coale-Demeny regions."
outcome_prop_plot <- endstates |>
  plot_category_proportion_stacked(
    x_var = "residence_rule",
    fill_var = "survival",
    facet_var = "coale_demeny_region",
    facet_ncol = 2,
    facet_scales = "fixed",
    y_axis_label = "Share of simulations"
  )
outcome_prop_plot
```

```{r}
#| label: outcome_prop_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    outcome_prop_plot,
    "v13_outcome_prop"
  )
}
```

```{r} 
#| label: outcome_prop_noextinction 
#| fig.cap: "Proportion of surviving simulations across residence rules and Coale-Demeny regions."
outcome_prop_plot_noextinction <- endstates |>
  filter(survival != "Extinction") |>
  plot_category_proportion_stacked(
    x_var = "residence_rule",
    fill_var = "survival",
    facet_var = "coale_demeny_region",
    facet_ncol = 2,
    facet_scales = "fixed",
    y_axis_label = "Share of simulations"
  )
outcome_prop_plot_noextinction
```

```{r}
#| label: outcome_prop_noextinction_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    outcome_prop_plot_noextinction,
    "v13_outcome_prop_plot_noextinction"
  )
}
```

As in previous versions, only a small fraction of simulations has populations that survive, with extinction remaining the dominant outcome across the explored parameter space. Compared with v1.1, persistence occurs across a wide range of population sizes, whereas growth cases are generally less frequent. However, in this version, growth cases are considerably more likely to occur under the patrilocal-patrilineal rule: c. 50% of surviving populations compared to c. 37% in cases under the matrilocal-matrilineal rule. Taken as the exploration as a whole, `coale-demeny-region` does not seem to have a strong influence on the overall distribution of end states.

### Temporal trajectories

```{r}
#| label: trajectories_load_1
rm(sampled_trajectories_symmetric, preprocessing_trajectories_cdmlt_level, sampled_trajectories_c1_fert)
sampled_trajectories <- readRDS("data/v13_trajectories.RData")
```

The parameter configuration use to generate the trajectories is the same across residence rules, except for the `residence_rule` parameter itself. This allows us to directly compare the population dynamics under different residence systems while controlling for other factors. The configuration used for this was:

```{r} 
#| label: trajectories_size_params_display 
#| tab.cap: "Parameter configuration for the population trajectories sampled."
knitr::kable(sampled_trajectories |> 
               mutate(
                 max_iterations_in_days = max_iterations * year_length_in_days,
                 max_iterations_in_days_cut = max(step)
               ) |>
               select(all_of(c("max_iterations",
                               "max_iterations_in_days",
                               "max_iterations_in_days_cut", # show this in case loading was partial
                               "max_population", 
                               names(parameter_ranges_v13), 
                               "coale_demeny_region", 
                               "maturity_age_in_years", "gestation_period_in_days", "year_length_in_days" ))) |> 
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

Dynamics of population trajectories across time steps, faceted by residence rule. The plot shows individual trajectories with low opacity and the median trajectory with higher opacity. The y-axis is on a logarithmic scale to better visualize differences in population sizes over time, especially when there are large variations between trajectories.

```{r} 
#| label: trajectories_size_plot 
#| fig.cap: "A sample of population trajectories across time steps, faceted by residence rule."
pop_trajectories_plot <- sampled_trajectories |>
  plot_trajectories(
    x_var = "currentYear",
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.08,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = "survival",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "max_population",
    thresholds_linetype = 2
  )
pop_trajectories_plot
```

```{r}
#| label: trajectories_size_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_plot,
    "v13_pop_trajectories"
  )
}
```

Compared with v1.1, the overall trajectory shapes are smoother and less punctuated, reflecting the transition from annual to daily demographic events. Population expansion is no longer driven by abrupt cohort jumps, but by the gradual accumulation of births and deaths across time. Despite this finer temporal structure, the asymmetry between residence rules remains visible, with patrilocal-patrilineal populations reaching higher median growth trajectories under the same parameter configuration.

In order to confirm that the above asymmetry is not caused by the asymmetry in the Peristera-Kostaki nuptiality parameters, the same experiment was performed with the following configuration:

```{r}
#| label: trajectories_symmetric_load
rm(sampled_trajectories, preprocessing_trajectories_cdmlt_level, sampled_trajectories_c1_fert)
sampled_trajectories_symmetric <- readRDS("data/v13_trajectories_symmetric.RData")
```

```{r} 
#| label: trajectories_size_params_symmetric_display 
#| tab.cap: "Parameter configuration for the population trajectories sampled."
knitr::kable(sampled_trajectories_symmetric |> 
               mutate(
                 max_iterations_in_days = max_iterations * year_length_in_days,
                 max_iterations_in_days_cut = max(step)
               ) |>
               select(all_of(c("max_iterations",
                               "max_iterations_in_days",
                               "max_iterations_in_days_cut", # show this in case loading was partial
                               "max_population", 
                               names(parameter_ranges_v13), 
                               "coale_demeny_region", 
                               "maturity_age_in_years", "gestation_period_in_days", "year_length_in_days" ))) |> 
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

Dynamics of population trajectories across time steps, faceted by residence rule. The plot shows individual trajectories with low opacity and the median trajectory with higher opacity. The y-axis is on a logarithmic scale to better visualize differences in population sizes over time, especially when there are large variations between trajectories.

```{r} 
#| label: trajectories_size_symmetric_plot 
#| fig.cap: "A sample of population trajectories across time steps, faceted by residence rule."
pop_trajectories_symmetric_plot <- sampled_trajectories_symmetric |>
  plot_trajectories(
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.08,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = NULL,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "max_population",
    thresholds_linetype = 2
  )
pop_trajectories_symmetric_plot
```

```{r}
#| label: trajectories_size_symmetric_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_symmetric_plot,
    "v13_pop_trajectories_symmetric"
  )
}
```

Again, we can confirm that the asymmetry between residence rules is not caused by the Peristera-Kostaki nuptiality parameters.

### Household composition

```{r}
#| label: household_structure_plot
#| fig.cap: ""
household_structure_plot <- endstates |> 
  plot_household_structure_faceted_fixed(
    color_var = "max_couple_count_distribution_max",
    point_alpha = 0.5,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "fixed",
    xy_log10 = TRUE,
    x_axis_label = "number of households",
    y_axis_label = "number of individuals",
    color_label = "Max. couple count allowed in a household\n(parametric maximum)",
    legend_position = "bottom"
  )
household_structure_plot
```

```{r}
#| label: household_structure_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    household_structure_plot,
    "v13_household_structure"
  )
}
```

The scatter plot immediately shows that the relationship between `log_totalIndividuals` and `log_totalHouseholds` is almost perfectly linear. It means that household formation scales approximately proportionally to population growth. The interesting information is therefore not in the total counts but in the residual variation around that relationship.

```{r}
#| label: trajectories_load_2
rm(sampled_trajectories_symmetric, preprocessing_trajectories_cdmlt_level, sampled_trajectories_c1_fert)
sampled_trajectories <- readRDS("data/v13_trajectories.RData")
```

```{r}
#| label: trajectories_household_structure_plot 
#| fig.cap: "A sample of household population trajectories across time steps, faceted by residence rule."
trajectories_household_structure_plot <- 
  sampled_trajectories |>
  group_by(residence_rule) |>
  mutate(
    household_structure_ratio = (totalIndividuals / totalHouseholds)
  ) |>
  ungroup() |>
  plot_trajectories(
    y_var = "household_structure_ratio",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = FALSE,
    y_axis_label = "Mean individual count per household",
    group_var = NULL,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = NULL
  )
trajectories_household_structure_plot
```

```{r}
#| label: trajectories_household_structure_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    trajectories_household_structure_plot,
    "v13_trajectories_household_structure",
    width = 8,
    height = 4
  )
}
```

```{r}
#| label: trajectories_household_structure_plot_peaks
trajectories_household_structure_plot_peaks <- 
  sampled_trajectories |>
  group_by(residence_rule) |>
  mutate(
    household_structure_ratio = if_else(totalHouseholds == 0, 0, (totalIndividuals / totalHouseholds)),
    # Find the absolute peak value within the entire residence rule
    global_max = max(household_structure_ratio, na.rm = TRUE)
  ) |>
  group_by(residence_rule, run_number) |>
  mutate(
    # Check if this specific run contains that global peak value
    selected_trajectory = any(household_structure_ratio == global_max)
  ) |>
  ungroup() |>
  plot_trajectories(
    y_var = "household_structure_ratio",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = FALSE,
    y_axis_label = "Mean individual count per household",
    group_var = NULL,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "fixed",
    y_threshold_var = NULL
  ) +
  # Add the highlighted layer on top
  geom_line(
    data = \(df) filter(df, selected_trajectory == TRUE),
    aes(
      x = step, 
      y = household_structure_ratio,
      group = run_number
    ),
    color = "red",
    linewidth = 1,
    alpha = 1
  )
```

```{r}
#| label: trajectories_household_structure_plot_peaks_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    trajectories_household_structure_plot_peaks,
    "v13_household_structure_trajectories_peaks",
    width = 8,
    height = 4
  )
}
```

```{r} 
#| label: trajectories_household_structure_plot_peaks_display 
#| fig.cap: "A sample of household population trajectories across time steps, faceted by residence rule, highlighting in red the trajectories that achieve the maximum value in each subset."
plotName = paste0("plots/v13_household_structure_trajectories_peaks.png")

knitr::include_graphics(plotName)
```

With the change to daily cycles, the two residence systems generate similar household structures, with patrilocal-patrilineal yielding a slightly lower average number of individuals per household. Discounting the greater stability and survivability of matrilocal-matrilineal populations, the general trend is to converge around a relatively stable ratio (c. 5, for this parameter configuration). 

The instability caused by escalating household size still seems associated with later collapse, consistent with what was observed in previous versions. As in v1.1 and in contrast to v1.2.x, patrilocal-patrilineal populations do not display higher ratio peaks.

### Age structure

#### Survival categories

```{r} 
#| label: age_structure_survival_pyramid_plot 
#| fig.cap: "Aggregate age structures of surviving populations across survival categories."
age_structures |>
  aggregate_age_structure(
    group_vars = c("survival"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "survival",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_survival_median_plot 
#| fig.cap: "Distribution of median age of surviving populations across survival categories."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival"
)
```

```{r} 
#| label: age-structure-survival-indicators-summary 
#| tab.cap: "Summary of age structure indicators for surviving populations across survival categories."
knitr::kable(indicators_summary(age_structure_indicators, group_vars = "survival"))
```

```{r} 
#| label: age_structure_survival_median_cohend 
#| tab.cap: "Cohen's d for median age across survival categories."
knitr::kable(
  interpret_cohens_d(
    cohens_d(
      median_age ~ survival,
      data = age_structure_indicators
    ),
    rules = "cohen1988"
  )
)
```

As in v1.1, populations that reach growth conditions tend to maintain substantially younger age structures than populations that merely persist. This distinction is stronger in v1.3, suggesting that the daily reformulation—especially the explicit modelling of gestation and postpartum amenorrhoea—amplifies the demographic effects of reproductive timing.
Because births are distributed more gradually rather than accumulating in abrupt annual cohorts, v1.3 reinforces the contrast between populations characterised by sustained renewal and those drifting toward ageing persistence.

The age structures observed in v1.3 end states differ noticeably from those produced by v1.1. In general, v1.3 generates smaller pre-adult cohorts (< 15 years old) and relatively larger adult and elder cohorts (15–75 years old), shifting median ages upward to approximately 21–25 years.
Despite these differences, the relationship between survival category and age structure remains clear and consistent with previous model versions: larger and growing populations are generally younger, whereas lower-level persistent populations tend to be older and more balanced.
This distinction is important because persistence does not always imply demographic security. Some persistent populations still display fragile age distributions that could lead to later collapse, including outlying cases in which the median age falls within elderly cohorts.

#### Residence rule and Coale-Demeny regions

```{r} 
#| label: age_structure_pyramid_plot_cdr_rr
pyramid_plot <- age_structures |>
  aggregate_age_structure(
    group_vars = c("residence_rule", "coale_demeny_region"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = c("coale_demeny_region", "residence_rule"),
    show_mean = FALSE
    )
```

```{r}
#| label: age_structure_pyramid_plot_cdr_rr_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pyramid_plot,
    "v13_pyramid_residence_rule_region",
    height = 10
  )
}
```

```{r} 
#| label: age_structure_pyramid_plot_cdr_rr_display 
#| fig.cap: "Aggregate age structures of surviving populations across residence rules and Coale-Demeny regions."
plotName = paste0("plots/v13_pyramid_residence_rule_region.png")

knitr::include_graphics(plotName)
```

```{r} 
#| label: age_structure_indicators_plot_median_cdr_rr 
#| fig.cap: "Distribution of median age of surviving populations across residence rules and Coale-Demeny regions."
violin_plot <- age_structure_indicators |>
  plot_age_structure_indicators(
    x_var = "survival", 
    fill_var = "residence_rule", 
    facet_var = "coale_demeny_region"
  )
```

```{r}
#| label: age_structure_indicators_plot_median_cdr_rr_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    violin_plot,
    "v13_violin_residence_rule_region",
    width = 6,
    height = 3
  )
}
```

```{r} 
#| label: age_structure_indicators_plot_median_cdr_rr_display 
#| fig.cap: "Aggregate age structures of surviving populations across residence rules and Coale-Demeny regions."
plotName = paste0("plots/v13_violin_residence_rule_region.png")

knitr::include_graphics(plotName)
```

```{r} 
#| label: age_structure_indicators_summary_cdr_rr 
#| tab.cap: "Summary of age structure indicators for surviving populations across residence rules and Coale-Demeny regions."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "coale_demeny_region")
    )
)
```

```{r} 
#| label: age_structure_cohens_d_cdr_rr 
#| tab.cap: "Cohen's d for median age across residence rules."
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ residence_rule,
      data = age_structure_indicators
    ),
    rules = "cohen1988"
  )
)
```

Cohen's d for median age across survival categories, by residence rule:

```{r} 
#| label: age_structure_cohens_d_cdr_rr_by_rr_matri 
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ survival,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "matrilocal-matrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_cohens_d_cdr_rr_by_rr_patri 
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
      effectsize::cohens_d(
        median_age ~ survival,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

As in v1.1, residence rules continue to shape the age composition of surviving populations, although their effect is now more clearly expressed through cumulative differences in reproductive tempo. Patrilocal-patrilineal populations tend to show slightly older median ages and broader upper-age tails, especially among persistent cases, suggesting slower replacement dynamics. By contrast, matrilocal-matrilineal populations maintain somewhat younger and more compact age distributions, consistent with higher demographic turnover.

## Sensitivity analysis

### Prepare data for sensitivity analysis

Prepare a dataframe out of `endstates` that contains only predictors (explored parameters) and response variables (`survival`, `log_totalIndividuals`, `median_age`), as well as splitting it according to `residence_rule`.

```{r} 
#| label: sensitivity_data
response_variables <- c(
  "log_totalIndividuals",
  "survival",
  "median_age"
)

sensitivity_data <- prepare_sensitivity_data_by_residence_rule(
  endstates,
  response_variables,
  parameter_metadata
)
```

### Fitting Random-forest models

```{r} 
#| label: sensitivity_rf
rf_models <- load_or_build_rf_models(
  model_version = MODEL_VERSION,
  sensitivity_data = sensitivity_data,
  response_variables = response_variables,
  SEED = 123,
  verbose = FALSE,
  overwrite = REMAKE_RF
)
```

Random forest models are fitted separately for each residence rule and response variable, using the same set of predictors. The primary response variable, `survival`, classifies end states as extinction, persistence, or growth, using parameter combinations as categorical outcomes. The more nuanced response variables `log_totalIndividuals` and `median_age` are used only after filtering the data to include cases where a population is present at the end state (i.e. `survival = "Extinction"`).

The models are built using a consistent random seed to ensure reproducibility, and the option to overwrite existing models is controlled by the `REMAKE_RF` flag.

### Parameter importance and partial dependence

#### Survival (probability of extinction end state)

```{r} 
#| label: sensitivity_survival_importance
rf_survival_importance <- bind_rows(
  
  extract_importance(
    rf_models$survival$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$survival$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)

param_color <- parameter_metadata$color[match(rf_survival_importance$parameter, parameter_metadata$parameter)]
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_survival_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population survival, by residence rule."
plot_importance(
  rf_survival_importance, 
  param_color = param_color, 
  facet_var = "residence_rule")
```

```{r} 
#| label: sensitivity_survival_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population survival, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_survival_importance, 
  groups = "residence_rule", 
  param_color = param_color
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_survival_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v13_survival_importance_dumbbell"
  )
}
```

Select top parameters in relation to survival:

```{r} 
#| label: sensitivity_survival_top_param
top_parameters_survival <- rf_survival_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(MeanDecreaseGini)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `survival` as response variable and no population ("Extinction") as response class. Plot each PD result with `log_totalIndividuals` and per `residence_rule`.

```{r} 
#| label: sensitivity_survival_pd
pd_survival <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunk) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_survival,
  response_variable = "survival",
  response_class = "Extinction" # used only with `survival` as response variable
)
```

```{r} 
#| label: sensitivity_survival_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_survival,
  top_parameters_survival,
  response_variable = "log_totalIndividuals",
  pd_label = "P(Extinction)",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_survival_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v13_survival_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_survival_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing population survival, by residence rule."
plotName = paste0("plots/v13_survival_param_pd.png")

knitr::include_graphics(plotName)
```

The parameter hierarchy closely mirrors that of v1.1, with mortality intensity (cdmlt_level) and fertility intensity (c1_fert) remaining the dominant predictors of extinction. However, v1.3 introduces a stronger influence of timing-related fertility parameters, indicating that reproductive spacing becomes more consequential once gestation and postpartum infertility are explicitly modelled. This reinforces the idea that temporal structure matters not only for realism but also for parameter interaction.

#### Population (`log_totalIndividuals`)

```{r} 
#| label: sensitivity_size_importance
rf_size_importance <- bind_rows(
  
  extract_importance(
    rf_models$log_totalIndividuals$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$log_totalIndividuals$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_size_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population size, by residence rule."
plot_importance(
  rf_size_importance, 
  param_color = param_color, 
  facet_var = "residence_rule",
  metric = "%IncMSE")
```

```{r} 
#| label: sensitivity_size_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population size, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_size_importance, 
  groups = "residence_rule", 
  param_color = param_color,
  metric = "%IncMSE"
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_size_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v13_size_importance_dumbbell"
  )
}
```

Select top parameters predicting survival:

```{r} 
#| label: sensitivity_size_top_param
top_parameters_size <- rf_size_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(`%IncMSE`)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `log_totalIndividuals` as response variable (Extinction cases are not included).

```{r} 
#| label: sensitivity_size_pd
pd_size <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunck) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_size,
  response_variable = "log_totalIndividuals"
)
```

```{r} 
#| label: sensitivity_size_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_size,
  top_parameters_size,
  response_variable = "log_totalIndividuals",
  pd_label = "Expected change in population size",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_size_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v13_size_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_size_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing population size, by residence rule."
plotName = paste0("plots/v13_size_param_pd.png")

knitr::include_graphics(plotName)
```

Among surviving populations, final size remains primarily shaped by the balance between mortality and fertility, as in previous versions. Yet the daily formulation produces more gradual and differentiated responses, particularly around threshold regions where small changes in fertility or mortality now generate larger cumulative effects over time. This suggests that population growth in v1.3 is more path-dependent than in the yearly model.

#### Median age in end state population

```{r} 
#| label: sensitivity_median_age_importance
rf_median_age_importance <- bind_rows(
  
  extract_importance(
    rf_models$median_age$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$median_age$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_median_age_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting median age, by residence rule."
plot_importance(
  rf_median_age_importance, 
  param_color = param_color, 
  facet_var = "residence_rule",
  metric = "%IncMSE")
```

```{r} 
#| label: sensitivity_median_age_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting median age, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_median_age_importance, 
  groups = "residence_rule", 
  param_color = param_color,
  metric = "%IncMSE"
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_median_age_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v13_median_age_importance_dumbbell"
  )
}
```

Select top parameters predicting median age:

```{r} 
#| label: sensitivity_median_age_top_param
top_parameters_median_age <- rf_median_age_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(`%IncMSE`)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `median_age` as response variable (Extinction cases are not included).

```{r} 
#| label: sensitivity_median_age_pd
pd_median_age <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunck) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_median_age,
  response_variable = "median_age"
)
```

```{r} 
#| label: sensitivity_median_age_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_median_age,
  top_parameters_median_age,
  response_variable = "median_age",
  pd_label = "Expected change in median age",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_median_age_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v13_median_age_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_median_age_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing median age, by residence rule."
plotName = paste0("plots/v13_median_age_param_pd.png")

knitr::include_graphics(plotName)
```

Median age appears to be especially sensitive to fertility-related parameters, confirming that reproductive tempo is the main determinant of demographic rejuvenation. Mortality still contributes indirectly by altering survival into older cohorts, but fertility intensity and timing now exert clearer and more direct effects than in v1.1. This reflects the stronger coupling between birth spacing and age-structure formation in the daily model.

### Age structure of surviving populations for selected parameters

Note: parameters are ordered in importance towards median age. The binomial categories created earlier in preprocessing, which split the parameter range according to the distribution median, are used.

#### `c1_fert`: high *versus* low fertility regimes

```{r} 
#| label: age_structure_c1_fert_pyramid_plot
#| fig.cap: "Population pyramids across fertility intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_intensity"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_intensity",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_fert_median_plot
#| fig.cap: "Median age of surviving populations across fertility regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_intensity"
)
```

```{r} 
#| label: age_structure_c1_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across fertility regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_intensity")
    )
)
```

Cohen's d for median age across fertility regimes, by residence rule:

```{r} 
#| label: age_structure_c1_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
    median_age ~ fertility_intensity,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_intensity,

    data =
        age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

High fertility regimes remain associated with younger populations. Additionally, as in v1.1, this effect is stronger under the patrilocal-patrilineal rule, given that these cases tend towards greater demographic instability.

#### `sigma2_fert`: broad *versus* narrow post-peak fertility regimes

```{r} 
#| label: age_structure_sigma2_fert_pyramid_plot
#| fig.cap: "Population pyramids across fertility post-peak spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_spread_post_peak"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_spread_post_peak",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_fert_median_plot
#| fig.cap: "Median age of surviving populations across fertility post-peak spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_spread_post_peak"
)
```

```{r} 
#| label: age_structure_sigma2_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across fertility post-peak spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_spread_post_peak")
    )
)
```

Cohen's d for median age across fertility spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_post_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_post_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader post-peak fertility regimes tend to generate older populations by extending reproductive activity into later ages, increasing the weight of adult cohorts and slowing demographic turnover. This effect is consistent with v1.1, but appears stronger in v1.3 because delayed births accumulate more gradually. Narrower post-peak fertility schedules concentrate reproduction earlier and contribute to younger end-state populations.

#### `cdmlt_level`: mortality regimes

```{r}
#| label: age_structure_cdmlt_level_pyramid_plot
#| fig.cap: "Population pyramids across mortality regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("mortality_regime"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "mortality_regime",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_cdmlt_level_median_plot
#| fig.cap: "Median age of surviving populations across mortality regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "mortality_regime"
)
```

```{r} 
#| label: age_structure_cdmlt_level_indicators_summary
#| tab.cap: "Summary of age structure indicators across mortality regimes and residence rules."
knitr::kable(
  indicators_summary(age_structure_indicators, group_vars = c("residence_rule", "mortality_regime"))
)
```

Cohen's d for median age across mortality regimes, by residence rule:

```{r} 
#| label: age_structure_cdmlt_level_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ mortality_regime,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_cdmlt_level_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ mortality_regime,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

Mortality regimes continue to exert a strong influence on age structure, with lower mortality producing older and more top-heavy populations. This follows directly from increased survivorship into later life stages. As in v1.1, mortality affects not only survival probabilities but also the internal demographic composition of surviving populations, making it one of the most structurally important parameters.

#### `c1_men`: high *versus* low male marriage intensity regimes

```{r} 
#| label: age_structure_c1_men_pyramid_plot
#| fig.cap: "Population pyramids across male marriage intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_intensity_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_intensity_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_men_median_plot
#| fig.cap: "Median age of surviving populations across male marriage intensity regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_intensity_men"
)
```

```{r} 
#| label: age_structure_c1_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male marriage intensity regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_intensity_men")
    )
)
```

Cohen's d for median age across male marriage intensity regimes, by residence rule:

```{r} 
#| label: age_structure_c1_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_intensity_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_intensity_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

Higher male marriage intensity has a similar but somewhat weaker rejuvenating effect, primarily by facilitating the formation of reproductive pairs and reducing delays in household formation. Compared with female intensity, its impact is more indirect, but still contributes to maintaining younger demographic profiles.

#### `sigma1_men`: broad *versus* narrow male pre-peak marriage spread regimes

```{r} 
#| label: age_structure_sigma1_men_pyramid_plot
#| fig.cap: "Population pyramids across male pre-peak marriage spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_spread_pre_peak_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_spread_pre_peak_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_men_median_plot
#| fig.cap: "Median age of surviving populations across male pre-peak marriage spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_spread_pre_peak_men"
)
```

```{r} 
#| label: age_structure_sigma1_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male pre-peak marriage spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_spread_pre_peak_men")
    )
)
```

Cohen's d for median age across male pre-peak marriage spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_pre_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_pre_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader pre-peak marriage spread for males continues to be associated with younger populations, this effect remaining relatively small. However, in contrast with v1.1, this effect is higher under the matrilocal-matrilineal rule. 

#### `c1_women`: high *versus* low female marriage intensity regimes

```{r} 
#| label: age_structure_c1_women_pyramid_plot
#| fig.cap: "Population pyramids across female marriage intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_intensity_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_intensity_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_women_median_plot
#| fig.cap: "Median age of surviving populations across female marriage intensity regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_intensity_women"
)
```

```{r} 
#| label: age_structure_c1_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across female marriage intensity regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_intensity_women")
    )
)
```

Cohen's d for median age across female marriage intensity regimes, by residence rule:

```{r} 
#| label: age_structure_c1_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_intensity_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_intensity_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

Higher female marriage intensity is associated with younger populations, reflecting faster coupling rates and earlier reproductive activation across the population. This effect is stronger under matrilocal-matrilineal residence, where household continuity depends more directly on the successful incorporation of women into reproductive units.

#### `mu_men`: late *versus* early male marriage regimes

```{r} 
#| label: age_structure_mu_men_pyramid_plot
#| fig.cap: "Population pyramids across male marriage timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_timing_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_timing_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_men_median_plot
#| fig.cap: "Median age of surviving populations across male marriage timing regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_timing_men"
)
```

```{r} 
#| label: age_structure_mu_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male marriage timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_timing_men")
    )
)
```

Cohen's d for median age across male marriage timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_timing_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_timing_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Later male marriage timing is associated with older age distributions, as delayed union formation reduces the onset of household reproduction and shifts birth cohorts forward in time. This effect is visible under both residence rules, but is stronger in patrilocal-patrilineal populations, where delayed male incorporation appears to slow household replacement more markedly.

#### `sigma1_fert`: broad *versus* narrow pre-peak fertility spread regimes

```{r} 
#| label: age_structure_sigma1_fert_pyramid_plot
#| fig.cap: "Population pyramids across pre-peak fertility spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_spread_pre_peak"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_spread_pre_peak",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_fert_median_plot
#| fig.cap: "Median age of surviving populations across pre-peak fertility spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_spread_pre_peak"
)
```

```{r} 
#| label: age_structure_sigma1_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across pre-peak fertility spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_spread_pre_peak")
    )
)
```

Cohen's d for median age across pre-peak fertility spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_pre_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_pre_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader pre-peak fertility schedules tend to produce younger populations by spreading reproductive entry across a wider age range while maintaining earlier births. This increases overlap between generations and supports more continuous replacement. Narrower schedules, by contrast, concentrate births into shorter reproductive windows, contributing to demographic ageing.

#### `sigma1_women`: broad *versus* narrow pre-peak fertility spread regimes

```{r} 
#| label: age_structure_sigma1_women_pyramid_plot
#| fig.cap: "Population pyramids across pre-peak fertility spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_spread_pre_peak_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_spread_pre_peak_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_women_median_plot
#| fig.cap: "Median age of surviving populations across pre-peak fertility spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_spread_pre_peak_women"
)
```

```{r} 
#| label: age_structure_sigma1_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across pre-peak fertility spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_spread_pre_peak_women")
    )
)
```

Cohen's d for median age across pre-peak fertility spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_pre_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_pre_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader female pre-peak marriage schedules are associated with younger and more balanced age distributions, likely because they allow earlier and more varied entry into reproduction. As with the equivalent male parameter, this broadening appears to stabilise demographic renewal rather than concentrating it into narrow cohort windows.

#### `amenorrhea_period_in_days`: long *versus* short amenorrhea period

```{r} 
#| label: age_structure_amenorrhea_pyramid_plot
#| fig.cap: "Population pyramids across amenorrhea period lengths, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("amenorrhea_length"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "amenorrhea_length",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_amenorrhea_median_plot
#| fig.cap: "Median age of surviving populations across amenorrhea period lengths, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "amenorrhea_length"
)
```

```{r} 
#| label: age_structure_amenorrhea_indicators_summary
#| tab.cap: "Summary of age structure indicators across amenorrhea period lengths and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "amenorrhea_length")
    )
)
```

Cohen's d for median age across amenorrhea period lengths, by residence rule:

```{r} 
#| label: age_structure_amenorrhea_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ amenorrhea_length,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_amenorrhea_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ amenorrhea_length,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Amenorrhea duration has a relatively weak but consistent ageing effect. Longer postpartum infertility slows reproductive tempo by increasing birth intervals, reducing juvenile recruitment over time. Unlike fertility intensity or timing, its influence is incremental rather than structural, acting mainly as a modulator of demographic pacing.

#### `sigma2_men`: broad *versus* narrow post-peak male marriage spread regimes

```{r} 
#| label: age_structure_sigma2_men_pyramid_plot
#| fig.cap: "Population pyramids across post-peak male marriage spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_spread_post_peak_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_spread_post_peak_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_men_median_plot
#| fig.cap: "Median age of surviving populations across post-peak male marriage spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_spread_post_peak_men"
)
```

```{r} 
#| label: age_structure_sigma2_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across post-peak male marriage spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_spread_post_peak_men")
    )
)
```

Cohen's d for median age across post-peak male marriage spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_post_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_post_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader post-peak male marriage schedules tend to increase median age by extending male participation in union formation into later life stages. This prolongs the reproductive contribution of older cohorts and slightly slows generational replacement. The effect is modest but consistent across residence systems.

#### `mu_fert`: early *versus* late fertility timing

```{r} 
#| label: age_structure_mu_fert_pyramid_plot
#| fig.cap: "Population pyramids across (female) fertility timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_timing"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_timing",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_fert_median_plot
#| fig.cap: "Median age of surviving populations across (female) fertility timing regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_timing"
)
```

```{r} 
#| label: age_structure_mu_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across (female) fertility timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_timing")
    )
)
```

Cohen's d for median age across female marriage timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_timing,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_timing,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Earlier fertility timing is strongly associated with younger age structures, as it accelerates generational turnover and increases the proportion of juveniles and young adults. This pattern is fully consistent with demographic theory and with v1.1, but appears more pronounced in v1.3 due to the explicit modelling of gestation and postpartum intervals.

#### `mu_women`: late *versus* early female marriage regimes

```{r} 
#| label: age_structure_mu_women_pyramid_plot
#| fig.cap: "Population pyramids across female marriage timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_timing_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_timing_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_women_median_plot
#| fig.cap: "Median age of surviving populations across female marriage timing regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_timing_women"
)
```

```{r} 
#| label: age_structure_mu_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across female marriage timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_timing_women")
    )
)
```

Cohen's d for median age across female marriage timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_timing_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_timing_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

As expected, later female marriage timing produces older end-state populations by postponing the beginning of reproductive careers. This effect remains one of the clearest pathways linking nuptiality to demographic ageing. Compared with v1.1, the daily structure makes this delay more cumulative, reinforcing its influence on age composition.

#### `sigma2_women`: broad *versus* narrow post-peak female marriage spread regimes

```{r} 
#| label: age_structure_sigma2_women_pyramid_plot
#| fig.cap: "Population pyramids across post-peak female marriage spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_spread_post_peak_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_spread_post_peak_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_women_median_plot
#| fig.cap: "Median age of surviving populations across post-peak female marriage spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_spread_post_peak_women"
)
```

```{r} 
#| label: age_structure_sigma2_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across post-peak female marriage spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_spread_post_peak_women")
    )
)
```

Cohen's d for median age across post-peak female marriage spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_post_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_post_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

As with fertility post-peak spread, broader female post-peak marriage schedules contribute to older age structures by delaying the closure of reproductive careers. This increases the representation of older adults while reducing the concentration of early-life births. The effect is clearer in persistence than in growth cases.

#### Interpretation

Taken together, these results confirm that the age structure of surviving populations in v1.3 is shaped primarily by fertility tempo and mortality intensity, with nuptiality parameters exerting secondary but still meaningful effects. Compared with v1.1, the daily iteration model sharpens these relationships by making demographic delays explicit. This produces older, smoother, and more differentiated age structures, particularly among populations that persist without sustained growth.

## Mortality intensity threshold experiment

Simulations using the same parameter configuration, except `residence-rule`, and varying `cdmlt_level` and `SEED`.

```{r}
#| label: trajectories_cdmlt_level_load
rm(sampled_trajectories, preprocessing_trajectories_symmetric, sampled_trajectories_c1_fert)
sampled_trajectories_cdmlt_level <- readRDS("data/v13_trajectories_cdmlt_level.RData")
```

```{r} 
#| label: trajectories_params_cdmlt_level_display
#| fig.cap: "Parameter configuration for the population trajectories sampled across different Coale-Demeny model levels."
knitr::kable(sampled_trajectories_cdmlt_level |> 
               mutate(
                 max_iterations_in_days = max_iterations * year_length_in_days,
                 max_iterations_in_days_cut = max(step)
               ) |>
               select(all_of(c("max_iterations",
                               "max_iterations_in_days",
                               "max_iterations_in_days_cut", # show this in case loading was partial
                               "max_population", 
                               names(parameter_ranges_v13), 
                               "coale_demeny_region", 
                               "maturity_age_in_years", "gestation_period_in_days", "year_length_in_days" )),
                      -c("cdmlt_level")) |>
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

```{r} 
#| label: trajectories_cdmlt_level
#| fig.cap: "Population trajectories across different Coale-Demeny model levels, faceted by residence rule."
log_max_population <- ceiling(max(sampled_trajectories_cdmlt_level$log_totalIndividuals))

pop_trajectories_plot <- sampled_trajectories_cdmlt_level |>
  mutate(
    cdmlt_level = factor(cdmlt_level)
  ) |>
  plot_trajectories(
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = "cdmlt_level",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "max_population",
    thresholds_linetype = 2
  )
pop_trajectories_plot
```

```{r}
#| label: trajectories_cdmlt_level_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_plot,
    "v13_pop_trajectories_cdmlt_level"
  )
}
```

Compared with v1.1 parallel results, the threshold value for this key parameter is placed considerably higher (i.e. c. 8-11 compared to c. 5-8 in v1.1). Still, the difference between the matrilineal-matrilocal and patrilineal-patrilocal rules continue within the same scale (i.e. 8 vs 11 or less in v1.3, compared to 5 and 8 or less in v1.1).

## Fertility intensity threshold experiment

Simulations using the same parameter configuration, except `residence-rule`, and varying `c1_fert` and `SEED`. `cdmlt_level` is kept just above the safe value under both residence rules.

```{r}
#| label: trajectories_c1_fert_load
rm(sampled_trajectories, preprocessing_trajectories_symmetric, sampled_trajectories_c1_fert)
sampled_trajectories_c1_fert <- readRDS("data/v13_trajectories_c1_fert.RData")
```

```{r} 
#| label: trajectories_params_c1_fert_display
#| fig.cap: "Parameter configuration for the population trajectories sampled across different fertility intensity levels."
knitr::kable(sampled_trajectories_c1_fert |> 
               mutate(
                 max_iterations_in_days = max_iterations * year_length_in_days,
                 max_iterations_in_days_cut = max(step)
               ) |>
               select(all_of(c("max_iterations",
                               "max_iterations_in_days",
                               "max_iterations_in_days_cut", # show this in case loading was partial
                               "max_population", 
                               names(parameter_ranges_v13), 
                               "coale_demeny_region", 
                               "maturity_age_in_years", "gestation_period_in_days", "year_length_in_days" )),
                      -c("c1_fert")) |>
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

```{r} 
#| label: trajectories_c1_fert
#| fig.cap: "Population trajectories across different fertility intensity levels, faceted by residence rule."
log_max_population <- ceiling(max(sampled_trajectories_c1_fert$log_totalIndividuals))

pop_trajectories_plot <- sampled_trajectories_c1_fert |>
  mutate(
    c1_fert = factor(c1_fert)
  ) |>
  plot_trajectories(
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = "c1_fert",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "max_population",
    thresholds_linetype = 2
  )
pop_trajectories_plot
```

```{r}
#| label: trajectories_c1_fert_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_plot,
    "v13_pop_trajectories_c1_fert"
  )
}
```

As in the case of `cdmlt_level`, `c1_fert` presents a local threshold value, less definitive but still very influential, that differs under the two residence rules. However, v1.3 presents a higher extinction probability overall, so this parameter setting already hardly allows for persistence, even under the most intense fertility. Still, the asymmetry between residence rules are hinted, by having only matrilocal-matrilinial cases, with 0.9 intensity, avoiding extinction. This confirms the importance of mortality intensity for survival, since a higher `c1_fert` can barely compasate the effect of a low `cdmlt_level` (higher mortality intensity). 

## Discussion notes

The key thing about v1.3 is that it is not just “v1.1 with smaller timesteps”; it inserts gestation and postpartum amenorrhea as explicit reproductive delays, which changes tempo and cohort spacing while preserving the broader demographic logic.

Overall, v1.3 preserves the same general demographic attractors observed in v1.1, indicating that the transition to daily iteration does not alter the model’s fundamental behaviour. Instead, it refines the internal timing of demographic events, producing smoother trajectories, more realistic reproductive spacing, and sharper differentiation in age structure between persistent and growing populations.

### Same hierarchy of demographic drivers, stronger temporal realism

Like v1.1, mortality (cdmlt_level) and fertility intensity (c1_fert) remain the dominant determinants of survival and final size. This is expected because daily scaling preserves annual cumulative probabilities. But the daily structure makes these effects unfold more gradually and with more path dependence. Therefore, v1.3 validates the structural robustness of v1.1 while increasing temporal fidelity.

### Growth becomes harder, persistence more differentiated

In v1.1, “growth” could emerge rapidly because births were instantaneous yearly events. In v1.3, births are delayed by pregnancy and followed by infertile postpartum periods. This introduces friction into expansion. Growth cases still exist, but become less explosive, persistence occupies a broader middle ground, and extinction still dominates most parameter space. This makes the model demographically more realistic.

### Age structure becomes smoother and older

The Cohen’s d values in the residence-rule comparison (0.56 matri, 0.85 patri across survival categories) suggest a stronger separation between persistence and growth than in v1.1. This likely means growth populations remain younger, persistence populations age more visibly, and temporal spacing between births reduces abrupt juvenile bulges. In other words: the demographic “pulse” becomes stretched.

### Residence-rule differences persist

Patrilocal-patrilineal populations continue to show older median ages and stronger divergence between survival outcomes, consistent with v1.1. This suggests kinship structure still mediates demographic resilience, daily iteration does not erase those structural asymmetries, but may amplify them through cumulative household turnover effects.

### Additional comments

The explicit incorporation of gestation and postpartum infertility reveals that reproductive timing parameters play a larger role than previously visible in annual models. This suggests that coarse temporal resolution may underestimate the cumulative effects of demographic delays. At the same time, the persistence of the same dominant parameter hierarchy confirms that the model remains structurally stable across scales, strengthening confidence in its underlying demographic logic.

## Chapter conclusion

Version 1.3 demonstrates that increasing temporal resolution can substantially improve demographic realism without disrupting the broader parameter sensitivities identified in the yearly model. This makes it an important intermediary step toward integrating ecological feedbacks at finer temporal scales in the subsequent v1.4.x models.
