---
execute:
  echo: true
  warning: false
  message: false
params:
  show_diagnostics: false
---

# Demand- and supply-weighted density dependent model (v.1.2.2)

## Setup

Set constant variables for this chapter.

```{r}
#| label: setup_constants
MODEL_VERSION <- "v1.2.2"
EXPORT_FIGURES <- TRUE
REMAKE_RF <- FALSE
REMAKE_DP <- FALSE
```

Load libraries, sources and set ggplot theme for the notebook.

```{r} 
#| label: setup_libraries
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(zoo)
library(stringr)
library(effectsize)
library(randomForest)
library(pdp)
library(forcats)
library(patchwork)
library(scales)
```

```{r} 
#| label: setup_load_sources
source("library/source_folder.R")
source_folder("library")
source_folder("library/structure")
source_folder("library/sensitivity")
source_folder("library/density")
```

```{r} 
#| label: setup_ggplot_theme
theme_set(
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold"),
      plot.title = element_text(face = "bold")
    )
)
```

## Experiment design

As version v1.2.2 introduce a new parameter, `labour_demand_per_capita`, an additional batch of Sobol sampled values is generated. As in v1.2.1, this model does require constants used to determine the correspondence between age and the adult-equivalent unit (i.e., 15, 45, and the exponential 10, which determines a steeper decline compared with the function for demand; see `get-labour-at-age`). These were generated previously with `library/generate_sobol.R`.

```{r} 
#| label: sobol_load
# Load object as Rdata file
load("data/v122_sobol_parameter_values.RData")
```

```{r} 
#| label: sobol_ranges_display
parameter_ranges_v122 <- as.data.frame(parameter_ranges[[MODEL_VERSION]])

knitr::kable(
  parameter_ranges_v122 |>
    as.matrix() |>
    t() |>
    as.data.frame() |>
    setNames(c("min", "max"))
)
```

```{r} 
#| label: sobol_pairs 
#| include: !expr params$show_diagnostics 
#| fig.cap: "Parameter ranges explored in the Sobol sampling for v1.2.2."
# Visualising parameter space with `pairs()`:
plotName = paste0("diagnostics/", 
  MODEL_VERSION, 
  "_sobol_param_pairsPlot.png"
  )

knitr::include_graphics(plotName)
```

## Data loading and preprocessing

Load simulation end states generated with the Sobol-sampled parameter combinations and sampled trajectories datasets. Preprocess datasets to adapt from NetLogo's specific format and categorise end states and trajectories.

```{r}
#| label: preprocessing_vars
overshoot_magnitude_threshold <- 0 # any value over carrying capacity is considered overshooting
burn_in_fraction = 0.5 # the second half of the simulation is considered representative of an equilibrium phase
overshoot_time_threshold <- 0.1 # 10% of time after "burn in" period
oscillatory_threshold <- 0.05 # coefficient of variance of population size after "burn in" period 
stress_threshold <- 0.5 # pressures over this value are considered as population under stress
```

```{r} 
#| label: preprocessing_endstates
endstates <- load_endstates(MODEL_VERSION) |>
  preprocess_simdata(
    model_version = MODEL_VERSION
  )
```

```{r} 
#| label: preprocessing_trajectories
sampled_trajectories <- load_trajectories(MODEL_VERSION) |>
  preprocess_simdata(
    model_version = MODEL_VERSION,
    full_trajectory = TRUE
  ) |>
  compute_density_metrics(
    burn_in_fraction = burn_in_fraction
  ) |>
  classify_regulation_regimes(
    overshoot_magnitude_threshold = overshoot_magnitude_threshold,
    overshoot_time_threshold = overshoot_time_threshold,
    oscillatory_threshold = oscillatory_threshold,
    stress_threshold = stress_threshold
  )
```

```{r}
#| label: preprocessing_trajectories_labour_demand_per_capita
sampled_trajectories_labour_demand_per_capita <- load_trajectories(
  MODEL_VERSION, 
  suffix = "labour-demand-per-capita"
) |>
  preprocess_simdata(
    model_version = MODEL_VERSION,
    full_trajectory = TRUE,
    trajectory_id_vars = c(
      # unique combinations of these variables id single runs
      "run_number",
      "residence_rule"
    ),
  ) |>
  compute_density_metrics(
    burn_in_fraction = burn_in_fraction
  ) |>
  classify_regulation_regimes(
    overshoot_magnitude_threshold = overshoot_magnitude_threshold,
    overshoot_time_threshold = overshoot_time_threshold,
    oscillatory_threshold = oscillatory_threshold,
    stress_threshold = stress_threshold
  )
```

A particularity in processing the data with age-weighted variables is that we need to use (or compute) the total carrying capacity demand (`carrying_capacity_demand`) and use it to calculate `pressure` instead of using `totalIndividuals` directly (i.e., `pressure = carrying_capacity_demand / carrying_capacity`). Although this variable can be exported through NetLogo's Behavior Space, it was not specified for all massive Sobol sampled experiments, so it is necessary to calculate it using `womenAgeStructure` and `menAgeStructure` (see "library/density/compute_carrying_capacity_demand.R").

### Surviving population structure

```{r} 
#| label: age_structure_prepare
group_vars = c(
  "survival",
  "residence_rule",
  "coale_demeny_region",
  "mortality_regime",
  "fertility_intensity",
  "fertility_timing",
  "fertility_spread_pre_peak",
  "fertility_spread_post_peak",
  "marriage_timing_women",
  "marriage_timing_men",
  "marriage_intensity_women",
  "marriage_intensity_men",
  "marriage_spread_pre_peak_women",
  "marriage_spread_pre_peak_men",
  "marriage_spread_post_peak_women",
  "marriage_spread_post_peak_men"
)

age_structures <- endstates |>

  filter(
    survival != "Extinction"
  ) |>
  
  characterise_mort_fert_nupt_regimes() |>

  prepare_age_structure(
    age_breaks = seq(0, 100, 5),
    normalize = TRUE,
    group_vars = group_vars
  )

age_structure_indicators <- age_structures |>
  compute_age_structure_indicators(group_vars = group_vars) |>
  # make a binomial classification for Cohen's d test
  mutate(
    pressure_regime = ifelse(
      survival == "Persistence\n(low pressure)",
      "Low pressure\nregimes",
      "High pressure\nregimes"
    )
  )

endstates <- endstates |>
  left_join(
    age_structure_indicators |>
      select(
        run_unique_id,
        median_age,
        youth_share,
        elderly_share,
        dependency_ratio
      ),
    by = "run_unique_id"
  )
```

## Overview of end states and dynamics

### Population persistence

```{r} 
#| label: endstates_distrib_size 
#| fig.cap: "Distribution of final population sizes across residence rules and Coale-Demeny regions."
log_max_population <- max(endstates$log_totalIndividuals)

endstates_distrib_size_plot <- endstates |>
  plot_endstate_variable_distribution(
    main_var = "log_totalIndividuals",
    fill_var = "survival",
    facet_var = c("coale_demeny_region", "residence_rule"),
    bins = 40,
    x_axis_label = "Final population size",
    y_axis_label = "Simulation count",
    threshold_values = log_max_population,
    thresholds_linetype = 2,
    x_axis_log10 = TRUE,
    x_max = log_max_population,
    y_axis_log10 = TRUE
  )
endstates_distrib_size_plot
```

```{r}
#| label: endstates_distrib_size_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    endstates_distrib_size_plot,
    "v122_endstates_size_distrib"
  )
}
```

```{r} 
#| label: endstates_distrib_pressure 
#| fig.cap: "Distribution of final population pressure in surviving populations across residence rules and Coale-Demeny regions."
endstates_pressure_plot <- endstates |>
  filter(survival != "Extinction") |>
  plot_endstate_variable_distribution(
    main_var = "pressure",
    fill_var = "survival",
    facet_var = c("coale_demeny_region", "residence_rule"),
    bins = 40,
    x_axis_label = "Final population pressure",
    y_axis_label = "Simulation count",
    threshold_values = c(0.5, 1),
    thresholds_linetype = c(2, 1),
    x_axis_log10 = FALSE,
    y_axis_log10 = FALSE
  )
endstates_pressure_plot
```

```{r}
#| label: endstates_distrib_pressure_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    endstates_pressure_plot,
    "v122_endstates_pressure"
  )
}
```

```{r} 
#| label: outcome_prop 
#| fig.cap: "Proportion of simulations across survival category, residence rules and Coale-Demeny regions."
outcome_prop_plot <- endstates |>
  plot_category_proportion_stacked(
    x_var = "residence_rule",
    fill_var = "survival",
    facet_var = "coale_demeny_region",
    facet_ncol = 2,
    facet_scales = "fixed",
    y_axis_label = "Share of simulations"
  )
outcome_prop_plot
```

```{r}
#| label: outcome_prop_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    outcome_prop_plot,
    "v122_outcome_prop"
  )
}
```

Although the overall structure of end-state outcomes remains comparable to that of v1.2.1, the introduction of labour-constrained carrying capacity substantially shifts the balance toward extinction. The proportion of non-extinction cases is considerably lower for all combination of `residence-rule` and `coale-demeny-region` (mean surviving proportion is ```r endstates |> filter(survival != "Extinction") |> nrow()```). This indicates that the additional supply-side constraint imposes a stronger viability filter on the system. As in previous versions, residence rule and mortality regime continue to shape survival asymmetrically, but their effects are now expressed under a narrower range of viable demographic configurations.

Still, patterns present in v1.2.1 continue, such as the small asymmetry between matrilineal-matrilocal rule (n = ```r endstates |> filter(residence_rule == "matrilocal-matrilineal", survival != "Extinction") |> nrow()```) and patrilineal-patrilocal rule (n = ```r endstates |> filter(residence_rule == "patrilocal-patrilineal", survival != "Extinction") |> nrow()```) and the slightly higher overshooting under patrilineal-patrilocal (n = ```r endstates |> filter(residence_rule == "patrilocal-patrilineal", survival == "Overshooting") |> nrow()``` compared to matrilineal-matrilocal's ```r endstates |> filter(residence_rule == "matrilocal-matrilineal", survival == "Overshooting") |> nrow()```).

### Temporal trajectories

The parameter configuration used to generate the trajectories is the same across residence rules, except for the `residence_rule` parameter itself. The configuration used for this was:

```{r} 
#| label: trajectories_params_display 
#| tab.cap: "Parameter configuration for the population trajectories sampled."
knitr::kable(sampled_trajectories |> 
               select(all_of(c("max_iterations", "max_population", 
                               names(parameter_ranges_v122), 
                               "coale_demeny_region", "maturity_age", 
                               "density_effect_steepness"
                               ))) |>
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

Dynamics of population trajectories across time steps, faceted by residence rule. The plot shows individual trajectories with low opacity and the median trajectory with higher opacity. The y-axis is on a logarithmic scale to better visualise differences in population sizes over time, especially when trajectories vary widely.

```{r} 
#| label: trajectories 
#| fig.cap: "A sample of population trajectories across time steps, faceted by residence rule."
pop_trajectories_plot <- sampled_trajectories |>
  plot_trajectories(
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = "regulation_regime",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "carrying_capacity",
    thresholds_linetype = 2
  )
pop_trajectories_plot
```

```{r}
#| label: trajectories_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_plot,
    "v122_pop_trajectories"
  )
}
```

Matrilocal-matrilineal systems clearly exhibit bifurcation: either stable high-pressure persistence near carrying capacity, or delayed collapse after prolonged demographic stress. The fact that collapse is here protracted is significant. Patrilocal-patrilineal systems show a much steeper and faster collapse gradient, suggesting lower resilience once the population enters under-replacement dynamics.

### Household composition

```{r}
#| label: household_structure_plot
#| fig.cap: ""
household_structure_plot <- endstates |> 
  plot_household_structure_faceted_fixed(
    color_var = "max_couple_count_distribution_max",
    point_alpha = 0.08,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "fixed",
    xy_log10 = TRUE,
    x_axis_label = "number of households",
    y_axis_label = "number of individuals",
    color_label = "Max. couple count allowed in a household\n(parametric maximum)",
    legend_position = "bottom"
  )
household_structure_plot
```

```{r}
#| label: household_structure_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    household_structure_plot,
    "v122_household_structure"
  )
}
```

```{r}
#| label: trajectories_household_structure_plot_regulation
trajectories_household_structure_plot_regulation <- 
  sampled_trajectories |>
  group_by(residence_rule) |>
  mutate(
    household_structure_ratio = (totalIndividuals / totalHouseholds)
  ) |>
  ungroup() |>
  plot_trajectories(
    y_var = "household_structure_ratio",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = FALSE,
    y_axis_label = "Mean individual count per household",
    group_var = "regulation_regime",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = NULL
  )
```

```{r}
#| label: trajectories_household_structure_plot_regulation_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    trajectories_household_structure_plot_regulation,
    "v122_trajectories_household_structure_regulation",
    width = 8,
    height = 4
  )
}
```

```{r} 
#| label: trajectories_household_structure_plot_regulation_display 
#| fig.cap: "A sample of household population trajectories across time steps, grouped by regulation_regime and faceted by residence rule."
plotName = paste0("plots/v122_trajectories_household_structure_regulation.png")

knitr::include_graphics(plotName)
```

```{r}
#| label: trajectories_household_structure_plot_peaks
trajectories_household_structure_plot_peaks <- 
  sampled_trajectories |>
  group_by(residence_rule) |>
  mutate(
    household_structure_ratio = if_else(totalHouseholds == 0, 0, (totalIndividuals / totalHouseholds)),
    # Find the absolute peak value within the entire residence rule
    global_max = max(household_structure_ratio, na.rm = TRUE)
  ) |>
  group_by(residence_rule, run_number) |>
  mutate(
    # Check if this specific run contains that global peak value
    selected_trajectory = any(household_structure_ratio == global_max)
  ) |>
  ungroup() |>
  plot_trajectories(
    y_var = "household_structure_ratio",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = FALSE,
    y_axis_label = "Mean individual count per household",
    group_var = NULL,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = NULL
  ) +
  # Add the highlighted layer on top
  geom_line(
    data = \(df) filter(df, selected_trajectory == TRUE),
    aes(
      x = step, 
      y = household_structure_ratio,
      group = run_number
    ),
    color = "red",
    linewidth = 1,
    alpha = 1
  )
```

```{r}
#| label: trajectories_household_structure_plot_peaks_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    trajectories_household_structure_plot_peaks,
    "v122_household_structure_trajectories_peaks",
    width = 8,
    height = 4
  )
}
```

```{r} 
#| label: trajectories_household_structure_plot_peaks_display
#| fig.cap: "A sample of household population trajectories across time steps, faceted by residence rule, highlighting in red the trajectories that achieve the maximum value in each subset."
plotName = paste0("plots/v122_household_structure_trajectories_peaks.png")

knitr::include_graphics(plotName)
```

Household-level outcomes remain broadly consistent with those observed in v1.2.1, suggesting that the labour constraint primarily modifies population-level viability rather than household-level social organisation. The mechanisms governing household formation, division, and persistence therefore remain structurally stable, while ecological and demographic constraints act mainly through aggregate labour availability and mortality feedback.

Worth noticing is that the trajectory with the maximum household size peak under the matrilineal-matrilocal rule collapses while those with more modest peaks remain persistent. This reinforces the still preliminary interpretation that a population peaking on this indicator might be associated with extinction.

### Effective carrying capacity and labour

```{r}
#| label: endstates_distrib_log_carrying_capacity_effective 
#| fig.cap: "Distribution of effective carrying capacity ratio over parametric carrying capacity in surviving populations across residence rules and Coale-Demeny regions."
endstates_distrib_log_carrying_capacity_effective_plot <- endstates |>
  filter(survival != "Extinction") |>
  plot_endstate_variable_distribution(
    main_var = "log_carrying_capacity_effective",
    fill_var = "survival",
    facet_var = c("coale_demeny_region", "residence_rule"),
    x_axis_label = "effective carrying capacity"
  )
endstates_distrib_log_carrying_capacity_effective_plot
```

```{r}
#| label: endstates_distrib_log_carrying_capacity_effective_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    endstates_distrib_log_carrying_capacity_effective_plot,
    "v122_endstates_distrib_log_carrying_capacity_effective"
  )
}
```

The plot is essentially reproducing the nominal carrying-capacity distribution. This means that surviving populations are almost always operating at full effective capacity.

```{r}
#| label: endstates_distrib_labour_ratio 
#| fig.cap: "Distribution of labour ratio (available / required) in surviving populations across residence rules and Coale-Demeny regions."
endstates_distrib_labour_ratio_plot <- endstates |>
  filter(survival != "Extinction") |>
  plot_endstate_variable_distribution(
    main_var = "labour_ratio",
    fill_var = "survival",
    facet_var = c("coale_demeny_region", "residence_rule"),
    x_axis_label = "Labour ratio",
    x_axis_log10 = FALSE
  )
endstates_distrib_labour_ratio_plot
```

```{r}
#| label: endstates_distrib_labour_ratio_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    endstates_distrib_labour_ratio_plot,
    "v122_endstates_distrib_laboured_share"
  )
}
```

```{r} 
#| label: labour_ratio_survival_plot 
#| fig.cap: "Labour ratio distribution of simulations across survival category, residence rules and Coale-Demeny regions."
labour_ratio_survival_plot <- endstates |>
  plot_labour_ratio_distrib(
    x_var = "survival",
    fill_var = "residence_rule",
    facet_var = NULL,
    pos_doge_width = 1.0
  )
labour_ratio_survival_plot
```

```{r}
#| label: labour_ratio_survival_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    labour_ratio_survival_plot,
    "v122_labour_ratio_survival"
  )
}
```

```{r}
#| label: labour_balance_plot
#| fig.cap: ""
labour_balance_plot <- endstates |>
  plot_labour_balance(
    color_var = "labour_demand_per_capita",
    point_alpha = 0.2,
    facet_var = "survival",
    facet_ncol = 1,
    facet_scales = "fixed",
    x_log10 = FALSE,
    x_axis_label = "Labour required",
    y_axis_label = "Labour ratio (available / required)",
    color_label = "Labour demand\nper capita",
    legend_position = "right"
  )
labour_balance_plot
```

```{r}
#| label: labour_balance_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    labour_balance_plot,
    "v122_labour_productivity"
  )
}
```

```{r}
knitr::kable(
  sampled_trajectories |>
    group_by(survival) |>
    summarise(
      mean_laboured_share = mean(laboured_share),
      median_laboured_share = median(laboured_share),
      min_laboured_share = min(laboured_share)
    )
)
```

```{r} 
#| label: trajectories_labour_ratio 
#| fig.cap: "The labour ratio of a sample of population trajectories across time steps, faceted by residence rule."
pop_trajectories_labour_plot <- sampled_trajectories |>
  plot_trajectories(
    y_var = "labour_ratio",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = FALSE,
    y_axis_label = "Labour ratio",
    group_var = "regulation_regime",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = ,
    thresholds_linetype = 2
  )
pop_trajectories_labour_plot
```

**_Labour limitation as a viability filter_**

The initial working questions focused on whether labour-dependent carrying capacity separates survival outcomes in a systematic way:

- Do low-pressure populations maintain higher effective carrying capacities? The results do not support this interpretation.  
- Are high-pressure populations labour constrained? No; surviving high-pressure populations generally maintain sufficient labour.  
- Do overshooting populations occupy a distinct parameter range? Not clearly, although the few observed cases appear unusual.  

Overall, labour does not appear to act as a limiting factor among surviving populations. Almost every surviving run ends with `laboured_share` close to 1, meaning that `labour_available` is greater than or equal to `labour_required`. As a result, `carrying_capacity_effective` is approximately equal to the nominal `carrying_capacity` for nearly all surviving populations.

This pattern suggests a strong filtering process. Populations that cannot maintain adequate labour availability tend to collapse, while those that survive have effectively escaped labour limitation. In this sense, labour limitation behaves less like a continuous regulator of population size and more like a viability constraint.

Introducing labour-dependent carrying capacity therefore does not produce a broad continuum of labour-constrained equilibria. Instead, it appears to operate primarily as a viability filter: populations that cannot meet labour requirements disappear, whereas surviving populations overwhelmingly retain enough labour to realise the nominal carrying capacity. Surviving populations are therefore not simply balancing labour supply and demand; they maintain an excess of productive capacity.

**_Main visual patterns_**

- Strong viability boundary. There is an almost empty region below the diagonal, indicating that `labour_available < labour_required` is rarely observed among surviving populations. This reinforces the interpretation of labour sufficiency as a threshold condition.  
- High-pressure populations extend further right. The “Persistence (high pressure)” panel includes runs with `labour_required > 3000` while still maintaining labour_available > labour_required. This suggests that high-pressure populations are not labour-limited; rather, they are labour-sufficient but ecologically stressed, which is precisely the distinction version 1.2.2 was designed to capture.  
- Overshooting populations are unusual. The overshooting panel is sparse, but the few observed points are informative. Several occupy extreme regions with high labour availability and/or high labour requirements, suggesting that overshooters may represent populations capable of maintaining unusually large productive cohorts. Given their rarity, however, this pattern should be interpreted cautiously.  

**_Interpretation of labour demand_**

The broad overlap in labour-demand-per-capita values across survival categories suggests that this parameter does not operate as a simple deterministic threshold under the full Sobol exploration. This is also one of the most striking features of the figure: although the points are coloured by labour-demand-per-capita, all three survival categories contain a wide range of values. The parameter is therefore important, but not because it directly determines collapse or persistence on its own.

Instead, labour-demand-per-capita appears to move populations closer to, or further from, the viability boundary defined by `labour_available = labour_required`. Whether a population ultimately collapses, persists under high pressure, or overshoots then depends on interactions with the demographic structure generated by fertility, mortality, nuptiality, and initial age composition. This interpretation is consistent with the random-forest results, where labour demand is influential because it modulates proximity to the viability boundary, while fertility and mortality parameters remain important even after labour demand is introduced.

### Density regulation dynamics

```{r}
#| label: regulation_dynamics_rr_prop
regulation_dynamics_rr_prop_plot <- sampled_trajectories |>
  group_by(run_unique_id) |>
  filter(row_number() == 1) |> 
  ungroup() |>
  
  plot_category_proportion(
    category_var = "regulation_regime",
    facet_var = "residence_rule",
    category_fill = TRUE,
    category_labs_size = 5,
    category_labs_angle = 45,
    category_labs_hjust = 0.5
  )
```

```{r}
#| label: regulation_dynamics_rr_prop_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    regulation_dynamics_rr_prop_plot,
    "v122_regulation_dynamics_rr_prop",
    width = 5,
    height = 3
  )
}
```

```{r} 
#| label: regulation_dynamics_rr_prop_plot_display 
#| fig.cap: "Distribution of density regulation regimes accross residence rules, for the trajectories sampled."
plotName = paste0("plots/v122_regulation_dynamics_rr_prop.png")

knitr::include_graphics(plotName)
```

```{r}
#| label: regulation_dynamics_rr_mean_cv_plot 
#| fig.cap: ""
regulation_dynamics_rr_mean_cv_plot <- sampled_trajectories |>
  plot_pressure_distributions(
    colour_var = "regulation_regime",
    shape_var = "residence_rule",
    stress_threshold = stress_threshold,
    oscillatory_threshold = oscillatory_threshold,
    thresholds_linetype = 2,
    point_alpha = 0.8
  )
```

```{r}
#| label: regulation_dynamics_rr_mean_cv_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    regulation_dynamics_rr_mean_cv_plot,
    "v122_regulation_dynamics_rr_mean_cv"
  )
}
```

```{r} 
#| label: regulation_dynamics_rr_mean_cv_plot_display 
#| fig.cap: "Distribution of mean and CV of population pressure accross density regulation regimes and residence rules, for the trajectories sampled."
plotName = paste0("plots/v122_regulation_dynamics_rr_mean_cv.png")

knitr::include_graphics(plotName)
```

In v1.2.2, the pressure metric becomes even more structurally informative than in v1.2.1. Pressure now reflects both age-weighted carrying-capacity demand and age-weighted labour supply, meaning that populations of equivalent size can experience very different effective pressures depending on their demographic composition. Ecological regulation is therefore no longer reducible to abundance alone, but emerges from the balance between demographic dependency and productive capacity.

### Age structure

#### Survival categories

```{r} 
#| label: age_structure_survival_pyramid_plot 
#| fig.cap: "Aggregate age structures of surviving populations across survival categories."
age_structures |>
  aggregate_age_structure(
    group_vars = c("survival"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet = "survival",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_survival_median_plot 
#| fig.cap: "Distribution of median age of surviving populations across survival categories."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival"
)
```

```{r} 
#| label: age-structure-survival-indicators-summary 
#| tab.cap: "Summary of age structure indicators for surviving populations across survival categories."
knitr::kable(indicators_summary(age_structure_indicators, group_vars = "survival"))
```

```{r} 
#| label: age_structure_pressure_median_cohend 
#| tab.cap: "Cohen's d for median age across low and high population pressure."
knitr::kable(
  interpret_cohens_d(
    cohens_d(
      median_age ~ pressure_regime,
      data = age_structure_indicators
    ),
    rules = "cohen1988"
  )
)
```

Surviving populations under lower pressure tend to exhibit older age structures and broader upper tails. This pattern does not necessarily indicate senescence or decline, but rather demographic efficiency: these populations maintain sufficient productive adults while avoiding the large dependent juvenile cohorts associated with high reproductive expansion. In this sense, lower pressure reflects not lower abundance, but a more balanced allocation between labour supply and carrying-capacity demand.

#### Residence rule and Coale-Demeny regions

```{r} 
#| label: age_structure_pyramid_plot_cdr_rr
pyramid_plot <- age_structures |>
  aggregate_age_structure(
    group_vars = c("residence_rule", "coale_demeny_region"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet = c("coale_demeny_region", "residence_rule"),
    show_mean = FALSE
    )
```

```{r}
#| label: age_structure_pyramid_plot_cdr_rr_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pyramid_plot,
    "v122_pyramid_residence_rule_region",
    height = 10
  )
}
```

```{r} 
#| label: age_structure_pyramid_plot_cdr_rr_display 
#| fig.cap: "Aggregate age structures of surviving populations across residence rules and Coale-Demeny regions."
plotName = paste0("plots/v122_pyramid_residence_rule_region.png")

knitr::include_graphics(plotName)
```

```{r} 
#| label: age_structure_indicators_plot_median_cdr_rr 
#| fig.cap: "Distribution of median age of surviving populations across residence rules and Coale-Demeny regions."
violin_plot <- age_structure_indicators |>
  plot_age_structure_indicators(
    x_var = "survival", 
    fill_var = "residence_rule", 
    facet_var = "coale_demeny_region"
  )
```

```{r}
#| label: age_structure_indicators_plot_median_cdr_rr_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    violin_plot,
    "v122_violin_residence_rule_region",
    width = 8,
    height = 3
  )
}
```

```{r} 
#| label: age_structure_indicators_plot_median_cdr_rr_display 
#| fig.cap: "Aggregate age structures of surviving populations across residence rules and Coale-Demeny regions."
plotName = paste0("plots/v122_violin_residence_rule_region.png")

knitr::include_graphics(plotName)
```

```{r} 
#| label: age_structure_indicators_summary_cdr_rr 
#| tab.cap: "Summary of age structure indicators for surviving populations across residence rules and Coale-Demeny regions."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "coale_demeny_region")
    )
)
```

```{r} 
#| label: age_structure_cohens_d_cdr_rr 
#| tab.cap: "Cohen's d for median age across residence rules."
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ residence_rule,
      data = age_structure_indicators
    ),
    rules = "cohen1988"
  )
)
```

Cohen's d for median age across pressure regimes, by residence rule:

```{r} 
#| label: age_structure_cohens_d_cdr_rr_by_rr_matri 
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ pressure_regime,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "matrilocal-matrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_cohens_d_cdr_rr_by_rr_patri 
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
      effectsize::cohens_d(
        median_age ~ pressure_regime,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

#### Interpretation

The age-structure analyses indicate that v1.2.2 introduces a tighter feedback between demographic composition and ecological regulation. In v1.2.1, age structure affected mortality through carrying-capacity demand; in v1.2.2 it additionally affects the productive capacity required to sustain that demand. Population viability therefore depends not only on dependency ratios but on the balance between consumers and effective labour contributors. This creates a more explicit coupling between reproduction, ageing, and ecological resilience.

## Sensitivity analysis

### Prepare data for sensitivity analysis

Prepare a dataframe out of `endstates` that contains only predictors (explored parameters) and response variables (`pressure`, `survival`, `median_age`), as well as splitting it according to `residence_rule`.

```{r} 
#| label: sensitivity_data
response_variables <- c(
  "pressure",
  "survival",
  "median_age",
  "log_carrying_capacity_effective"
)

sensitivity_data <- prepare_sensitivity_data_by_residence_rule(
  endstates,
  response_variables,
  parameter_metadata
)
```

### Fitting Random-forest models

```{r} 
#| label: sensitivity_rf
rf_models <- load_or_build_rf_models(
  model_version = MODEL_VERSION,
  sensitivity_data = sensitivity_data,
  response_variables = response_variables,
  SEED = 123,
  verbose = FALSE,
  overwrite = REMAKE_RF
)
```

Random forest models are fitted separately for each residence rule, using the same set of predictors. The response variables include `pressure`, `survival`, `median_age`, and `log_carrying_capacity_effective` allowing us to assess the importance of predictors for different aspects of population dynamics and structure. Notice that the model built for `median_age` uses only cases with surviving populations.

The models are built using a consistent random seed to ensure reproducibility, and the option to overwrite existing models is controlled by the `REMAKE_RF` flag.

### Parameter importance and partial dependence

#### Survival (probability of extinction end state)

```{r} 
#| label: sensitivity_survival_importance
rf_survival_importance <- bind_rows(
  
  extract_importance(
    rf_models$survival$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$survival$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)

param_color <- parameter_metadata$color[match(rf_survival_importance$parameter, parameter_metadata$parameter)]
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality, density).

```{r} 
#| label: sensitivity_survival_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population survival, by residence rule."
plot_importance(
  rf_survival_importance, 
  param_color = param_color, 
  facet_var = "residence_rule")
```

```{r} 
#| label: sensitivity_survival_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population survival, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_survival_importance, 
  groups = "residence_rule", 
  param_color = param_color
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_survival_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v122_survival_importance_dumbbell"
  )
}
```

Select top parameters in relation to survival:

```{r} 
#| label: sensitivity_survival_top_param
top_parameters_survival <- rf_survival_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(MeanDecreaseGini)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `survival` as response variable and no population ("Extinction") as response class. Plot each PD result with `pressure` and per `residence_rule`.

```{r} 
#| label: sensitivity_survival_pd
pd_survival <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunk) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_survival,
  response_variable = "survival",
  response_class = "Extinction" # used only with `survival` as response variable
)
```

```{r} 
#| label: sensitivity_survival_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_survival,
  top_parameters_survival,
  response_variable = "pressure",
  pd_label = "P(Extinction)",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_survival_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v122_survival_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_survival_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing population survival, by residence rule."
plotName = paste0("plots/v122_survival_param_pd.png")

knitr::include_graphics(plotName)
```

#### Population pressure

```{r} 
#| label: sensitivity_pressure_importance
rf_pressure_importance <- bind_rows(
  
  extract_importance(
    rf_models$pressure$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$pressure$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_pressure_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population pressure, by residence rule."
plot_importance(
  rf_pressure_importance, 
  param_color = param_color, 
  facet_var = "residence_rule",
  metric = "%IncMSE")
```

```{r} 
#| label: sensitivity_pressure_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population pressure, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_pressure_importance, 
  groups = "residence_rule", 
  param_color = param_color,
  metric = "%IncMSE"
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_pressure_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v122_pressure_importance_dumbbell"
  )
}
```

Select top parameters predicting pressure:

```{r} 
#| label: sensitivity_pressure_top_param
top_parameters_pressure <- rf_pressure_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(`%IncMSE`)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `pressure` as response variable. Plot each PD result with `pressure` and per `residence_rule`.

```{r} 
#| label: sensitivity_pressure_pd
pd_pressure <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunk) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = c(top_parameters_pressure, "cdmlt_level"), # force add cdmlt_level because it is not a top 10 in this version (so it is available for manuscript figure)
  response_variable = "pressure"
)
```

```{r} 
#| label: sensitivity_pressure_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_pressure,
  top_parameters_pressure,
  response_variable = "pressure",
  pd_label = "Expected change in population pressure",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_pressure_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v122_pressure_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_pressure_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing population pressure, by residence rule."
plotName = paste0("plots/v122_pressure_param_pd.png")

knitr::include_graphics(plotName)
```

The overall parameter ranking in relation to pressure remains structurally comparable to v1.2.1, but the introduction of labour supply introduces an additional layer of demographic mediation. Fertility and mortality parameters now affect pressure not only by shaping total population size, but by shaping the age composition that determines both demand and productive capacity.

#### Median age in end state population

```{r} 
#| label: sensitivity_median_age_importance
rf_median_age_importance <- bind_rows(
  
  extract_importance(
    rf_models$median_age$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$median_age$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_median_age_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting median age, by residence rule."
plot_importance(
  rf_median_age_importance, 
  param_color = param_color, 
  facet_var = "residence_rule",
  metric = "%IncMSE")
```

```{r} 
#| label: sensitivity_median_age_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting median age, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_median_age_importance, 
  groups = "residence_rule", 
  param_color = param_color,
  metric = "%IncMSE"
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_median_age_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v122_median_age_importance_dumbbell"
  )
}
```

Select top parameters predicting median age:

```{r} 
#| label: sensitivity_median_age_top_param
top_parameters_median_age <- rf_median_age_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(`%IncMSE`)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `median_age` as response variable (Extinction cases are not included).

```{r} 
#| label: sensitivity_median_age_pd
pd_median_age <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunck) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_median_age,
  response_variable = "median_age"
)
```

```{r} 
#| label: sensitivity_median_age_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_median_age,
  top_parameters_median_age,
  response_variable = "median_age",
  pd_label = "Expected change in median age",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_median_age_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v122_median_age_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_median_age_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing median age, by residence rule."
plotName = paste0("plots/v122_median_age_param_pd.png")

knitr::include_graphics(plotName)
```

The determinants of median age remain broadly similar to those observed in previous versions, indicating that labour constraints do not fundamentally alter the demographic mechanisms generating age structure. Instead, they selectively filter which age structures remain viable.

#### Effective carrying capacity

```{r} 
#| label: sensitivity_log_carrying_capacity_effective_importance
rf_log_carrying_capacity_effective_importance <- bind_rows(
  
  extract_importance(
    rf_models$log_carrying_capacity_effective$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$log_carrying_capacity_effective$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_log_carrying_capacity_effective_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting effective carrying capacity, by residence rule."
plot_importance(
  rf_log_carrying_capacity_effective_importance, 
  param_color = param_color, 
  facet_var = "residence_rule",
  metric = "%IncMSE")
```

```{r} 
#| label: sensitivity_log_carrying_capacity_effective_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting effective carrying capacity, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_log_carrying_capacity_effective_importance, 
  groups = "residence_rule", 
  param_color = param_color,
  metric = "%IncMSE"
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_log_carrying_capacity_effective_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v122_log_carrying_capacity_effective_importance_dumbbell"
  )
}
```

Select top parameters predicting effective carrying capacity:

```{r} 
#| label: sensitivity_log_carrying_capacity_effective_top_param
top_parameters_log_carrying_capacity_effective <- rf_log_carrying_capacity_effective_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(`%IncMSE`)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `log_carrying_capacity_effective` as response variable (Extinction cases are not included).

```{r} 
#| label: sensitivity_log_carrying_capacity_effective_pd
pd_log_carrying_capacity_effective <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunck) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_log_carrying_capacity_effective,
  response_variable = "log_carrying_capacity_effective"
)
```

```{r} 
#| label: sensitivity_log_carrying_capacity_effective_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_log_carrying_capacity_effective,
  top_parameters_log_carrying_capacity_effective,
  response_variable = "log_carrying_capacity_effective",
  pd_label = "Expected change in effective carrying capacity",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_log_carrying_capacity_effective_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v122_log_carrying_capacity_effective_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_log_carrying_capacity_effective_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing effective carrying capacity, by residence rule."
plotName = paste0("plots/v122_log_carrying_capacity_effective_param_pd.png")

knitr::include_graphics(plotName)
```

The importance ranking for effective carrying capacity differs from that of survival and pressure. Notably, labour-demand-per-capita has comparatively weaker direct effects here. This apparent paradox reflects the fact that the parameter primarily determines whether populations persist at all, rather than the equilibrium effective capacity they occupy once survival has been achieved.

### Age structure of surviving populations for selected parameters

Note: parameters are ordered in importance towards median age. The binomial categories created earlier in preprocessing, which split the parameter range according to the distribution median, are used.

#### `c1-fert`: high *versus* low fertility regimes

```{r} 
#| label: age_structure_c1_fert_pyramid_plot
#| fig.cap: "Population pyramids across fertility intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_intensity"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_intensity",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_fert_median_plot
#| fig.cap: "Median age of surviving populations across fertility regimes, by residence rule."
#| fig.width: 10
#| fig.height: 4
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_intensity"
)
```

```{r} 
#| label: age_structure_c1_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across fertility regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_intensity")
    )
)
```

Cohen's d for median age across fertility regimes, by residence rule:

```{r} 
#| label: age_structure_c1_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
    median_age ~ fertility_intensity,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_intensity,

    data =
        age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

#### `cdmlt-level`: mortality regimes

```{r}
#| label: age_structure_cdmlt_level_pyramid_plot
#| fig.cap: "Population pyramids across mortality regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("mortality_regime"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "mortality_regime",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_cdmlt_level_median_plot
#| fig.cap: "Median age of surviving populations across mortality regimes, by residence rule."
#| fig.width: 10
#| fig.height: 4
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "mortality_regime"
)
```

```{r} 
#| label: age_structure_cdmlt_level_indicators_summary
#| tab.cap: "Summary of age structure indicators across mortality regimes and residence rules."
knitr::kable(
  indicators_summary(age_structure_indicators, group_vars = c("residence_rule", "mortality_regime"))
)
```

Cohen's d for median age across mortality regimes, by residence rule:

```{r} 
#| label: age_structure_cdmlt_level_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ mortality_regime,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_cdmlt_level_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ mortality_regime,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

#### `sigma2_fert`: broad *versus* narrow post-peak fertility regimes

```{r} 
#| label: age_structure_sigma2_fert_pyramid_plot
#| fig.cap: "Population pyramids across fertility spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_spread_post_peak"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_spread_post_peak",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_fert_median_plot
#| fig.cap: "Median age of surviving populations across fertility spread regimes, by residence rule."
#| fig.width: 10
#| fig.height: 4
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_spread_post_peak"
)
```

```{r} 
#| label: age_structure_sigma2_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across fertility spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_spread_post_peak")
    )
)
```

Cohen's d for median age across fertility spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_post_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_post_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

#### `c1_women`: high *versus* low female marriage intensity regimes

```{r} 
#| label: age_structure_c1_women_pyramid_plot
#| fig.cap: "Population pyramids across female marriage intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_intensity_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_intensity_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_women_median_plot
#| fig.cap: "Median age of surviving populations across female marriage intensity regimes, by residence rule."
#| fig.width: 10
#| fig.height: 4
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_intensity_women"
)
```

```{r} 
#| label: age_structure_c1_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across female marriage intensity regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_intensity_women")
    )
)
```

Cohen's d for median age across female marriage intensity regimes, by residence rule:

```{r} 
#| label: age_structure_c1_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_intensity_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_intensity_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

#### `sigma1_men`: broad *versus* narrow male pre-peak marriage spread regimes

```{r} 
#| label: age_structure_sigma1_men_pyramid_plot
#| fig.cap: "Population pyramids across male pre-peak marriage spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_spread_pre_peak_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_spread_pre_peak_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_men_median_plot
#| fig.cap: "Median age of surviving populations across male pre-peak marriage spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_spread_pre_peak_men"
)
```

```{r} 
#| label: age_structure_sigma1_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male pre-peak marriage spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_spread_pre_peak_men")
    )
)
```

Cohen's d for median age across male pre-peak marriage spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_pre_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_spread_pre_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

#### `mu_men`: late *versus* early male marriage regimes

```{r} 
#| label: age_structure_mu_men_pyramid_plot
#| fig.cap: "Population pyramids across male marriage timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_timing_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_timing_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_men_median_plot
#| fig.cap: "Median age of surviving populations across male marriage timing regimes, by residence rule."
#| fig.width: 10
#| fig.height: 4
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_timing_men"
)
```

```{r} 
#| label: age_structure_mu_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male marriage timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_timing_men")
    )
)
```

Cohen's d for median age across male marriage timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_timing_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_timing_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

#### `sigma1_women`: broad *versus* narrow pre-peak fertility spread regimes

```{r} 
#| label: age_structure_sigma1_women_pyramid_plot
#| fig.cap: "Population pyramids across pre-peak fertility spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_timing"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_timing",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_women_median_plot
#| fig.cap: "Median age of surviving populations across pre-peak fertility spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_timing"
)
```

```{r} 
#| label: age_structure_sigma1_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across pre-peak fertility spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_timing")
    )
)
```

Cohen's d for median age across pre-peak fertility spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_timing,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_timing,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

#### `c1_men`: high *versus* low male marriage intensity regimes

```{r} 
#| label: age_structure_c1_men_pyramid_plot
#| fig.cap: "Population pyramids across male marriage intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("marriage_intensity_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "marriage_intensity_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_men_median_plot
#| fig.cap: "Median age of surviving populations across male marriage intensity regimes, by residence rule."
#| fig.width: 10
#| fig.height: 4
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "marriage_intensity_men"
)
```

```{r} 
#| label: age_structure_c1_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male marriage intensity regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "marriage_intensity_men")
    )
)
```

Cohen's d for median age across male marriage intensity regimes, by residence rule:

```{r} 
#| label: age_structure_c1_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_intensity_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ marriage_intensity_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

### Interpretation

As the effect of labour constraints on median age already indicated, the patterns in age structure are broadly similar to those observed in previous versions.

## Labour-demand threshold experiment

To better isolate the effect of labour demand per capita, an additional targeted experiment was conducted in which this parameter was varied while all others were held constant. Unlike the Sobol exploration, this threshold experiment was designed specifically to identify viability boundaries under changing labour requirements.

Simulations using the same parameter configuration, except `residence-rule`, and varying `labour_demand_per_capita` and `SEED`.

Strategy:

- Fix all parameters at a viable baseline.  
- Vary only labour_demand_per_capita.  
- Compare trajectories across residence rules.  
- Identify collapse thresholds.  

```{r} 
#| label: trajectories_params_labour_demand_per_capita_display
#| fig.cap: "Parameter configuration for the population trajectories sampled across different Coale-Demeny model levels."
knitr::kable(sampled_trajectories_labour_demand_per_capita |> 
  select(all_of(c("max_iterations", "max_population", 
                  names(parameter_ranges_v122), 
                  "coale_demeny_region", "maturity_age")),
         -c("labour_demand_per_capita")) |> 
  distinct() |>
  t())
```

```{r} 
#| label: trajectories_labour_demand_per_capita
#| fig.cap: "Population trajectories across different labour demand per capita, faceted by residence rule."
log_max_population <- ceiling(max(sampled_trajectories_labour_demand_per_capita$log_totalIndividuals))

pop_trajectories_plot <- sampled_trajectories_labour_demand_per_capita |>
  mutate(
    labour_demand_per_capita = factor(labour_demand_per_capita)
  ) |>
  plot_trajectories(
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = "labour_demand_per_capita",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "carrying_capacity",
    thresholds_linetype = 2
  )
  
pop_trajectories_plot
```

```{r}
#| label: trajectories_labour_demand_per_capita_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_plot,
    "v122_pop_trajectories_labour_demand_per_capita"
  )
}
```

```{r}
#| label: trajectories_labour_demand_per_capita_survival_plot
# collapse trajectories into endstates (one row per run)
survival_by_labour_demand <- sampled_trajectories_labour_demand_per_capita |>
  group_by(
    run_unique_id,
    residence_rule,
    labour_demand_per_capita
  ) |>
  slice_tail(n = 1) |>
  ungroup() |>
  mutate(
    survived = totalIndividuals > 0
  ) |>
  group_by(
    residence_rule,
    labour_demand_per_capita
  ) |>
  summarise(
    n_runs = n(),
    n_survived = sum(survived),
    p_survived = n_survived / n_runs,
    .groups = "drop"
  )

# plot survival curve
survival_threshold_plot <- survival_by_labour_demand |>
  ggplot(
    aes(
      x = labour_demand_per_capita,
      y = p_survived,
      color = residence_rule
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_viridis_d(
    option = "viridis",
    guide = guide_legend(title = "")
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.25)
  ) +
  labs(
    x = "Labour demand per capita",
    y = "Proportion surviving"
  ) +
  geom_vline(
    xintercept = c(0.3, 0.45),
    linetype = 2,
    alpha = 0.5
  )

survival_threshold_plot
```

```{r}
#| label: trajectories_labour_demand_per_capita_survival_table
knitr::kable(
  sampled_trajectories_labour_demand_per_capita |>
    group_by(labour_demand_per_capita, survival) |>
    summarise(
      mean_laboured_share = mean(laboured_share),
      median_laboured_share = median(laboured_share),
      min_laboured_share = min(laboured_share)
    )
)
```

**_Nonlinear labour-demand thresholds_**

The targeted threshold experiment shows that labour_demand_per_capita has a strongly nonlinear effect on demographic viability. In both residence systems, values below approximately 0.2 produce near-stable persistence. Beyond this range, however, survival declines rapidly, especially between 0.3 and 0.6. The pattern is therefore not one of gradual linear attrition, but of a viability threshold: once labour demand passes a critical point, collapse accelerates.

**Matrilocal-matrilineal system: delayed but abrupt collapse**

- 0.0–0.4: effectively full persistence  
- 0.5: near-total collapse  
- 0.6: total collapse  

This pattern indicates a relatively wide viable labour-demand envelope. The approximate threshold lies between 0.4 and 0.5, after which persistence gives way rapidly to collapse.

**Patrilocal-patrilineal system: earlier and more gradual decline**

The patrilocal-patrilineal system begins to lose viability much sooner. Its decline is evident even at low levels of labour demand:

- Survival is already below full persistence at 0.0.  
- Substantial losses occur by 0.2.  
- Collapse is nearly complete by 0.3.  
- No persistence is observed thereafter.  

The approximate threshold is therefore lower, around 0.2–0.3. Unlike the matrilocal-matrilineal system, however, the transition is smoother: less a sudden cliff than a progressive landslide. This suggests lower resilience, but also a more graded pathway into failure.

**_Residence-rule comparison_**

These results support a stronger interpretation of version 1.2.2 than simply stating that labour demand affects survival. Labour-dependent carrying capacity introduces a regime-dependent viability threshold, and both the position and shape of that threshold differ systematically by residence rule.

| System	| Threshold position | Transition shape |
| ------- | ------------------ | ---------------- |
| Matrilocal-matrilineal | Higher, around 0.4–0.5	| Abrupt
| Patrilocal-patrilineal | Lower, around 0.2–0.3 | Gradual

The key implication is that kinship structure mediates tolerance to labour scarcity. The two systems do not merely differ in how much labour demand they can absorb; they also display different geometries of failure. Matrilocal-matrilineal organisation increases tolerance to labour inefficiency before demographic viability is compromised, whereas patrilocal-patrilineal organisation shows earlier vulnerability.

The patrilocal-patrilineal system is already below full survival when labour demand is negligible. This indicates that labour demand is not the sole determinant of failure. Rather, it acts as a stress amplifier: higher labour requirements intensify existing demographic vulnerabilities, but do not create them alone.

**_Mechanism_**

The mechanism can be summarised as follows:

- Higher labour demand reduces effective carrying capacity.  
- Lower effective carrying capacity increases density-related mortality pressure.  
- Higher mortality pressure produces population decline and, beyond the threshold, collapse.  

However, the same perturbation propagates differently depending on kinship structure. A targeted threshold experiment varying labour demand per capita while holding all other parameters constant showed that the two residence systems differ not only in their critical labour-demand threshold but also in the shape of the transition from persistence to collapse. Matrilocal-matrilineal systems exhibit a higher and sharper viability threshold, whereas patrilocal-patrilineal systems display an earlier but more gradual decline in survival probability.

## Discussion notes

The introduction of age-dependent carrying-capacity demand in v1.2.1 allows a fixed carrying capacity to sustain larger nominal populations by reducing the effective burden imposed by children and elders. In v1.2.2, this same demographic logic is extended to labour supply, making productive capacity itself age-dependent. While this preserves the structural advantages of demographic efficiency, it also imposes stricter viability constraints, substantially increasing the frequency of extinction and collapse-prone trajectories. Across all model variants from v1.1 to v1.2.2, populations following the matrilocal-matrilineal rule consistently exhibit higher survivability, suggesting that their demographic organisation provides greater resilience under increasingly complex ecological constraints.

## Chapter conclusion

Version 1.2.2 completes the transition from purely abundance-based ecological regulation to a demographically mediated model of population viability. Whereas earlier versions treated carrying capacity as a fixed external constraint, the combined demand- and supply-weighted formulation makes ecological pressure an emergent property of age structure itself. Population persistence is therefore no longer determined simply by how many individuals exist, but by who those individuals are in demographic terms and how they balance consumption against productive labour. This substantially deepens the ecological realism of the model and establishes the conceptual foundations for the daily-cycle versions introduced in the following chapter.
