---
title: "Household Demography model: sensitivity analysis on base model (v.1.1)"
output: html_notebook
---

## Preparation

```{r, message=FALSE}
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(gridExtra)
library(scales)
library(randomForest)
```

```{r}
plotDirectory <- "plots/v11_"
plot_file_format <- c("png", "eps")[1] # modify index number to change format
```

## Experiment design with stochastic parameter settings

### Generate parameter values with Sobol Sampling

```{r}
sobol_parameter_sample_dataFileName <- "data/v11_sobol_parameter_values.RData"

sobol_parameter_ranges <- list(
  `initial-num-households` = c(2, 50),         
  `household_initial_age_distribution_min` = c(0, 5),        
  `household_initial_age_distribution_max` = c(6, 60),
  `max_couple_count_distribution_min` = c(1, 2),
  `max_couple_count_distribution_max` = c(3, 8),
  `acceptable-kinship-degree-for-couples` = c(0, 9),
  `cdmlt-level` = c(1, 25),
  `c1-fert` = c(0, 0.5),
  `mu-fert` = c(15, 30),
  `sigma1-fert` = c(1, 10),
  `sigma2-fert` = c(5, 20), 
  `c1-women` = c(0, 1),
  `mu-women` = c(15, 40),
  `sigma1-women` = c(1, 20),
  `sigma2-women` = c(1, 20), 
  `c1-men` = c(0, 1),
  `mu-men` = c(15, 40), 
  `sigma1-men` = c(1, 20), 
  `sigma2-men` = c(1, 20)
)
```

```{r, eval=FALSE}
library(randtoolbox)
set.seed(123)
n <- 2000  # Number of samples
k <- length(sobol_parameter_ranges)  # Number of parameters

# Generate Sobol sequence
sobol_sample_matrix <- sobol(n, dim = k, scrambling = 3)

get_values_from_sobol <- function(param_index){
  qunif(sobol_sample_matrix[,param_index], 
        min = sobol_parameter_ranges[[param_index]][1], 
        max = sobol_parameter_ranges[[param_index]][2])
}

```

```{r, eval=FALSE}
# Transform to specific distributions
params_sobol <- data.frame(
  `initial-num-households` = round(get_values_from_sobol(1)),
  `household_initial_age_distribution_min` = round(get_values_from_sobol(2)),
  `household_initial_age_distribution_max` = round(get_values_from_sobol(3)),
  `max_couple_count_distribution_min` = round(get_values_from_sobol(4)),
  `max_couple_count_distribution_max` = round(get_values_from_sobol(5)),
  `acceptable-kinship-degree-for-couples` = round(get_values_from_sobol(6)),
  `cdmlt-level` = round(get_values_from_sobol(7)),
      
  `c1-fert` = get_values_from_sobol(8),
  `mu-fert` = get_values_from_sobol(9),
  `sigma1-fert` = get_values_from_sobol(10),
  `sigma2-fert` = get_values_from_sobol(11),
      
  `c1-women` = get_values_from_sobol(12),
  `mu-women` = get_values_from_sobol(13),
  `sigma1-women` = get_values_from_sobol(14),
  `sigma2-women` = get_values_from_sobol(15),
  `c1-men` = get_values_from_sobol(16),
  `mu-men` = get_values_from_sobol(17),
  `sigma1-men` = get_values_from_sobol(18),
  `sigma2-men` = get_values_from_sobol(19)
)

head(params_sobol)
```

```{r, eval=FALSE}
save(params_sobol, file = sobol_parameter_sample_dataFileName)
```

```{r}
load(sobol_parameter_sample_dataFileName)
```

Visualising parameter space with `pairs()`:

```{r, eval=FALSE}
plotName = paste0(plotDirectory, "sobol_param_pairsPlot.png")

plotScale = 10

png(plotName,
    width = 300 * plotScale, height = 300 * plotScale)
pairs(params_sobol, pch='·')
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

Saving parameter configurations (experiments) in separate files:

```{r, eval=FALSE}
save_experiment <- function(experiment_df, experiment_row, version_name) {
  write(paste(unlist(experiment_df[experiment_row,], use.names = FALSE)), 
        file = paste0(dirname(getwd()), "/analysis/experiments/", version_name, "/", experiment_row, ".txt"))
}

for (i in 1:nrow(params_sobol)) {
  save_experiment(params_sobol, i, "v1.1")
}
```

## Analysis on population end states

>Run experiments in NetLogo Behaviour Space using the parameters `type-of-experiment` and `exp-number`.

### Data loading and preprocessing

Define usable reference to parameter names:

```{r}
param_names_netlogo <- names(sobol_parameter_ranges)
param_names_r <- gsub("-", ".", names(sobol_parameter_ranges))
```

Loading simulation experiments' end states:

```{r}
endstates_matri <- read.csv("simdata/v1.1/household-demography-main exp-endstates-matri-table.csv", skip = 6)
endstates_patri <- read.csv("simdata/v1.1/household-demography-main exp-endstates-patri-table.csv", skip = 6)
```

Preprocess dataset adapting from NetLogo'S specific format. Categorise end states according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

```{r}
MAX_POPULATION = endstates_matri$max.population[1]

extract_values_from_NetLogo_list <- function(char_vector) {
  # Remove brackets and split each string into two parts
  split_values <- gsub("\\[|\\]", "", char_vector) |>
    strsplit(" ")

  # Convert to numeric vectors
  first_values <- sapply(split_values, function(x) as.numeric(x[1]))
  second_values <- sapply(split_values, function(x) as.numeric(x[2]))

  # Return as a named list
  list(
    first = first_values,
    second = second_values
  )
}

preprocess_params <- function(endstate_df) {
  household_initial_age_distribution_range <- extract_values_from_NetLogo_list(endstate_df$household.initial.age.distribution)
  max_couple_count_distribution_range <- extract_values_from_NetLogo_list(endstate_df$max.couple.count.distribution)
  
  cbind(
    endstate_df,
    household_initial_age_distribution_min = household_initial_age_distribution_range[[1]],
    household_initial_age_distribution_max = household_initial_age_distribution_range[[2]],
    max_couple_count_distribution_min = max_couple_count_distribution_range[[1]],
    max_couple_count_distribution_max = max_couple_count_distribution_range[[2]]
  )
}

set_survival_category <- function(endstate_df) {
  endstate_df |>
    mutate(
      survival = case_when(
        totalIndividuals <= 0 ~ "extinction",
        totalIndividuals > MAX_POPULATION ~ "survival with growth",
        TRUE ~ "survival"
      )
    )
}
```

```{r}
endstates_matri <- endstates_matri |>
  preprocess_params() |>
  set_survival_category()
endstates_patri <- endstates_patri |>
  preprocess_params() |>
  set_survival_category()
```

### Overview of simulation results (population size)

Visualise `totalIndividuals`:

```{r}
plotName = paste0(plotDirectory, paste0("endstates_totalIndividuals.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1800, height = graphic_scale * 800, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- ggplot(endstates_matri, aes(x = totalIndividuals)) +
  geom_histogram() +
  ggtitle("'matrilocal-matrilineal'")
plot_2 <- ggplot(endstates_patri, aes(x = totalIndividuals)) +
  geom_histogram() +
  ggtitle("'patrilocal-patrilineal'") +
  theme(axis.title.y = element_blank())
grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1))
#layout(matrix(1:2, ncol = 2))
#hist(endstates_matri$totalIndividuals, breaks = 50, main = "'matrilocal-matrilineal'")
#hist(endstates_patri$totalIndividuals, breaks = 50, main = "'patrilocal-patrilineal'")
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

```{r}
table(endstates_matri$survival)
table(endstates_patri$survival)
```

```{r}
plotName = paste0(plotDirectory, paste0("endstates_survival.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1800, height = graphic_scale * 800, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- ggplot(endstates_matri, aes(x = survival)) +
  geom_bar() +
  ggtitle("'matrilocal-matrilineal'") +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_2 <- ggplot(endstates_patri, aes(x = survival)) +
  geom_bar() +
  ggtitle("'patrilocal-patrilineal'") +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1.2))
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

### Visualising population age structure

```{r}
source("library/age_structure.R")
```

```{r aggregated-pyramid}
pyramid_matri <- build_population_pyramid(
  endstates_matri,
  age_breaks = seq(0, 80, 5),
  normalize = FALSE
)
pyramid_patri <- build_population_pyramid(
  endstates_patri,
  age_breaks = seq(0, 80, 5),
  normalize = FALSE
)
```


```{r aggregated-pyramid}
plotName = paste0(plotDirectory, paste0("endstates_age-pyramid_stats.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1800, height = graphic_scale * 800, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- plot_population_pyramid(pyramid_matri, show_range = TRUE)
plot_2 <- plot_population_pyramid(pyramid_patri, show_range = TRUE, show_legend = FALSE)
grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1.2))
dev.off()
```

```{r}
plot_variability(plot_df)
```

```{r}
age_structure_sample <- sample_runs(endstates_matri, normalize = FALSE)
plot_sample_runs(age_structure_sample)
```

```{r}
matri_patri_diff <- diff_population_pyramids(pyramid_matri, pyramid_patri)
```

```{r}
plot_diff_population_pyramids(matri_patri_diff)
```

```{r}
extract_age_structure <- function(char_value) {
  split_values <- gsub("\\[|\\]", "", char_value) |>
    strsplit(" ")
  as.numeric(split_values[[1]])
}

build_age_structure_df <- function(endstate_df, selected_case = 500, max_age = 110) {
  survival_cases <- endstate_df |> filter(survival != "extinction")
  
  age_structure_men <- extract_age_structure(survival_cases[selected_case, "menAgeStructure"])
  age_structure_women <- extract_age_structure(survival_cases[selected_case, "womenAgeStructure"])
  
  # Put data into long format
  age_structure_df <- tibble(
    age = c(age_structure_men, age_structure_women),
    sex = c(rep("Male", length(age_structure_men)), 
            rep("Female", length(age_structure_women)))
  )
  
  # Bin ages into groups (5-year bins, adjust as needed)
  age_structure_df |>
    mutate(age_group = cut(age, breaks = seq(0, max_age, 5), right = FALSE))
}

summarise_age_structure_df <- function(age_structure_df) {
  age_structure_df |>
    count(sex, age_group) |> # Summarize counts per age group and sex
    mutate(count = ifelse(sex == "Male", -n, n)) # For pyramid: make men negative
}
```

```{r}
age_structure_matri <- build_age_structure_df(endstates_matri)
age_structure_patri <- build_age_structure_df(endstates_patri)
```

```{r}
age_structure_matri_summary <- summarise_age_structure_df(age_structure_matri)
age_structure_patri_summary <- summarise_age_structure_df(age_structure_patri)
```

```{r}
create_age_pyramid <- function(age_structure, title, x_label, y_label, fill_guide = FALSE, axis.title.y = TRUE) {
  
  p <- ggplot(age_structure, aes(x = age_group, y = count, fill = sex)) +
    geom_col(width = 1, color = "white") +
    coord_flip() +
    scale_y_continuous(labels = abs) +
    labs(
      title = title,
      x = y_label,
      y = x_label
    ) +
    theme_minimal()
  
  if (!fill_guide) {
    p <- p + guides(fill = fill_guide)
  }
  if (!axis.title.y) {
    p <- p + theme(axis.title.y = element_blank())
  }
  p
}

create_age_pyramid_pair <- function(age_structure_1, age_structure_2, 
                                    title_1, title_2, x_label, y_label,
                                    plot_name, plot_scale = 10, width = 80, height = 150)
{
  if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plot_name, width = graphic_scale * 1800, height = graphic_scale * 800, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plot_name ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }
  
  plot_1 <- create_age_pyramid(age_structure_1, title = title_1, x_label = x_label, y_label = y_label)
  
  plot_2 <- create_age_pyramid(age_structure_2, title = title_2, x_label = x_label, y_label = y_label, fill_guide = TRUE, axis.title.y = FALSE)
  
  grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1.4))
}
```

```{r}
plotName = paste0(plotDirectory, paste0("endstates_age-pyramid-example.png"))

create_age_pyramid_pair(age_structure_matri_summary, age_structure_patri_summary,
                   title_1 = "'matrilocal-matrilineal'",
                   title_2 = "'patrilocal-patrilineal'",
                   x_label = "Population count",
                   y_label = "Age group",
                   plot_name = plotName)

```

```{r}
knitr::include_graphics(plotName)
```

### Identifying survival parameter space

Calculate the Peristera-Kostaki parameter ranges associated to survival:
```{r}
compute_param_envelope <- function(param_df, low=0.025, high=0.975){
  apply(param_df, 2, quantile, probs = c(low, high), na.rm=TRUE) |>
    t() |> as.data.frame() #|> rename(lower = `2.5%`, upper = `97.5%`)
}

compute_survival_param_space <- function(endstate_df, param_names) {
  compute_param_envelope(endstate_df |> filter(survival != "extinction") |> select(all_of(param_names)))
}

param_names_pk <- param_names_r[7:19]

survival_param_space_matri <- compute_survival_param_space(endstates_matri, param_names_pk)
survival_param_space_patri <- compute_survival_param_space(endstates_patri, param_names_pk)
```

With residential-rule = 'matrilocal-matrilineal':
```{r}
knitr::kable(survival_param_space_matri)
```

With residential-rule = 'patrilocal-patrilineal':
```{r}
knitr::kable(survival_param_space_patri)
```

### Identifying relevant factors of survival

```{r}
covert_to_long_by_param <- function(df, param_names) {
  df |>
    pivot_longer(cols = all_of(param_names)) |>
    mutate(name = factor(name, levels = param_names))  # Preserve order
}
endstates_matri_long <- covert_to_long_by_param(endstates_matri, param_names_r)
endstates_patri_long <- covert_to_long_by_param(endstates_patri, param_names_r)
```

```{r}
source("library/bifurcationPlot.R")

create_bifurcation_plot_multiple <- function(df_long, plot_name, plot_scale = 10, num_columns = 3, width = 80, height = 150, strip_text_size = 1.2) {
  png(plot_name,
      width = width * plot_scale, height = height * plot_scale)
  
  ggplot(data = df_long, aes(x = value, y = totalIndividuals)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "glm") +
    facet_wrap(vars(name), ncol = num_columns, scales = "free_x", as.table = TRUE) +
    theme(
      strip.text = element_text(size = strip_text_size * plot_scale)
    )
  #dev.off()
}

plotName_matri = paste0(plotDirectory, "endstates_matri_bifurcationPlot.png")
create_bifurcation_plot_multiple(endstates_matri_long, plot_name = plotName_matri)

plotName_patri = paste0(plotDirectory, "endstates_patri_bifurcationPlot.png")
create_bifurcation_plot_multiple(endstates_patri_long, plot_name = plotName_patri)
```

```{r}
knitr::include_graphics(plotName_matri)
```

```{r}
knitr::include_graphics(plotName_patri)
```

Train a Random Forest model to assess parameter importance for each dataset:

```{r}
run_rf_for_totalIndividuals <- function(endstates_df, param_names, seed = 123) {
  endstates_df <- endstates_df |> select(all_of(c(param_names, "totalIndividuals")))
  
  # Train Random Forest to predict simulation output
  set.seed(seed)
  randomForest(totalIndividuals ~ ., data = endstates_df, importance = TRUE, ntree = 500)
}
```

```{r}
rf_matri <- run_rf_for_totalIndividuals(endstates_matri, param_names_r)
print(rf_matri)
```

```{r}
rf_patri <- run_rf_for_totalIndividuals(endstates_patri, param_names_r)
print(rf_patri)
```

Extract importance values and print as table:

```{r}
extract_rf_param_importance <- function(rf_model) {
  # Extract importance values
  importance_values <- importance(rf_model)
  
  # Convert to a data frame
  data.frame(Parameter = rownames(importance_values),
             MDA = importance_values[, 1], 
             MDG = importance_values[, 2])
}
```

```{r}
rf_matri_importance <- extract_rf_param_importance(rf_matri)
print(rf_matri_importance)
```

```{r}
rf_patri_importance <- extract_rf_param_importance(rf_patri)
print(rf_patri_importance)
```

Plot feature importance colouring parameters by domain (general, mortality, fertility, nuptiality):

```{r}
source("library/variableImportancePlot.R")

plotName = paste0(plotDirectory, "variableImportancePlot.png")

color_variable_map <- rev(c(
  rep("#0000a2", 6),
  "#bc272d", 
  rep("#db9b35", 4), 
  rep("#50ad9f", 8)
))

# png
png(filename = plotName, width = 2 * 129, height = 168, units="mm", res=1200)

variableImportancePlotPair(rf_matri, rf_patri,
                           title_extra = c("residential-rule = 'matrilocal-matrilineal'", "residential-rule = 'patrilocal-patrilineal'"),
                           plot_scale = 0.5, 
                           color_variable_map = color_variable_map, 
                           range_standardised = TRUE, 
                           log_scale = FALSE,
                           layout_type = 'horizontal')

dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

#### Initial household population

This parameter has very small positive effect on population survival:

```{r}
cat("matri | extinct |")
endstates_matri |>
  filter(totalIndividuals == 0) |>
  select(initial.num.households) |>
  summary()
cat("matri | survive |")
endstates_matri |>
  filter(totalIndividuals > 0) |>
  select(initial.num.households) |>
  summary()
cat("patri | extinct |")
endstates_patri |>
  filter(totalIndividuals == 0) |>
  select(initial.num.households) |>
  summary()
cat("patri | survive |")
endstates_patri |>
  filter(totalIndividuals > 0) |>
  select(initial.num.households) |>
  summary()
```

```{r}
plotName = paste0(plotDirectory, paste0("endstates_survival_initial-num-households.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1800, height = graphic_scale * 800, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- ggplot(endstates_matri, aes(x = survival, y = initial.num.households)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank()) +
  ggtitle("'matrilocal-matrilineal'")
plot_2 <- ggplot(endstates_patri, aes(x = survival, y = initial.num.households)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("'patrilocal-patrilineal'")
grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1))
# layout(matrix(1:2, ncol = 2))
# boxplot(endstates_matri$initial.num.households ~ (endstates_matri$totalIndividuals > 0))
# boxplot(endstates_patri$initial.num.households ~ (endstates_patri$totalIndividuals > 0))
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

#### Mortality factors

Coale-Demeny Life Tables Model level parameter has great effect:

```{r}
cat("matri | extinct |")
endstates_matri |>
  filter(totalIndividuals == 0) |>
  select(cdmlt.level) |>
  summary()
cat("matri | survive |")
endstates_matri |>
  filter(totalIndividuals > 0) |>
  select(cdmlt.level) |>
  summary()
cat("patri | extinct |")
endstates_patri |>
  filter(totalIndividuals == 0) |>
  select(cdmlt.level) |>
  summary()
cat("patri | survive |")
endstates_patri |>
  filter(totalIndividuals > 0) |>
  select(cdmlt.level) |>
  summary()
```

```{r}
plotName = paste0(plotDirectory, paste0("endstates_survival_cdmlt-level.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1800, height = graphic_scale * 800, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- ggplot(endstates_matri, aes(x = survival, y = cdmlt.level)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank()) +
  ggtitle("'matrilocal-matrilineal'")
plot_2 <- ggplot(endstates_patri, aes(x = survival, y = cdmlt.level)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("'patrilocal-patrilineal'")
grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1))
# layout(matrix(1:2, ncol = 2))
# boxplot(endstates_matri$cdmlt.level ~ (endstates_matri$totalIndividuals > 0))
# boxplot(endstates_patri$cdmlt.level ~ (endstates_patri$totalIndividuals > 0))
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

but region parameter has a smaller effect. East tends to survive more often.  

```{r}
cat("matri | survive? |")
table(endstates_matri$coale.demeny.region, endstates_matri$totalIndividuals > 0)
cat("patri | survive? |")
table(endstates_patri$coale.demeny.region, endstates_patri$totalIndividuals > 0)
```

#### Fertility factors

All have a visible positive effect, except `mu.fert`. `c1.fert` is the most influential. 

```{r}
plotName = paste0(plotDirectory, paste0("endstates_survival_pk-fert-params.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1600, height = graphic_scale * 2000, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- ggplot(endstates_matri, aes(x = survival, y = c1.fert)) +
  geom_boxplot() +
  ggtitle("'matrilocal-matrilineal'") +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_2 <- ggplot(endstates_matri, aes(x = survival, y = mu.fert)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_3 <- ggplot(endstates_matri, aes(x = survival, y = sigma1.fert)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_4 <- ggplot(endstates_matri, aes(x = survival, y = sigma2.fert)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())

plot_5 <- ggplot(endstates_patri, aes(x = survival, y = c1.fert)) +
  geom_boxplot() +
  ggtitle("'patrilocal-patrilineal'") +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
plot_6 <- ggplot(endstates_patri, aes(x = survival, y = mu.fert)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
plot_7 <- ggplot(endstates_patri, aes(x = survival, y = sigma1.fert)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
plot_8 <- ggplot(endstates_patri, aes(x = survival, y = sigma2.fert)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

grid.arrange(grobs = c(plot_1, plot_5, plot_2, plot_6, plot_3, plot_7, plot_4, plot_8), ncol = 2)

dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

#### Nuptiality factors

Residential rule showing only a small, with "matri" rule being slightly associated with survival.

```{r}
cat("matri | survive? |")
table(endstates_matri$residence.rule, endstates_matri$totalIndividuals > 0)
cat("patri | survive? |")
table(endstates_patri$residence.rule, endstates_patri$totalIndividuals > 0)
```

The effect of the acceptable kinship degree is not significant. However, there could be a very small effect, positive under "patri" and negative under "matri". 

```{r}
plotName = paste0(plotDirectory, paste0("endstates_survival_acceptable-kinship-degree-for-couples.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1800, height = graphic_scale * 800, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- ggplot(endstates_matri, aes(x = survival, y = acceptable.kinship.degree.for.couples)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8)) +
  ggtitle("'matrilocal-matrilineal'")
plot_2 <- ggplot(endstates_patri, aes(x = survival, y = acceptable.kinship.degree.for.couples)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("'patrilocal-patrilineal'")
grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1))
# layout(matrix(1:2, ncol = 2))
# boxplot(endstates_matri$cdmlt.level ~ (endstates_matri$totalIndividuals > 0))
# boxplot(endstates_patri$cdmlt.level ~ (endstates_patri$totalIndividuals > 0))
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

All nuptiality parameters for both men and women present a small, yet apparently consistent effect: c1 and sigma parameters have a positive effect while mu has a smaller negative effect. Effects are stronger for women's parameters when under the "matri" rule, while men's parameters are highlighted under the "patri" rule.

```{r}
plotName = paste0(plotDirectory, paste0("endstates_survival_pk-nupt-women-params.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1600, height = graphic_scale * 2000, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- ggplot(endstates_matri, aes(x = survival, y = c1.women)) +
  geom_boxplot() +
  ggtitle("'matrilocal-matrilineal'") +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_2 <- ggplot(endstates_matri, aes(x = survival, y = mu.women)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_3 <- ggplot(endstates_matri, aes(x = survival, y = sigma1.women)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_4 <- ggplot(endstates_matri, aes(x = survival, y = sigma2.women)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())

plot_5 <- ggplot(endstates_patri, aes(x = survival, y = c1.women)) +
  geom_boxplot() +
  ggtitle("'patrilocal-patrilineal'") +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
plot_6 <- ggplot(endstates_patri, aes(x = survival, y = mu.women)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
plot_7 <- ggplot(endstates_patri, aes(x = survival, y = sigma1.women)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
plot_8 <- ggplot(endstates_patri, aes(x = survival, y = sigma2.women)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

grid.arrange(grobs = c(plot_1, plot_5, plot_2, plot_6, plot_3, plot_7, plot_4, plot_8), ncol = 2)

dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

```{r}
plotName = paste0(plotDirectory, paste0("endstates_survival_pk-nupt-men-params.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1600, height = graphic_scale * 2000, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- ggplot(endstates_matri, aes(x = survival, y = c1.men)) +
  geom_boxplot() +
  ggtitle("'matrilocal-matrilineal'") +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_2 <- ggplot(endstates_matri, aes(x = survival, y = mu.men)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_3 <- ggplot(endstates_matri, aes(x = survival, y = sigma1.men)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())
plot_4 <- ggplot(endstates_matri, aes(x = survival, y = sigma2.men)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank())

plot_5 <- ggplot(endstates_patri, aes(x = survival, y = c1.men)) +
  geom_boxplot() +
  ggtitle("'patrilocal-patrilineal'") +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
plot_6 <- ggplot(endstates_patri, aes(x = survival, y = mu.men)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
plot_7 <- ggplot(endstates_patri, aes(x = survival, y = sigma1.men)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
plot_8 <- ggplot(endstates_patri, aes(x = survival, y = sigma2.men)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

grid.arrange(grobs = c(plot_1, plot_5, plot_2, plot_6, plot_3, plot_7, plot_4, plot_8), ncol = 2)

dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

### Pace of growth

A generally equivalent result for both residential rules regarding the pace of growth. The "matri" rule produces slightly more cases of rapid exponential (those cases reaching `max-population` within the first hundreds of simulation steps).

```{r}
plotName = paste0(plotDirectory, paste0("endstates_survival_pk-nupt-men-params.png"))

if (plot_file_format == "png") {
    graphic_scale <- 1
    #legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
    #axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5
    
    png(plotName, width = graphic_scale * 1800, height = graphic_scale * 800, res = 300, pointsize = 4)
  } else if (plot_file_format == "eps") {
    graphic_scale = 1.2
    # font_rescale = -0.2
    # axis_text_rescale = -0.1
    # margin_text_rescale = -0.5
    
    extrafont::loadfonts(device = "postscript")
    grDevices::cairo_ps(filename = plotName ,
                        pointsize = 12,
                        width = graphic_scale * 8,
                        height = graphic_scale * 4,
                        onefile = FALSE,
                        family = "sans"
    )
  }

plot_1 <- ggplot(endstates_matri, aes(x = X.step., color = survival)) +
  geom_histogram() +
  ggtitle("'matrilocal-matrilineal'") + 
  xlab("endstate step") +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(size = 5))
plot_2 <- ggplot(endstates_patri, aes(x = X.step., color = survival)) +
  geom_histogram() +
  ggtitle("'patrilocal-patrilineal'") + 
  xlab("endstate step") +
  theme(axis.text.x = element_text(size = 5),
        axis.title.y = element_blank())
grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1.6))

dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

```{r}
endstates_matri_long_survived <- endstates_matri_long |>
  filter(totalIndividuals > 0)
endstates_patri_long_survived <- endstates_patri_long |>
  filter(totalIndividuals > 0)
```

```{r}
source("library/bifurcationPlot.R")

create_bifurcation_plot_multiple <- function(df_long, plot_name, plot_scale = 10, num_columns = 3, width = 80, height = 150, strip_text_size = 1.2) {
  png(plot_name,
      width = width * plot_scale, height = height * plot_scale)
  
  ggplot(data = df_long, aes(x = value, y = X.step.)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "glm") +
    facet_wrap(vars(name), ncol = num_columns, scales = "free_x", as.table = TRUE) +
    theme(
      strip.text = element_text(size = strip_text_size * plot_scale)
    )
  #dev.off()
}

plotName_matri = paste0(plotDirectory, "endstates_matri_bifurcationPlot_growth.png")
create_bifurcation_plot_multiple(endstates_matri_long_survived, plot_name = plotName_matri)

plotName_patri = paste0(plotDirectory, "endstates_patri_bifurcationPlot_growth.png")
create_bifurcation_plot_multiple(endstates_patri_long_survived, plot_name = plotName_patri)
```

```{r}
knitr::include_graphics(plotName_matri)
```

```{r}
knitr::include_graphics(plotName_patri)
```

## Stochasticity and trajectory variation

### Using a fixed parameter setting

>Run simulations in NetLogo Behaviour Space using the same parameter configuration, except `residence-rule`, and varying `SEED`.

Loading simulation experiments' end states:

```{r}
trajectories_v11 <- read.csv("simdata/v1.1/household-demography-main exp-trajectories-table.csv", skip = 6)
```

Preprocess dataset adapting from NetLogo'S specific format. Categorise trajectories according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

```{r}
MAX_POPULATION = trajectories_v11$max.population[1]

extract_values_from_NetLogo_list <- function(char_vector) {
  # Remove brackets and split each string into two parts
  split_values <- gsub("\\[|\\]", "", char_vector) |>
    strsplit(" ")

  # Convert to numeric vectors
  first_values <- sapply(split_values, function(x) as.numeric(x[1]))
  second_values <- sapply(split_values, function(x) as.numeric(x[2]))

  # Return as a named list
  list(
    first = first_values,
    second = second_values
  )
}

preprocess_params <- function(trajectories_df) {
  household_initial_age_distribution_range <- extract_values_from_NetLogo_list(trajectories_df$household.initial.age.distribution)
  max_couple_count_distribution_range <- extract_values_from_NetLogo_list(trajectories_df$max.couple.count.distribution)
  
  cbind(
    trajectories_df,
    household_initial_age_distribution_min = household_initial_age_distribution_range[[1]],
    household_initial_age_distribution_max = household_initial_age_distribution_range[[2]],
    max_couple_count_distribution_min = max_couple_count_distribution_range[[1]],
    max_couple_count_distribution_max = max_couple_count_distribution_range[[2]]
  )
}

set_survival_category <- function(trajectories_df) {
  trajectories_df |>
    group_by(X.run.number.) |>
    mutate(
      endstate_total_individuals = dplyr::last(totalIndividuals),
      survival = case_when(
        endstate_total_individuals <= 0 ~ "extinction",
        endstate_total_individuals > MAX_POPULATION ~ "survival with growth",
        TRUE ~ "survival"
      )
    ) |>
    ungroup()
}

trajectories_v11 <- trajectories_v11 |>
  preprocess_params() |>
  set_survival_category()
```

```{r}
params_v11 <- trajectories_v11 |> 
  select(all_of(param_names_r)) |> 
  distinct() |> 
  t() |>
  apply(1, unique)
params_v11
```

```{r}
plot_name <- file.path(paste0(plotDirectory, "trajectories.", plot_file_format))

if (plot_file_format == "png") {
  graphic_scale <- 1
  legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
  axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5

  png(plot_name, width = graphic_scale * 1600, height = graphic_scale * 800, res = 300, pointsize = 4)
} else if (plot_file_format == "eps") {
  graphic_scale = 1.2
  # font_rescale = -0.2
  # axis_text_rescale = -0.1
  # margin_text_rescale = -0.5

  extrafont::loadfonts(device = "postscript")
  grDevices::cairo_ps(filename = plot_name ,
                      pointsize = 12,
                      width = graphic_scale * 8,
                      height = graphic_scale * 4,
                      onefile = FALSE,
                      family = "sans"
                      )
}

ggplot(trajectories_v11, aes(x = X.step., y = totalIndividuals, group = X.run.number., color = as.factor(X.run.number.))) +
  geom_line(linewidth = .1) +
  facet_wrap(~ residence.rule, scales = "free") +
  xlab("time (years)") +
  ylab("population (individuals)") +
  #ylim(c(0, MAX_POPULATION)) +
  guides(color = "none") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10 * (graphic_scale + axis_title_rescale)),
        axis.text = element_text(size = 10 * (graphic_scale + axis_text_rescale)))

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plot_name)
```

### Exploring `cdmlt-level` with a fixed parameter setting

>Run simulations in NetLogo Behaviour Space using the same parameter configuration, except `residence-rule`, and varying `cdmlt-level` and `SEED`.

Loading simulation experiments' end states:

```{r}
trajectories_v11_cdmlt_level <- read.csv("simdata/v1.1/household-demography-main exp-trajectories_cdmlt-level-table.csv", skip = 6)
```

Preprocess dataset adapting from NetLogo'S specific format. Categorise trajectories according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

```{r}
MAX_POPULATION = trajectories_v11_cdmlt_level$max.population[1]

trajectories_v11_cdmlt_level <- trajectories_v11_cdmlt_level |>
  preprocess_params() |>
  set_endstate_survival_category()
```

```{r}
plot_name <- file.path(paste0(plotDirectory, "trajectories_cdmlt-level.", plot_file_format))

if (plot_file_format == "png") {
  graphic_scale <- 1
  legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
  axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5

  png(plot_name, width = graphic_scale * 1600, height = graphic_scale * 800, res = 300, pointsize = 4)
} else if (plot_file_format == "eps") {
  graphic_scale = 1.2
  # font_rescale = -0.2
  # axis_text_rescale = -0.1
  # margin_text_rescale = -0.5

  extrafont::loadfonts(device = "postscript")
  grDevices::cairo_ps(filename = plot_name ,
                      pointsize = 12,
                      width = graphic_scale * 8,
                      height = graphic_scale * 4,
                      onefile = FALSE,
                      family = "sans"
                      )
}

ggplot(trajectories_v11_cdmlt_level, aes(x = X.step., y = totalIndividuals, group = X.run.number., color = as.factor(cdmlt.level))) +
  geom_line(linewidth = .1) +
  facet_wrap(~ residence.rule, scales = "free") +
  xlab("time (years)") +
  ylab("population (individuals)") +
  #ylim(c(0, MAX_POPULATION)) +
  guides(color = guide_legend(title = "cdmlt-level")) +
  theme_minimal() +
  theme(axis.title = element_text(size = 10 * (graphic_scale + axis_title_rescale)),
        axis.text = element_text(size = 10 * (graphic_scale + axis_text_rescale)),
        legend.position = "bottom")

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plot_name)
```

### Exploring `c1-fert` Using a fixed parameter setting

>Run simulations in NetLogo Behaviour Space using the same parameter configuration, except `residence-rule`, and varying `c1-fert` and `SEED`.

Loading simulation experiments' end states:

```{r}
trajectories_v11_c1_fert <- read.csv("simdata/v1.1/household-demography-main exp-trajectories_c1-fert-table.csv", skip = 6)
```

Preprocess dataset adapting from NetLogo'S specific format. Categorise trajectories according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

```{r}
MAX_POPULATION = trajectories_v11_c1_fert$max.population[1]

trajectories_v11_c1_fert <- trajectories_v11_c1_fert |>
  preprocess_params() |>
  set_endstate_survival_category()
```

```{r}
plot_name <- file.path(paste0(plotDirectory,"trajectories_c1_fert.", plot_file_format))

if (plot_file_format == "png") {
  graphic_scale <- 1
  legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
  axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5

  png(plot_name, width = graphic_scale * 1600, height = graphic_scale * 800, res = 300, pointsize = 4)
} else if (plot_file_format == "eps") {
  graphic_scale = 1.2
  # font_rescale = -0.2
  # axis_text_rescale = -0.1
  # margin_text_rescale = -0.5

  extrafont::loadfonts(device = "postscript")
  grDevices::cairo_ps(filename = plot_name ,
                      pointsize = 12,
                      width = graphic_scale * 8,
                      height = graphic_scale * 4,
                      onefile = FALSE,
                      family = "sans"
                      )
}

ggplot(trajectories_v11_c1_fert, aes(x = X.step., y = totalIndividuals, group = X.run.number., color = as.factor(c1.fert))) +
  geom_line(linewidth = .1) +
  facet_wrap(~ residence.rule, scales = "free") +
  xlab("time (years)") +
  ylab("population (individuals)") +
  #ylim(c(0, MAX_POPULATION)) +
  guides(color = guide_legend(title = "c1-fert")) +
  theme_minimal() +
  theme(axis.title = element_text(size = 10 * (graphic_scale + axis_title_rescale)),
        axis.text = element_text(size = 10 * (graphic_scale + axis_text_rescale)),
        legend.position = "bottom")

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plot_name)
```
