---
execute:
  echo: true
  warning: false
  message: false
params:
  show_diagnostics: false
---

# Base model with daily iteration (v1.3)

## Setup

Set constant variables for this chapter.

```{r}
#| label: setup_constants
MODEL_VERSION <- "v1.3"
EXPORT_FIGURES <- TRUE
REMAKE_RF <- FALSE
REMAKE_DP <- FALSE
```

Load libraries, sources and set ggplot theme for the notebook.

```{r} 
#| label: setup_libraries
library(data.table)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(ggtext)
library(zoo)
library(stringr)
library(effectsize)
library(randomForest)
library(randtoolbox)
library(rngWELL)
library(pdp)
library(forcats)
library(patchwork)
library(scales)
```

```{r} 
#| label: setup_load_sources
source("library/source_folder.R")
source_folder("library")
source_folder("library/structure")
source_folder("library/sensitivity")
```

```{r} 
#| label: setup_ggplot_theme
theme_set(
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold"),
      plot.title = element_text(face = "bold")
    )
)
```

## Experiment design

Sobol sampling is used to generate a diverse set of parameter combinations across the defined parameter space. This allows us to explore how different parameter configurations affect the model's outcomes and to identify which parameters are most influential in driving population survival or extinction. The actual sampling was done with `library/generate_sobol.R` script, which generates a sample with 2,000 parameter combinations across the parameter ranges defined in `parameter_ranges.R`. Note that given the higher computational cost of daily iterations, a nested subset of 512 combinations were subsequently selected to generate sufficiently-robust simulation results that can be compared to the yearly versions.

```{r} 
#| label: sobol_load
# Load object as Rdata file
load("data/v13_sobol_parameter_values.RData")
```

Sobol sampling is applied to the following parameter ranges:

```{r} 
#| label: sobol_ranges_display
parameter_ranges_v13 <- as.data.frame(parameter_ranges$v1.3)

knitr::kable(
  parameter_ranges_v13 |>
    as.matrix() |>
    t() |>
    as.data.frame() |>
    setNames(c("min", "max"))
)
```

```{r} 
#| label: sobol_pairs 
#| include: !expr params$show_diagnostics 
#| fig.cap: "Parameter ranges explored in the Sobol sampling for v1.3."
# Visualising parameter space with `pairs()`:
plotName = paste0("diagnostics/", MODEL_VERSION, "_sobol_param_pairsPlot.png")

knitr::include_graphics(plotName)
```

The 512-subset of the generated parameter combinations were then used to run simulations in NetLogo's Behaviour Space, for each combination of `residence-rule` ("matrilineal-matrilocal" and "patrilineal-patrilocal") and `coale-demeny-region` ("north", "south", "east", and "west"), varying the `SEED` parameter (n = 3) to introduce stochasticity, for a total of 12,288 simulations. Moreover, all runs assume a fixed `maturity-age-in-years` of 15 years, a year length of 365 days (no leap years), and a gestation period of 280 days (40 weeks). Remember that `maturity-age-in-years` determines a conservative age solely for social maturity, used as a criterion for household viability, while union-formation and reproductive maturity are regulated probabilistically through the Peristera-Kostaki parameters in the fertility and nuptiality submodels.

## Data loading and preprocessing

Load simulation end states generated with the Sobol-sampled parameter combinations and sampled trajectories datasets. Preprocess datasets to adapt from NetLogo's specific format and categorise end states and trajectories. 

```{r}
#| label: preprocessing_vars
growth_threshold = 5000 # endstates with population equal or higher than this are considered "Growth" cases
```

```{r} 
#| label: preprocessing
endstates <- load_endstates(MODEL_VERSION) |>
  preprocess_simdata(
    model_version = MODEL_VERSION,
    growth_threshold = growth_threshold
  )
```

```{r} 
#| label: preprocessing_trajectories
file_name <- "data/v13_trajectories.RData"
if (!file.exists(file_name))
{
  sampled_trajectories <- load_and_preprocess_trajectories_v13_v14x_single(
    MODEL_VERSION,
    growth_threshold = growth_threshold,
    sample_interval_days = 365,
    step_start = 1,
    step_end = 182500, # max-iterations is 500 years or 182500 days
    ignored_variables = c(
        "womenAgeStructure",
        "menAgeStructure"
    ),
    keep_variables = c(
        "model_version",
        "run_unique_id",
        "run_number",
        "residence_rule",
        "max_iterations",
        "max_iterations_in_days",
        "max_population", 
        names(parameter_ranges_v13), 
        "coale_demeny_region", 
        "maturity_age_in_years", 
        "gestation_period_in_days", 
        "year_length_in_days",
        "step",
        "currentYear",
        "currentDayOfYear",
        "log_totalIndividuals",
        "totalIndividuals",
        "totalHouseholds",
        "survival"
    )
  )
  
  save_trajectories(
    sampled_trajectories,
    model_version = MODEL_VERSION
  )
}
```

```{r}
#| label: preprocessing_trajectories_symmetric
rm(sampled_trajectories)
file_name <- "data/v13_trajectories_symmetric.RData"
if (file.exists(file_name))
{
  sampled_trajectories_symmetric <- readRDS(file_name)
} else {
  sampled_trajectories_symmetric <- load_and_preprocess_trajectories_v13_v14x_single(
    MODEL_VERSION,
    suffix = "symmetric",
    growth_threshold = growth_threshold,
    sample_interval_days = 365,
    step_start = 1,
    step_end = 182500, # max-iterations is 500 years or 182500 days
    ignored_variables = c(
        "womenAgeStructure",
        "menAgeStructure"
    ),
    keep_variables = c(
        "model_version",
        "run_unique_id",
        "run_number",
        "residence_rule",
        "max_iterations",
        "max_iterations_in_days",
        "max_population", 
        names(parameter_ranges_v13), 
        "coale_demeny_region", 
        "maturity_age_in_years", 
        "gestation_period_in_days", 
        "year_length_in_days",
        "step",
        "currentYear",
        "currentDayOfYear",
        "log_totalIndividuals",
        "totalIndividuals",
        "totalHouseholds",
        "survival"
    )
  )
  
  save_trajectories(
    sampled_trajectories_symmetric,
    model_version = MODEL_VERSION,
    suffix = "symmetric"
  )
}
```

```{r}
#| label: preprocessing_trajectories_cdmlt_level
rm(sampled_trajectories_symmetric)
file_name <- "data/v13_trajectories_cdmlt_level.RData"
if (file.exists(file_name))
{
  sampled_trajectories_cdmlt_level <- readRDS(file_name)
} else {
  sampled_trajectories_cdmlt_level <- load_and_preprocess_trajectories_v13_v14x_single(
    MODEL_VERSION,
    suffix = "cdmlt-level",
    growth_threshold = growth_threshold,
    sample_interval_days = 365,
    step_start = 1,
    step_end = 182500, # max-iterations is 500 years or 182500 days
    ignored_variables = c(
        "womenAgeStructure",
        "menAgeStructure"
    ),
    keep_variables = c(
        "model_version",
        "run_unique_id",
        "run_number",
        "residence_rule",
        "max_iterations",
        "max_iterations_in_days",
        "max_population", 
        names(parameter_ranges_v13), 
        "coale_demeny_region", 
        "maturity_age_in_years", 
        "gestation_period_in_days", 
        "year_length_in_days",
        "step",
        "currentYear",
        "currentDayOfYear",
        "log_totalIndividuals",
        "totalIndividuals",
        "totalHouseholds",
        "survival"
    )
  )
  
  save_trajectories(
    sampled_trajectories_cdmlt_level,
    model_version = MODEL_VERSION,
    suffix = "cdmlt_level"
  )
}
```

```{r}
#| label: preprocessing_trajectories_c1_fert
rm(sampled_trajectories_cdmlt_level)
file_name <- "data/v13_trajectories_c1_fert.RData"
if (file.exists(file_name))
{
  sampled_trajectories_c1_fert <- readRDS(file_name)
} else {
  sampled_trajectories_c1_fert <- load_and_preprocess_trajectories_v13_v14x_single(
    MODEL_VERSION,
    suffix = "c1-fert",
    growth_threshold = growth_threshold,
    sample_interval_days = 365,
    step_start = 1,
    step_end = 182500, # max-iterations is 500 years or 182500 days
    ignored_variables = c(
        "womenAgeStructure",
        "menAgeStructure"
    ),
    keep_variables = c(
        "model_version",
        "run_unique_id",
        "run_number",
        "residence_rule",
        "max_iterations",
        "max_iterations_in_days",
        "max_population", 
        names(parameter_ranges_v13), 
        "coale_demeny_region", 
        "maturity_age_in_years", 
        "gestation_period_in_days", 
        "year_length_in_days",
        "step",
        "currentYear",
        "currentDayOfYear",
        "log_totalIndividuals",
        "totalIndividuals",
        "totalHouseholds",
        "survival"
    )
  )
  
  save_trajectories(
    sampled_trajectories_c1_fert,
    model_version = MODEL_VERSION,
    suffix = "c1_fert"
  )
}
```

As with v1.1, the variable `survival` is created during preprocessing to classify whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

Because of the sheer number of daily iterations required to approach comparable time scales of previous versions, trajectories datasets are simply too large to be loaded conventionally in R. For this reason, we need to use the function argument `ignored_variables` (alternatively also `selected_variables` is available) to skip loading unnecessary variables, as well as potentially restricting the span of the simulation data to a subset of iteration steps (`start_step`, `end_step`). The different trajectory-type data frames must be loaded, preprocessed, saved as RData, and clear, so that only one is in memory simultaneously.

### Surviving population structure

```{r} 
#| label: age_structure_prepare
group_vars = c(
  "survival",
  "residence_rule",
  "coale_demeny_region",
  "mortality_regime",
  "fertility_intensity",
  "fertility_timing",
  "fertility_spread_pre_peak",
  "fertility_spread_post_peak",
  "union_timing_women",
  "union_timing_men",
  "union_intensity_women",
  "union_intensity_men",
  "union_spread_pre_peak_women",
  "union_spread_pre_peak_men",
  "union_spread_post_peak_women",
  "union_spread_post_peak_men",
  "amenorrhea_length"
)

age_structures <- endstates |>

  filter(
    survival != "Extinction"
  ) |>
  
  characterise_mort_fert_nupt_regimes() |>

  prepare_age_structure(
    age_breaks = seq(0, 100, 5),
    normalize = TRUE,
    group_vars = group_vars
  )

age_structure_indicators <- age_structures |>
  compute_age_structure_indicators(group_vars = group_vars)

endstates <- endstates |>
  left_join(
    age_structure_indicators |>
      select(
        run_unique_id,
        median_age,
        youth_share,
        elderly_share,
        dependency_ratio
      ),
    by = "run_unique_id"
  )
```

Age structures of surviving populations are prepared by categorising them according to the same grouping variables used in the sensitivity analysis, including categorical parameters (i.e., `residence_rule`, `coale_demeny_region`) and parameters converted into binary categories based on their median value, exclusively those regulating demographic probabilities (e.g., `mortality_regime`, `fertility_intensity`, etc.).

Furthermore, this selection includes the categorical variable `survival` that distinguishes between populations that became extinct, persisted without reaching the maximum population size, and reached the maximum population size. 
This allows us to compare the age structures of surviving populations across different demographic regimes and residence systems, and to identify patterns or differences associated with population persistence or growth.

Moreover, a set of indicators is computed for each age structure, including median age, dependency ratios, and measures of age-structure irregularity, which can provide further insights into the demographic characteristics of surviving populations and how they relate to different parameter regimes and residence systems. 
These indicators can help identify specific features of the age structure associated with population persistence or growth, and understand how different demographic regimes and residence systems may influence the age structure of surviving populations. Namely, the indicators computed include:

- `median_age`: the age that divides the population into two halves, with 50% of the population being younger and 50% being older.  
- `youth_share`: the ratio of the population aged 0-14 to the population aged 15-64, which indicates the proportion of dependents in the population.  
- `reproductive_share`: the ratio of the population aged 15-64 to the population aged 65 and over, which indicates the proportion of reproductive-age individuals in the population.  
- `elderly_share`: the ratio of the population aged 65 and over to the population aged 15-64, which indicates the proportion of elderly dependents in the population.  
- `total_dependency_ratio`: the ratio of the population aged 0-14 and 65 and over to the population aged 15-64, which indicates the overall proportion of dependent individuals in the population.  

## Overview of end states and dynamics

### Population persistence

We start with an overview of end states across residence rules and the categorical "regions" of the Coale-Demeny model (`coale-demeny-region`), which are the two categorical parameters explored outside the Sobol sampling. The histogram shows the distribution of final population sizes for each residence rule, with different colours indicating whether the population became extinct or survived. The vertical line represents the maximum population size defined in the model, which helps visually assess how many simulations reached or exceeded this threshold. The x-axis can be set to a logarithmic scale to better visualise differences in population sizes among those that survived.

```{r} 
#| label: endstates_distrib 
#| fig.cap: "Distribution of final population sizes across residence rules and Coale-Demeny regions."
log_max_population <- max(endstates$log_totalIndividuals)

endstates_distrib_size_plot <- endstates |>
  plot_endstate_variable_distribution(
    main_var = "log_totalIndividuals",
    fill_var = "survival",
    facet_var = c("coale_demeny_region", "residence_rule"),
    bins = 40,
    x_axis_label = "Final population size",
    y_axis_label = "Simulation count",
    threshold_values = log_max_population,
    thresholds_linetype = 2,
    x_axis_log10 = TRUE,
    x_max = log_max_population,
    y_axis_log10 = TRUE
  )
endstates_distrib_size_plot
```

```{r}
#| label: endstates_distrib_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    endstates_distrib_size_plot,
    "v13_endstates_distrib"
  )
}
```

```{r} 
#| label: outcome_prop 
#| fig.cap: "Proportion of simulations across survival category, residence rules and Coale-Demeny regions."
outcome_prop_plot <- endstates |>
  plot_category_proportion_stacked(
    x_var = "residence_rule",
    fill_var = "survival",
    facet_var = "coale_demeny_region",
    facet_ncol = 2,
    facet_scales = "fixed",
    y_axis_label = "Share of simulations"
  )
outcome_prop_plot
```

```{r}
#| label: outcome_prop_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    outcome_prop_plot,
    "v13_outcome_prop"
  )
}
```

```{r} 
#| label: outcome_prop_noextinction 
#| fig.cap: "Proportion of surviving simulations across residence rules and Coale-Demeny regions."
outcome_prop_plot_noextinction <- endstates |>
  filter(survival != "Extinction") |>
  plot_category_proportion_stacked(
    x_var = "residence_rule",
    fill_var = "survival",
    facet_var = "coale_demeny_region",
    facet_ncol = 2,
    facet_scales = "fixed",
    y_axis_label = "Share of simulations"
  )
outcome_prop_plot_noextinction
```

```{r}
#| label: outcome_prop_noextinction_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    outcome_prop_plot_noextinction,
    "v13_outcome_prop_plot_noextinction"
  )
}
```

The transition to daily iteration shifts the distribution of demographic outcomes relative to v1.1, but extinction does not dominate the explored parameter space as strongly as previously inferred. Approximately half of the simulations survive to the end state, with a modestly higher survival frequency under matrilocal-matrilineal residence. Among surviving populations, Growth remains the most common outcome under both residence rules, accounting for roughly three quarters of survivors, while Persistence represents a substantially larger share than in v1.1.

Daily demographic scheduling therefore appears to reduce the frequency and tempo of unconstrained expansion without eliminating it. Explicit gestation and postpartum amenorrhoea introduce delays into reproductive replacement, causing more viable populations to remain below the technical `max_population` threshold within the simulated time horizon.

Differences associated with `coale_demeny_region` remain small, and residence-rule contrasts are modest relative to the overall effects of the demographic parameters explored.

### Temporal trajectories

```{r}
#| label: trajectories_load_1
rm(sampled_trajectories_symmetric, preprocessing_trajectories_cdmlt_level, sampled_trajectories_c1_fert)
sampled_trajectories <- readRDS("data/v13_trajectories.RData")
```

The parameter configuration used to generate the trajectories is the same across residence rules, except for the `residence_rule` parameter itself. This allows us to directly compare the population dynamics under different residence systems while controlling for other factors. The configuration used for this was:

```{r} 
#| label: trajectories_size_params_display 
#| tab.cap: "Parameter configuration for the population trajectories sampled."
knitr::kable(sampled_trajectories |> 
               mutate(
                 max_iterations_in_days = max_iterations * year_length_in_days,
                 max_iterations_in_days_cut = max(step)
               ) |>
               select(all_of(c("max_iterations",
                               "max_iterations_in_days",
                               "max_iterations_in_days_cut", # show this in case loading was partial
                               "max_population", 
                               names(parameter_ranges_v13), 
                               "coale_demeny_region", 
                               "maturity_age_in_years", "gestation_period_in_days", "year_length_in_days" ))) |> 
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

Dynamics of population trajectories across time steps, faceted by residence rule. The plot shows individual trajectories with low opacity and the median trajectory with higher opacity. The y-axis is on a logarithmic scale to better visualize differences in population sizes over time, especially when there are large variations between trajectories.

```{r} 
#| label: trajectories_size_plot 
#| fig.cap: "A sample of population trajectories across time steps, faceted by residence rule."
pop_trajectories_plot <- sampled_trajectories |>
  plot_trajectories(
    x_var = "currentYear",
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.08,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = "survival",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "max_population",
    thresholds_linetype = 2
  )
pop_trajectories_plot
```

```{r}
#| label: trajectories_size_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_plot,
    "v13_pop_trajectories"
  )
}
rm(pop_trajectories_plot)
```

Relative to the annual baseline, daily scheduling produces smooth, continuous population trajectories in which demographic gains accumulate progressively through time. Under this favourable parameter configuration, all trajectories grow towards `max_population`, but patrilocal-patrilineal populations tend to reach the ceiling somewhat earlier than matrilocal-matrilineal populations.

In order to confirm that the above asymmetry is not caused by the asymmetry in the Peristera-Kostaki nuptiality parameters, the same experiment was performed with the following configuration:

```{r}
#| label: trajectories_symmetric_load
rm(sampled_trajectories, preprocessing_trajectories_cdmlt_level, sampled_trajectories_c1_fert)
sampled_trajectories_symmetric <- readRDS("data/v13_trajectories_symmetric.RData")
```

```{r} 
#| label: trajectories_size_params_symmetric_display 
#| tab.cap: "Parameter configuration for the population trajectories sampled."
knitr::kable(sampled_trajectories_symmetric |> 
               mutate(
                 max_iterations_in_days = max_iterations * year_length_in_days,
                 max_iterations_in_days_cut = max(step)
               ) |>
               select(all_of(c("max_iterations",
                               "max_iterations_in_days",
                               "max_iterations_in_days_cut", # show this in case loading was partial
                               "max_population", 
                               names(parameter_ranges_v13), 
                               "coale_demeny_region", 
                               "maturity_age_in_years", "gestation_period_in_days", "year_length_in_days" ))) |> 
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

Dynamics of population trajectories across time steps, faceted by residence rule. The plot shows individual trajectories with low opacity and the median trajectory with higher opacity. The y-axis is on a logarithmic scale to better visualize differences in population sizes over time, especially when there are large variations between trajectories.

```{r} 
#| label: trajectories_size_symmetric_plot 
#| fig.cap: "A sample of population trajectories across time steps, faceted by residence rule."
pop_trajectories_symmetric_plot <- sampled_trajectories_symmetric |>
  plot_trajectories(
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.08,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = NULL,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "max_population",
    thresholds_linetype = 2
  )
pop_trajectories_symmetric_plot
```

```{r}
#| label: trajectories_size_symmetric_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_symmetric_plot,
    "v13_pop_trajectories_symmetric"
  )
}
rm(pop_trajectories_symmetric_plot)
```

Equalising the male and female Peristera-Kostaki union schedules reduces one potential source of asymmetry but does not eliminate the difference in growth tempo. The targeted experiment therefore indicates that the residence-rule contrast is not generated solely by sex differences in the union-formation schedules. It must also arise from the way residence rules determine household continuity and movements associated with union formation.

This trajectory contrast remains conditional on the selected demographic configuration. The global exploration shows only modest differences in overall survival between residence rules, so faster growth in this experiment should not be interpreted as a general patrilocal survival advantage.

### Household composition

```{r}
#| label: household_structure_plot
#| fig.cap: ""
household_structure_plot <- endstates |> 
  plot_household_structure_faceted_fixed(
    color_var = "max_couple_count_distribution_max",
    point_alpha = 0.5,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "fixed",
    xy_log10 = TRUE,
    x_axis_label = "number of households",
    y_axis_label = "number of individuals",
    color_label = "Max. couple count allowed in a household\n(parametric maximum)",
    legend_position = "bottom"
  )
household_structure_plot
```

```{r}
#| label: household_structure_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    household_structure_plot,
    "v13_household_structure"
  )
}
```

Household number remains strongly related to total population size under both residence systems, indicating that household formation broadly scales with demographic expansion. Nevertheless, considerable residual variation remains around this relationship. Across the Sobol exploration, mean household size is somewhat larger under matrilocal-matrilineal than patrilocal-patrilineal residence.

```{r}
#| label: trajectories_load_2
rm(sampled_trajectories_symmetric, preprocessing_trajectories_cdmlt_level, sampled_trajectories_c1_fert)
sampled_trajectories <- readRDS("data/v13_trajectories.RData")
```

```{r}
#| label: trajectories_household_structure_plot 
#| fig.cap: "A sample of household population trajectories across time steps, faceted by residence rule."
trajectories_household_structure_plot <- 
  sampled_trajectories |>
  group_by(residence_rule) |>
  mutate(
    household_structure_ratio = (totalIndividuals / totalHouseholds)
  ) |>
  ungroup() |>
  plot_trajectories(
    y_var = "household_structure_ratio",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = FALSE,
    y_axis_label = "Mean individual count per household",
    group_var = NULL,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = NULL
  )
trajectories_household_structure_plot
```

```{r}
#| label: trajectories_household_structure_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    trajectories_household_structure_plot,
    "v13_trajectories_household_structure",
    width = 8,
    height = 4
  )
}
rm(trajectories_household_structure_plot)
```

```{r}
#| label: trajectories_household_structure_plot_peaks
trajectories_household_structure_plot_peaks <- 
  sampled_trajectories |>
  group_by(residence_rule) |>
  mutate(
    household_structure_ratio = if_else(totalHouseholds == 0, 0, (totalIndividuals / totalHouseholds)),
    # Find the absolute peak value within the entire residence rule
    global_max = max(household_structure_ratio, na.rm = TRUE)
  ) |>
  group_by(residence_rule, run_number) |>
  mutate(
    # Check if this specific run contains that global peak value
    selected_trajectory = any(household_structure_ratio == global_max)
  ) |>
  ungroup() |>
  plot_trajectories(
    y_var = "household_structure_ratio",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = FALSE,
    y_axis_label = "Mean individual count per household",
    group_var = NULL,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "fixed",
    y_threshold_var = NULL
  ) +
  # Add the highlighted layer on top
  geom_line(
    data = \(df) filter(df, selected_trajectory == TRUE),
    aes(
      x = step, 
      y = household_structure_ratio,
      group = run_number
    ),
    color = "red",
    linewidth = 1,
    alpha = 1
  )
```

```{r}
#| label: trajectories_household_structure_plot_peaks_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    trajectories_household_structure_plot_peaks,
    "v13_household_structure_trajectories_peaks",
    width = 8,
    height = 4
  )
}
rm(trajectories_household_structure_plot_peaks)
```

```{r} 
#| label: trajectories_household_structure_plot_peaks_display 
#| fig.cap: "A sample of household population trajectories across time steps, faceted by residence rule, highlighting in red the trajectories that achieve the maximum value in each subset."
plotName = paste0("plots/v13_household_structure_trajectories_peaks.png")

knitr::include_graphics(plotName)
```

The targeted trajectories show a common temporal pattern: mean household size increases during the early phase of population expansion and subsequently declines towards a comparatively stable late-run range. Patrilocal-patrilineal trajectories display a somewhat larger early peak in this particular configuration, but both systems converge towards similar household sizes thereafter.

The sampled trajectories therefore provide little evidence that either residence system generates a sustained escalation of household size. Residence organisation appears to modify transient household growth and fission dynamics more strongly than long-run household size.

### Age structure

#### Survival categories

```{r} 
#| label: age_structure_survival_pyramid_plot 
#| fig.cap: "Aggregate age structures of surviving populations across survival categories."
age_structures |>
  aggregate_age_structure(
    group_vars = c("survival"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "survival",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_survival_median_plot 
#| fig.cap: "Distribution of median age of surviving populations across survival categories."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival"
)
```

```{r} 
#| label: age-structure-survival-indicators-summary 
#| tab.cap: "Summary of age structure indicators for surviving populations across survival categories."
knitr::kable(indicators_summary(age_structure_indicators, group_vars = "survival"))
```

```{r} 
#| label: age_structure_survival_median_cohend 
#| tab.cap: "Cohen's d for median age across survival categories."
knitr::kable(
  interpret_cohens_d(
    cohens_d(
      median_age ~ survival,
      data = age_structure_indicators
    ),
    rules = "cohen1988"
  )
)
```

Growth and Persistence remain strongly differentiated by age structure. Persistent populations have a mean median age of 30.8 years, compared with 21.5 years among populations reaching `max_population`, and they contain substantially smaller juvenile cohorts. The separation is large (\(d=1.73\)), although slightly weaker than the corresponding contrast in v1.1.

The principal change is instead an upward shift in age structure. Both persistent and growing populations are older in v1.3 than in the annual baseline, consistent with the slower reproductive tempo introduced by explicit gestation and postpartum amenorrhoea. Daily scheduling therefore alters the quantitative expression of demographic renewal without changing the fundamental association between rapid growth and younger populations.

Persistent populations also exhibit broader upper tails, but these should not be interpreted as intrinsically unstable. In this density-independent model, Persistence simply identifies viable configurations that do not reach the technical population ceiling within the simulated horizon.

#### Residence rule and Coale-Demeny regions

```{r} 
#| label: age_structure_pyramid_plot_cdr_rr
pyramid_plot <- age_structures |>
  aggregate_age_structure(
    group_vars = c("residence_rule", "coale_demeny_region"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = c("coale_demeny_region", "residence_rule"),
    show_mean = FALSE
    )
```

```{r}
#| label: age_structure_pyramid_plot_cdr_rr_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pyramid_plot,
    "v13_pyramid_residence_rule_region",
    height = 10
  )
}
```

```{r} 
#| label: age_structure_pyramid_plot_cdr_rr_display 
#| fig.cap: "Aggregate age structures of surviving populations across residence rules and Coale-Demeny regions."
plotName = paste0("plots/v13_pyramid_residence_rule_region.png")

knitr::include_graphics(plotName)
```

```{r} 
#| label: age_structure_indicators_plot_median_cdr_rr 
#| fig.cap: "Distribution of median age of surviving populations across residence rules and Coale-Demeny regions."
violin_plot <- age_structure_indicators |>
  plot_age_structure_indicators(
    x_var = "survival", 
    fill_var = "residence_rule", 
    facet_var = "coale_demeny_region"
  )
```

```{r}
#| label: age_structure_indicators_plot_median_cdr_rr_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    violin_plot,
    "v13_violin_residence_rule_region",
    width = 6,
    height = 3
  )
}
```

```{r} 
#| label: age_structure_indicators_plot_median_cdr_rr_display 
#| fig.cap: "Distribution of median age of surviving populations across residence rules and Coale-Demeny regions."
plotName = paste0("plots/v13_violin_residence_rule_region.png")

knitr::include_graphics(plotName)
```

```{r} 
#| label: age_structure_indicators_summary_cdr_rr 
#| tab.cap: "Summary of age structure indicators for surviving populations across residence rules and Coale-Demeny regions."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "coale_demeny_region")
    )
)
```

```{r} 
#| label: age_structure_cohens_d_cdr_rr 
#| tab.cap: "Cohen's d for median age across residence rules."
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ residence_rule,
      data = age_structure_indicators
    ),
    rules = "cohen1988"
  )
)
```

Cohen's d for median age across survival categories, by residence rule:

```{r} 
#| label: age_structure_cohens_d_cdr_rr_by_rr_matri 
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ survival,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "matrilocal-matrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_cohens_d_cdr_rr_by_rr_patri 
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
      effectsize::cohens_d(
        median_age ~ survival,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

Residence rule has only a very small aggregate association with age structure. Matrilocal-matrilineal populations have slightly higher median ages than patrilocal-patrilineal populations across all four Coale-Demeny regions, but the overall effect is negligible (\(d=0.055\)). Regional differences are similarly modest, with South generally producing the oldest and West the youngest surviving populations.

By contrast, the distinction between Persistence and Growth remains large under both residence systems. It is slightly stronger under matrilocal-matrilineal residence (\(d=1.81\)) than under patrilocal-patrilineal residence (\(d=1.66\)). Thus, demographic growth regime remains much more consequential for age composition than residence organisation itself.

Daily iteration therefore does not produce a strong residence-specific age-structure divergence. Its principal structural effect is the general ageing of surviving populations and the slower renewal associated with explicit reproductive delays.

## Sensitivity analysis

### Prepare data for sensitivity analysis

Prepare a dataframe out of `endstates` that contains only predictors (explored parameters) and response variables (`survival`, `log_totalIndividuals`, `median_age`), as well as splitting it according to `residence_rule`.

```{r} 
#| label: sensitivity_data
response_variables <- c(
  "log_totalIndividuals",
  "survival",
  "median_age"
)

sensitivity_data <- prepare_sensitivity_data_by_residence_rule(
  endstates,
  response_variables,
  parameter_metadata
)
```

### Fitting Random-forest models

```{r} 
#| label: sensitivity_rf
rf_models <- load_or_build_rf_models(
  model_version = MODEL_VERSION,
  sensitivity_data = sensitivity_data,
  response_variables = response_variables,
  SEED = 123,
  verbose = FALSE,
  overwrite = REMAKE_RF
)
```

Random forest models are fitted separately for each residence rule and response variable, using the same set of predictors. The primary response variable, `survival`, classifies end states as extinction, persistence, or growth, using parameter combinations as categorical outcomes. The more nuanced response variables `log_totalIndividuals` and `median_age` are used only after filtering the data to include cases where a population is present at the end state (i.e. `survival != "Extinction"`).

The models are built using a consistent random seed to ensure reproducibility, and the option to overwrite existing models is controlled by the `REMAKE_RF` flag.

### Parameter importance and partial dependence

#### Survival (probability of extinction end state)

```{r} 
#| label: sensitivity_survival_importance
rf_survival_importance <- bind_rows(
  
  extract_importance(
    rf_models$survival$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$survival$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)

param_color <- parameter_metadata$color[match(rf_survival_importance$parameter, parameter_metadata$parameter)]
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_survival_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population survival, by residence rule."
plot_importance(
  rf_survival_importance, 
  param_color = param_color, 
  facet_var = "residence_rule")
```

```{r} 
#| label: sensitivity_survival_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population survival, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_survival_importance, 
  groups = "residence_rule", 
  param_color = param_color
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_survival_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v13_survival_importance_dumbbell"
  )
}
```

Select top parameters in relation to survival:

```{r} 
#| label: sensitivity_survival_top_param
top_parameters_survival <- rf_survival_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(MeanDecreaseGini)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `survival` as response variable and no population ("Extinction") as response class. Plot each PD result with `log_totalIndividuals` and per `residence_rule`.

```{r} 
#| label: sensitivity_survival_pd
pd_survival <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunk) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_survival,
  response_variable = "survival",
  response_class = "Extinction" # used only with `survival` as response variable
)
```

```{r} 
#| label: sensitivity_survival_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_survival,
  top_parameters_survival,
  response_variable = "log_totalIndividuals",
  pd_label = "P(Extinction)",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_survival_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v13_survival_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_survival_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing population survival, by residence rule."
plotName = paste0("plots/v13_survival_param_pd.png")

knitr::include_graphics(plotName)
```

Fertility peak intensity (`c1_fert`) remains overwhelmingly the strongest predictor of extinction in v1.3 under both residence rules. Its partial dependence retains the strongly nonlinear form observed in the annual baseline: extinction probability falls rapidly as fertility rises through the lower part of the explored range and then declines more gradually at higher values.

A secondary group of predictors includes female union timing (`mu_women`), baseline mortality intensity (`cdmlt_level`), post-peak fertility spread (`sigma2_fert`), and female pre-peak union spread and peak intensity. Higher Coale-Demeny levels reduce extinction risk, but mortality no longer approaches fertility peak intensity in predictive importance.

Residence-specific asymmetry remains clearest among union-formation parameters, with female schedules generally receiving greater importance under matrilocal-matrilineal residence and male schedules under patrilocal-patrilineal residence. This reproduces the interaction between local retention and sex-specific union formation already observed in the annual baseline.

The daily-specific `amenorrhea_period_in_days` parameter contributes to survival, but it is not among the dominant predictors. Explicit reproductive spacing therefore modifies demographic viability without replacing fertility peak intensity as the principal control.

#### Final population size (`log_totalIndividuals`)

```{r} 
#| label: sensitivity_size_importance
rf_size_importance <- bind_rows(
  
  extract_importance(
    rf_models$log_totalIndividuals$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$log_totalIndividuals$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_size_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population size, by residence rule."
plot_importance(
  rf_size_importance, 
  param_color = param_color, 
  facet_var = "residence_rule",
  metric = "%IncMSE")
```

```{r} 
#| label: sensitivity_size_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population size, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_size_importance, 
  groups = "residence_rule", 
  param_color = param_color,
  metric = "%IncMSE"
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_size_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v13_size_importance_dumbbell"
  )
}
```

Select top parameters predicting population size:

```{r} 
#| label: sensitivity_size_top_param
top_parameters_size <- rf_size_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(`%IncMSE`)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `log_totalIndividuals` as response variable (Extinction cases are not included).

```{r} 
#| label: sensitivity_size_pd
pd_size <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunck) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_size,
  response_variable = "log_totalIndividuals"
)
```

```{r} 
#| label: sensitivity_size_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_size,
  top_parameters_size,
  response_variable = "log_totalIndividuals",
  pd_label = "Expected change in population size",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_size_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v13_size_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_size_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing population size, by residence rule."
plotName = paste0("plots/v13_size_param_pd.png")

knitr::include_graphics(plotName)
```

Among surviving populations, fertility peak intensity is again the strongest predictor of final population size. Baseline mortality occupies a clear but substantially lower second position, followed by union-formation and fertility-spread parameters. This hierarchy broadly preserves the demographic foundations observed in v1.1 while shifting their quantitative importance under daily scheduling.

Partial dependence indicates that higher fertility and lower mortality increase expected final population size, but interpretation remains constrained by the technical `max_population` stopping rule. Many successful trajectories accumulate at the same upper limit, so this response distinguishes slower-growing from ceiling-reaching configurations more effectively than it measures unconstrained long-term population size.

The daily formulation therefore primarily changes the tempo with which demographic configurations approach the ceiling rather than introducing a qualitatively new determinant of population abundance.

#### Median age

```{r} 
#| label: sensitivity_median_age_importance
rf_median_age_importance <- bind_rows(
  
  extract_importance(
    rf_models$median_age$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$median_age$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_median_age_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting median age, by residence rule."
plot_importance(
  rf_median_age_importance, 
  param_color = param_color, 
  facet_var = "residence_rule",
  metric = "%IncMSE")
```

```{r} 
#| label: sensitivity_median_age_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting median age, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_median_age_importance, 
  groups = "residence_rule", 
  param_color = param_color,
  metric = "%IncMSE"
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_median_age_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v13_median_age_importance_dumbbell"
  )
}
```

Select top parameters predicting median age:

```{r} 
#| label: sensitivity_median_age_top_param
top_parameters_median_age <- rf_median_age_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(`%IncMSE`)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `median_age` as response variable (Extinction cases are not included).

```{r} 
#| label: sensitivity_median_age_pd
pd_median_age <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunck) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_median_age,
  response_variable = "median_age"
)
```

```{r} 
#| label: sensitivity_median_age_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_median_age,
  top_parameters_median_age,
  response_variable = "median_age",
  pd_label = "Expected change in median age",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_median_age_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v13_median_age_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_median_age_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing median age, by residence rule."
plotName = paste0("plots/v13_median_age_param_pd.png")

knitr::include_graphics(plotName)
```

Fertility peak intensity is again the dominant determinant of median age, with higher values producing substantially younger surviving populations. Several parameters describing the breadth of fertility and union-formation schedules form the secondary tier, while `amenorrhea_period_in_days` also emerges as a meaningful daily-specific predictor.

Baseline mortality has comparatively little importance for median age. Daily scheduling therefore does not strengthen mortality control over age composition; instead, it makes reproductive pacing explicit through gestation, postpartum amenorrhoea, and the temporal distribution of fertility and union formation.

The new amenorrhoea parameter is particularly informative in this respect. Its importance is well below fertility peak intensity but comparable to several established demographic schedule parameters, showing that birth spacing can alter age composition even when the broad fertility schedule remains unchanged.

### Age structure of surviving populations for selected parameters

Note: parameters are grouped by submodels and meaning, not by importance. The binary categories created earlier in preprocessing, which split the parameter range according to the distribution median, are used.

#### `c1_fert`: high *versus* low fertility peak intensity regimes

```{r} 
#| label: age_structure_c1_fert_pyramid_plot
#| fig.cap: "Population pyramids across fertility peak intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_intensity"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_intensity",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_fert_median_plot
#| fig.cap: "Median age of surviving populations across fertility peak intensity regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_intensity"
)
```

```{r} 
#| label: age_structure_c1_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across fertility peak intensity regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_intensity")
    )
)
```

Cohen's d for median age across fertility peak intensity regimes, by residence rule:

```{r} 
#| label: age_structure_c1_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
    median_age ~ fertility_intensity,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_intensity,

    data =
        age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

High fertility peak intensity produces markedly younger populations under both residence systems. The effect is large under matrilocal-matrilineal residence (\(d=-0.81\)) and medium under patrilocal-patrilineal residence (\(d=-0.72\)). The effect is therefore somewhat stronger under matrilocal-matrilineal, not patrilocal-patrilineal, residence, and is weaker than in v1.1 despite remaining the largest parameter-specific age-structure effect.

#### `mu_fert`: early *versus* late fertility timing

```{r} 
#| label: age_structure_mu_fert_pyramid_plot
#| fig.cap: "Population pyramids across (female) fertility timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_timing"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_timing",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_fert_median_plot
#| fig.cap: "Median age of surviving populations across (female) fertility timing regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_timing"
)
```

```{r} 
#| label: age_structure_mu_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across (female) fertility timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_timing")
    )
)
```

Cohen's d for median age across fertility timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_timing,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_timing,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Fertility peak timing has little influence on median age. Earlier peak fertility is associated with only a very small decrease in median age under matrilocal-matrilineal residence (\(d=-0.09\)), while no detectable difference is present under patrilocal-patrilineal residence (\(d=0.02\)). Explicit gestation and postpartum amenorrhoea therefore do not translate into a stronger structural effect of the fertility peak location itself.

#### `sigma1_fert`: broad *versus* narrow pre-peak fertility spread regimes

```{r} 
#| label: age_structure_sigma1_fert_pyramid_plot
#| fig.cap: "Population pyramids across pre-peak fertility spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_spread_pre_peak"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_spread_pre_peak",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_fert_median_plot
#| fig.cap: "Median age of surviving populations across pre-peak fertility spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_spread_pre_peak"
)
```

```{r} 
#| label: age_structure_sigma1_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across pre-peak fertility spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_spread_pre_peak")
    )
)
```

Cohen's d for median age across pre-peak fertility spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_pre_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_pre_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader pre-peak fertility schedules are associated with slightly younger surviving populations, but the effects are very small (\(d=-0.18\) and \(-0.10\)). Earlier reproductive participation therefore contributes modestly to demographic rejuvenation, much less strongly than fertility peak intensity.

#### `sigma2_fert`: broad *versus* narrow post-peak fertility regimes

```{r} 
#| label: age_structure_sigma2_fert_pyramid_plot
#| fig.cap: "Population pyramids across fertility post-peak spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_spread_post_peak"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_spread_post_peak",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_fert_median_plot
#| fig.cap: "Median age of surviving populations across fertility post-peak spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_spread_post_peak"
)
```

```{r} 
#| label: age_structure_sigma2_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across fertility post-peak spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_spread_post_peak")
    )
)
```

Cohen's d for median age across fertility spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_post_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_post_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader post-peak fertility schedules produce younger, not older, surviving populations. The effect is small under both residence rules and somewhat stronger under patrilocal-patrilineal residence (\(d=-0.47\) versus \(-0.40\)). Extending fertility across later reproductive ages therefore sustains cohort renewal and lowers median age, rather than ageing the population. This effect is stronger than in the annual baseline.

#### `cdmlt_level`: mortality regimes

```{r}
#| label: age_structure_cdmlt_level_pyramid_plot
#| fig.cap: "Population pyramids across mortality regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("mortality_regime"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "mortality_regime",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_cdmlt_level_median_plot
#| fig.cap: "Median age of surviving populations across mortality regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "mortality_regime"
)
```

```{r} 
#| label: age_structure_cdmlt_level_indicators_summary
#| tab.cap: "Summary of age structure indicators across mortality regimes and residence rules."
knitr::kable(
  indicators_summary(age_structure_indicators, group_vars = c("residence_rule", "mortality_regime"))
)
```

Cohen's d for median age across mortality regimes, by residence rule:

```{r} 
#| label: age_structure_cdmlt_level_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ mortality_regime,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_cdmlt_level_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ mortality_regime,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

Lower-mortality regimes produce slightly older surviving populations, but the structural effect is weak in v1.3. It is effectively absent under matrilocal-matrilineal residence (\(d=-0.04\), confidence interval including zero) and very small under patrilocal-patrilineal residence (\(d=-0.12\)). Mortality remains relevant to survival and population growth, but it is not a major determinant of age composition among surviving populations.

#### `c1_women`: high *versus* low female union peak intensity regimes

```{r} 
#| label: age_structure_c1_women_pyramid_plot
#| fig.cap: "Population pyramids across female union peak intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_intensity_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_intensity_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_women_median_plot
#| fig.cap: "Median age of surviving populations across female union peak intensity regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_intensity_women"
)
```

```{r} 
#| label: age_structure_c1_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across female union peak intensity regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_intensity_women")
    )
)
```

Cohen's d for median age across female union peak intensity regimes, by residence rule:

```{r} 
#| label: age_structure_c1_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_intensity_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_intensity_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

Higher female union peak intensity is associated with slightly younger populations under both residence rules. The effect is very small and nearly identical in magnitude (\(d=-0.14\) and \(-0.15\)), providing little evidence for a meaningful residence-specific difference.

#### `mu_women`: late *versus* early female union regimes

```{r} 
#| label: age_structure_mu_women_pyramid_plot
#| fig.cap: "Population pyramids across female union timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_timing_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_timing_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_women_median_plot
#| fig.cap: "Median age of surviving populations across female union timing regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_timing_women"
)
```

```{r} 
#| label: age_structure_mu_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across female union timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_timing_women")
    )
)
```

Cohen's d for median age across female union timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_timing_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_timing_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Earlier female union timing is associated with younger populations under matrilocal-matrilineal residence, but the effect is very small (\(d=-0.14\)). Under patrilocal-patrilineal residence the effect is essentially absent (\(d\approx0\)). Female union timing therefore retains some structural importance where women form the locally retained line, but daily scheduling does not amplify this effect across both systems.

#### `sigma1_women`: broad *versus* narrow pre-peak female union spread regimes

```{r} 
#| label: age_structure_sigma1_women_pyramid_plot
#| fig.cap: "Population pyramids across pre-peak female union spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_spread_pre_peak_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_spread_pre_peak_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_women_median_plot
#| fig.cap: "Median age of surviving populations across pre-peak female union spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_spread_pre_peak_women"
)
```

```{r} 
#| label: age_structure_sigma1_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across pre-peak female union spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_spread_pre_peak_women")
    )
)
```

Cohen's d for median age across pre-peak female union spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_pre_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_pre_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader pre-peak female union spread is associated with slightly younger populations under matrilocal-matrilineal residence (\(d=-0.17\)), but has no detectable effect under patrilocal-patrilineal residence. Its influence is therefore weak and strongly conditional on residence organisation.

#### `sigma2_women`: broad *versus* narrow post-peak female union spread regimes

```{r} 
#| label: age_structure_sigma2_women_pyramid_plot
#| fig.cap: "Population pyramids across post-peak female union spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_spread_post_peak_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_spread_post_peak_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_women_median_plot
#| fig.cap: "Median age of surviving populations across post-peak female union spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_spread_post_peak_women"
)
```

```{r} 
#| label: age_structure_sigma2_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across post-peak female union spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_spread_post_peak_women")
    )
)
```

Cohen's d for median age across post-peak female union spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_post_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_post_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Post-peak female union spread has little influence on age structure. The effect is effectively absent under matrilocal-matrilineal residence, while broader schedules are associated with only a very small reduction in median age under patrilocal-patrilineal residence (\(d=-0.10\)). Results therefore provide no evidence that extending female union formation into later ages systematically ages surviving populations.

#### `c1_men`: high *versus* low male union peak intensity regimes

```{r} 
#| label: age_structure_c1_men_pyramid_plot
#| fig.cap: "Population pyramids across male union peak intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_intensity_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_intensity_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_men_median_plot
#| fig.cap: "Median age of surviving populations across male union peak intensity regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_intensity_men"
)
```

```{r} 
#| label: age_structure_c1_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male union peak intensity regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_intensity_men")
    )
)
```

Cohen's d for median age across male union peak intensity regimes, by residence rule:

```{r} 
#| label: age_structure_c1_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_intensity_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_intensity_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

Higher male union peak intensity is associated with slightly younger populations under both residence systems, but the effect is very small (\(d=-0.12\) and \(-0.11\)). Its structural influence is therefore modest and similar in magnitude to that of female union peak intensity.

#### `mu_men`: late *versus* early male union regimes

```{r} 
#| label: age_structure_mu_men_pyramid_plot
#| fig.cap: "Population pyramids across male union timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_timing_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_timing_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_men_median_plot
#| fig.cap: "Median age of surviving populations across male union timing regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_timing_men"
)
```

```{r} 
#| label: age_structure_mu_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male union timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_timing_men")
    )
)
```

Cohen's d for median age across male union timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_timing_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_timing_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Earlier male union timing produces younger populations under both residence systems, with small effects of \(d=-0.41\) under matrilocal-matrilineal and \(d=-0.38\) under patrilocal-patrilineal residence. The effect is therefore marginally stronger under matrilocal-matrilineal residence, although the difference between systems is small.

#### `sigma1_men`: broad *versus* narrow male pre-peak union spread regimes

```{r} 
#| label: age_structure_sigma1_men_pyramid_plot
#| fig.cap: "Population pyramids across male pre-peak union spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_spread_pre_peak_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_spread_pre_peak_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_men_median_plot
#| fig.cap: "Median age of surviving populations across male pre-peak union spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_spread_pre_peak_men"
)
```

```{r} 
#| label: age_structure_sigma1_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male pre-peak union spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_spread_pre_peak_men")
    )
)
```

Cohen's d for median age across male pre-peak union spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_pre_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_pre_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader pre-peak male union spread is associated with younger surviving populations under both residence systems. The effect is very small under matrilocal-matrilineal residence (\(d=-0.19\)) but substantially stronger under patrilocal-patrilineal residence (\(d=-0.38\)). This is one of the clearer residence-dependent union-formation effects in v1.3.

#### `sigma2_men`: broad *versus* narrow post-peak male union spread regimes

```{r} 
#| label: age_structure_sigma2_men_pyramid_plot
#| fig.cap: "Population pyramids across post-peak male union spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_spread_post_peak_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_spread_post_peak_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_men_median_plot
#| fig.cap: "Median age of surviving populations across post-peak male union spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_spread_post_peak_men"
)
```

```{r} 
#| label: age_structure_sigma2_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across post-peak male union spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_spread_post_peak_men")
    )
)
```

Cohen's d for median age across post-peak male union spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_post_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_post_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader post-peak male union spread is associated with marginally older populations under both residence systems, but the effect is extremely small (\(d\approx0.06\)–0.08) and the matrilocal-matrilineal confidence interval includes zero. No strong demographic interpretation is therefore warranted.

#### `amenorrhea_period_in_days`: long *versus* short amenorrhea period

```{r} 
#| label: age_structure_amenorrhea_pyramid_plot
#| fig.cap: "Population pyramids across amenorrhea period lengths, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("amenorrhea_length"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "amenorrhea_length",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_amenorrhea_median_plot
#| fig.cap: "Median age of surviving populations across amenorrhea period lengths, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "amenorrhea_length"
)
```

```{r} 
#| label: age_structure_amenorrhea_indicators_summary
#| tab.cap: "Summary of age structure indicators across amenorrhea period lengths and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "amenorrhea_length")
    )
)
```

Cohen's d for median age across amenorrhea period lengths, by residence rule:

```{r} 
#| label: age_structure_amenorrhea_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ amenorrhea_length,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_amenorrhea_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ amenorrhea_length,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Longer postpartum amenorrhoea is associated with slightly older surviving populations under both residence systems (\(d=0.18\) and \(0.16\)). The effect is very small but consistent with longer birth intervals reducing the rate of juvenile recruitment. Amenorrhoea therefore acts as a secondary regulator of demographic tempo rather than a dominant structural control.

#### Interpretation

Age structure in v1.3 remains dominated by reproductive recruitment rather than mortality. Fertility peak intensity produces the largest structural effect, while fertility spread, male union timing and pre-peak union spread, and several female union parameters provide weaker secondary effects. Postpartum amenorrhoea adds a new, explicitly temporal influence by modulating birth spacing.

Baseline mortality has comparatively little direct effect on median age among survivors, despite remaining relevant to survival and population growth. Daily scheduling therefore separates the roles of demographic mechanisms more clearly: fertility and union formation shape cohort renewal, amenorrhoea regulates reproductive pacing, and mortality primarily influences whether populations persist and how rapidly they expand.

Relative to v1.1, surviving populations are generally older and growth is slower, but the principal fertility-driven architecture of age composition remains intact.

## Mortality intensity sweep experiment

Simulations using the same parameter configuration, except `residence-rule`, and varying `cdmlt_level` and `SEED`.

```{r}
#| label: trajectories_cdmlt_level_load
rm(sampled_trajectories, preprocessing_trajectories_symmetric, sampled_trajectories_c1_fert)
sampled_trajectories_cdmlt_level <- readRDS("data/v13_trajectories_cdmlt_level.RData")
```

```{r} 
#| label: trajectories_params_cdmlt_level_display
#| fig.cap: "Parameter configuration for the population trajectories sampled across different Coale-Demeny model levels."
knitr::kable(sampled_trajectories_cdmlt_level |> 
               mutate(
                 max_iterations_in_days = max_iterations * year_length_in_days,
                 max_iterations_in_days_cut = max(step)
               ) |>
               select(all_of(c("max_iterations",
                               "max_iterations_in_days",
                               "max_iterations_in_days_cut", # show this in case loading was partial
                               "max_population", 
                               names(parameter_ranges_v13), 
                               "coale_demeny_region", 
                               "maturity_age_in_years", "gestation_period_in_days", "year_length_in_days" )),
                      -c("cdmlt_level")) |>
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

```{r} 
#| label: trajectories_cdmlt_level
#| fig.cap: "Population trajectories across different Coale-Demeny model levels, faceted by residence rule."
log_max_population <- ceiling(max(sampled_trajectories_cdmlt_level$log_totalIndividuals))

pop_trajectories_plot <- sampled_trajectories_cdmlt_level |>
  mutate(
    cdmlt_level = factor(cdmlt_level)
  ) |>
  plot_trajectories(
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = "cdmlt_level",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "max_population",
    thresholds_linetype = 2
  )
pop_trajectories_plot
```

```{r}
#| label: trajectories_cdmlt_level_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_plot,
    "v13_pop_trajectories_cdmlt_level"
  )
}
rm(pop_trajectories_plot)
```

The mortality sweep does not reveal a viability threshold within the range examined. Populations under both residence systems grow across all tested Coale-Demeny levels, and the median trajectories eventually approach or reach `max_population`.

Mortality intensity instead controls growth tempo. Higher Coale-Demeny levels, corresponding to lower mortality, produce more rapid population expansion and reach the technical ceiling substantially earlier. At the lowest levels, growth is much slower, particularly under patrilocal-patrilineal residence, but remains positive.

The targeted experiment therefore complements the global sensitivity analysis: mortality remains an important secondary determinant of demographic performance, but favourable fertility and union parameters can compensate for high mortality within this particular configuration. No universal mortality threshold should be inferred from these trajectories.

## Fertility peak intensity sweep experiment

Simulations using the same parameter configuration, except `residence-rule`, and varying `c1_fert` and `SEED`.

```{r}
#| label: trajectories_c1_fert_load
rm(sampled_trajectories, preprocessing_trajectories_symmetric, sampled_trajectories_c1_fert)
sampled_trajectories_c1_fert <- readRDS("data/v13_trajectories_c1_fert.RData")
```

```{r} 
#| label: trajectories_params_c1_fert_display
#| fig.cap: "Parameter configuration for the population trajectories sampled across different fertility peak intensity levels."
knitr::kable(sampled_trajectories_c1_fert |> 
               mutate(
                 max_iterations_in_days = max_iterations * year_length_in_days,
                 max_iterations_in_days_cut = max(step)
               ) |>
               select(all_of(c("max_iterations",
                               "max_iterations_in_days",
                               "max_iterations_in_days_cut", # show this in case loading was partial
                               "max_population", 
                               names(parameter_ranges_v13), 
                               "coale_demeny_region", 
                               "maturity_age_in_years", "gestation_period_in_days", "year_length_in_days" )),
                      -c("c1_fert")) |>
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

```{r} 
#| label: trajectories_c1_fert
#| fig.cap: "Population trajectories across different fertility peak intensity levels, faceted by residence rule."
log_max_population <- ceiling(max(sampled_trajectories_c1_fert$log_totalIndividuals))

pop_trajectories_plot <- sampled_trajectories_c1_fert |>
  mutate(
    c1_fert = factor(c1_fert)
  ) |>
  plot_trajectories(
    y_var = "log_totalIndividuals",
    trajectory_id = "run_unique_id",
    line_alpha = 0.035,
    median_alpha = 0.8,
    median_linewidth = 1,
    y_axis_log10 = TRUE,
    y_axis_label = "number of individuals",
    group_var = "c1_fert",
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "free_x",
    y_threshold_var = "max_population",
    thresholds_linetype = 2
  )
pop_trajectories_plot
```

```{r}
#| label: trajectories_c1_fert_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_plot,
    "v13_pop_trajectories_c1_fert"
  )
}
rm(pop_trajectories_plot)
```

The targeted fertility sweep shows a strong monotonic effect on demographic growth rate but no extinction threshold within the range examined. Populations grow under every tested `c1_fert` value from 0.3 to 0.9 in both residence systems.

Higher fertility peak intensity substantially shortens the time required to reach `max_population`. At intermediate and high values, populations reach the technical ceiling increasingly rapidly, whereas at `c1_fert = 0.3` growth remains positive but is too slow to reach the ceiling within the 500-year simulation horizon.

The experiment therefore clarifies the dominant random-forest importance of `c1_fert`: fertility peak intensity regulates replacement and growth tempo continuously rather than imposing a universal critical value separating survival from extinction. Whether low fertility produces extinction in the broader Sobol exploration depends on its interaction with mortality, union formation, amenorrhoea, initial conditions, and stochastic history.

Differences between residence rules are modest compared with the effect of fertility itself.

## Discussion notes

### Same demographic foundations, altered quantitative hierarchy

Version 1.3 preserves the density-independent demographic architecture of v1.1 while replacing annual event scheduling with daily demographic processes and explicit gestation and postpartum amenorrhoea. Fertility peak intensity remains the dominant predictor of survival, final population size, and median age, demonstrating strong continuity in the model's underlying demographic logic.

Mortality remains an important secondary predictor of survival and final size, but it no longer approaches fertility peak intensity in importance and has little direct effect on median age. The daily model therefore preserves the main causal architecture of v1.1 while changing its quantitative expression rather than simply reproducing the same parameter hierarchy.

### Growth becomes slower and persistence more common

Daily reproductive scheduling reduces demographic tempo. Growth remains the most common outcome among surviving populations, but Persistence occupies a noticeably larger share than in v1.1, and overall extinction is more frequent. Explicit gestation and postpartum amenorrhoea delay the conversion of reproductive opportunity into births, so viable populations take longer to accumulate demographic gains.

The targeted mortality and fertility sweeps reinforce this interpretation. Neither displays a sharp viability threshold under favourable conditions; instead, both parameters primarily alter the time required to approach the technical population ceiling.

### Age structure becomes older, but Growth-Persistence separation does not strengthen

Surviving populations are systematically older in v1.3 than in v1.1. Growth populations have a mean median age of about 21.5 years and persistent populations about 30.8 years, compared with approximately 18.6 and 29.6 years in the annual baseline. The Persistence-Growth contrast nevertheless becomes slightly weaker in standardised terms (\(d=1.73\) versus \(1.92\)), not stronger.

Daily scheduling therefore shifts demographic composition towards older ages while preserving the strong association between rapid recruitment and youthful age structure.

### Residence-rule differences remain secondary

Aggregate residence-rule effects remain small. Matrilocal-matrilineal populations are marginally older on average and show a slightly larger Persistence-Growth age contrast, while patrilocal-patrilineal populations grow somewhat faster in the targeted trajectory experiments. These differences are conditional and modest compared with the influence of fertility and reproductive scheduling.

The clearest residence effects continue to occur through sex-specific union parameters rather than through a general survival or age-structure advantage.

### Explicit reproductive delays add information without replacing the baseline

Postpartum amenorrhoea provides the clearest genuinely new demographic control introduced by v1.3. Longer amenorrhoea modestly increases median age by extending birth intervals, but its influence remains secondary to fertility peak intensity and fertility spread.

The daily formulation therefore demonstrates that temporal resolution is a substantive modelling choice. It changes reproductive tempo, cohort structure, and the time required for population expansion while preserving the broad fertility-driven demographic logic of the annual baseline.

## Chapter conclusion

Version 1.3 demonstrates that introducing daily demographic scheduling changes the quantitative expression of the Household Demography model without replacing its underlying demographic architecture. Fertility peak intensity remains the principal determinant of survival, population growth, and age composition, while explicit gestation and postpartum amenorrhoea slow reproductive turnover, increase the frequency of persistent outcomes, and produce generally older surviving populations.

Daily iteration therefore contributes more than finer numerical resolution: it makes reproductive delays and birth spacing explicit and changes the tempo through which demographic outcomes emerge. At the same time, the persistence of the baseline fertility-driven hierarchy provides a clear point of continuity with v1.1 and establishes the appropriate daily baseline for the density-regulated v1.4 series.
