---
title: "Household Demography model: sensitivity analysis on density-dependent (v.1.2)"
output: html_notebook
---

## Preparation

```{r, message=FALSE}
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(gridExtra)
library(scales)
library(randomForest)
```

```{r}
plotDirectory <- "plots/v12_"
plot_file_format <- c("png", "eps")[1] # modify index number to change format
```

## Experiment design with stochastic parameter settings

### Generate parameter values with Sobol Sampling

```{r}
sobol_parameter_sample_dataFileName <- "data/v12_sobol_parameter_values.RData"

sobol_parameter_ranges <- list(
  `initial-num-households` = c(2, 50),         
  `household_initial_age_distribution_min` = c(0, 5),        
  `household_initial_age_distribution_max` = c(6, 60),
  `max_couple_count_distribution_min` = c(1, 2),
  `max_couple_count_distribution_max` = c(3, 8),
  `acceptable-kinship-degree-for-couples` = c(0, 9),
  `cdmlt-level` = c(1, 25),
  `c1-fert` = c(0, 0.5),
  `mu-fert` = c(15, 30),
  `sigma1-fert` = c(1, 10),
  `sigma2-fert` = c(5, 20), 
  `c1-women` = c(0, 1),
  `mu-women` = c(15, 40),
  `sigma1-women` = c(1, 20),
  `sigma2-women` = c(1, 20), 
  `c1-men` = c(0, 1),
  `mu-men` = c(15, 40), 
  `sigma1-men` = c(1, 20), 
  `sigma2-men` = c(1, 20),
  `carrying-capacity` = c(100, 10000),
  `density-effect-scaling-factor` = c(0.01, 0.5)
)
```

```{r, eval=FALSE}
library(randtoolbox)
set.seed(123)
n <- 2000  # Number of samples
k <- length(sobol_parameter_ranges)  # Number of parameters

# Generate Sobol sequence
sobol_sample_matrix <- sobol(n, dim = k, scrambling = 3)

get_values_from_sobol <- function(param_index){
  qunif(sobol_sample_matrix[,param_index], 
        min = sobol_parameter_ranges[[param_index]][1], 
        max = sobol_parameter_ranges[[param_index]][2])
}

```

```{r, eval=FALSE}
# Transform to specific distributions
params_sobol <- data.frame(
  `initial-num-households` = round(get_values_from_sobol(1)),
  `household_initial_age_distribution_min` = round(get_values_from_sobol(2)),
  `household_initial_age_distribution_max` = round(get_values_from_sobol(3)),
  `max_couple_count_distribution_min` = round(get_values_from_sobol(4)),
  `max_couple_count_distribution_max` = round(get_values_from_sobol(5)),
  `acceptable-kinship-degree-for-couples` = round(get_values_from_sobol(6)),
  `cdmlt-level` = round(get_values_from_sobol(7)),
      
  `c1-fert` = get_values_from_sobol(8),
  `mu-fert` = get_values_from_sobol(9),
  `sigma1-fert` = get_values_from_sobol(10),
  `sigma2-fert` = get_values_from_sobol(11),
      
  `c1-women` = get_values_from_sobol(12),
  `mu-women` = get_values_from_sobol(13),
  `sigma1-women` = get_values_from_sobol(14),
  `sigma2-women` = get_values_from_sobol(15),
  `c1-men` = get_values_from_sobol(16),
  `mu-men` = get_values_from_sobol(17),
  `sigma1-men` = get_values_from_sobol(18),
  `sigma2-men` = get_values_from_sobol(19),
  `carrying-capacity` = get_values_from_sobol(20),
  `density-effect-scaling-factor` = get_values_from_sobol(21)
)

head(params_sobol)
```

```{r, eval=FALSE}
save(params_sobol, file = sobol_parameter_sample_dataFileName)
```

```{r, eval=FALSE}
load(sobol_parameter_sample_dataFileName)
```

Visualising parameter space with `pairs()`:

```{r, eval=FALSE}
plotName = paste0(plotDirectory, "sobol_param_pairsPlot.png")

plotScale = 10

png(plotName,
    width = 300 * plotScale, height = 300 * plotScale)
pairs(params_sobol, pch='·')
dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

Saving parameter configurations (experiments) in separate files:

```{r, eval=FALSE}
save_experiment <- function(experiment_df, experiment_row, version_name) {
  write(paste(unlist(experiment_df[experiment_row,], use.names = FALSE)),
        file = paste0(dirname(getwd()), "/analysis/experiments/", version_name, "/", experiment_row, ".txt"))
}

for (i in 1:nrow(params_sobol)) {
  save_experiment(params_sobol, i, "v1.2")
}
```

## Analysis on population end states

>Run experiments in NetLogo Behaviour Space using the parameters `type-of-experiment` and `exp-number`.

### Data loading and preprocessing

Define usable reference to parameter names:

```{r}
param_names_netlogo <- names(sobol_parameter_ranges)
param_names_r <- gsub("-", ".", names(sobol_parameter_ranges))
```

Loading simulation experiments' end states:

```{r}
endstates_matri <- read.csv("simdata/v1.2/household-demography-main_density exp-endstates-matri-table.csv", skip = 6)
endstates_patri <- read.csv("simdata/v1.2/household-demography-main_density exp-endstates-patri-table.csv", skip = 6)
```

Preprocess dataset adapting from NetLogo'S specific format. Categorise end states according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

```{r}
MAX_POPULATION = endstates_matri$max.population[1]

extract_values_from_NetLogo_list <- function(char_vector) {
  # Remove brackets and split each string into two parts
  split_values <- gsub("\\[|\\]", "", char_vector) |>
    strsplit(" ")

  # Convert to numeric vectors
  first_values <- sapply(split_values, function(x) as.numeric(x[1]))
  second_values <- sapply(split_values, function(x) as.numeric(x[2]))

  # Return as a named list
  list(
    first = first_values,
    second = second_values
  )
}

set_survival_category <- function(population_count, carrying_capacity) {
  survival_category <- ""
  if (population_count > 0) {
    survival_category <- "survival"
    if (population_count > carrying_capacity) {
      survival_category <- paste0(survival_category, " with overshot")
    }
  } else {
    survival_category <- "extinction"
  }
  survival_category
}

preprocess_endstates <- function(endstate_df) {
  household_initial_age_distribution_range <- extract_values_from_NetLogo_list(endstate_df$household.initial.age.distribution)
  max_couple_count_distribution_range <- extract_values_from_NetLogo_list(endstate_df$max.couple.count.distribution)
  
  endstates <- cbind(
    endstate_df,
    household_initial_age_distribution_min = household_initial_age_distribution_range[[1]],
    household_initial_age_distribution_max = household_initial_age_distribution_range[[2]],
    max_couple_count_distribution_min = max_couple_count_distribution_range[[1]],
    max_couple_count_distribution_max = max_couple_count_distribution_range[[2]]
  )
  
  endstates |>
    rowwise() |>
    mutate(survival = set_survival_category(totalIndividuals, carrying.capacity)) |>
    ungroup()
}

endstates_matri <- preprocess_endstates(endstates_matri)
endstates_patri <- preprocess_endstates(endstates_patri)
```

### Overview of simulation results (population size)

Visualise `totalIndividuals`:

```{r}
layout(matrix(1:2, ncol = 2))
hist(endstates_matri$totalIndividuals, breaks = 50, main = "residential-rule = 'matrilocal-matrilineal'")
hist(endstates_patri$totalIndividuals, breaks = 50, main = "residential-rule = 'patrilocal-patrilineal'")
```

```{r}
table(endstates_matri$survival)
table(endstates_patri$survival)
```

```{r}
layout(matrix(1:2, ncol = 2))
barplot(table(endstates_matri$survival), main = "residential-rule = 'matrilocal-matrilineal'")
barplot(table(endstates_patri$survival), main = "residential-rule = 'patrilocal-patrilineal'")
```

Visualise `totalIndividuals` in surviving configurations:
```{r}
layout(matrix(1:2, ncol = 2))
hist(endstates_matri$totalIndividuals[endstates_matri$survival != "extinction"], breaks = 50, main = "residential-rule = 'matrilocal-matrilineal'")
hist(endstates_patri$totalIndividuals[endstates_patri$survival != "extinction"], breaks = 50, main = "residential-rule = 'patrilocal-patrilineal'")
```

### Visualising population age structure

```{r}
extract_age_structure <- function(char_value) {
  split_values <- gsub("\\[|\\]", "", char_value) |>
    strsplit(" ")
  as.numeric(split_values[[1]])
}

build_age_structure_df <- function(endstate_df, selected_case = 500, max_age = 110) {
  survival_cases <- endstate_df %>% filter(survival != "extinction")
  
  age_structure_men <- extract_age_structure(survival_cases[selected_case, "menAgeStructure"])
  age_structure_women <- extract_age_structure(survival_cases[selected_case, "womenAgeStructure"])
  
  # Put data into long format
  age_structure_df <- tibble(
    age = c(age_structure_men, age_structure_women),
    sex = c(rep("Male", length(age_structure_men)), 
            rep("Female", length(age_structure_women)))
  )
  
  # Bin ages into groups (5-year bins, adjust as needed)
  age_structure_df %>%
    mutate(age_group = cut(age, breaks = seq(0, max_age, 5), right = FALSE))
}

summarise_age_structure_df <- function(age_structure_df) {
  age_structure_df %>%
    count(sex, age_group) %>% # Summarize counts per age group and sex
    mutate(count = ifelse(sex == "Male", -n, n)) # For pyramid: make men negative
}
```

```{r}
age_structure_matri <- build_age_structure_df(endstates_matri)
age_structure_patri <- build_age_structure_df(endstates_patri)
```

```{r}
max_age = max(age_structure_matri$age, age_structure_patri$age)
plot_1 <- ggplot(age_structure_matri, aes(x = age, fill = sex)) +
  geom_histogram() +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  lims(x = c(0, max_age + 5)) +
  guides(fill=FALSE) +
  labs(
    title = "Population Age Pyramid \n residential-rule = 'matrilocal-matrilineal'",
    x = "Age",
    y = "Population count"
  ) +
  theme_minimal()

plot_2 <- ggplot(age_structure_patri, aes(x = age, fill = sex)) +
  geom_histogram() +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  lims(x = c(0, max_age + 5)) +
  labs(
    title = "Population Age Pyramid \n residential-rule = 'patrilocal-patrilineal'",
    x = "Age",
    y = "Population count"
  ) +
  theme_minimal()
# 
# layout(matrix(1:4, nrow = 2, byrow = TRUE))
# hist(age_structure_matri[sex == "women",], breaks = 101, main = "")
# hist(age_structure_matri_men, breaks = 101, main = "")
# hist(age_structure_patri_women, breaks = 101, main = "")
# hist(age_structure_patri_men, breaks = 101, main = "")
grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1.2))
```

```{r}
age_structure_matri_summary <- summarise_age_structure_df(age_structure_matri)
age_structure_patri_summary <- summarise_age_structure_df(age_structure_patri)
```

```{r}
plot_1 <- ggplot(age_structure_matri_summary, aes(x = age_group, y = count, fill = sex)) +
  geom_col(width = 1, color = "white") +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  guides(fill=FALSE) +
  labs(
    title = "Population Age Pyramid \n residential-rule = 'matrilocal-matrilineal'",
    x = "Age group",
    y = "Population count"
  ) +
  theme_minimal()

plot_2 <- ggplot(age_structure_patri_summary, aes(x = age_group, y = count, fill = sex)) +
  geom_col(width = 1, color = "white") +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  labs(
    title = "Population Age Pyramid \n residential-rule = 'patrilocal-patrilineal'",
    x = "Age group",
    y = "Population count"
  ) +
  theme_minimal()

grid.arrange(grobs = c(plot_1, plot_2), ncol = 2, widths = c(1, 1.2))
```

### Identifying survival parameter space

Calculate the Peristera-Kostaki parameter ranges associated to survival:
```{r}
compute_param_envelope <- function(param_df, low=0.025, high=0.975){
  apply(param_df, 2, quantile, probs = c(low, high), na.rm=TRUE) %>%
    t() %>% as.data.frame() #%>% rename(lower = `2.5%`, upper = `97.5%`)
}

compute_survival_param_space <- function(endstate_df, param_names) {
  compute_param_envelope(endstate_df %>% filter(survival != "extinction") %>% select(all_of(param_names)))
}

param_names_pk <- param_names_r[7:19]

survival_param_space_matri <- compute_survival_param_space(endstates_matri, param_names_pk)
survival_param_space_patri <- compute_survival_param_space(endstates_patri, param_names_pk)
```

With residential-rule = 'matrilocal-matrilineal':
```{r}
knitr::kable(survival_param_space_matri)
```

With residential-rule = 'patrilocal-patrilineal':
```{r}
knitr::kable(survival_param_space_patri)
```

Identify the ranges for the version-specific parameters:

```{r}
param_names_v12 <- param_names_r[20:21]

survival_param_space_matri_v12 <- compute_survival_param_space(endstates_matri, param_names_v12)
survival_param_space_patri_v12 <- compute_survival_param_space(endstates_patri, param_names_v12)
```

With residential-rule = 'matrilocal-matrilineal':
```{r}
knitr::kable(survival_param_space_matri_v12)
```

With residential-rule = 'patrilocal-patrilineal':
```{r}
knitr::kable(survival_param_space_patri_v12)
```

### Identifying relevant factors of survival

```{r}
param_names <- c("initial.num.households", 
                 "household_initial_age_distribution_min",
                 "household_initial_age_distribution_max",
                 "max_couple_count_distribution_min",
                 "max_couple_count_distribution_max",
                 param_names_pk,
                 param_names_v12)

covert_to_long_by_param <- function(df, param_names) {
  df %>%
    pivot_longer(cols = all_of(param_names)) %>%
    mutate(name = factor(name, levels = param_names))  # Preserve order
}
endstates_matri_long <- covert_to_long_by_param(endstates_matri, param_names)
endstates_patri_long <- covert_to_long_by_param(endstates_patri, param_names)
```

```{r}
source("library/bifurcationPlot.R")

create_bifurcation_plot_multiple <- function(df_long, plot_name, plot_scale = 10, num_columns = 3, width = 80, height = 150, strip_text_size = 1.2) {
  png(plot_name,
      width = width * plot_scale, height = height * plot_scale)
  
  ggplot(data = df_long, aes(x = value, y = totalIndividuals)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "glm") +
    facet_wrap(vars(name), ncol = num_columns, scales = "free_x", as.table = TRUE) +
    theme(
      strip.text = element_text(size = strip_text_size * plot_scale)
    )
  #dev.off()
}

plotName_matri = paste0(plotDirectory, "endstates_matri_bifurcationPlot.png")
create_bifurcation_plot_multiple(endstates_matri_long, plot_name = plotName_matri)

plotName_patri = paste0(plotDirectory, "endstates_patri_bifurcationPlot.png")
create_bifurcation_plot_multiple(endstates_patri_long, plot_name = plotName_patri)
```

```{r}
knitr::include_graphics(plotName_matri)
```

```{r}
knitr::include_graphics(plotName_patri)
```

Train a Random Forest model to assess parameter importance for each dataset:

```{r}
run_rf_for_totalIndividuals <- function(endstates_df, param_names, seed = 123) {
  endstates_df <- endstates_df %>% select(all_of(c(param_names, "totalIndividuals")))
  
  # Train Random Forest to predict simulation output
  set.seed(seed)
  randomForest(totalIndividuals ~ ., data = endstates_df, importance = TRUE, ntree = 500)
}
```

```{r}
rf_matri <- run_rf_for_totalIndividuals(endstates_matri, param_names)
print(rf_matri)
```

```{r}
rf_patri <- run_rf_for_totalIndividuals(endstates_patri, param_names)
print(rf_patri)
```

Extract importance values and print as table:

```{r}
extract_rf_param_importance <- function(rf_model) {
  # Extract importance values
  importance_values <- importance(rf_model)
  
  # Convert to a data frame
  data.frame(Parameter = rownames(importance_values),
             MDA = importance_values[, 1], 
             MDG = importance_values[, 2])
}
```

```{r}
rf_matri_importance <- extract_rf_param_importance(rf_matri)
print(rf_matri_importance)
```

```{r}
rf_patri_importance <- extract_rf_param_importance(rf_patri)
print(rf_patri_importance)
```

Plot feature importance colouring parameters by domain (general, mortality, fertility, nuptiality):

```{r}
source("library/variableImportancePlot.R")

plotName = paste0(plotDirectory, "variableImportancePlot.png")

color_variable_map <- rev(c(
  rep("#0000a2", 5),
  "#bc272d", 
  rep("#db9b35", 4), 
  rep("#50ad9f", 8),
  rep("#7953b8", 2)
))

# png
png(filename = plotName, width = 2 * 129, height = 168, units="mm", res=1200)

variableImportancePlotPair(rf_matri, rf_patri,
                           title_extra = c("residential-rule = 'matrilocal-matrilineal'", "residential-rule = 'patrilocal-patrilineal'"),
                           plot_scale = 0.5, 
                           color_variable_map = color_variable_map, 
                           range_standardised = TRUE, 
                           log_scale = FALSE,
                           layout_type = 'horizontal')

# layout(matrix(1:2, ncol = 2))
# 
# op <- par(no.readonly = TRUE)
# 
# par(mar = c(5, 4, 4, 2) + 0.1)
# variableImportancePlot(rf_matri, title_extra = "residential-rule = 'matrilocal-matrilineal'", plot_scale = 0.5, color_variable_map = color_variable_map, range_standardised = TRUE)
# 
# par(mar = c(5, 4, 4, 2) + 0.1)
# variableImportancePlot(rf_patri, title_extra = "residential-rule = 'patrilocal-patrilineal'", plot_scale = 0.5, color_variable_map = color_variable_map, range_standardised = TRUE)

dev.off()
```

```{r}
knitr::include_graphics(plotName)
```

#### Initial household population

This parameter has very small positive effect on population survival:

```{r}
cat("matri | extinct |")
endstates_matri %>%
  filter(totalIndividuals == 0) %>%
  select(initial.num.households) %>%
  summary()
cat("matri | survive |")
endstates_matri %>%
  filter(totalIndividuals > 0) %>%
  select(initial.num.households) %>%
  summary()
cat("patri | extinct |")
endstates_patri %>%
  filter(totalIndividuals == 0) %>%
  select(initial.num.households) %>%
  summary()
cat("patri | survive |")
endstates_patri %>%
  filter(totalIndividuals > 0) %>%
  select(initial.num.households) %>%
  summary()
```

```{r}
layout(matrix(1:2, ncol = 2))
boxplot(endstates_matri$initial.num.households ~ (endstates_matri$totalIndividuals > 0))
boxplot(endstates_patri$initial.num.households ~ (endstates_patri$totalIndividuals > 0))
```

#### Mortality factors

Coale-Demeny Life Tables Model level parameter has great effect:

```{r}
cat("matri | extinct |")
endstates_matri %>%
  filter(totalIndividuals == 0) %>%
  select(cdmlt.level) %>%
  summary()
cat("matri | survive |")
endstates_matri %>%
  filter(totalIndividuals > 0) %>%
  select(cdmlt.level) %>%
  summary()
cat("patri | extinct |")
endstates_patri %>%
  filter(totalIndividuals == 0) %>%
  select(cdmlt.level) %>%
  summary()
cat("patri | survive |")
endstates_patri %>%
  filter(totalIndividuals > 0) %>%
  select(cdmlt.level) %>%
  summary()
```

```{r}
layout(matrix(1:2, ncol = 2))
boxplot(endstates_matri$cdmlt.level ~ (endstates_matri$totalIndividuals > 0))
boxplot(endstates_patri$cdmlt.level ~ (endstates_patri$totalIndividuals > 0))
```

but region parameter has a smaller effect. East tends to survive more often.  

```{r}
cat("matri | survive? |")
table(endstates_matri$coale.demeny.region, endstates_matri$totalIndividuals > 0)
cat("patri | survive? |")
table(endstates_patri$coale.demeny.region, endstates_patri$totalIndividuals > 0)
```

#### Fertility factors

All have a visible effect, but particularly c1.fert and sigma1.fert. 

```{r}
layout(matrix(1:4, nrow = 1, byrow = F))
boxplot(endstates_matri$c1.fert ~ (endstates_matri$totalIndividuals > 0), main = "matri - c1.fert")
boxplot(endstates_matri$mu.fert ~ (endstates_matri$totalIndividuals > 0), main = "mu.fert")
boxplot(endstates_matri$sigma1.fert ~ (endstates_matri$totalIndividuals > 0), main = "sigma1.fert")
boxplot(endstates_matri$sigma2.fert ~ (endstates_matri$totalIndividuals > 0), main = "sigma2.fert")
```

```{r}
layout(matrix(1:4, nrow = 1, byrow = F))
boxplot(endstates_patri$c1.fert ~ (endstates_patri$totalIndividuals > 0), main = "patri - c1.fert")
boxplot(endstates_patri$mu.fert ~ (endstates_patri$totalIndividuals > 0), main = "mu.fert")
boxplot(endstates_patri$sigma1.fert ~ (endstates_patri$totalIndividuals > 0), main = "sigma1.fert")
boxplot(endstates_patri$sigma2.fert ~ (endstates_patri$totalIndividuals > 0), main = "sigma2.fert")
```

#### Nuptiality factors

Residential rule showing no clear effect:

```{r}
cat("matri | survive? |")
table(endstates_matri$residence.rule, endstates_matri$totalIndividuals > 0)
cat("patri | survive? |")
table(endstates_patri$residence.rule, endstates_patri$totalIndividuals > 0)
```

The effect of the acceptable kinship degree is also unclear: 

```{r}
layout(matrix(1:2, ncol = 2))
boxplot(endstates_matri$acceptable.kinship.degree.for.couples ~ (endstates_matri$totalIndividuals > 0))
boxplot(endstates_patri$acceptable.kinship.degree.for.couples ~ (endstates_patri$totalIndividuals > 0))
```

All nuptiality parameters for both men and women present a small, yet apparently consistent effect. c1 and sigma parameters have a positive effect while mu has a smaller negative effect.

```{r}
layout(matrix(1:8, nrow = 2, byrow = F))
boxplot(endstates_matri$c1.women ~ (endstates_matri$totalIndividuals > 0), main = "matri- c1.women")
boxplot(endstates_matri$mu.women ~ (endstates_matri$totalIndividuals > 0), main = "mu.women")
boxplot(endstates_matri$sigma1.women ~ (endstates_matri$totalIndividuals > 0), main = "sigma1.women")
boxplot(endstates_matri$sigma2.women ~ (endstates_matri$totalIndividuals > 0), main = "sigma2.women")
boxplot(endstates_matri$c1.men ~ (endstates_matri$totalIndividuals > 0), main = "c1.men")
boxplot(endstates_matri$mu.men ~ (endstates_matri$totalIndividuals > 0), main = "mu.men")
boxplot(endstates_matri$sigma1.men ~ (endstates_matri$totalIndividuals > 0), main = "sigma1.men")
boxplot(endstates_matri$sigma2.men ~ (endstates_matri$totalIndividuals > 0), main = "sigma2.men")
```

```{r}
layout(matrix(1:8, nrow = 2, byrow = F))
boxplot(endstates_patri$c1.women ~ (endstates_patri$totalIndividuals > 0), main = "patri - c1.women")
boxplot(endstates_patri$mu.women ~ (endstates_patri$totalIndividuals > 0), main = "mu.women")
boxplot(endstates_patri$sigma1.women ~ (endstates_patri$totalIndividuals > 0), main = "sigma1.women")
boxplot(endstates_patri$sigma2.women ~ (endstates_patri$totalIndividuals > 0), main = "sigma2.women")
boxplot(endstates_patri$c1.men ~ (endstates_patri$totalIndividuals > 0), main = "c1.men")
boxplot(endstates_patri$mu.men ~ (endstates_patri$totalIndividuals > 0), main = "mu.men")
boxplot(endstates_patri$sigma1.men ~ (endstates_patri$totalIndividuals > 0), main = "sigma1.men")
boxplot(endstates_patri$sigma2.men ~ (endstates_patri$totalIndividuals > 0), main = "sigma2.men")
```

### Pace of growth

```{r}
layout(matrix(1:2, ncol = 2))
hist(endstates_matri$X.step.[endstates_matri$totalIndividuals > 0], main = "matri")
hist(endstates_patri$X.step.[endstates_patri$totalIndividuals > 0], main = "patri")
```

```{r}
endstates_matri_long_survived <- endstates_matri_long %>%
  filter(totalIndividuals > 0)
endstates_patri_long_survived <- endstates_patri_long %>%
  filter(totalIndividuals > 0)
```

```{r}
source("library/bifurcationPlot.R")

create_bifurcation_plot_multiple <- function(df_long, plot_name, plot_scale = 10, num_columns = 3, width = 80, height = 150, strip_text_size = 1.2) {
  png(plot_name,
      width = width * plot_scale, height = height * plot_scale)
  
  ggplot(data = df_long, aes(x = value, y = X.step.)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "glm") +
    facet_wrap(vars(name), ncol = num_columns, scales = "free_x", as.table = TRUE) +
    theme(
      strip.text = element_text(size = strip_text_size * plot_scale)
    )
  #dev.off()
}

plotName_matri = paste0(plotDirectory, "endstates_matri_bifurcationPlot_growth.png")
create_bifurcation_plot_multiple(endstates_matri_long_survived, plot_name = plotName_matri)

plotName_patri = paste0(plotDirectory, "endstates_patri_bifurcationPlot_growth.png")
create_bifurcation_plot_multiple(endstates_patri_long_survived, plot_name = plotName_patri)
```

```{r}
knitr::include_graphics(plotName_matri)
```

```{r}
knitr::include_graphics(plotName_patri)
```

## Stochasticity and trajectory variation

>Run experiments in NetLogo Behaviour Space using parameter configurations where populations tend to survive, according to the exploration above.

Loading simulation experiments' end states:

```{r}
trajectories_v11 <- read.csv("simdata/v1.1/household-demography-main exp-trajectories-table.csv", skip = 6)
trajectories_v122 <- read.csv("simdata/v1.2/household-demography-main_density exp-trajectories-table.csv", skip = 6)
```

Preprocess dataset adapting from NetLogo'S specific format. Categorise end states according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

```{r}
MAX_POPULATION = endstates_matri$max.population[1]

extract_values_from_NetLogo_list <- function(char_vector) {
  # Remove brackets and split each string into two parts
  split_values <- gsub("\\[|\\]", "", char_vector) |>
    strsplit(" ")

  # Convert to numeric vectors
  first_values <- sapply(split_values, function(x) as.numeric(x[1]))
  second_values <- sapply(split_values, function(x) as.numeric(x[2]))

  # Return as a named list
  list(
    first = first_values,
    second = second_values
  )
}

set_survival_category <- function(population_count) {
  survival_category <- ""
  if (population_count > 0) {
    survival_category <- "survival"
    if (population_count > MAX_POPULATION) {
      survival_category <- paste0(survival_category, " with growth")
    }
  } else {
    survival_category <- "extinction"
  }
  survival_category
}

preprocess_endstates <- function(trajectories_df) {
  household_initial_age_distribution_range <- extract_values_from_NetLogo_list(trajectories_df$household.initial.age.distribution)
  max_couple_count_distribution_range <- extract_values_from_NetLogo_list(endstate_df$max.couple.count.distribution)
  
  cbind(
    trajectories_df,
    household_initial_age_distribution_min = household_initial_age_distribution_range[[1]],
    household_initial_age_distribution_max = household_initial_age_distribution_range[[2]],
    max_couple_count_distribution_min = max_couple_count_distribution_range[[1]],
    max_couple_count_distribution_max = max_couple_count_distribution_range[[2]],
    survival = sapply(trajectories_df$totalIndividuals, set_survival_category)
  )
}

trajectories_matri <- preprocess_endstates(trajectories_matri)
trajectories_patri <- preprocess_endstates(trajectories_patri)
```

Process parameters in list format in NetLogo:

```{r}
extract_values_from_NetLogo_list <- function(char_vector) {
  # Remove brackets and split each string into two parts
  split_values <- gsub("\\[|\\]", "", char_vector) |>
    strsplit(" ")

  # Convert to numeric vectors
  first_values <- sapply(split_values, function(x) as.numeric(x[1]))
  second_values <- sapply(split_values, function(x) as.numeric(x[2]))

  # Return as a named list
  list(
    first = first_values,
    second = second_values
  )
}

household_initial_age_distribution_range_v11 <- extract_values_from_NetLogo_list(trajectories_v11$household.initial.age.distribution)
max_couple_count_distribution_range_v11 <- extract_values_from_NetLogo_list(trajectories_v11$max.couple.count.distribution)

household_initial_age_distribution_range_v122 <- extract_values_from_NetLogo_list(trajectories_v122$household.initial.age.distribution)
max_couple_count_distribution_range_v122 <- extract_values_from_NetLogo_list(trajectories_v122$max.couple.count.distribution)

trajectories_v11 <- cbind(
  trajectories_v11,
  household_initial_age_distribution_min = household_initial_age_distribution_range_v11[[1]],
  household_initial_age_distribution_max = household_initial_age_distribution_range_v11[[2]],
  max_couple_count_distribution_min = max_couple_count_distribution_range_v11[[1]],
  max_couple_count_distribution_max = max_couple_count_distribution_range_v11[[2]]
)

trajectories_v122 <- cbind(
  trajectories_v122,
  household_initial_age_distribution_min = household_initial_age_distribution_range_v122[[1]],
  household_initial_age_distribution_max = household_initial_age_distribution_range_v122[[2]],
  max_couple_count_distribution_min = max_couple_count_distribution_range_v122[[1]],
  max_couple_count_distribution_max = max_couple_count_distribution_range_v122[[2]]
)
```

```{r}
params_indexes <- c(3:21)
param_names <- names(trajectories_v11)[params_indexes]
params_v11 <- trajectories_v11 %>% 
  select(all_of(param_names)) %>% 
  distinct() %>% 
  t %>%
  apply(1, unique)
params_v11
```

```{r}
params_indexes <- c(3:21, 22:24)
param_names <- names(trajectories_v122)[params_indexes]
params_v122 <- trajectories_v122 %>% 
  select(all_of(param_names)) %>% 
  distinct() %>% 
  t %>%
  apply(1, unique)
params_v122
```

```{r}
plot_name <- file.path(plotDirectory, paste0("HouseholdDemography-trajectories_v11-v122.", plot_file_format))

if (plot_file_format == "png") {
  graphic_scale <- 1
  legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
  axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5

  png(plot_name, width = graphic_scale * 1600, height = graphic_scale * 800, res = 300, pointsize = 4)
} else if (plot_file_format == "eps") {
  graphic_scale = 1.2
  # font_rescale = -0.2
  # axis_text_rescale = -0.1
  # margin_text_rescale = -0.5

  extrafont::loadfonts(device = "postscript")
  grDevices::cairo_ps(filename = plot_name ,
                      pointsize = 12,
                      width = graphic_scale * 8,
                      height = graphic_scale * 4,
                      onefile = FALSE,
                      family = "sans"
                      )
}

p1 <- ggplot(trajectories_v11[trajectories_v11$X.step. < 600,], aes(x = X.step., y = totalIndividuals, group = X.run.number., color = cdmlt.level)) +
  geom_line(linewidth = .1) +
  title("v.1.1") +
  xlab("time (years)") +
  ylab("population (individuals)") +
  ylim(c(0, 7000)) +
  guides(color = "none") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10 * (graphic_scale + axis_title_rescale)),
        axis.text = element_text(size = 10 * (graphic_scale + axis_text_rescale)))

p2 <- ggplot(trajectories_v122[trajectories_v122$X.step. < 600,], aes(x = X.step., y = totalIndividuals, group = X.run.number., color = cdmlt.level)) +
  geom_line(linewidth = .1) +
  geom_hline(yintercept = as.numeric(params_v122$par_carrying.capacity)) +
  title("v.1.2.2") +
  xlab("time (years)") +
  ylab("") +
  ylim(c(0, 7000)) +
  guides(color = guide_legend(title="Coale-Demeny\nLife Table Level", override.aes = list(linewidth = 2), reverse = TRUE)) +
  theme_minimal() +
  theme(axis.title = element_text(size = 10 * (graphic_scale + axis_title_rescale)),
        axis.text = element_text(size = 10 * (graphic_scale + axis_text_rescale)),
        legend.title = element_text(size = 10 * (graphic_scale + legend_text_rescale)),
        legend.text = element_text(size = 10 * (graphic_scale + legend_text_rescale)),
        legend.key.width = unit(2 * (graphic_scale + legend_key_rescale), 'lines'),
        legend.key.height = unit(2 * (graphic_scale + legend_key_rescale), 'lines'))

p3 <- grid.arrange(p1, p2, ncol = 2, widths = c(4, 5.1))

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plot_name)
```

