---
title: "Household Demography model: sensitivity analysis on base model (v.1.1)"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
VERSION <- "v1.1"
SHOW_DIAGNOSTICS <- FALSE
EXPORT_FIGURES <- TRUE
RUN_RF_MODELS <- FALSE
```

## Introduction

### Setup

Load libraries, sources and set ggplot theme for the notebook.

```{r setup_libraries}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(randomForest)
library(pdp)
library(forcats)
library(patchwork)
library(scales)
```

```{r setup_load_sources}
source("library/parameter_ranges.R")
source("library/load_endstates.R")
source("library/load_trajectories.R")
source("library/preprocess_simdata.R")

source("library/export_figure.R")
source("library/bifurcationPlot.R")

source("library/sensitivity/parameter_metadata.R")
source("library/sensitivity/prepare_sensitivity_data.R")
source("library/sensitivity/load_or_build_rf_models.R")
source("library/sensitivity/extract_importance.R")
source("library/sensitivity/plot_importance.R")
source("library/sensitivity/load_or_build_pd_results.R")
source("library/sensitivity/plot_parameter_dependence.R")
```

```{r setup_ggplot-theme}
theme_set(
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold"),
      plot.title = element_text(face = "bold")
    )
)
```

### Purpose

This notebook analyses the behaviour of the baseline Household Demography model (v1.1) under Sobol-sampled parameter combinations. The analysis focuses on:

- overall demographic outcomes,
- persistence versus extinction dynamics,
- variation between residence systems,
- and parameter influence on population trajectories and end states.

## Experiment design

### Generate parameter values with Sobol Sampling

Sobol sampling is used to generate a diverse set of parameter combinations across the defined parameter space. This allows us to explore how different parameter configurations affect the model's outcomes and to identify which parameters are most influential in driving population survival or extinction. The actual sampling was done with `generate_sobol.R` script, which generates a sample with 2000 parameter combinations across the parameter ranges defined in `parameter_ranges.R`. The generated parameter combinations were then used to run simulations in NetLogo's Behaviour Space, for each combination of `residential-rule` and `cdmlt-region` (2 x 4), varying the `SEED` parameter (n = 3) to introduce stochasticity, for a total of 48000 simulations.

```{r sobol_load}
# Load object as Rdata file
load("data/v11_sobol_parameter_values.RData")
```

Sobol sampling is applied to the following parameter ranges:

```{r sobol_ranges-display}
parameter_ranges_v11 <- as.data.frame(parameter_ranges$v1.1)

knitr::kable(t(as.matrix(cbind("parameter" = c("min", "max"), parameter_ranges_v11))))
```

```{r sobol_pairs, include=SHOW_DIAGNOSTICS}
# Visualising parameter space with `pairs()`:
plotName = paste0("diagnostics/", VERSION, "_sobol_param_pairsPlot.png")

knitr::include_graphics(plotName)
```

## Data loading and preprocessing

Load simulation end states generated with the Sobol-sampled parameter combinations and sampled trajectories datasets. Preprocess datasets to adapt from NetLogo's specific format and categorise trajectories according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

```{r preprocessing}
endstates <- load_endstates("v1.1") |>
  preprocess_simdata("v1.1")
```

```{r preprocessing_trajectories}
sampled_trajectories <- load_trajectories("v1.1") |>
  preprocess_simdata("v1.1", full_trajectory = TRUE)
```

## Population dynamics

### End-state distributions

Overview of end states across residence rules. The histogram shows the distribution of final population sizes for each residence rule, with different colors indicating whether the population went extinct or survived. The vertical line represents the maximum population size defined in the model, which can help to visually assess how many simulations reached or exceeded this threshold. The x-axis can be set to a logarithmic scale to better visualize differences in population sizes among those that survived.

```{r fig-endstates-distrib}
endstates_distrib_plot <- ggplot(
  endstates,
  aes(x = totalIndividuals,
      fill = survival)) +
  
  geom_histogram(bins = 40) +
  
  geom_vline(aes(xintercept = max_population)) +
  
  facet_wrap(~ residence_rule, ncol = 1) +
  
  #scale_x_log10(labels = comma) + # to exclude extinction cases
  
  labs(
    x = "Final population size",
    y = "Simulation count",
    fill = "Outcome"
  )

endstates_distrib_plot

if (EXPORT_FIGURES) {
  export_figure(
    endstates_distrib_plot,
    "v11_endstates_distrib"
  )
}
```

Notice that only a small fraction of simulations reach the maximum population size, and that extinction predominate considering the entirity of the explored parameter space.

```{r fig-outcome-prop}
outcome_prop_plot <- endstates |>
  count(residence_rule, survival) |>
  group_by(residence_rule) |>
  mutate(freq = n / sum(n)) |>
  ggplot(aes(residence_rule, freq, fill = survival)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent) +
  labs(
    x = NULL,
    y = "Share of simulations",
    fill = "Outcome"
  )

outcome_prop_plot

if (EXPORT_FIGURES) {
  export_figure(
    outcome_prop_plot,
    "v11_outcome_prop"
  )
}
```

Taken as the exploration as a whole, `residential-rule` does not seem to have a strong influence on the overall distribution of end states, with both residence rules showing a similar predominance of extinction outcomes. However, there is a slightly higher proportion of survival outcomes under the "matri" rule compared to the "patri" rule, which could suggest a small influence of residence system on population persistence. This observation will be further explored in the subsequent analyses.

### Temporal trajectories

Dynamics of population trajectories across time steps, faceted by residence rule. The plot shows individual trajectories with low opacity and the median trajectory with higher opacity. The y-axis is on a logarithmic scale to better visualize differences in population sizes over time, especially when there are large variations between trajectories.

```{r trajectories}
pop_trajectories_plot <- ggplot(
  sampled_trajectories,
  aes(step, totalIndividuals,
    group = run_number)) +

  geom_line(alpha = 0.08) +

  stat_summary(
    aes(group = 1),
    fun = median,
    geom = "line",
    linewidth = 1.2
  ) +

  facet_wrap(~ residence_rule) +

  scale_y_log10()

pop_trajectories_plot

if (EXPORT_FIGURES) {
  export_figure(
    pop_trajectories_plot,
    "v11_pop_trajectories"
  )
}
```

The parameter configuration use to generate the trajectories is the same across residence rules, except for the `residence_rule` parameter itself. This allows us to directly compare the population dynamics under different residence systems while controlling for other factors. The configuration used for this was:

```{r trajectories-params-display}
knitr::kable(sampled_trajectories |> 
  select(all_of(names(parameter_ranges_v11))) |> 
  distinct() |>
  t())
```

For this specific parameter configuration, we can see that the population under the "matri" rule shows a more consistent growth trajectory, with most runs showing an increase in population size over time. In contrast, the population under the "patri" rule shows more variability, with some runs showing growth while others show decline or extinction. This suggests that, at least for this parameter configuration, the "matri" residence system may be more conducive to population persistence and growth compared to the "patri" residence system. The asymmetry explored here is focused on the parameters regulating nuptiality in the Peristera-Kostaki model; generally prescribing a scenario where women marry slightly earlier and more often than men. However, it's important to note that this is just one parameter configuration, and the overall influence of residence system on population dynamics may vary across different parameter combinations. Further analysis will be needed to assess the robustness of this observation across the broader parameter space explored in the simulations. 

## Sensitivity analysis

```{r}
sensitivity_data <- list(

  matri =
    endstates |>
    filter(
      residence_rule ==
        "matrilocal-matrilineal"
    ) |>
    prepare_sensitivity_predictors(
      parameter_metadata
    ),

  patri =
    endstates |>
    filter(
      residence_rule ==
        "patrilocal-patrilineal"
    ) |>
    prepare_sensitivity_predictors(
      parameter_metadata
    )
)
```


### Fitting Random-forest models

```{r}
rf_models <- load_or_build_rf_models(
  model_version = "v1.1",
  sensitivity_data = sensitivity_data
)
```

primary response variable: survival

Classification of end states as extinction, persistence and growth as categorical outcomes.

secondary, more nuanced response variable: log_totalIndividuals

### Parameter importance

```{r}
rf_survival_importance <- bind_rows(
  
  extract_importance(
    rf_models$class_survival$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$class_survival$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```



```{r}
plot_importance(
  rf_survival_importance, 
  param_color = parameter_metadata$color[match(rf_survival_importance$parameter, parameter_metadata$parameter)], 
  facet = "residence_rule")
```

associated colors for parameters based on their domain (general, mortality, fertility, nuptiality)

```{r}
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_survival_importance, 
  groups = "residence_rule", 
  param_color = parameter_metadata$color[match(rf_survival_importance$parameter, parameter_metadata$parameter)]
  )

importance_dumbbell_plot

if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v11_importance_dumbbell"
  )
}
```

### Partial dependence

Select top parameters

```{r}
top_parameters <- rf_survival_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(MeanDecreaseGini)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PD for RFs using `survival` as response variable and no population ("Extinction") as response class. Plot each PD result with `log_totalIndividuals` and per `residence_rule`.

```{r}
pd_results <- load_or_build_pd_results(
  model_version = "v1.1",
  overwrite = TRUE,
  sensitivity_data,
  rf_models$class_survival,
  parameters = top_parameters,
  response_class = "Extinction"
)
```

Plot results

```{r}
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_results,
  top_parameters,
  response_variable = "log_totalIndividuals",
  pd_label = "P(Extinction)",
  ncol = 2,
  plot_title_margin_bottom = -5
)

if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v11_param_pd",
    height = 10
  )
}
```

```{r}
plotName = paste0("plots/main/v11_param_pd.png")

knitr::include_graphics(plotName)
```

### Comparing trajectories for the top parameters

#### `cdmlt-level`

>Run simulations in NetLogo Behaviour Space using the same parameter configuration, except `residence-rule`, and varying `cdmlt-level` and `SEED`.

Loading simulation experiments' end states:

```{r}
trajectories_v11_cdmlt_level <- read.csv("simdata/v1.1/household-demography-main exp-trajectories_cdmlt-level-table.csv", skip = 6)
```

Preprocess dataset adapting from NetLogo'S specific format. Categorise trajectories according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

```{r}
MAX_POPULATION = trajectories_v11_cdmlt_level$max.population[1]

trajectories_v11_cdmlt_level <- trajectories_v11_cdmlt_level |>
  preprocess_params() |>
  set_endstate_survival_category()
```

```{r}
plot_name <- file.path(paste0(plotDirectory, "trajectories_cdmlt-level.", plot_file_format))

if (plot_file_format == "png") {
  graphic_scale <- 1
  legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
  axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5

  png(plot_name, width = graphic_scale * 1600, height = graphic_scale * 800, res = 300, pointsize = 4)
} else if (plot_file_format == "eps") {
  graphic_scale = 1.2
  # font_rescale = -0.2
  # axis_text_rescale = -0.1
  # margin_text_rescale = -0.5

  extrafont::loadfonts(device = "postscript")
  grDevices::cairo_ps(filename = plot_name ,
                      pointsize = 12,
                      width = graphic_scale * 8,
                      height = graphic_scale * 4,
                      onefile = FALSE,
                      family = "sans"
                      )
}

ggplot(trajectories_v11_cdmlt_level, aes(x = X.step., y = totalIndividuals, group = X.run.number., color = as.factor(cdmlt.level))) +
  geom_line(linewidth = .1) +
  facet_wrap(~ residence.rule, scales = "free") +
  xlab("time (years)") +
  ylab("population (individuals)") +
  #ylim(c(0, MAX_POPULATION)) +
  guides(color = guide_legend(title = "cdmlt-level")) +
  theme_minimal() +
  theme(axis.title = element_text(size = 10 * (graphic_scale + axis_title_rescale)),
        axis.text = element_text(size = 10 * (graphic_scale + axis_text_rescale)),
        legend.position = "bottom")

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plot_name)
```

### `c1-fert`

>Run simulations in NetLogo Behaviour Space using the same parameter configuration, except `residence-rule`, and varying `c1-fert` and `SEED`.

Loading simulation experiments' end states:

```{r}
trajectories_v11_c1_fert <- read.csv("simdata/v1.1/household-demography-main exp-trajectories_c1-fert-table.csv", skip = 6)
```

Preprocess dataset adapting from NetLogo'S specific format. Categorise trajectories according to whether populations get extinct, survive and/or grow within the maximum number of simulation steps (`max-iterations`).

```{r}
MAX_POPULATION = trajectories_v11_c1_fert$max.population[1]

trajectories_v11_c1_fert <- trajectories_v11_c1_fert |>
  preprocess_params() |>
  set_endstate_survival_category()
```

```{r}
plot_name <- file.path(paste0(plotDirectory,"trajectories_c1_fert.", plot_file_format))

if (plot_file_format == "png") {
  graphic_scale <- 1
  legend_text_rescale <- legend_key_rescale <- graphic_scale * -0.5
  axis_text_rescale <- axis_title_rescale <- graphic_scale * -0.5

  png(plot_name, width = graphic_scale * 1600, height = graphic_scale * 800, res = 300, pointsize = 4)
} else if (plot_file_format == "eps") {
  graphic_scale = 1.2
  # font_rescale = -0.2
  # axis_text_rescale = -0.1
  # margin_text_rescale = -0.5

  extrafont::loadfonts(device = "postscript")
  grDevices::cairo_ps(filename = plot_name ,
                      pointsize = 12,
                      width = graphic_scale * 8,
                      height = graphic_scale * 4,
                      onefile = FALSE,
                      family = "sans"
                      )
}

ggplot(trajectories_v11_c1_fert, aes(x = X.step., y = totalIndividuals, group = X.run.number., color = as.factor(c1.fert))) +
  geom_line(linewidth = .1) +
  facet_wrap(~ residence.rule, scales = "free") +
  xlab("time (years)") +
  ylab("population (individuals)") +
  #ylim(c(0, MAX_POPULATION)) +
  guides(color = guide_legend(title = "c1-fert")) +
  theme_minimal() +
  theme(axis.title = element_text(size = 10 * (graphic_scale + axis_title_rescale)),
        axis.text = element_text(size = 10 * (graphic_scale + axis_text_rescale)),
        legend.position = "bottom")

dev.off()
```

```{r, out.width = "\\textwidth"}
knitr::include_graphics(plot_name)
```

## Discussion notes

household demographic persistence in v1.1 is shaped at least as strongly by marriage-system structure, taken as a whole, as by mortality and fertility.

Most important overarching finding is that v1.1 persistence is jointly structured by mortality intensity (Coale-Demeny model's "level"), fertility intensity, and especially female nuptiality dynamics. All other parameters are much lower in importance in this regard. 

The top 10 parameters in importance regarding population survival are:

| Group               | Parameters          | Effect on P(Extinction) (direction/importance rank) |
| ------------------- | ------------------- | --------------------------------------------------------- |
| Mortality regime    | `cdmlt_level`       | - / 1
| Fertility schedule  | `c1_fert`           | - / 2 |                                                             
| Fertility schedule  | `sigma2_fert`       | - / 3 |                                   
| Fertility schedule  | `mu_fert`           | - / 8 |
| Nuptiality schedule | `c1_women`          | - / 5 |
| Nuptiality schedule | `mu_women`          | + / 6 |
| Nuptiality schedule | `sigma1_women`      | - / 4 |
| Nuptiality schedule | `c1_men`            | - / 7 |
| Nuptiality schedule | `mu_men`            | - / 10 |
| Nuptiality schedule | `sigma1_men`        | - / 9 |

1. `cdmlt_level`. Mortality remains the dominant threshold parameter. Higher Coale-Demeny levels correspond to lower mortality, greater population survival, and therefore lower extinction probability. What is particularly important is the shape of the curve: a long plateau, followed by a relatively abrupt transition. This suggests that demographic persistence is buffered up to a critical mortality regime, after which household replacement processes fail rapidly. The difference between residence systems remains important: the patrilocal-patrilineal system appears slightly more sensitive to mortality deterioration.

2. `c1_fert`. This parameter reflects the maximum intensity of fertility. Peak fertility probability strongly regulates persistence: higher peak fertility substantially reduces extinction probability. Importantly: the effect is smoother than mortality, suggesting fertility operates as a gradual stabilising force, rather than a threshold mechanism. This distinction between mortality thresholds and fertility compensation is analytically very valuable.

3. `sigma2_fert`. The spread of post-peak fertility matters substantially. Increasing it broadens the fertility schedule after the peak, lowering extinction probability by extending reproductive opportunities across a wider age range, and thus improving demographic resilience. It implies persistence depends not only on fertility intensity, but also on reproductive temporal flexibility.

4. `c1_women`. Female nuptiality intensity strongly affects persistence. The peak probability of women marrying displays a strong directional effect that implies household persistence is highly sensitive to female marriage participation. Marriage formation regulates household continuity, reproduction, and lineage transmission simultaneously. Importantly, the effect is stronger than the corresponding male parameter.

5. `mu_women`. Timing of female marriage, specifically the age at which women reach peak marriage probability, strongly shapes persistence. The positive slope suggests that later peak female marriage ages increase extinction probability. Delaying marriage compresses or shifts reproductive windows, which reduces demographic replacement efficiency. Notably, it appears stronger than the equivalent male timing parameter. Again, the system seems disproportionately sensitive to female nuptial timing.

6. `sigma1_women`. The breadth of early female marriage transition, governing the spread of the rising portion of the female nuptiality schedule, affects population resilience. Broader transitions increase extinction probability modestly, meaning that less concentrated marriage timing may weaken household formation efficiency, possibly by delaying stable couple formation, eating out female fertility time and decreasing the aggregate birth frequency.

7. `c1_men`. Male marriage intensity matters, but less strongly. The same general mechanism applied to women appears: higher male marriage probability improves persistence. However, the effect is weaker than for women. This asymmetry may reflect the centrality of female marriage timing for reproductive continuity, especially in household-based demographic systems.

8. `mu_fert`. Fertility peak timing has relatively weak influence, mattering less than fertility intensity, fertility spread, or nuptiality structure.

9. `sigma1_men`. Male marriage-transition spread has weak but detectable effects, suggesting that variation in the onset of male marriage timing affects persistence only indirectly. Again, the system appears much more sensitive to female nuptial structure.

10. `mu_men`. Delayed male peak marriage age mildly destabilises populations, but its effect is weaker than the female counterpart.

Interpreting these results in light of the parameters domains, they indicate three main points:

A. Mortality buffering and collapse thresholds (`cdmlt_level`). Mortality intensity takes the lead in importance over survival, unsurprisingly, as it is represented by a single parameter in the Coale-Demeny model. However, it is noteworthy that it is the only parameter presenting a clear threshold (between levels 15 and 16).

B. Fertility intensity versus fertility timing (`c1_fert` vs. `sigma2_fert`, `mu_fert`). The share of females giving birth on the peak age is much more defining to population survival than the peak age and spread of fertility. In contrast with `cdmlt_level`, `c1_fert` presents a much smoother, incremental effect over population survival.

C. Nuptiality as a demographic stabilisation mechanism. Delayed male marriage contributes to lower persistence, but not nearly as strongly as delayed female marriage. The model appears consistently more sensitive to female marriage intensity (`c1_women`) and female marriage timing (`mu_women`, `sigma1_women`) than to the equivalent male parameters. The high importance of reproductive scheduling (`sigma2_fert`, `mu_fert`) adds greatly the importance of female indicated by this asymmetry. That is extremely interesting because it emerges from household-demographic interaction structure, rather than being directly imposed analytically.

### Residential rule

Residential rule shows a moderate, but consistent effect over the probability of extinction, with "matrilineal-matrilocal" rule being slightly associated with survival. Consequence of this is that, when trajectories on the same parameter conditions are compared, the "matri" rule produces slightly more cases of rapid exponential, reaching `max-population` within the first hundreds of simulation steps.

This suggests kinship structure modifies demographic sensitivity, without completely overriding the underlying demographic mechanisms. Mortality, fertility, and nuptiality remain the core drivers, while residence rules modulate their effects. Not surprisingly, the effects of nuptiality parameters are generally stronger for women's parameters when under the "matri" rule, while men's parameters are highlighted under the "patri" rule.

### Other parameters

Initial household population

This parameter has very small positive effect on population survival:

Mortality model: region parameter has a smaller effect. East tends to survive more often.  

```{r}
cat("matri | survive? |")
table(endstates_matri$coale.demeny.region, endstates_matri$totalIndividuals > 0)
cat("patri | survive? |")
table(endstates_patri$coale.demeny.region, endstates_patri$totalIndividuals > 0)
```


