---
execute:
  echo: true
  warning: false
  message: false
params:
  show_diagnostics: false
---

# Demand-weighted density dependent model with daily iteration (v1.4.1)

## Setup

Set constant variables for this chapter.

```{r}
#| label: setup_constants
MODEL_VERSION <- "v1.4.1"
EXPORT_FIGURES <- TRUE
REMAKE_RF <- FALSE
REMAKE_DP <- FALSE
REMAKE_TRAJECTORY_PLOTS <- FALSE
```

Load libraries, sources and set ggplot theme for the notebook.

```{r} 
#| label: setup_libraries
library(data.table)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(ggtext)
library(zoo)
library(stringr)
library(effectsize)
library(randomForest)
library(randtoolbox)
library(rngWELL)
library(pdp)
library(forcats)
library(patchwork)
library(scales)
```

```{r} 
#| label: setup_load_sources
source("library/source_folder.R")
source_folder("library")
source_folder("library/structure")
source_folder("library/sensitivity")
source_folder("library/density")
```

```{r} 
#| label: setup_ggplot_theme
theme_set(
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold"),
      plot.title = element_text(face = "bold")
    )
)
```

## Experiment design

As v1.4.1 does not introduce new parameters explored in relation to v1.4, the same Sobol-sampled values are used here. Notice however that this model does introduce new constants used to determine the correspondence between age and the adult-equivalent unit (i.e., 15, 45, and the exponents 1.2 and 2; see `get-carrying-capacity-demand-at-age`).

As one of the daily iteration versions, only 512 configurations of the original Sobol exploration sample (\(n=2,000\)) was used to obtain sufficiently-robust results. The end states data set then contains 12,288 runs (512 configurations × 2 residence rules × 4 Coale-Demeny regions × 3 replicates). 

```{r} 
#| label: sobol_load
# Load object as Rdata file
load("data/v14_sobol_parameter_values.RData")
```

Sobol sampling is applied to the following parameter ranges:

```{r} 
#| label: sobol_ranges_display
parameter_ranges_v141 <- as.data.frame(parameter_ranges$v1.4)

knitr::kable(
  parameter_ranges_v141 |>
    as.matrix() |>
    t() |>
    as.data.frame() |>
    setNames(c("min", "max"))
)
```

```{r} 
#| label: sobol_pairs 
#| include: !expr params$show_diagnostics 
#| fig.cap: "Parameter ranges explored in the Sobol sampling for v1.4.1."
# Visualising parameter space with `pairs()`:
plotName = paste0("diagnostics/v1.4_sobol_param_pairsPlot.png")

knitr::include_graphics(plotName)
```

## Data loading and preprocessing

Load simulation end states generated with the Sobol-sampled parameter combinations and sampled trajectories datasets. Preprocess datasets to adapt from NetLogo's specific format and categorise end states and trajectories. 

```{r}
#| label: preprocessing_vars
overshoot_magnitude_threshold <- 0 # any value over carrying capacity is considered overshooting
burn_in_fraction = 0.5 # the second half of the simulation is considered representative of an equilibrium phase
overshoot_time_threshold <- 0.1 # 10% of time after "burn in" period
oscillatory_threshold <- 0.05 # coefficient of variance of population size after "burn in" period 
stress_threshold <- 0.5 # pressures over this value are considered as population under stress
```

```{r} 
#| label: preprocessing
endstates <- load_endstates(MODEL_VERSION) |>
  preprocess_simdata(MODEL_VERSION)
```

```{r} 
#| label: preprocessing_trajectories
file_name <- "data/v141_trajectories.RData"
if (!file.exists(file_name))
{
  sampled_trajectories <- load_trajectories(
    MODEL_VERSION,
    ignored_variables = c("womenAgeStructure", "menAgeStructure"),
    step_start = 1,
    step_end = 182500 # max-iterations is 500 years or 182500 days; cut down this number to save RAM
  ) |>
    preprocess_simdata(
      model_version = MODEL_VERSION,
      full_trajectory = TRUE
    ) |>
    compute_density_metrics(
      burn_in_fraction = burn_in_fraction
    ) |>
    classify_regulation_regimes(
      overshoot_magnitude_threshold = overshoot_magnitude_threshold,
      overshoot_time_threshold = overshoot_time_threshold,
      oscillatory_threshold = oscillatory_threshold,
      stress_threshold = stress_threshold
    )
  
  save_trajectories(
    sampled_trajectories,
    model_version = MODEL_VERSION
  )
}
```

A particularity in processing the data with age-weighted variables is that we need to use (or compute) the total carrying capacity demand (`carrying_capacity_demand`) and use it to calculate `pressure` instead of using `totalIndividuals` directly (i.e., `pressure = carrying_capacity_demand / carrying_capacity`). Although this variable can be exported through NetLogo's Behavior Space, it was not specified for all massive Sobol sampled experiments, so it is necessary to calculate it using `womenAgeStructure` and `menAgeStructure` (see "library/density/compute_carrying_capacity_demand.R").

Because of the sheer number of daily iterations required to approach comparable time scales of previous versions, trajectories datasets are simply too large to be loaded conventionally in R. For this reason, we need to use the function argument `ignored_variables` (alternatively also `selected_variables` is available) to skip loading unnecessary variables, as well as potentially restricting the span of the simulation data to a subset of iteration steps (`start_step`, `end_step`).

### Surviving population structure

```{r} 
#| label: age_structure_prepare
group_vars = c(
  "survival",
  "residence_rule",
  "coale_demeny_region",
  "mortality_regime",
  "fertility_intensity",
  "fertility_timing",
  "fertility_spread_pre_peak",
  "fertility_spread_post_peak",
  "union_timing_women",
  "union_timing_men",
  "union_intensity_women",
  "union_intensity_men",
  "union_spread_pre_peak_women",
  "union_spread_pre_peak_men",
  "union_spread_post_peak_women",
  "union_spread_post_peak_men",
  "amenorrhea_length"
)

age_structures <- endstates |>

  filter(
    survival != "Extinction"
  ) |>
  
  characterise_mort_fert_nupt_regimes() |>

  prepare_age_structure(
    age_breaks = seq(0, 100, 5),
    normalize = TRUE,
    group_vars = group_vars
  )

age_structure_indicators <- age_structures |>
  compute_age_structure_indicators(group_vars = group_vars) |>
  # make a binary classification for Cohen's d test
  mutate(
    pressure_regime = ifelse(
      survival == "Persistence\n(low pressure)",
      "Low pressure\nregimes",
      "High pressure\nregimes"
    )
  )

endstates <- endstates |>
  left_join(
    age_structure_indicators |>
      select(
        run_unique_id,
        median_age,
        youth_share,
        elderly_share,
        dependency_ratio
      ),
    by = "run_unique_id"
  )
```

## Overview of end states and dynamics

### Population persistence

```{r} 
#| label: endstates_distrib 
#| fig.cap: "Distribution of final population sizes across residence rules and Coale-Demeny regions."
log_max_population <- max(endstates$log_totalIndividuals)

endstates_distrib_size_plot <- endstates |>
  plot_endstate_variable_distribution(
    main_var = "log_totalIndividuals",
    fill_var = "survival",
    facet_var = c("coale_demeny_region", "residence_rule"),
    bins = 40,
    x_axis_label = "Final population size",
    y_axis_label = "Simulation count",
    threshold_values = log_max_population,
    thresholds_linetype = 2,
    x_axis_log10 = TRUE,
    x_max = log_max_population,
    y_axis_log10 = TRUE
  )
endstates_distrib_size_plot
```

```{r}
#| label: endstates_distrib_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    endstates_distrib_size_plot,
    "v141_endstates_distrib"
  )
}
```

```{r} 
#| label: endstates_distrib_pressure 
#| fig.cap: "Distribution of final population pressure in surviving populations across residence rules and Coale-Demeny regions."
endstates_pressure_plot <- endstates |>
  filter(survival != "Extinction") |>
  plot_endstate_variable_distribution(
    main_var = "pressure",
    fill_var = "survival",
    facet_var = c("coale_demeny_region", "residence_rule"),
    bins = 40,
    x_axis_label = "Final population pressure",
    y_axis_label = "Simulation count",
    threshold_values = c(0.5, 1),
    thresholds_linetype = c(2, 1),
    x_axis_log10 = FALSE,
    y_axis_log10 = FALSE
  )
endstates_pressure_plot
```

```{r}
#| label: endstates_distrib_pressure_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    endstates_pressure_plot,
    "v141_endstates_pressure"
  )
}
```
```{r} 
#| label: outcome_prop 
#| fig.cap: "Proportion of simulations across survival category, residence rules and Coale-Demeny regions."
outcome_prop_plot <- endstates |>
  plot_category_proportion_stacked(
    x_var = "residence_rule",
    fill_var = "survival",
    facet_var = "coale_demeny_region",
    facet_ncol = 2,
    facet_scales = "fixed",
    y_axis_label = "Share of simulations"
  )
outcome_prop_plot
```

```{r}
#| label: outcome_prop_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    outcome_prop_plot,
    "v141_outcome_prop"
  )
}
```

The distribution of end states is broadly similar across Coale-Demeny regions, with extinction and low-pressure persistence accounting for most outcomes and high-pressure persistence forming a smaller share. Overshooting is rare. Age-weighted demand therefore produces a predominantly regulated system in which surviving populations generally remain below the nominal carrying-capacity threshold in demand-equivalent terms.

Residence rule introduces a modest asymmetry. Patrilocal-patrilineal populations show a slightly higher frequency of survival than matrilocal-matrilineal populations, although the overall distributions remain similar. The small magnitude of this difference indicates that residence organisation is a secondary influence relative to the continuously varied demographic and regulatory parameters.

### Temporal trajectories

The parameter configuration used to generate the trajectories is the same across residence rules, except for the `residence_rule` parameter itself. The configuration used for this was:

```{r}
#| label: trajectories_load
rm(sampled_trajectories_labour_demand_per_capita)
sampled_trajectories <- readRDS("data/v141_trajectories.RData")
```

```{r} 
#| label: trajectories_size_params_display 
#| tab.cap: "Parameter configuration for the population trajectories sampled."
knitr::kable(sampled_trajectories |> 
               mutate(
                 max_iterations_in_days = max_iterations * year_length_in_days,
                 max_iterations_in_days_cut = max(step)
               ) |>
               select(all_of(c("max_iterations",
                               "max_iterations_in_days",
                               "max_iterations_in_days_cut", # show this in case loading was partial
                               "max_population", 
                               names(parameter_ranges_v141), 
                               "coale_demeny_region", 
                               "maturity_age_in_years", "gestation_period_in_days", "year_length_in_days" ))) |> 
               distinct() |>
               t() |>
               as.data.frame() |>
               rename(value = 1)
)
```

Dynamics of population trajectories across time steps, faceted by residence rule. The plot shows individual trajectories with low opacity and the median trajectory with higher opacity. The y-axis is on a logarithmic scale to better visualize differences in population sizes over time, especially when there are large variations between trajectories.

```{r} 
#| label: trajectories_size_plot 
#| fig.cap: "A sample of population trajectories across time steps, faceted by residence rule."
if (REMAKE_TRAJECTORY_PLOTS)
{
  pop_trajectories_plot <- sampled_trajectories |>
    plot_trajectories(
      y_var = "log_totalIndividuals",
      trajectory_id = "run_unique_id",
      line_alpha = 0.04,
      median_alpha = 0.8,
      median_linewidth = 1,
      y_axis_log10 = TRUE,
      y_axis_label = "number of individuals",
      group_var =  "regulation_regime",
      facet_var = "residence_rule",
      facet_ncol = 2,
      facet_scales = "fixed",
      y_threshold_var = "carrying_capacity",
      thresholds_linetype = 2
    )
  pop_trajectories_plot
} else {
  plotName = paste0("plots/v141_pop_trajectories.png")

  knitr::include_graphics(plotName)
}
```

```{r}
#| label: trajectories_size_plot_export
#| echo: false
if (EXPORT_FIGURES & REMAKE_TRAJECTORY_PLOTS) {
  export_figure(
    pop_trajectories_plot,
    "v141_pop_trajectories"
  )
}
rm(pop_trajectories_plot)
```

Under the selected parameter configuration, both residence systems rapidly approach bounded population levels and remain viable throughout the simulation horizon. Individual trajectories fluctuate around their long-run range, but neither residence system exhibits sustained expansion or systematic collapse.

Population levels are also remarkably similar between residence rules. Age-weighted density regulation therefore produces a robust high-pressure equilibrium under this favourable demographic configuration, with residence organisation contributing comparatively little to the realised abundance.

Because pressure is based on age-weighted carrying-capacity demand rather than headcount, the population can remain numerically above the value of `carrying_capacity` without necessarily representing an equivalent ecological overshoot. The regulatory state must therefore be interpreted through demand-weighted pressure rather than population size alone.

### Household composition

```{r}
#| label: household_structure_plot
#| fig.cap: ""
household_structure_plot <- endstates |> 
  plot_household_structure_faceted_fixed(
    color_var = "max_couple_count_distribution_max",
    point_alpha = 0.08,
    facet_var = "residence_rule",
    facet_ncol = 2,
    facet_scales = "fixed",
    xy_log10 = TRUE,
    x_axis_label = "number of households",
    y_axis_label = "number of individuals",
    color_label = "Max. couple count allowed in a household\n(parametric maximum)",
    legend_position = "bottom"
  )
household_structure_plot
```

```{r}
#| label: household_structure_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    household_structure_plot,
    "v141_household_structure"
  )
}
```

Population size and household number remain closely coupled under both residence systems, while mean household size stays within a comparatively narrow range despite substantial differences in total population abundance. Household formation and fission therefore continue to scale with population growth without producing persistent enlargement of households.

```{r}
#| label: trajectories_household_structure_plot 
#| fig.cap: "A sample of household population trajectories across time steps, faceted by residence rule."
if (REMAKE_TRAJECTORY_PLOTS)
{
  trajectories_household_structure_plot <- 
    sampled_trajectories |>
    group_by(residence_rule) |>
    mutate(
      household_structure_ratio = (totalIndividuals / totalHouseholds)
    ) |>
    ungroup() |>
    plot_trajectories(
      y_var = "household_structure_ratio",
      trajectory_id = "run_unique_id",
      line_alpha = 0.035,
      median_alpha = 0.8,
      median_linewidth = 1,
      y_axis_log10 = FALSE,
      y_axis_label = "Mean individual count per household",
      group_var = "regulation_regime",
      facet_var = "residence_rule",
      facet_ncol = 2,
      facet_scales = "free_x",
      y_threshold_var = NULL
    )
  trajectories_household_structure_plot
} else {
  plotName = paste0("plots/v141_trajectories_household_structure.png")

  knitr::include_graphics(plotName)
}
```

```{r}
#| label: trajectories_household_structure_plot_export
#| echo: false
if (EXPORT_FIGURES & REMAKE_TRAJECTORY_PLOTS) {
  export_figure(
    trajectories_household_structure_plot,
    "v141_trajectories_household_structure",
    width = 8,
    height = 4
  )
}
rm(trajectories_household_structure_plot)
```

```{r}
#| label: trajectories_household_structure_plot_peaks
#| fig.cap: "A sample of household population trajectories across time steps, faceted by residence rule, highlighting in red the trajectories that achieve the maximum value in each subset."
if (REMAKE_TRAJECTORY_PLOTS)
{
  trajectories_household_structure_plot_peaks <- 
    sampled_trajectories |>
    group_by(residence_rule) |>
    mutate(
      household_structure_ratio = if_else(totalHouseholds == 0, 0, (totalIndividuals / totalHouseholds)),
      # Find the absolute peak value within the entire residence rule
      global_max = max(household_structure_ratio, na.rm = TRUE)
    ) |>
    group_by(residence_rule, run_number) |>
    mutate(
      # Check if this specific run contains that global peak value
      selected_trajectory = any(household_structure_ratio == global_max)
    ) |>
    ungroup() |>
    plot_trajectories(
      y_var = "household_structure_ratio",
      trajectory_id = "run_unique_id",
      line_alpha = 0.035,
      median_alpha = 0.8,
      median_linewidth = 1,
      y_axis_log10 = FALSE,
      y_axis_label = "Mean individual count per household",
      group_var = NULL,
      facet_var = "residence_rule",
      facet_ncol = 2,
      facet_scales = "fixed",
      y_threshold_var = NULL
    ) +
    # Add the highlighted layer on top
    geom_line(
      data = \(df) filter(df, selected_trajectory == TRUE),
      aes(
        x = step, 
        y = household_structure_ratio,
        group = run_number
      ),
      color = "red",
      linewidth = 1,
      alpha = 1
    )
  trajectories_household_structure_plot_peaks
} else {
  plotName = paste0("plots/v141_household_structure_trajectories_peaks.png")

  knitr::include_graphics(plotName)
}
```

```{r}
#| label: trajectories_household_structure_plot_peaks_export
#| echo: false
if (EXPORT_FIGURES & REMAKE_TRAJECTORY_PLOTS) {
  export_figure(
    trajectories_household_structure_plot_peaks,
    "v141_household_structure_trajectories_peaks",
    width = 8,
    height = 4
  )
}
rm(trajectories_household_structure_plot_peaks)
```

The trajectory experiment shows an initial increase in mean household size followed by prolonged fluctuation around a stable range. Occasional trajectories reach substantially larger household sizes, but these excursions are temporary and are not followed systematically by population collapse. They are better interpreted as stochastic episodes of household aggregation than as indicators of demographic instability.

Differences between residence systems are modest, although the largest transient household-size peaks occur in the matrilocal-matrilineal sample.

### Density regulation dynamics

```{r}
#| label: regulation_dynamics_rr_prop
regulation_dynamics_rr_prop_plot <- sampled_trajectories |>
  group_by(run_unique_id) |>
  filter(row_number() == 1) |> 
  ungroup() |>
  plot_category_proportion(
    category_var = "regulation_regime",
    facet_var = "residence_rule",
    category_fill = TRUE,
    category_labs_size = 5,
    category_labs_angle = 45,
    category_labs_hjust = 0.5
  )
```

```{r}
#| label: regulation_dynamics_rr_prop_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    regulation_dynamics_rr_prop_plot,
    "v141_regulation_dynamics_rr_prop",
    width = 5,
    height = 3
  )
}
```

```{r} 
#| label: regulation_dynamics_rr_prop_plot_display 
#| fig.cap: "Distribution of density regulation regimes across residence rules, for the trajectories sampled."
plotName = paste0("plots/v141_regulation_dynamics_rr_prop.png")

knitr::include_graphics(plotName)
```

```{r}
#| label: regulation_dynamics_rr_mean_cv_plot 
#| fig.cap: ""
regulation_dynamics_rr_mean_cv_plot <- sampled_trajectories |>
  plot_pressure_distributions(
    colour_var = "regulation_regime",
    shape_var = "residence_rule",
    stress_threshold = stress_threshold,
    oscillatory_threshold = oscillatory_threshold,
    thresholds_linetype = 2,
    point_alpha = 0.8
  )
```

```{r}
#| label: regulation_dynamics_rr_mean_cv_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    regulation_dynamics_rr_mean_cv_plot,
    "v141_regulation_dynamics_rr_mean_cv"
  )
}
```

```{r} 
#| label: regulation_dynamics_rr_mean_cv_plot_display 
#| fig.cap: "Distribution of mean and CV of population pressure across density regulation regimes and residence rules, for the trajectories sampled."
plotName = paste0("plots/v141_regulation_dynamics_rr_mean_cv.png")

knitr::include_graphics(plotName)
```

The targeted trajectories occupy an exceptionally consistent regulatory regime. All sampled runs under both residence systems are classified as Stress-prone, with late-run mean pressures close to the upper part of the persistence range and very low coefficients of variation.

This combination indicates stable regulation under sustained high ecological pressure rather than oscillatory instability. Although the daily population trajectories exhibit visible short-term stochastic fluctuations, these do not translate into large long-run variation in pressure.

Residence rule has little effect on the regulatory regime under this configuration. Matrilocal-matrilineal and patrilocal-patrilineal populations occupy almost the same region of mean-CV space, showing that demand-weighted density dependence can produce highly similar stable equilibria despite differences in household organisation.

### Age structure

#### Survival categories

```{r} 
#| label: age_structure_survival_pyramid_plot 
#| fig.cap: "Aggregate age structures of surviving populations across survival categories."
age_structures |>
  aggregate_age_structure(
    group_vars = c("survival"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "survival",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_survival_median_plot 
#| fig.cap: "Distribution of median age of surviving populations across survival categories."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival"
)
```

```{r} 
#| label: age-structure-survival-indicators-summary 
#| tab.cap: "Summary of age structure indicators for surviving populations across survival categories."
knitr::kable(indicators_summary(age_structure_indicators, group_vars = "survival"))
```

```{r} 
#| label: age_structure_survival_median_cohend 
#| tab.cap: "Cohen's d for median age across survival categories; low vs high pressure (and overshooting)."
knitr::kable(
  interpret_cohens_d(
    cohens_d(
      median_age ~ pressure_regime,
      data = age_structure_indicators
    ),
    rules = "cohen1988"
  )
)
```

Population-pressure regime is strongly associated with demographic composition. Low-pressure populations have a mean median age of approximately 24.9 years, compared with 20.2 years among high-pressure populations, and they also contain smaller juvenile shares and lower dependency ratios. The low- versus high-pressure difference is medium in magnitude (\(d=-0.71\)).

High ecological pressure is therefore associated with more rapid demographic renewal. Parameter combinations that sustain larger juvenile cohorts also generate greater age-weighted demand relative to carrying capacity, linking demographic composition directly to the ecological feedback mechanism.

The relatively small number of overshooting populations occupies an intermediate position in median age and should not be interpreted as a distinct demographic regime without further evidence. The clearest structural division remains that between low- and high-pressure persistence.

#### Residence rule and Coale-Demeny regions

```{r} 
#| label: age_structure_pyramid_plot_cdr_rr
pyramid_plot <- age_structures |>
  aggregate_age_structure(
    group_vars = c("residence_rule", "coale_demeny_region"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = c("coale_demeny_region", "residence_rule"),
    show_mean = FALSE
    )
```

```{r}
#| label: age_structure_pyramid_plot_cdr_rr_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    pyramid_plot,
    "v141_pyramid_residence_rule_region",
    height = 10
  )
}
```

```{r} 
#| label: age_structure_pyramid_plot_cdr_rr_display 
#| fig.cap: "Aggregate age structures of surviving populations across residence rules and Coale-Demeny regions."
plotName = paste0("plots/v141_pyramid_residence_rule_region.png")

knitr::include_graphics(plotName)
```

```{r} 
#| label: age_structure_indicators_plot_median_cdr_rr 
#| fig.cap: "Distribution of median age of surviving populations across residence rules and Coale-Demeny regions."
violin_plot <- age_structure_indicators |>
  plot_age_structure_indicators(
    x_var = "survival", 
    fill_var = "residence_rule", 
    facet_var = "coale_demeny_region"
  )
```

```{r}
#| label: age_structure_indicators_plot_median_cdr_rr_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    violin_plot,
    "v141_violin_residence_rule_region",
    width = 6,
    height = 3
  )
}
```

```{r} 
#| label: age_structure_indicators_plot_median_cdr_rr_display 
#| fig.cap: "Distribution of median age of surviving populations across residence rules and Coale-Demeny regions."
plotName = paste0("plots/v141_violin_residence_rule_region.png")

knitr::include_graphics(plotName)
```

```{r} 
#| label: age_structure_indicators_summary_cdr_rr 
#| tab.cap: "Summary of age structure indicators for surviving populations across residence rules and Coale-Demeny regions."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "coale_demeny_region")
    )
)
```

```{r} 
#| label: age_structure_cohens_d_cdr_rr 
#| tab.cap: "Cohen's d for median age across residence rules."
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ residence_rule,
      data = age_structure_indicators
    ),
    rules = "cohen1988"
  )
)
```

Cohen's d for median age across survival categories, by residence rule:

```{r} 
#| label: age_structure_cohens_d_cdr_rr_by_rr_matri 
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ pressure_regime,

      data =
          age_structure_indicators |>
          filter(
              residence_rule ==
                  "matrilocal-matrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_cohens_d_cdr_rr_by_rr_patri 
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
      effectsize::cohens_d(
        median_age ~ pressure_regime,

      data =
          age_structure_indicators |>
          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

Residence organisation produces a consistent but very small direct difference in age structure. Matrilocal-matrilineal populations are slightly older and contain somewhat smaller juvenile shares across all Coale-Demeny regions, but the aggregate effect is only \(d=0.16\). Regional mortality families likewise produce comparatively modest variation.

Population-pressure regime generates much larger differences. The low- versus high-pressure contrast is medium under matrilocal-matrilineal residence (\(d=-0.60\)) and large under patrilocal-patrilineal residence (\(d=-0.82\)). High-pressure populations are therefore substantially younger under both systems, with the separation somewhat sharper under patrilocal-patrilineal organisation.

Pressure regime, rather than residence rule or mortality-region category, is consequently the principal structural distinction among surviving populations.

## Sensitivity analysis

### Prepare data for sensitivity analysis

Prepare a dataframe out of `endstates` that contains only predictors (explored parameters) and response variables (`pressure`, `survival`, `median_age`), as well as splitting it according to `residence_rule`.

```{r} 
#| label: sensitivity_data
response_variables <- c(
  "pressure",
  "survival",
  "median_age"
)

sensitivity_data <- prepare_sensitivity_data_by_residence_rule(
  endstates,
  response_variables,
  parameter_metadata
)
```

### Fitting Random-forest models

```{r} 
#| label: sensitivity_rf
rf_models <- load_or_build_rf_models(
  model_version = MODEL_VERSION,
  sensitivity_data = sensitivity_data,
  response_variables = response_variables,
  SEED = 123,
  verbose = FALSE,
  overwrite = REMAKE_RF
)
```

Random forest models are fitted separately for each residence rule, using the same set of predictors. The response variables include `pressure`, `survival`, and `median_age`, allowing us to assess the importance of predictors for different aspects of population dynamics and structure. Notice that the model built for `median_age` uses only cases with surviving populations.

The models are built using a consistent random seed to ensure reproducibility, and the option to overwrite existing models is controlled by the `REMAKE_RF` flag.

### Parameter importance and partial dependence

#### Survival (probability of extinction end state)

```{r} 
#| label: sensitivity_survival_importance
rf_survival_importance <- bind_rows(
  
  extract_importance(
    rf_models$survival$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$survival$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)

param_color <- parameter_metadata$color[match(rf_survival_importance$parameter, parameter_metadata$parameter)]
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_survival_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population survival, by residence rule."
plot_importance(
  rf_survival_importance, 
  param_color = param_color, 
  facet_var = "residence_rule")
```

```{r} 
#| label: sensitivity_survival_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population survival, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_survival_importance, 
  groups = "residence_rule", 
  param_color = param_color
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_survival_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v141_survival_importance_dumbbell"
  )
}
```

Select top parameters in relation to survival:

```{r} 
#| label: sensitivity_survival_top_param
top_parameters_survival <- rf_survival_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(MeanDecreaseGini)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `survival` as response variable and no population ("Extinction") as response class. Plot each PD result with `survival` and per `residence_rule`.

```{r} 
#| label: sensitivity_survival_pd
pd_survival <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunk) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_survival,
  response_variable = "survival",
  response_class = "Extinction" # used only with `survival` as response variable
)
```

```{r} 
#| label: sensitivity_survival_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_survival,
  top_parameters_survival,
  response_variable = "pressure",
  pd_label = "P(Extinction)",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_survival_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v141_survival_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_survival_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing population survival, by residence rule."
plotName = paste0("plots/v141_survival_param_pd.png")

knitr::include_graphics(plotName)
```

Fertility peak intensity (`c1_fert`) is overwhelmingly the strongest predictor of extinction under both residence systems. Its partial dependence shows a steep decline in extinction probability as fertility rises through the lower part of its range, confirming reproductive replacement as the principal determinant of demographic viability.

Female union-formation parameters, particularly pre-peak spread (`sigma1_women`) and modal timing (`mu_women`), form an important secondary tier, together with post-peak fertility spread (`sigma2_fert`). `density_effect_scaling_factor`, nominal carrying capacity, female union peak intensity, and baseline mortality contribute at lower but still appreciable levels.

Survival in v1.4.1 is therefore governed primarily by reproductive recruitment and union formation, with ecological regulation providing an additional constraint. Baseline mortality is influential but does not dominate the sensitivity hierarchy.

#### Population pressure

```{r} 
#| label: sensitivity_pressure_importance
rf_pressure_importance <- bind_rows(
  
  extract_importance(
    rf_models$pressure$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$pressure$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_pressure_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population pressure, by residence rule."
plot_importance(
  rf_pressure_importance, 
  param_color = param_color, 
  facet_var = "residence_rule",
  metric = "%IncMSE")
```

```{r} 
#| label: sensitivity_pressure_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting population pressure, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_pressure_importance, 
  groups = "residence_rule", 
  param_color = param_color,
  metric = "%IncMSE"
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_pressure_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v141_pressure_importance_dumbbell"
  )
}
```

Select top parameters predicting pressure:

```{r} 
#| label: sensitivity_pressure_top_param
top_parameters_size <- rf_pressure_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(`%IncMSE`)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `pressure` as response variable. Plot each PD result with `pressure` and per `residence_rule`.

```{r} 
#| label: sensitivity_pressure_pd
pd_size <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunck) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_size,
  response_variable = "pressure"
)
```

```{r} 
#| label: sensitivity_pressure_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_size,
  top_parameters_size,
  response_variable = "pressure",
  pd_label = "Expected change in population size",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_pressure_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v141_pressure_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_pressure_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing population pressure, by residence rule."
plotName = paste0("plots/v141_pressure_param_pd.png")

knitr::include_graphics(plotName)
```

Population pressure is controlled jointly by demographic recruitment and ecological feedback. Fertility peak intensity is the strongest predictor, with higher fertility increasing age-weighted demand relative to carrying capacity. `density_effect_scaling_factor` follows closely and has the opposite effect: stronger density-dependent mortality lowers the pressure maintained by surviving populations.

Baseline mortality and nominal carrying capacity form the next tier of predictors, while male and female union parameters contribute secondary variation. Higher Coale-Demeny levels generally permit greater pressure by reducing baseline mortality, whereas stronger ecological feedback suppresses it.

Demand weighting adds an important compositional dimension to this response. Pressure is not simply a function of population headcount but of the age-specific consumption demand generated by that population. Fertility and union formation can therefore influence ecological load through both population abundance and demographic composition.

#### Median age

```{r} 
#| label: sensitivity_median_age_importance
rf_median_age_importance <- bind_rows(
  
  extract_importance(
    rf_models$median_age$matri
  ) |>
    mutate(
      residence_rule = "matrilocal-matrilineal"
    ),
  
  extract_importance(
    rf_models$median_age$patri
  ) |>
    mutate(
      residence_rule = "patrilocal-patrilineal"
    )
)
```

Note: associated colours for parameters based on their domain (general, mortality, fertility, nuptiality).

```{r} 
#| label: sensitivity_median_age_importance_barplot 
#| fig.cap: "Variable importance for the Random Forest models predicting median age, by residence rule."
plot_importance(
  rf_median_age_importance, 
  param_color = param_color, 
  facet_var = "residence_rule",
  metric = "%IncMSE")
```

```{r} 
#| label: sensitivity_median_age_importance_dumbbellplot 
#| fig.cap: "Variable importance for the Random Forest models predicting median age, by residence rule, with a dumbbell plot to directly compare the importance of parameters across residence rules."
importance_dumbbell_plot <- plot_importance_dumbbell(
  rf_median_age_importance, 
  groups = "residence_rule", 
  param_color = param_color,
  metric = "%IncMSE"
  )

importance_dumbbell_plot
```

```{r}
#| label: sensitivity_median_age_importance_dumbbellplot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    importance_dumbbell_plot,
    "v141_median_age_importance_dumbbell"
  )
}
```

Select top parameters predicting median age:

```{r} 
#| label: sensitivity_median_age_top_param
top_parameters_median_age <- rf_median_age_importance |>
  group_by(parameter) |>
  summarise(
    mean_importance =
      mean(`%IncMSE`)
  ) |>
  slice_max(
    mean_importance,
    n = 10
  ) |>
  pull(parameter)
```

For each top parameter, calculate PDs for the corresponding RFs using `median_age` as response variable (Extinction cases are not included).

```{r} 
#| label: sensitivity_median_age_pd
pd_median_age <- load_or_build_pd_results(
  model_version = MODEL_VERSION,
  overwrite = REMAKE_DP, # set REMAKE_DP = TRUE (first code chunck) to overwrite any previous results
  sensitivity_data,
  rf_models,
  parameters = top_parameters_median_age,
  response_variable = "median_age"
)
```

```{r} 
#| label: sensitivity_median_age_pd_plot
param_pd_plots <- plot_parameter_dependence_composite(
  sensitivity_data,
  pd_median_age,
  top_parameters_median_age,
  response_variable = "median_age",
  pd_label = "Expected change in median age",
  ncol = 2,
  plot_title_margin_bottom = -5
)
```

```{r}
#| label: sensitivity_median_age_pd_plot_export
#| echo: false
if (EXPORT_FIGURES) {
  export_figure(
    param_pd_plots,
    "v141_median_age_param_pd",
    height = 10
  )
}
```

```{r} 
#| label: sensitivity_median_age_pd_plot_display 
#| fig.cap: "Partial dependence plots for the top parameters influencing median age, by residence rule."
plotName = paste0("plots/v141_median_age_param_pd.png")

knitr::include_graphics(plotName)
```

Fertility peak intensity is by far the strongest predictor of median age among surviving populations. Male union peak intensity, post-peak fertility spread, female union peak intensity, male pre-peak union spread, and postpartum amenorrhoea form the principal secondary tier.

Baseline mortality has comparatively little importance for median age, despite its relevance to survival and population pressure. The age composition of surviving populations is therefore determined primarily by reproductive recruitment, union formation, and birth spacing rather than by differences in mortality intensity.

Demand-weighted regulation does not overturn this demographic hierarchy. Instead, it links the age structure produced by these mechanisms more directly to ecological pressure by assigning different carrying-capacity demands to individuals of different ages.

### Age structure of surviving populations for selected parameters

Note: parameters are grouped by submodels and meaning, not by importance. The binary categories created earlier in preprocessing, which split the parameter range according to the distribution median, are used.

#### `c1_fert`: high *versus* low fertility regimes

```{r} 
#| label: age_structure_c1_fert_pyramid_plot
#| fig.cap: "Population pyramids across fertility intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_intensity"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_intensity",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_fert_median_plot
#| fig.cap: "Median age of surviving populations across fertility regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_intensity"
)
```

```{r} 
#| label: age_structure_c1_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across fertility regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_intensity")
    )
)
```

Cohen's d for median age across fertility regimes, by residence rule:

```{r} 
#| label: age_structure_c1_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
    median_age ~ fertility_intensity,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_intensity,

    data =
        age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

High fertility peak intensity produces substantially younger populations under both residence systems, with medium effects (\(d=-0.77\) and \(-0.73\)). The effect is marginally stronger under matrilocal-matrilineal residence, although the two estimates are very similar. Fertility peak intensity therefore remains the dominant parameter-specific determinant of cohort renewal.

#### `mu_fert`: early *versus* late fertility timing

```{r} 
#| label: age_structure_mu_fert_pyramid_plot
#| fig.cap: "Population pyramids across (female) fertility timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_timing"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_timing",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_fert_median_plot
#| fig.cap: "Median age of surviving populations across (female) fertility timing regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_timing"
)
```

```{r} 
#| label: age_structure_mu_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across (female) fertility timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_timing")
    )
)
```

Cohen's d for median age across fertility timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_timing,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_timing,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Earlier fertility timing is associated with slightly younger populations under both residence rules, but the effect is very small and virtually identical in magnitude (\(d\approx-0.12\)). Modal fertility timing therefore modifies demographic pacing only modestly compared with fertility peak intensity and fertility spread.

#### `sigma1_fert`: broad *versus* narrow pre-peak fertility spread regimes

```{r} 
#| label: age_structure_sigma1_fert_pyramid_plot
#| fig.cap: "Population pyramids across pre-peak fertility spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_spread_pre_peak"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_spread_pre_peak",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_fert_median_plot
#| fig.cap: "Median age of surviving populations across pre-peak fertility spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_spread_pre_peak"
)
```

```{r} 
#| label: age_structure_sigma1_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across pre-peak fertility spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_spread_pre_peak")
    )
)
```

Cohen's d for median age across pre-peak fertility spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_pre_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_pre_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader pre-peak fertility schedules produce somewhat younger surviving populations, particularly under matrilocal-matrilineal residence. The effect remains very small (\(d=-0.19\) and \(-0.10\)), indicating a secondary contribution from earlier and more dispersed reproductive participation.

#### `sigma2_fert`: broad *versus* narrow post-peak fertility regimes

```{r} 
#| label: age_structure_sigma2_fert_pyramid_plot
#| fig.cap: "Population pyramids across fertility post-peak spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("fertility_spread_post_peak"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "fertility_spread_post_peak",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_fert_median_plot
#| fig.cap: "Median age of surviving populations across fertility post-peak spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "fertility_spread_post_peak"
)
```

```{r} 
#| label: age_structure_sigma2_fert_indicators_summary
#| tab.cap: "Summary of age structure indicators across fertility post-peak spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "fertility_spread_post_peak")
    )
)
```

Cohen's d for median age across fertility spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_fert_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_post_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_fert_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ fertility_spread_post_peak,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader post-peak fertility schedules are associated with younger populations under both residence systems, with small effects (\(d=-0.43\) and \(-0.35\)). Extended reproductive opportunity therefore sustains demographic renewal over a wider age range, although its influence remains substantially below that of fertility peak intensity.

#### `cdmlt_level`: mortality regimes

```{r}
#| label: age_structure_cdmlt_level_pyramid_plot
#| fig.cap: "Population pyramids across mortality regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("mortality_regime"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "mortality_regime",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_cdmlt_level_median_plot
#| fig.cap: "Median age of surviving populations across mortality regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "mortality_regime"
)
```

```{r} 
#| label: age_structure_cdmlt_level_indicators_summary
#| tab.cap: "Summary of age structure indicators across mortality regimes and residence rules."
knitr::kable(
  indicators_summary(age_structure_indicators, group_vars = c("residence_rule", "mortality_regime"))
)
```

Cohen's d for median age across mortality regimes, by residence rule:

```{r} 
#| label: age_structure_cdmlt_level_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ mortality_regime,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_cdmlt_level_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ mortality_regime,

      data =
          age_structure_indicators |>

          filter(
              residence_rule ==
                  "patrilocal-patrilineal"
          )
    ),
    rules = "cohen1988"
  )
)
```

Mortality regime has virtually no direct effect on median age among surviving populations. Mean median ages differ by only a few tenths of a year, and both effect sizes are close to zero with confidence intervals spanning zero. Baseline mortality therefore influences whether populations survive and the pressure they can maintain more strongly than it influences the age structure of populations once they have survived.

#### `c1_women`: high *versus* low female union peak intensity regimes

```{r} 
#| label: age_structure_c1_women_pyramid_plot
#| fig.cap: "Population pyramids across female union peak intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_intensity_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_intensity_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_women_median_plot
#| fig.cap: "Median age of surviving populations across female union peak intensity regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_intensity_women"
)
```

```{r} 
#| label: age_structure_c1_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across female union peak intensity regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_intensity_women")
    )
)
```

Cohen's d for median age across female union peak intensity regimes, by residence rule:

```{r} 
#| label: age_structure_c1_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_intensity_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_intensity_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

Higher female union peak intensity is associated with slightly younger populations under both residence systems, but the effect is very small (\(d=-0.14\) and \(-0.16\)). Female union intensity therefore contributes modestly to cohort renewal while remaining secondary to the principal fertility parameters.

#### `mu_women`: late *versus* early female union regimes

```{r} 
#| label: age_structure_mu_women_pyramid_plot
#| fig.cap: "Population pyramids across female union timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_timing_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_timing_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_women_median_plot
#| fig.cap: "Median age of surviving populations across female union timing regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_timing_women"
)
```

```{r} 
#| label: age_structure_mu_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across female union timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_timing_women")
    )
)
```

Cohen's d for median age across female union timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_timing_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_timing_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Female union timing has weak and residence-dependent effects. Earlier union formation is associated with slightly younger populations under matrilocal-matrilineal residence (\(d=-0.11\)), whereas the relationship reverses under patrilocal-patrilineal residence (\(d=0.12\)). Given the very small magnitudes, the result is better interpreted as context dependence than as evidence for a common directional effect of female union timing.

#### `sigma1_women`: broad *versus* narrow pre-peak female union spread regimes

```{r} 
#| label: age_structure_sigma1_women_pyramid_plot
#| fig.cap: "Population pyramids across pre-peak female union spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_spread_pre_peak_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_spread_pre_peak_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_women_median_plot
#| fig.cap: "Median age of surviving populations across pre-peak female union spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_spread_pre_peak_women"
)
```

```{r} 
#| label: age_structure_sigma1_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across pre-peak female union spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_spread_pre_peak_women")
    )
)
```

Cohen's d for median age across pre-peak female union spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_pre_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_pre_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader pre-peak female union spread is associated with slightly younger populations under matrilocal-matrilineal residence (\(d=-0.18\)), whereas no detectable effect appears under patrilocal-patrilineal residence. Its structural influence is therefore weak and strongly dependent on residence organisation.

#### `sigma2_women`: broad *versus* narrow post-peak female union spread regimes

```{r} 
#| label: age_structure_sigma2_women_pyramid_plot
#| fig.cap: "Population pyramids across post-peak female union spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_spread_post_peak_women"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_spread_post_peak_women",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_women_median_plot
#| fig.cap: "Median age of surviving populations across post-peak female union spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_spread_post_peak_women"
)
```

```{r} 
#| label: age_structure_sigma2_women_indicators_summary
#| tab.cap: "Summary of age structure indicators across post-peak female union spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_spread_post_peak_women")
    )
)
```

Cohen's d for median age across post-peak female union spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_women_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_post_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_women_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_post_peak_women,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Post-peak female union spread has very little influence under matrilocal-matrilineal residence, while broader schedules are associated with slightly younger populations under patrilocal-patrilineal residence (\(d=-0.14\)). The overall effect remains small enough that no general directional relationship should be inferred across residence systems.

#### `c1_men`: high *versus* low male union peak intensity regimes

```{r} 
#| label: age_structure_c1_men_pyramid_plot
#| fig.cap: "Population pyramids across male union peak intensity regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_intensity_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_intensity_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_c1_men_median_plot
#| fig.cap: "Median age of surviving populations across male union peak intensity regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_intensity_men"
)
```

```{r} 
#| label: age_structure_c1_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male union peak intensity regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_intensity_men")
    )
)
```

Cohen's d for median age across male union peak intensity regimes, by residence rule:

```{r} 
#| label: age_structure_c1_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_intensity_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_c1_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_intensity_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

Higher male union peak intensity produces younger populations under both residence systems, with small and nearly identical effects (\(d\approx-0.21\)). Male union intensity therefore contributes consistently, though modestly, to demographic renewal.

#### `mu_men`: late *versus* early male union timing regimes

```{r} 
#| label: age_structure_mu_men_pyramid_plot
#| fig.cap: "Population pyramids across male union timing regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_timing_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_timing_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_mu_men_median_plot
#| fig.cap: "Median age of surviving populations across male union timing regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_timing_men"
)
```

```{r} 
#| label: age_structure_mu_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male union timing regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_timing_men")
    )
)
```

Cohen's d for median age across male union timing regimes, by residence rule:

```{r} 
#| label: age_structure_mu_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_timing_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_mu_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_timing_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Earlier male union timing is associated with younger populations under both residence rules. The effect is small and slightly stronger under patrilocal-patrilineal residence (\(d=-0.38\)) than under matrilocal-matrilineal residence (\(d=-0.36\)). Male union timing is consequently one of the clearer secondary controls on age structure.

#### `sigma1_men`: broad *versus* narrow male pre-peak union spread regimes

```{r} 
#| label: age_structure_sigma1_men_pyramid_plot
#| fig.cap: "Population pyramids across male pre-peak union spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_spread_pre_peak_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_spread_pre_peak_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma1_men_median_plot
#| fig.cap: "Median age of surviving populations across male pre-peak union spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_spread_pre_peak_men"
)
```

```{r} 
#| label: age_structure_sigma1_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across male pre-peak union spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_spread_pre_peak_men")
    )
)
```

Cohen's d for median age across male pre-peak union spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma1_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_pre_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma1_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_pre_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Broader pre-peak male union spread produces younger surviving populations under both residence systems, but the effect is strongly residence-dependent. It is very small under matrilocal-matrilineal residence (\(d=-0.20\)) and small under patrilocal-patrilineal residence (\(d=-0.43\)). Earlier and more broadly distributed male union formation therefore has substantially greater demographic consequences under the latter residence rule.

#### `sigma2_men`: broad *versus* narrow post-peak male union spread regimes

```{r} 
#| label: age_structure_sigma2_men_pyramid_plot
#| fig.cap: "Population pyramids across post-peak male union spread regimes, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("union_spread_post_peak_men"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "union_spread_post_peak_men",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_sigma2_men_median_plot
#| fig.cap: "Median age of surviving populations across post-peak male union spread regimes, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "union_spread_post_peak_men"
)
```

```{r} 
#| label: age_structure_sigma2_men_indicators_summary
#| tab.cap: "Summary of age structure indicators across post-peak male union spread regimes and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "union_spread_post_peak_men")
    )
)
```

Cohen's d for median age across post-peak male union spread regimes, by residence rule:

```{r} 
#| label: age_structure_sigma2_men_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_post_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_sigma2_men_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ union_spread_post_peak_men,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Post-peak male union spread has effectively no influence on median age under either residence system. Effect sizes are close to zero and their confidence intervals include zero. Late-age variation in male union formation therefore contributes little to long-term age composition in v1.4.1.

#### `amenorrhea_period_in_days`: long *versus* short amenorrhea period

```{r} 
#| label: age_structure_amenorrhea_pyramid_plot
#| fig.cap: "Population pyramids across amenorrhea period lengths, by residence rule."
age_structures |>
  aggregate_age_structure(
    group_vars = c("amenorrhea_length"),
    smooth = FALSE
    ) |>
  prepare_pyramid_plot() |>
  plot_population_pyramid(
    facet_var = "amenorrhea_length",
    show_mean = TRUE
    )
```

```{r} 
#| label: age_structure_amenorrhea_median_plot
#| fig.cap: "Median age of surviving populations across amenorrhea period lengths, by residence rule."
plot_age_structure_indicators(
  age_structure_indicators, 
  x_var = "survival", 
  fill_var = "residence_rule", 
  facet_var = "amenorrhea_length"
)
```

```{r} 
#| label: age_structure_amenorrhea_indicators_summary
#| tab.cap: "Summary of age structure indicators across amenorrhea period lengths and residence rules."
knitr::kable(
  indicators_summary(
    age_structure_indicators, 
    group_vars = c("residence_rule", "amenorrhea_length")
    )
)
```

Cohen's d for median age across amenorrhea period lengths, by residence rule:

```{r} 
#| label: age_structure_amenorrhea_median_cohend_matri
#| tab.cap: "matrilocal-matrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ amenorrhea_length,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "matrilocal-matrilineal"
        )
    ),
    rules = "cohen1988"
  )
)
```

```{r} 
#| label: age_structure_amenorrhea_median_cohend_patri
#| tab.cap: "patrilocal-patrilineal"
knitr::kable(
  interpret_cohens_d(
    effectsize::cohens_d(
      median_age ~ amenorrhea_length,

    data =
        age_structure_indicators |>

        filter(
            residence_rule ==
                "patrilocal-patrilineal"
        )
  ),
    rules = "cohen1988"
  )
)
```

Longer postpartum amenorrhoea is associated with slightly older populations under both residence systems, with very small but consistent effects (\(d\approx0.18\)). The mean difference is about 1.2 years, reflecting slower juvenile recruitment as birth intervals lengthen. Postpartum amenorrhoea therefore acts as a modest regulator of reproductive tempo rather than a negligible one.

#### Interpretation

Fertility peak intensity remains the principal determinant of age structure in v1.4.1. Fertility spread, male union timing and pre-peak union spread, and union peak intensities provide secondary influences, while postpartum amenorrhoea contributes a smaller but consistent ageing effect through birth spacing.

Baseline mortality has virtually no direct association with median age among surviving populations. The different response variables therefore reveal a useful division of demographic roles: fertility and union formation primarily shape cohort composition, whereas mortality and ecological feedback are more consequential for survival and population pressure.

Demand weighting adds a direct connection between these demographic structures and ecological load, because the age composition generated by reproductive processes now determines the total carrying-capacity demand exerted by the population.

## Discussion notes

Version 1.4.1 combines daily demographic scheduling with age-weighted ecological demand. Population pressure is therefore determined not simply by the number of individuals but by the demand generated by their age composition, while births, deaths, union formation, gestation, and postpartum amenorrhoea unfold through the daily demographic cycle.

The model produces a broad range of end states, dominated by extinction and low-pressure persistence, with high-pressure persistence less frequent and overshooting rare. Under the favourable reference configuration, however, both residence systems converge on highly stable Stress-prone trajectories. Demand-weighted density dependence can therefore maintain populations close to ecological constraint without producing substantial long-run oscillation.

Fertility peak intensity remains the dominant demographic control. It strongly determines survival, population pressure, and median age. Density-effect strength provides the principal ecological counterweight, reducing realised pressure as feedback intensifies. Baseline mortality contributes to survival and pressure but has little direct effect on age structure among survivors.

Age weighting makes demographic composition ecologically consequential. The same reproductive mechanisms that determine juvenile recruitment and median age also alter total carrying-capacity demand, creating a direct pathway from reproductive schedules to ecological pressure.

Residence rule remains secondary at the aggregate level. Patrilocal-patrilineal populations show a small survival advantage in the global exploration, while matrilocal-matrilineal populations are slightly older on average. In the targeted trajectory experiment, however, both residence systems occupy virtually identical Stress-prone regimes. Residence effects are therefore conditional and are expressed more clearly through particular sex-specific union parameters than through a general difference in resilience.

The comparison with the yearly demand-weighted formulation indicates that temporal resolution changes the quantitative expression of demographic turnover and pressure, but the daily model should not be regarded simply as a smoother version of the yearly one. Explicit reproductive delays and event timing alter the pathway through which demographic demand accumulates, even though the principal fertility-driven architecture remains recognisable.

## Chapter conclusion

Version 1.4.1 establishes the daily demand-weighted density-regulated formulation of the Household Demography model. Fertility peak intensity remains the dominant demographic driver, while density-effect strength provides the principal ecological regulation of population pressure. Age-weighted demand adds an explicit compositional dimension by making ecological load depend on who comprises the population rather than on headcount alone.

Surviving populations are usually maintained below the nominal demand threshold, although favourable parameter combinations can sustain stable high-pressure states for centuries. Age structure remains governed primarily by fertility, union formation, and birth spacing, whereas baseline mortality and density regulation act more strongly on viability and ecological pressure.

The model therefore provides the appropriate daily demand-side foundation for v1.4.2, where productive labour is added to the supply side and effective carrying capacity becomes endogenous to demographic composition.
